### 2014-02-12 AO_bondedSi.tex
first commit of existing AO_bondedSi file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                         File: osa-revtex4-1.tex                             %
%                        Date: April 15, 2013                                 %
%                                                                             %
%                              BETA VERSION!                                  %
%                   JOSA A, JOSA B, Applied Optics, Optics Letters            %
%                                                                             %
%            This file requires the substyle file osajnl4-1.rtx,              %
%                   running under REVTeX 4.1 and LaTeX 2e                     %
%                                                                             %
%                   USE THE FOLLOWING REVTeX 4-1 OPTIONS:                     %
% \documentclass[osajnl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1}%
%                    %% Use 11pt for Applied Optics                           %
%                                                                             %
%               (c) 2013 The Optical Society of America                       %
%                                                                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[osajnl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1} %% use 11pt for Applied Optics
\documentclass[osajnl,preprint,showpacs,superscriptaddress,12pt]{revtex4-1} %% use 12pt for preprint option
\usepackage{amsmath,amssymb,graphicx}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{longtable}

\newcommand{\procspie}{{Proc. ~SPIE}}
\begin{document}

\title{Characterization of gaps in directly bonded Si compound optics}

\author{Michael Gully-Santiago}\email{gully@astro.as.utexas.edu}
\author{Daniel T. Jaffe}
\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX, 78712, USA}

\author{Victor White}
\affiliation{NASA Jet Propulsion Laboratory, Pasadena, CA}

\begin{abstract}
Si direct bonding offers flexibility in the design and development of large Si optics.  The bonding process presents challenges in meeting the performance demands and physical robustness, in particular for cryogenic and space applications.  Air cavities in between the Si interface will have large Fresnel reflections which will interfere as a low-Finesse Fabry-P\`erot etalon.  Even small 35 nm gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = $ 1.15 $\mu$m at normal incidence.  Stock near infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% photometric precision.  We can detect measure gaps as small as 8 nm by measuring transmission as a function of wavelength.  Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.  The detection and morphology of minuscule gaps provides valuable feedback in the design of Si direct bonding strategies.
\end{abstract}

\ocis{(030.1670) Coherent optical effects; (050.2230) Fabry-Perot; (120.2230)   Fabry-Perot; (120.2830) Height measurements; (120.4610) Optical fabrication; (120.7000) Transmission; (120.7000)   Transmission; (220.4840) Testing; (230.4170) Multilayers; (240.1485) Buried interfaces;  (300.6340) Spectroscopy, infrared; (310.6628)  Subwavelength structures, nanostructures}

\maketitle %% required

\section{Introduction}

Direct optical bonding of silicon allows manufacturers to combine parts with no Fresnel loss at the interface.  The high refractive index of Si makes this advantage particularly useful.  Traditional figures of merit for the optical audits of bonds, however, are either insensitive to gaps that are small yet have a significant impact on throughput and phase coherence, or do not provide quantitative information about the extent of the gaps along the direction of propagation.

Issues with the quality of the bonds become particular acute when the optics are used in double-pass and/or when the incidence angle at the bonds is far from the normal.

The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J, 2010SPIE.7739E.146W}.  These gratings are used either in transmission (i.e. as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e. as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W, 2012SPIE.8450E..2SG}.  These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B, 2010SPIE.7739E.123G}) applications.  To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.  The next generation of astronomy and earth science spectrographs, however, will need optical devices larger and more precise than what can be made with existing technology.  The proposed Giant Magellan Telescope\cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph\cite{2006SPIE.6269E.143J, 2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.  Substrates of the required size cannot fit into conventional processing equipment like reactive ion etchers and electron-beam lithography tools.  Large, thick Si substrates are also prohibitively expensive to experiment on.  Silicon direct bonding\cite{1986JAP....60.2987S, 2012SPIE.8450E..2TV} offers advantages by reducing the thickness and mass needed for the development of a thin (0.5$-$10 mm) Silicon diffractive optic that can later be stuck to a fine polished Si substrate, from which an immersion grating prism can be made.  Other groups producing SI diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).

While it is possible that proper precautions to eliminate all micron-sized particles from both SI surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments (\cite{doi:10.1080/02786828608959094}) means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.

There are two major concerns with Si direct bonding: the performance and robustness.  Specifically, what happens if the bonding process is imperfect and there are small gaps between the Si parts?  Does a gap degrade the performance of a final device, either through optical aberrations or transmission loss?  How mechanically robust is the final unit?  What is the strength of the bond?  Can the interface be broken by shearing, torquing, or pulling?  In practice, the figure of merit for robustness is whether the part can survive a shake test simulating the forcing experienced during a rocket launch or transport, or during cryogenic cycling, if the part will be used at low temperature.  Optically, the relevant parameters are the loss caused by the gap and the effective distortion of the transmitted wavefront.

We describe a technique for directly measuring gaps at silicon-silicon interfaces.  Other techniques for measuring and mapping gaps have already been described elsewhere\cite{1992JEMat..21..669M}, including IR interference, X-ray topography, and ultrasound microscopy.  The standard infrared interference technique is to examine the bond in reflection for the presence of Newton rings.  The presence and number of such rings can yield information about both the lateral and axial extend of any gap, as long as the gap along the direction of propagation is of the order of the vacuum wavelength or larger.  X-ray topography \cite{1992JEMat..21..669M, 1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provide only coarse estimates of the axial extent.  Ultrasound microscopy exploits the reflections at the gap interface to produce reflections of sound waves and map the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.

Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared wavelength range to measure gaps at the few nanometer size scale.  In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.  Because the size scales probed are so small, they are comparable to the range over which molecular (van der Waals) forces are acting.  It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore the optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.  By measuring both the spatial and axial extent of any gaps the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.

Our new technique for measurement of the axial extent of gaps exploits wavelength-dependence of reflection losses at the gap and the availability of commercial scanning monochromator systems with both high precision and good amplitudes stability.  We first describe the theory behind the technique and predicted outcomes for various gap dimensions.  We then evaluate the technique on a variety of bonded samples.

\section{History of the IR imaging technique}

Serious research on semiconductor wafer bonding started around 1989, with Lehmann et al. \cite{1989JaJAP..28L2141L} bonded GaAs and InP wafers to Si.  They detected no gaps with IR imaging.  They measured the surface energy of the bond and found a range of 40$-$600 dyne/cm.

See the table below for a list of papers on Si wafer bonding and a note on their contribution.

\begin{spacing}{1.0}

\begin{center}
\begin{longtable}{ |c | p{8cm} | p{8cm} |}
%\caption{History of Si direct bonding and metrology \label{tab:litbonding}}

%    \begin{tabular}{ |c | p{8cm} | p{8cm} |}
    \hline
    Ref & Title & Notes \\ 
        \hline
	\cite{1986JAP....60.2987S} &  \emph{Silicon-to-silicon direct bonding method} &  Foundation paper  \\
         \hline
     \cite{1988JaJAP..27L2364S} &  \emph{Bubble-Free Silicon Wafer Bonding in a Non-Cleanroom Environment} &  Sequence of IR images showing bond propagation  \\
         \hline
     \cite{1989JaJAP..28L2141L} &  \emph{Bubble-Free Wafer Bonding of GaAs and InP on Silicon in a Microcleanroom} &  No gaps detected with IR imaging; measured surface energy of the bond 40$-$600 dyne/cm.  \\
         \hline
     \cite{Mitani1990} &  \emph{Bubble formation during silicon wafer bonding: causes and remedies} &  Bubble elimination at temperature greater than 800 C. \\
         \hline
	\cite{1992JEMat..21..669M} &  \emph{Wafer bonding technology for silicon-on-lnsulator applications: A review} &  Excellent paper.  Shows infrared imaging detected bubbles and the bubble density as a function of anneal temperature.  Directly compares the three bubble detection techniques\\
         \hline
     \cite{feijoo1994} &  \emph{Silicon wafer bonding studied by infrared absorption spectroscopy} &  Works out the chemistry of the interface, finds distinct temperature phases, hydrogen bonds above   \\          
         \hline
     \cite{1994JaJAP..33....6H} &  \emph{Silicon Wafer Direct Bonding without Hydrophilic Native Oxides} &  Shows x-ray topograph technique, hydrophobic is strong bond\\
              \hline         
     \cite{1995ApPhA..61..101R} &  \emph{Multiple internal reflection spectroscopy of bonded silicon wafers} &  Chemistry of the bond interface- Evidence for Si$-$H bonds\\
              \hline
     \cite{1995ApPhL..67.3614G} &  \emph{Self-propagating room-temperature silicon wafer bonding in ultrahigh vacuum} &  IR image of wafers bonding with gaps visible  \\
              \hline     
     \cite{1995ApPhL..67..863G} &  \emph{What determines the lateral bonding speed in silicon wafer bonding?} &  Spread speed is 10 mm/s, decreases with increasing ambient pressure; thickness has no effect.  \\
              \hline
     \cite{1996ApPhL..68.2222T} &  \emph{Surface activated bonding of silicon wafers at room temperature} &  Shows that Ar etching makes stronger bonds, shows no void from IR microscopy or KOH etching  \\     
              \hline
     \cite{1996JaJAP..35.2102R} &  \emph{Characterization of Interfaces of Directly Bonded Silicon Wafers: A Comparative Study of Secondary Ion Mass Spectroscopy Multiple Internal Reflection Spectroscopy, and Transmission Electron Microscopy} &  They figure out what is the chemistry participating in Si bond formation\\               
              \hline
     \cite{1998AnRMS..28..215G} &  \emph{Semiconductor wafer bonding} &  Annual Reviews article summarizing the state of the subfield\\ 
              \hline
     \cite{JJAP.37.4197} &  \emph{Effect of Surface Roughness on Room-Temperature Wafer Bonding by Ar Beam Surface Activation} &  Appears to only explore surface roughness of less than 1 nm RMS, which seems low compared to our wafers.\\ 
              \hline
     \cite{2000JAP....88.4404H} &  \emph{Modeling the dynamics of Si wafer bonding during annealing} &  Prescription for how long to run annealing experiments\\     
              \hline
     \cite{2000RScI...71.1869G} &  \emph{Photoacoustic imaging of voids in direct wafer bonding} &  Compares photoacoustic technique to the IR technique \\     
              \hline
     \cite{2001JOptA...3...85G} &  \emph{Optical contact and van der Waals interactions: the role of the surface topography in determining the bonding strength of thick glass plates} &  What is the effect of surface roughness on optical bonding; uses 1 nm roughness.\\
              \hline
     \cite{2001JAP....89.6013L} &  \emph{Modeling of hydrophilic wafer bonding by molecular dynamics simulations} &  Examines the effect of moisture and defects in forming bonds\\ 
          \cite{2007ApOpt..46.6793H} &  \emph{Direct bonding and beyond} &  Excellent paper.  Demonstrates Si direct bonding on thick substrates.\\

    \hline
%    \end{tabular}
\end{longtable}
\end{center}

\end{spacing}

\section{Silicon-gap-silicon Fabry-P\`{e}rot theory}

The measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.  The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm\cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.  The Fresnel reflection is accordingly about 30\% per surface.  Popular optical materials with refractive indices in the range 1.5$-$1.7 with Fresnel reflection 4$-$7\% per surface will suffer less than Si from the effects described here.

Consider the following two conceivable pitfalls in Si direct bonding.  First, particles at the tens of nanometer size scale are trapped between the Si pucks during bonding.  This potential problem is especially acute for silicon direct bonding compared to bonding of popular low-index optical materials.  A typical particle surface flux is shown in Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094}.  The flux of particles is a step function of particle size.  For every micron sized particle causing Newton's rings, there will be as many as 100,000 times more ten-nanometer sized particles.

The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap. Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197, 2001JOptA...3...85G}.  The details of this aspect of the bonding process are outside the scope of this paper.

\section{Substrate preparation}
The substrates were silicon wafers, some had unknown pedigree.  The surface roughness was typically about 2 nm, as measured with a Wyko 2D noncontact microscope.  The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will be more conformal than thick ones.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a peak to valley surface flatness of 2 waves over the central 2 inch diameter.

The detailed bonding procedure will be chronicled in a future paper.  Briefly, we prepared the substrates with standard cleaning procedures.  We used MHz frequency oxygen plasma ashing.  We soaked the wafers in DI water then dried with an N$_2$ gun.  We pressed the wafers together from the center.

\section{Imbedding gaps of known sizes}
We evaluated our metrology method by imbedding gaps of known sizes in the interface.  We bored divots in silicon with inductively coupled plasma etching.  The tool was a Unaxis.  We used two different gases with low and high etch rates.  CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.  The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.  We protected the unexposed areas with thick photoresist.  We achieved etch depths of 20$\pm$5 nm and 4000$\pm$100 nm, as measured with Wyko 2D noncontact white light microscopy for the small depth, and Dektak stylus profilometry for the large depth.

Table \ref{tab:bondexper} lists part numbers and descriptions for our bonded substrates.

\begin{table}[h!]
\caption{UTexas Si bonding experiments \label{tab:bondexper}}
\begin{center}
    \begin{tabular}{ c c c c c c c}
    \hline
    Name & Bonded/DSP & Diameter & Thickness & Anneal Temp & Date & Pattern notes \\ 
    - & - & mm & mm & $^\circ$C & MM/DD/YYY & - \\ 
        \hline
    VG01 & Bonded  & 100  &$1$ $\times$ 2 & 23  & 2/13/2013 &  \\
    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  23  & 2/13/2013 &  \\
    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  23 & 2/14/2013 & 4 $\mu$m hole \\    
    VG04 & Bonded & 100 & $1$ $\times$ 2 &  23 & 2/14/2013 & 15 nm gap \\        
    VG05 & DSP & 100 & 1&  23 & 2/15/2013 & cleaved  \\
    VG06 & DSP & 100 & 3.5 &  23 & 2/15/2013 & \\
    VG07 & - & - & - &  - & - & \\
    VG08 & - & - & - &  - & - & \\
    VG09 & Bonded & 75 & 1.5? & 23 & 6/21/2013 & mesh C, 49 $\pm$6 nm \\
    VG15 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 14 $\pm$2 nm\\
    VG17 & Bonded & 75 & 6.0? &  23 & 6/21/2013 & mesh C, 64 $\pm$6 nm\\
    VS20 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 95 $\pm$5 nm\\
    \hline
    \end{tabular}
\end{center}
\end{table}

\section{Transmission measurement theory}
The Si-air interfaces of the air gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.  The transmission through a Fabry-Perot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the index of air is 1.0, and optics are at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si-air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}.  The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  The coefficient of Finesse\cite{2007fuph.book.....S} encapsulates the Fabry-P\`erot etalon's dependence on the reflectivity.  We compute a coefficient of Finesse in the range of about 2.8$\pm$0.1 depending on the wavelength.

\subsection{Consideration of multiple incoherent reflections}
Multiple incoherent reflections occur within the Si substrate.  Treatment for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  I adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with an air gap thickness $d$.  We normalize the transmission with a gap to the transmission with no gap to isolate the effect of the gap.  We call this normalized transmission the etalon transmission $T_{e}$.  It is derived in Equation \ref{eqn:Tetalon} in the Appendix:

\begin{eqnarray}
T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \nonumber
\end{eqnarray}

It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.  In the next section we compare measurements to by-eye fits of this equation.

\section{Measurements}
The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \url{cary5000web}.  The typical measurement parameters and tool performance are listed below.  More details about the tool are available at the vendor's website.  Briefly, the tool has a double monochromator.  The double monochromator feature reduces scattered light, with the vendor claiming 40 dB of linearity.  We independently checked the linearity in the following way.  We separately measured the transmission spectrum of 5 neutral density filters.  We computed the mean transmission of each filter, predicting the net transmission of a combination of filters up to all five filters.  We then measured the net spectrum and compared to prediction.  The measured mean transmission of the stacks were indistinguishable from the predicted mean transmission within the measurement uncertainty over 24 dB.

The Cary 5000 has four operation modes: single front, single back, double, and double reverse.  We experimented with single front and double modes.  We estimated the uncertainty and measurement repeatability of single and double modes in the following way.  We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a the beam with no sample in a series of spectra taken immediately after a 100\% baseline scan.  The predicted mean level should naturally be 100\%.  We found that the mean value in single mode drifted by about 0.2\%.  The standard deviation of the featureless spectrum was typically 0.02\%, with some portions of the spectrum demonstrating increased uncertainty, perhaps attributable to atmospheric absorption. In double beam mode the spectrograph performance was much better.  We put a double side polished Si wafer in the reference beam, and no sample in the target beam during the baseline scan.  In this way the transmission spectrum of a target Si wafer is automatically normalized.  The double beam mode effectively divides out variations in the mean transmission levels attributable to lamp drift at 0.2\% in single beam measurements.  The drift in mean transmission level is merely 0.02\% in double beam mode.  These values are accurate for measurements over a half hour timescale or less.  Over periods of hours the drift increases for both single and double beam mode, so a repeat baseline scan is sometimes necessary.  The need for repeat baselines is more acute for single mode.

The choice of spectral sampling was straightforward.  We expect smoothly varying spectral features for small ($<$ 1 $\mu$m) size gaps in Si bonded wafers.  Sampling theory dictates a few spectral resolution elements over the spectral feature of interest.  The Fabry-P\`erot fringes will be well-sampled so long as the gap is less than $\sim$100 $\mu$m.  Such large gaps would be easily detectable by other means.  So we simply pick the largest slit width, and therefore the lowest spectral resolution that the Cary 5000 permits: 5 nm.  We then sample the spectral resolution with $\sim$2.5 samples per spectral resolution element, 2.0 nm samples.

\begin{table}[h!]
\caption{Summary of Cary 5000 Measurement parameters}
\begin{center}
    \begin{tabular}{ c c }
    \hline
    Parameter (Units) & Value \\ 
        \hline
       Spectral sampling interval (nm) & 2.0 \\
       Spectral resolution (nm) & 5 \\
       Time per sample (s) & 0.1 \\
       Available measurement range (nm) & 175$-$3300 \\
       Typical measurement range (nm) & 1000$-$2500 \\
	Single mode RMS  (\%) & 0.02 \\
	Double mode RMS  (\%) & 0.002 \\
	Single mode mean drift (\%) & 0.20 \\
	Double mode mean drift (\%) & 0.02 \\
	Mean 0\% baseline correction (\%) & 0.03 \\
	Beam size (mm $\times$ mm) & 2 $\times$8 \\
	\multirow{2}{*}{Wavelengths of instrumental artifacts (nm)} &  \\
      &1200 \\
      &1350 \\
      &1360$-$1450\\
      &1850 \\
      &1800$-$1950 \\
      &2000 \\
      Vendor-reported linearity range (dB) & 40 \\
      Directly measured linearity range (dB) & $>24$ \\
    \hline
    \end{tabular}
\end{center}
\end{table}

We also looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with the coarse IR imaging technique.  The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.  We used a 50 mm F/2 lens, with a fiber optic illumination source.  The lens was spaced from the detector plane by about XX mm, with the illumination source at a distance XXX from the lens.  The illumination source was gently diverging.  The result was a vignetted beam.  The target was located a distance XXX from the lens to provide imaging of the Si interface, with a field of view of roughly X $\times$ X degrees, or X $\times$ X mm.  This field of view sub-sampled the 4 inch diameter wafer, so we dithered the sample, collecting images.

%\begin{figure}[htbp]
%\centerline{\includegraphics[width=.8\columnwidth]{IMG_7662.JPG}}
%\caption{Photo of part VG15 with locations of defects marked.}
%\end{figure}

\section{Results}

\appendix
\section{Incoherent Multiple Reflections Transfer Matrix Method}
\label{sec:Append-IMRTMM}

The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.  Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
\begin{eqnarray}
\boldsymbol{S} = \left(
\begin{array}{cc}
 t_{12} & r_{21} \\
 r_{12} & t_{21} \\
\end{array}
\right)
\end{eqnarray}
Where $t$ and $r$ stand for transmission and reflection respectively.  The order of subscripts is the order of origin and destination of the wave in regards to the interface, so we know something about the direction of the wave and how it got there.  The $\boldsymbol{S}$ matrix encapsulates all the information we need to know about how light waves interact with the interface.  The real power of the matrix technique comes from a cousin of the scattering matrix, the so-called wave transfer matrix, $\boldsymbol{M}$.  The matrix $\boldsymbol{M}$ has the convenient property that its output can be used as the input for another matrix.  In other words, the input vector to $\boldsymbol{M}$ is made of the left and right moving components directly before the interface; the outputs are the left and right moving components directly after the interface:
\begin{eqnarray}
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)=\boldsymbol{S} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right) \\
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right)=\boldsymbol{M} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)
\end{eqnarray}

The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\cite{2007fuph.book.....S}.  Much of the literature on the transfer matrix method deals with applications related to thin films, in which the wavelength is comparable to the size of the dielectric layer.  For thin films the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.  The polarization state can be encapsulated in scattering matrix components\cite{2007fuph.book.....S}.  The intensities of the emergent spectrum can be computed from the absolute square of the complex amplitudes.  For thick films, the vector components are the intensities of the emergent spectrum, since waves are incoherent.  \cite{2002ApOpt..41.3978K} work out the general transfer-matrix method for optical multilayer systems with incoherent interference.  The key idea for the incoherent transfer matrix method approach is to populate a scattering matrix with elements equal to the (wavelength dependent) transmitted $T_i$ and reflected $R_i$ intensities of an interface or set of interfaces that act together.  Then, use the geometric transformations to construct the $\boldsymbol{M}$ matrix.

The matrix for the Fabry-P\`erot etalon is calculated in the following way.  The first key idea is that I had to treat the entire air gap as an abstraction, a Fabry-P\`erot etalon.  We don't need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon.

\begin{eqnarray}
 \delta = \frac{2\pi}{\lambda}2d \\
  F \equiv \frac{4R}{(1-R)^2} \\
 T_e = \frac{1}{1+F\sin^2(\delta/2)}  \label{eq:FabPerot}
\end{eqnarray}

with $\lambda$ the vacuum wavelength, $R$ the Fresnel reflection of silicon, and $F$ the coefficient of finesse.  The coefficient of finesse $F$ and the phase $\delta$ are the two parameters of the Fabry-P\`erot etalon.  The coefficient of Finesse encapsulates the Fresnel reflection and depends only on the Si refractive index which has a minute wavelength (and temperature) dependence.  The phase depends on the wavelength $\lambda$ and $d$ the air gap spacing: $\delta=\frac{2\pi}{\lambda}2d$ .  We assume the etalon is lossless, i.e. the etalon has no absorption $T_e+R_e=1$.  So the incoherent scattering and transfer matrices for the etalon are:

\begin{eqnarray}
\boldsymbol{S_e}&=&\frac{1}{1+F\sin^2{\delta/2}} \left(
\begin{array}{cc}
1 & F \sin ^2(\delta/2) \\
F \sin ^2(\delta/2) & 1 \\
\end{array}
\right) \nonumber \\
\nonumber \\
\boldsymbol{M_e}&=&\left(
\begin{array}{cc}
 1-F \sin ^2(\delta/2) & F \sin ^2(\delta/2) \\
 -F \sin ^2(\delta/2) & 1+F \sin ^2(\delta/2) \\
\end{array}
\right)
\label{eqn:EtalonMatrix}
\end{eqnarray}

I assembled the matrix for the Air-Si Fresnel interface in the following way.  First it is important to note that the matrix is the same whether the transmission is from Si to air or air to Si.  This reciprocity is not necessarily true for the complex amplitudes matrix, but our approach employs intensities not complex amplitudes.  The Fresnel interface is lossless.  The transmission and reflection are given by the Fresnel equation for normal incidence:
\begin{eqnarray}
T_n&=&\frac{4n_{Si}}{(n_{Si}+1)^2} \\
R_n&=&\frac{(n_{Si}-1)^2}{(n_{Si}+1)^2} \label{eq:FresnelTrans}
\end{eqnarray}
For clarity I will drop the subscripts from $n_{Si}$, since we have already set $n_{air}=1$ and there are no other dielectric interfaces to think about.  So the scattering and transfer matrices for the air-Si Fresnel boundary are:

\begin{eqnarray}
\boldsymbol{S_n}&=&\frac{1}{(n+1)^2} \left(
\begin{array}{cc}
4n & (n-1)^2 \\
(n-1)^2 & 4n \\
\end{array}
\right)  \nonumber \\
\nonumber \\
\boldsymbol{M_n}&=&\frac{1}{4n}\left(
\begin{array}{cc}
 -n^2+6  n-1 & ( n-1)^2 \\
 -( n-1)^2 & ( n+1)^2 \\
\end{array}
\right)
\label{eqn:SiAirMatrix}
\end{eqnarray}

Finally, we cascade the matrices together to compute the net transmission through the stack of abstractions.  The result is a $2\times2$ transfer matrix: $$\boldsymbol{M_{net}}=\boldsymbol{M_n}\boldsymbol{M_e}\boldsymbol{M_n}$$  From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.  Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:

\begin{eqnarray}
T_{net}=\frac{2 n}{1+ 2n F\sin ^2(\delta/2)+n^2} \label{eqn:FPmatTrans}
\end{eqnarray}

Compare equations \ref{eqn:FPmatTrans} and \ref{eq:FabPerot}.  The revised transmission has picked up a few factors of 2 and $n$.  While we are at it, let's compute the matrix for the scenario with no intermediate etalon: there are simply two Fresnel interfaces with which light interacts incoherently.  This scenario is the model for a single DSP reference puck. The revised matrix multiplication is simply $\boldsymbol{M_{DSP}}=\boldsymbol{M_n}\boldsymbol{M_n}$.  Taking the inverse of the $M_{22}$ element, we find: 
\begin{eqnarray}
T_{DSP}=\frac{2 n}{n^2+1}\label{eqn:EqofSummedSlab}
\end{eqnarray}
which is identical to the result obtained by directly summing the intensities from multiple reflections:
\begin{eqnarray}
T_{DSP}=T^2 \sum_{i=0}^{N}R^{2i} \label{eqn:multsum}
\end{eqnarray}

It is informative to isolate the effect of the gap by dividing the measured bonded wafer transmission by the transmission of a reference DSP Si part.  We call this normalized transmission the etalon transmission $T_{e}$:
\begin{eqnarray}
T_{e} = T_{net}/T_{DSP} \\
T_{e} = \frac{n^2+1}{2 n F \sin ^2(\delta/2)+n^2+1} \label{eqn:Tetalon}
\end{eqnarray}

\bibliographystyle{osajnl}
\bibliography{AO_bondedSi}

%\begin{thebibliography}{99}

%% Do not include separate BibTeX files; if BibTeX is used,
%% paste the output (contents of .bbl file) here.

%\end{thebibliography}

\end{document}

### 2014-02-13 AO_bondedSi.tex
add prose on IR imaging

@@
 
-The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J, 2010SPIE.7739E.146W}.
+The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J,2010SPIE.7739E.146W}.
 These gratings are used either in transmission (i.e.
-as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e.
-as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W, 2012SPIE.8450E..2SG}.
-These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B, 2010SPIE.7739E.123G}) applications.
+as grisms \cite{2008SPIE.7014E..77D,2010SPIE.7739E.123G}), or in reflection (i.e.
+as immersion gratings, \cite{2007ApOpt..46.3400M,2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.
+These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y,2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.
 To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.
 The next generation of astronomy and earth science spectrographs, however, will need optical devices larger and more precise than what can be made with existing technology.
-The proposed Giant Magellan Telescope\cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph\cite{2006SPIE.6269E.143J, 2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.
+The proposed Giant Magellan Telescope \cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph \cite{2006SPIE.6269E.143J,2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.
 Substrates of the required size cannot fit into conventional processing equipment like reactive ion etchers and electron-beam lithography tools.
 Large, thick Si substrates are also prohibitively expensive to experiment on.
-Silicon direct bonding\cite{1986JAP....60.2987S, 2012SPIE.8450E..2TV} offers advantages by reducing the thickness and mass needed for the development of a thin (0.5$-$10 mm) Silicon diffractive optic that can later be stuck to a fine polished Si substrate, from which an immersion grating prism can be made.
+Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by reducing the thickness and mass needed for the development of a thin (0.5$-$10 mm) Silicon diffractive optic that can later be bonded to a fine polished Si prism substrate.
 Other groups producing SI diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
 
-While it is possible that proper precautions to eliminate all micron-sized particles from both SI surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments (\cite{doi:10.1080/02786828608959094}) means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
+While it is possible that proper precautions to eliminate all micron-sized particles from both SI surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments (\cite{Cooper_1986}) means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
+Small particles could be a concern if they act as "tent poles", ti create gaps where the bonding is incomplete.
 
@@
 What is the strength of the bond?
-Can the interface be broken by shearing, torquing, or pulling?
+Can the interface be broken by shearing, torquing, pulling, or thermal stress?
 In practice, the figure of merit for robustness is whether the part can survive a shake test simulating the forcing experienced during a rocket launch or transport, or during cryogenic cycling, if the part will be used at low temperature.
@@
 We describe a technique for directly measuring gaps at silicon-silicon interfaces.
-Other techniques for measuring and mapping gaps have already been described elsewhere\cite{1992JEMat..21..669M}, including IR interference, X-ray topography, and ultrasound microscopy.
+Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}, including IR interference, X-ray topography, and ultrasound microscopy.
 The standard infrared interference technique is to examine the bond in reflection for the presence of Newton rings.
 The presence and number of such rings can yield information about both the lateral and axial extend of any gap, as long as the gap along the direction of propagation is of the order of the vacuum wavelength or larger.
-X-ray topography \cite{1992JEMat..21..669M, 1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provide only coarse estimates of the axial extent.
-Ultrasound microscopy exploits the reflections at the gap interface to produce reflections of sound waves and map the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
+X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.
+Ultrasound microscopy exploits the density change at the gap interface to produce reflections of sound waves and map the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
 
-Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared wavelength range to measure gaps at the few nanometer size scale.
+Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared wavelength range to measure the axial extent of gaps at the few nanometer size scale.
 In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.
@@
 It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore the optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.
-By measuring both the spatial and axial extent of any gaps the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
+By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
 
-Our new technique for measurement of the axial extent of gaps exploits wavelength-dependence of reflection losses at the gap and the availability of commercial scanning monochromator systems with both high precision and good amplitudes stability.
+Our new technique for measurement of the axial extent of gaps exploits wavelength-dependence of reflection losses at the gap and the availability of commercial scanning monochromator systems with both high precision and good amplitude stability.
 We first describe the theory behind the technique and predicted outcomes for various gap dimensions.
-We then evaluate the technique on a variety of bonded samples.
+We then evaluate the technique experimentally on a variety of bonded samples.
 
-\section{History of the IR imaging technique}
+\section{History of Si direct bonding}
+\ref{secHistory}
 
@@
 
+The majority of work on bonding was carried out from 1990-2000.
+The key advances included studies that probed the surface energy as a function of anneal temperature, and various probes of the bond chemistry through multiply reflection infrared spectroscopy and mass spectroscopy.
+A fairly complete story was then constructed which tied together all the disparate measurements.
+No where in the literature did we find a technique similar to the one presented here.
 See the table below for a list of papers on Si wafer bonding and a note on their contribution.
 
+%\caption{Literature review}
 \begin{spacing}{1.0} 
-
 \begin{center}
 \begin{longtable}{ |c | p{8cm} | p{8cm} |}
-%\caption{History of Si direct bonding and metrology \label{tab:litbonding}}
-
-%    \begin{tabular}{ |c | p{8cm} | p{8cm} |}
     \hline
@@
           \cite{2007ApOpt..46.6793H} &  \emph{Direct bonding and beyond} &  Excellent paper.  Demonstrates Si direct bonding on thick substrates.\\ 
-
     \hline
-%    \end{tabular}
 \end{longtable}
 \end{center}
-
 \end{spacing}
@@
 
-The measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm\cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
+The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
 The Fresnel reflection is accordingly about 30\% per surface.
@@
 This potential problem is especially acute for silicon direct bonding compared to bonding of popular low-index optical materials.
-A typical particle surface flux is shown in Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094}.
+A typical particle surface flux is shown in Figure 5 of Cooper 1986\cite{Cooper_1986}.
 The flux of particles is a step function of particle size.
-For every micron sized particle causing Newton's rings, there will be as many as 100,000 times more ten-nanometer sized particles.
+For every micron sized particle causing Newton's rings, there will be as many as 100,000 more ten-nanometer sized particles.
 
 The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197, 2001JOptA...3...85G}.
+Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.
 The details of this aspect of the bonding process are outside the scope of this paper.
-
-\section{Substrate preparation}
-The substrates were silicon wafers, some had unknown pedigree.
-The surface roughness was typically about 2 nm, as measured with a Wyko 2D noncontact microscope.
-The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
-We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
-The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will be more conformal than thick ones.
-There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
-We measured the large scale surface flatness of the pucks.
-The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.
-We found a peak to valley surface flatness of 2 waves over the central 2 inch diameter.
-
-The detailed bonding procedure will be chronicled in a future paper.
-Briefly, we prepared the substrates with standard cleaning procedures.
-We used MHz frequency oxygen plasma ashing.
-We soaked the wafers in DI water then dried with an N$_2$ gun.
-We pressed the wafers together from the center.
 
@@
 \begin{table}[h!]
-\caption{UTexas Si bonding experiments \label{tab:bondexper}}
+%\caption{UTexas Si bonding experiments \label{tab:bondexper}}
 \begin{center}
@@
 \section{Transmission measurement theory}
-The Si-air interfaces of the air gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.
+The Si-air\footnote{We imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.} interfaces of the in the bond gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.
 The transmission through a Fabry-Perot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We assume transmission is at normal incidence, the index of air is 1.0, and optics are at room temperature.
+We assume transmission is at normal incidence, the index of the gap is 1.0, and the material is at room temperature.
 We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.
 We compute the reflectivity for an Si-air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}.
+The computation is simply Fresnel's law at normal incidence:
+
+$$
+R = \frac{(n-1)^2}{(n+1)^2}
+$$
+
 The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
+The modest wavelength dependence is shown in the bottom curve of the figure below.
 The coefficient of Finesse\cite{2007fuph.book.....S} encapsulates the Fabry-P\`erot etalon's dependence on the reflectivity.
-We compute a coefficient of Finesse in the range of about 2.8$\pm$0.1 depending on the wavelength.
+We compute a coefficient of Finesse in the range of about 2.7$-$2.9 depending on the wavelength.
 
@@
 The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with an air gap thickness $d$.
-We normalize the transmission with a gap to the transmission with no gap to isolate the effect of the gap.
+We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.
+The DSP Si puck with no air gap has a transmission equal to:
+
+$$
+T_{DSP} = \frac{2n}{1+n^2}
+$$
+
+which has an average value of about 53\%, as shown in the top curve of the Figure above.
+Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
+If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+We work out this behavior in the Appendix section X and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
+
+For the case of a pair of bonded Si pucks with a gap thickness $d$, we normalize the transmission with a gap to the transmission with no gap to isolate the effect of the gap.
 We call this normalized transmission the etalon transmission $T_{e}$.
@@
 
-\begin{eqnarray}
-T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \nonumber
-\end{eqnarray}
+$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \nonumber $$
 
@@
 
-\section{Measurements}
-The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \url{cary5000web}.
+\section{Substrate preparation}
+It is important to clean the wafers to minimize the particle density, since particles cause gaps.
+The substrates were silicon wafers.
+The surface roughness was typically about 2 nm, as measured with a Wyko 2D noncontact microscope.
+The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
+We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
+The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
+There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+We measured the large scale surface flatness of the pucks.
+The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.
+We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
+
+The detailed bonding procedure will be chronicled in a future paper.
+Briefly, we prepared the substrates with standard cleaning procedures.
+We used MHz frequency oxygen plasma ashing.
+We soaked the wafers in DI water then dried with an N$_2$ gun.
+We pressed the wafers together from the center.
+
+\section{Measurement strategy, conditions, and details}
+The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \nocite{cary5000web}.
 The typical measurement parameters and tool performance are listed below.
@@
 Briefly, the tool has a double monochromator.
-The double monochromator feature reduces scattered light, with the vendor claiming 40 dB of linearity.
-We independently checked the linearity in the following way.
-We separately measured the transmission spectrum of 5 neutral density filters.
-We computed the mean transmission of each filter, predicting the net transmission of a combination of filters up to all five filters.
-We then measured the net spectrum and compared to prediction.
-The measured mean transmission of the stacks were indistinguishable from the predicted mean transmission within the measurement uncertainty over 24 dB.
+The double monochromator feature reduces scattered light, with the vendor reported linearity exceeding 40 dB.
 
@@
 We experimented with single front and double modes.
+Double mode takes a reference spectrum simultaneously to the target spectrum.
+In both single and double beam mode a baseline scan is taken of air.
+The baseline spectrum is the raw spectral response of the system, which is automatically divided out of subsequent spectra.
+In single beam mode the baseline measurements must be repeated about every half hour.
 We estimated the uncertainty and measurement repeatability of single and double modes in the following way.
-We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a the beam with no sample in a series of spectra taken immediately after a 100\% baseline scan.
+We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of the beam with no sample in a series of spectra taken immediately after a 100\% baseline scan.
 The predicted mean level should naturally be 100\%.
@@
 
-\begin{table}[h!]
-\caption{Summary of Cary 5000 Measurement parameters}
+Table \ref{tab:tablenum3} shows some properties of the Cary5000.
+
+%\caption{Summary of Cary 5000 Measurement parameters \label{tab:tablenum3}}
 \begin{center}
-    \begin{tabular}{ c c }
-    \hline
+    \begin{tabular}{l l}
         Parameter (Units) & Value \\ 
-        \hline
         Spectral sampling interval (nm) & 2.0 \\
@@
         Time per sample (s) & 0.1 \\
-       Available measurement range (nm) & 175$-$3300 \\
-       Typical measurement range (nm) & 1000$-$2500 \\
-	Single mode RMS  (\%) & 0.02 \\
-	Double mode RMS  (\%) & 0.002 \\
-	Single mode mean drift (\%) & 0.20 \\
-	Double mode mean drift (\%) & 0.02 \\
-	Mean 0\% baseline correction (\%) & 0.03 \\
-	Beam size (mm $\times$ mm) & 2 $\times$8 \\
-	\multirow{2}{*}{Wavelengths of instrumental artifacts (nm)} &  \\
-      &1200 \\
-      &1350 \\
-      &1360$-$1450\\
-      &1850 \\
-      &1800$-$1950 \\
-      &2000 \\
+        Available measurement range (nm) & 175-3300 \\
+        Typical measurement range (nm) & 1000-2500 \\
+	Single mode RMS  (percent) & 0.02 \\
+	Double mode RMS  (percent) & 0.002 \\
+	Single mode mean drift (percent) & 0.20 \\
+	Double mode mean drift (percent) & 0.02 \\
+	Mean 0\% baseline correction (percent) & 0.03 \\
+ 	Beam size (mm x mm) & 2 x 8 \\
+	Wavelengths of instrumental artifacts (nm) & stuff \\
+        num1 & 1200 \\
+        num2 & 1350 \\
+        num3 & 1360-1450\\
+        num4 & 1850 \\
+        num5 & 1800-1950 \\
+        num6 & 2000 \\
         Vendor-reported linearity range (dB) & 40 \\
-      Directly measured linearity range (dB) & $>24$ \\
-    \hline
+        Directly measured linearity range (dB) & 24 \\
       \end{tabular}
 \end{center}
-\end{table}
 
@@
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
-We used a 50 mm F/2 lens, with a fiber optic illumination source.
-The lens was spaced from the detector plane by about XX mm, with the illumination source at a distance XXX from the lens.
-The illumination source was gently diverging.
-The result was a vignetted beam.
-The target was located a distance XXX from the lens to provide imaging of the Si interface, with a field of view of roughly X $\times$ X degrees, or X $\times$ X mm.
-This field of view sub-sampled the 4 inch diameter wafer, so we dithered the sample, collecting images.
+We used a 50 mm F/2 lens, with a fiber optic white light illumination source.
+We estimate that the lamp-detector combination was sensitive to wavelengths in the range 1.1-1.6 $\mu$m.
+The field of view sub-sampled the 4 inch diameter wafers, so we dithered the sample.
+No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.
+It was easy to detect IR bubbles despite the artifacts since the bubbles translate with the wafer during dithering.
+We detected IR bubbles in all 3 of the bonded Si pucks on which we recorded IR images.
+The bubble areal density varied from about 4 per wafer (100 mm diameter) to over 20 per wafer.
+Naturally, the sample of control solitary DSP (unbonded) wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrtes as discussed in the literature (see Section \label{secHistory}).
+Figure \ref{IRbubble} shows an example of one of the IR bubbles in which several bright and dark fringes are clearly discernible.
 
-%\begin{figure}[htbp]
-%\centerline{\includegraphics[width=.8\columnwidth]{IMG_7662.JPG}}
-%\caption{Photo of part VG15 with locations of defects marked.}
-%\end{figure}
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.8\columnwidth]{figs/IR_bubble.pdf}}
+\caption{Photo of part VG09 with transmission image of a bubble highlighted\label{IRbubble}.
+The right side photo shows the 75 mm diameter wafer with blue Sharpie rectangles indicating the position of different spectral measurements (as detailed in the next section).
+The yellow circle in the photo highlights the outline of the location of an IR bubble we detected.
+The bubble shows a couple of fringes over the effective wavelength range 1.1-1.6 $\mu$m.
+This fringe pattern is consistent with a gap comparable to this measurement wavelength.
+The pointlike measurement artifacts, Sharpie marker, and vignetting are perceptible in the raw IR image.
+Our interpretation of the IR bubbles is robust against these artifacts.
+The area surrounding the IR bubble appears to be mostly devoid of large (wavelength-scale) gaps, since the transmission is fairly uniform, and consistent with the brightest fringe of the IR bubble.
+}
+\end{figure}
 
-\section{Results}
+\section{Results of infrared spectroscopy of bonded Si pucks}
 
 \appendix
+
+\section{Appendix: X-ray Computed Tomography}
+We acquired x-ray computed tomography of sample VG03.  The measurement details are listed in Table\ref{tab:tablenum4}.
+
+%\caption{Summary of X-ray CT measurements for VG03 \label{tab:tablenum4}}
+\begin{center}
+    \begin{tabular}{ll}
+    \emph{Parameter (Units)} & \emph{Value} \\ 
+    Instrument& Xradia MicroCT \\
+     Objective & 0.7X \\
+    Accelerating voltage (kV) & 100 \\
+        Power (W) & 10 \\
+        Acquisition time (s) & 3.5 \\
+        Number of views & 2521 \\
+        Total slices & 469 \\
+        Field of view diameter (pixels) & 988 \\
+        Voxel lateral pixel resolution ($\mu$m) & 38.02 \\
+        Estimate field of view (mm) & 37.5 \\
+        Voxel axial pixel resolution ($\mu$m) & 0.5 \\
+        Scan date & 2013 August 8 \\
+        Sample thickness (mm) & 6.6 \\
+     \end{tabular}
+\end{center}
+
+The axial resolution of the X-ray CT system is about 500 nm, which is much larger than the small ($\sim$ 15$-$90 nm) patterned gaps on many of our bonded samples.
+Part VG03 has a cylindrical patterned gap with a $\sim$20 mm diameter and 4000 nm height.
+We expected to see the cylindrical gap in about 8 resolution elements, given that the pattern was centered in the 25 mm diameter field of view.
+We saw a circular feature that extended over 297 pixels, which corresponds to 11.3 mm, assuming the voxel lateral dimension from the table above, which is reported in the X-ray CT lab measurement details.
+The feature...
+
 \section{Incoherent Multiple Reflections Transfer Matrix Method}
@@
 
+\section{Appendix: High spectral resolution coherent interference of thin substrates}
+
+Throughout this article we have assumed that $L >> \lambda^2 / \delta \lambda$, so that the interference of the substrate is safely incoherent.
+Another way of stating this criterion is that the spectral bandwidth is large enough that high spectral frequency fringes average out in a composite spectrum of low spectral resolution.
+If we relax this assumption so that the Si substrate thickness is comparable to or smaller than the coherence length, we would begin to observe high frequency fringes in the transmission spectra.
+In this appendix section we work out the transmission spectrum for two scenarios.
+First, we work out the case of a DSP Si wafer of thickness $L$, with no gap.
+Second we work out the case of two bonded Si wafers of thicknesses $L_1$ and $L_2$.
+For both cases we consider infinite spectral resolution, then convolve the spectra with a finite spectral resolution to show that the spectrum tends to that derived using an incoherent assumption.
+
+\subsection{DSP Si wafer at high spectral resolution}
+The Fabry-P\`erot etalon transmission is given in Equation \# X (currently, 5).  What we want to evaluate is:
+
+$$
+ <T_e> = \lim_{\sin^2(\delta/2) \to 0.5} \frac{1}{1+F\sin^2(\delta/2)} \\
+ <T_e> = \frac{1}{1+F/2}
+$$
+
 \bibliographystyle{osajnl}

### 2014-02-13 AO_bondedSi.tex
lots of good additions to the IR spectroscopy section

@@
 We achieved etch depths of 20$\pm$5 nm and 4000$\pm$100 nm, as measured with Wyko 2D noncontact white light microscopy for the small depth, and Dektak stylus profilometry for the large depth.
+For measurements taken with the Wyko, uncertainties were constructed by inspecting the histogram of topology.
+The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
+The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
 
-Table \ref{tab:bondexper} lists part numbers and descriptions for our bonded substrates.
+Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.
 
 \begin{table}[h!]
-%\caption{UTexas Si bonding experiments \label{tab:bondexper}}
+\caption{UTexas Si bonding experiments \label{tabbondexper}}
 \begin{center}
@@
 \subsection{Consideration of multiple incoherent reflections}
+\label{secTheory}
 Multiple incoherent reflections occur within the Si substrate.
@@
 
-$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \nonumber $$
+$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}$$
 
@@
 Figure \ref{IRbubble} shows an example of one of the IR bubbles in which several bright and dark fringes are clearly discernible.
+No attempt at quantitative measurement was made on the IR images due to anecdotal evidence that the detector is non-linear.
+The morphology was clear regardless of the detector response.
 
@@
 \section{Results of infrared spectroscopy of bonded Si pucks}
+In this section we describe the spectroscopy of bonded Si pucks.
+In some instances IR imaging was available and was leveraged to target IR bubbles.
+In other cases the IR imaging was not available or simply ignored, and spectra were taken at random positions on the wafer surface.
+
+\subsection{Constructing the predicted transmission}
+We constructed the predicted transmission in the following way.
+If IR imaging was unavailable, our optimistic prior belief was simply that the bond should be perfect and indistinguishable from a solitary DSP Si wafer, (i.e.
+100\% relative transmission).
+For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section ref{secTheory}, specifically equation \label{eqFP}.
+For non-unit fill factor, we simply average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor.
+Notably the meshes we designed were all 50\% fill factor.
+No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we take the design fill factor as gospel.
+The depth of the meshes has an associated uncertainty listed in \label{tabbondexper}.
+We incorporated the uncertainty in the mesh depth in the following way.
+We computed the transmission curve for the 1$\sigma$ lower and upper bounds, and plot a filled band marked as the prediction band.
+The uncertainties in the mesh depths were low enough that the band spread includes all intermediate lines of predicted transmission between plus and minus 1$\sigma$.
+In cases where the IR imaging was unavailable the mesh position was only coarsely known, so our prediction band had to incorporate our uncertainty about whether or not our measurement beam was on or off or overlapping the mesh border.
+If the beam was off the mesh we assumed 100\% relative transmission, so we simply set the lower bound of the predicted transmission to the equivalent of a zero gap.
+
+\subsection{Spectra of pucks with known gap sizes}
+Figures \ref{1} to \ref{2} show examples of the predicted and measured IR spectra.  Figure \ref{1} shows
 

### 2014-02-14 AO_bondedSi.tex

--- 2014-02-13 AO_bondedSi.tex
@@
 \section{History of Si direct bonding}
-\ref{secHistory}
 
@@
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
-We used a 50 mm F/2 lens, with a fiber optic white light illumination source.
-We estimate that the lamp-detector combination was sensitive to wavelengths in the range 1.1-1.6 $\mu$m.
-The field of view sub-sampled the 4 inch diameter wafers, so we dithered the sample.
-No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.
-It was easy to detect IR bubbles despite the artifacts since the bubbles translate with the wafer during dithering.
-We detected IR bubbles in all 3 of the bonded Si pucks on which we recorded IR images.
-The bubble areal density varied from about 4 per wafer (100 mm diameter) to over 20 per wafer.
-Naturally, the sample of control solitary DSP (unbonded) wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrtes as discussed in the literature (see Section \label{secHistory}).
-Figure \ref{IRbubble} shows an example of one of the IR bubbles in which several bright and dark fringes are clearly discernible.
+We used a 50 mm F/2 lens, with a fiber optic illumination source.
+The lens was spaced from the detector plane by about XX mm, with the illumination source at a distance XXX from the lens.
+The illumination source was gently diverging.
+The result was a vignetted beam.
+The target was located a distance XXX from the lens to provide imaging of the Si interface, with a field of view of roughly X $\times$ X degrees, or X $\times$ X mm.
+This field of view sub-sampled the 4 inch diameter wafer, so we dithered the sample, collecting images.
+%\begin{figure}[htbp]
+%\centerline{\includegraphics[width=.8\columnwidth]{IMG_7662.JPG}}
+%\caption{Photo of part VG15 with locations of defects marked.}
+%\end{figure}
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/IR_bubble.pdf}}
-\caption{Photo of part VG09 with transmission image of a bubble highlighted\label{IRbubble}.
-The right side photo shows the 75 mm diameter wafer with blue Sharpie rectangles indicating the position of different spectral measurements (as detailed in the next section).
-The yellow circle in the photo highlights the outline of the location of an IR bubble we detected.
-The bubble shows a couple of fringes over the effective wavelength range 1.1-1.6 $\mu$m.
-This fringe pattern is consistent with a gap comparable to this measurement wavelength.
-The pointlike measurement artifacts, Sharpie marker, and vignetting are perceptible in the raw IR image.
-Our interpretation of the IR bubbles is robust against these artifacts.
-The area surrounding the IR bubble appears to be mostly devoid of large (wavelength-scale) gaps, since the transmission is fairly uniform, and consistent with the brightest fringe of the IR bubble.
-}
-\end{figure}
-
-\section{Results of infrared spectroscopy of bonded Si pucks}
+\section{Results}
 

### 2014-02-17 AO_bondedSi.tex
add section on spectra, begin the table of measurement outcomes

--- 2014-02-13 AO_bondedSi.tex
@@
 \subsection{Spectra of pucks with known gap sizes}
-Figures \ref{1} to \ref{2} show examples of the predicted and measured IR spectra.  Figure \ref{1} shows
+Figures \ref{1} to \ref{2} show examples of the predicted and measured IR spectra.
+Figure \ref{VS21specm1} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
+Meanwhile, Figure \ref{VS21specf1} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 2 \sigma$).
+Other parts show less consistency with the prediction.
+A complete description of the diversity of the measured spectra requires an appeal to optical phenomena beyond the scope of this paper.
+Briefly we classify the diversity of measurements into the categories laid out in the table below.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1.eps}}
+\caption{IR transmission spectrum of VS21 (M1) showing consistent gap\label{VS21specm1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1.eps}}
+\caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
+\end{figure}
 

### 2014-02-17 AO_bondedSi.tex
add table of measurement outcomes interpretation

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1.eps}}
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1}}
 \caption{IR transmission spectrum of VS21 (M1) showing consistent gap\label{VS21specm1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1.eps}}
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1}}
 \caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
 \end{figure}
+
+\begin{table}[h!]
+  \caption{A simple table}
+  \begin{center}
+    \begin{tabular}{ccc}
+    \hline
+    Appearance of spectra & Interpretation & Exemplar \\
+    \hline
+    Measurement within the prediction band & Life is good & VS21-M1, Figure \ref{VS21specm1}\\
+    Measured spectrum below the prediction band but shape is similar & Gap was slightly larger than predicted or bonding was incomplete& VS21-F1, Figure \ref{VS21specf1}\\
+    Measured spectrum above the prediction band but shape is similar & Gap was slightly smaller than predicted & VS21-M2\\
+    Measured spectrum exceeds DSP reference puck & unanticipated anti-reflection coating on the surface of the Si sample (e.g. Sharpie marker residue) & -\\
+    High frequency, low amplitude wiggles superimposed upon smooth underlying trend & Large, unanticipated gap, with a small fill factor  & VS21-C3\\   
+     																			     & wavelength dependent diffraction from the mesh patterns & VS21-C3\\       
+     																			     & measurement artifact  & VS21-C3\\       												
+    Non-unit transmission constant with wavelength (gray spectrum) & Artifact of measurement- vignetting of beam, absorptive coating, zero-point drift & VS21-C4 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
 

### 2014-02-17 AO_bondedSi.tex
minor tweaks to table

--- 2014-02-17 AO_bondedSi.tex
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1.eps}}
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1}}
 \caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
 \end{figure}
+
+\begin{table}[h!]
+  \caption{A simple table}
+  \begin{center}
+    \begin{tabular}{ |p{8cm}| p{8cm} | p{2.2cm} |}
+    \hline
+    Appearance of spectra & Interpretation & Exemplars \\
+    \hline
+    Measurement within the prediction band & Life is good & VS21-M1, Figure \ref{VS21specm1}, VG15, VG07\\
+    \hline
+    Measured spectrum below the prediction band but shape is similar & Gap was slightly larger than predicted or bonding was incomplete& VS21-F1, Figure \ref{VS21specf1}, VS09, VG15, \\
+    \hline
+    Measured spectrum above the prediction band but shape is similar & Gap was slightly smaller than predicted & VS21-M2\\
+    \hline
+    Measured spectrum exceeds DSP reference puck & unanticipated anti-reflection coating on the surface of the Si sample (e.g. Sharpie marker residue) & VG08\\
+    \hline
+    High frequency, low amplitude wiggles superimposed upon smooth underlying trend & Large, unanticipated gap, with a small fill factor  & VS21-C3\\   
+     																			     & wavelength dependent diffraction from the mesh patterns & VS21-C3\\       
+     																			     & measurement artifact  & VS21-C3\\       												
+    \hline
+    Non-unit transmission constant with wavelength (gray spectrum) & Artifact of measurement- vignetting of beam, absorptive coating, zero-point drift & VS21-C4 \\
+    \hline
+    Completely different spectrum than anticipated with low absolute transmission & Large, inhomogeneous gap which should appear as an IR bubble & VG09 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
 

### 2014-02-17 AO_bondedSi.tex
minor tweaks, just pushing to remote

@@
 \begin{table}[h!]
-  \caption{A simple table}
+  \caption{Interpretation of the diversity of IR spectra}
   \begin{center}
@@
     \hline
-    Completely different spectrum than anticipated with low absolute transmission & Large, inhomogeneous gap which should appear as an IR bubble & VG09 \\
+    Completely different spectrum than anticipated with low absolute transmission & Large, inhomogeneous gap consistent with an IR bubble & VG09 \\
     \hline

### 2014-02-18 AO_bondedSi.tex
lots of good additions to results

@@
 \subsection{Spectra of pucks with known gap sizes}
-Figures \ref{1} to \ref{2} show examples of the predicted and measured IR spectra.
+Figures \ref{VS21specm1} to \ref{VS21specf1} show examples of the predicted and measured IR spectra.
 Figure \ref{VS21specm1} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
@@
 \end{table}
+
+A key result of this work is to show that gaps of known dimensions are recovered in our spectroscopy.
+The cases in which predicted spectroscopy differs from measurement are due to either a bona-fide detection of a gap above and beyond the implanted gap, or some measurement artifacts.
+In the next section we address mitigation of artifacts.
+
+\subsection{Mitigation of measurement artifacts}
+The principle artifacts and their interpretation are listed in table \ref{1}.
+We mitigated the artifacts in the following ways.
+For the Sharpie marker behaving as an anti-reflection (AR) coating, we simply removed the sharpie with IPA.
+We have mostly been able to rule out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.
+Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.
+We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature (see the work of Mitani and collaborators from the mid 1990s).
+Namely, Mitani showed that the gas pressure of bonded wafers grew as the fill factor of voids decreased.
+In other words, what matters is the fill factor of voids, not their relative sizes.
+Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the voids, and therefore no difference in the bonding properties.
+
+Our sensitive measurement technique introduces a few challenges at the parts per thousand level baseline level.
+Specifically, we have been ambivalent about whether or not to use the double-beam capability of the Cary5000.
+We have shown that this mode reduces by a factor of 10, the random flicker attributable to lamp noise.
+In practice, the double beam mode can cause some systematic problems near the portions of wavelengths near lamp and filter changes.
+So of these issues might be related to polarization differences, and a way to mitigate these techniques is to follow the advice of the vendor and employ a depolarizer in the beam path.
+A depolarizer strategy is critical for applications in which a non-normal incidence angle is desired.
+A half-hearted attempt to measure non-normal incidence without the polarization accessory resulted in failure.
+In the next section we describe our interest in non-normal incidence.
+
+\subsection{Non-normal incidence angles}
+For many applications, normal incidence performance is the quantity of interest.
+The technique described here outlines an approach for measuring the gap size of bonded Si wafers at normal incidence.
+In principle, from there the performance at non-normal incidence can be worked out either by simulation or calculation.
+Here we briefly motivate the challenge of experiment and simulation at non-normal incidence.
+
+Our final application that motivated this project is to make diffraction gratings on the hypotenuse of a silicon prism.
+The diffraction occurs inside the silicon material and is dispersed and reflected back along its path.
+This configuration is called an immersion grating.
+Our group has made monolithic immersion gratings for over a decade, but challenges in processing large (30 mm thick) Si wafers motivated an approach of patterning and processing thin (1-10 mm thick) Si wafers, and then bonding them to shaped and polished prisms.
+This technique has been pioneered at SRON.
+
+The consequence of our design is a non-normal incidence interface in double pass.
+In this case, the framework developed here fails to predict the spectrum of the interface loss attributable to a finite albeit small (few nm) gap.
+Specifically, at large incidence angles complete internal reflection and evanescent waves will matter.
+Polarization effects are important.
+
+\subsection{Spectra of pucks with no implanted gap}
+We have shown that the IR spectroscopy technique combined with the Fabry-P\'{e}rot framework provide accurate knowledge of the gap size.
+In the cases where we expected no gap (i.e.
+perfect Si-Si bond), we have seen spectra consistent with 100\% transmission relative to a DSP Si wafer.
+In other cases we have seen evidence for gaps, which is simply an indication that the bonding process introduces gaps.
+In a followup paper we will describe ways to mitigate bond formation in bonded Si pucks.
+
+\section{Conclusions}
+We have shown a method for indirectly measuring the size of voids in bonded Si wafers.
+The technique broadly applies to any compound Si optics measured at normal incidence.
+We showed lab IR spectra with consistent predictions, and addressed ways to mitigate measurement artifacts.
 

### 2014-03-04 AO_bondedSi.tex
revise History section- replace table of papers with narative

@@
 These gratings are used either in transmission (i.e.
-as grisms \cite{2008SPIE.7014E..77D,2010SPIE.7739E.123G}), or in reflection (i.e.
-as immersion gratings, \cite{2007ApOpt..46.3400M,2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.
-These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y,2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.
+as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e.
+as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.
+These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.
 To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.
@@
 
-We describe a technique for directly measuring gaps at silicon-silicon interfaces.
-Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}, including IR interference, X-ray topography, and ultrasound microscopy.
+There are two main solutions for dealing with bubbles.
+We will refer to the two approaches as the \emph{strength approach} and the \emph{resilience approach}.
+The \emph{strength approach} is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
+Typically this approach demands disciplined handling protocols, and rigorous multi-stage cleaning steps, although unconventional micro-cleanroom approaches exist \cite{1989JaJAP..28L2141L}.
+The \emph{resilience approach} is simply to deal with the uncomfortable likelihood that bubbles will occur, and then do something to remove them.
+The most cited way to get rid of IR bubbles is to anneal them\cite{Mitani1990}.
+We describe the current understanding of both approaches in section \ref{secHistory}.
+
+This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.
+Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}, including IR interference\cite{Mitani1990}, X-ray topography\cite{1994JaJAP..33....6H}, and ultrasound microscopy\cite{2000RScI...71.1869G}.
 The standard infrared interference technique is to examine the bond in reflection for the presence of Newton rings.
@@
 
-Serious research on semiconductor wafer bonding started around 1989, with Lehmann et al.
-\cite{1989JaJAP..28L2141L} bonded GaAs and InP wafers to Si.
-They detected no gaps with IR imaging.
-They measured the surface energy of the bond and found a range of 40$-$600 dyne/cm.
-
-The majority of work on bonding was carried out from 1990-2000.
-The key advances included studies that probed the surface energy as a function of anneal temperature, and various probes of the bond chemistry through multiply reflection infrared spectroscopy and mass spectroscopy.
-A fairly complete story was then constructed which tied together all the disparate measurements.
-No where in the literature did we find a technique similar to the one presented here.
-See the table below for a list of papers on Si wafer bonding and a note on their contribution.
-
-%\caption{Literature review}
-\begin{spacing}{1.0} 
-\begin{center}
-\begin{longtable}{ |c | p{8cm} | p{8cm} |}
-    \hline
-    Ref & Title & Notes \\ 
-        \hline
-	\cite{1986JAP....60.2987S} &  \emph{Silicon-to-silicon direct bonding method} &  Foundation paper  \\
-         \hline
-     \cite{1988JaJAP..27L2364S} &  \emph{Bubble-Free Silicon Wafer Bonding in a Non-Cleanroom Environment} &  Sequence of IR images showing bond propagation  \\
-         \hline
-     \cite{1989JaJAP..28L2141L} &  \emph{Bubble-Free Wafer Bonding of GaAs and InP on Silicon in a Microcleanroom} &  No gaps detected with IR imaging; measured surface energy of the bond 40$-$600 dyne/cm.
-\\
-         \hline
-     \cite{Mitani1990} &  \emph{Bubble formation during silicon wafer bonding: causes and remedies} &  Bubble elimination at temperature greater than 800 C. \\
-         \hline
-	\cite{1992JEMat..21..669M} &  \emph{Wafer bonding technology for silicon-on-lnsulator applications: A review} &  Excellent paper.
-Shows infrared imaging detected bubbles and the bubble density as a function of anneal temperature.
-Directly compares the three bubble detection techniques\\
-         \hline
-     \cite{feijoo1994} &  \emph{Silicon wafer bonding studied by infrared absorption spectroscopy} &  Works out the chemistry of the interface, finds distinct temperature phases, hydrogen bonds above   \\
-         \hline
-     \cite{1994JaJAP..33....6H} &  \emph{Silicon Wafer Direct Bonding without Hydrophilic Native Oxides} &  Shows x-ray topograph technique, hydrophobic is strong bond\\
-              \hline         
-     \cite{1995ApPhA..61..101R} &  \emph{Multiple internal reflection spectroscopy of bonded silicon wafers} &  Chemistry of the bond interface- Evidence for Si$-$H bonds\\
-              \hline
-     \cite{1995ApPhL..67.3614G} &  \emph{Self-propagating room-temperature silicon wafer bonding in ultrahigh vacuum} &  IR image of wafers bonding with gaps visible  \\
-              \hline     
-     \cite{1995ApPhL..67..863G} &  \emph{What determines the lateral bonding speed in silicon wafer bonding?} &  Spread speed is 10 mm/s, decreases with increasing ambient pressure; thickness has no effect.
-\\
-              \hline
-     \cite{1996ApPhL..68.2222T} &  \emph{Surface activated bonding of silicon wafers at room temperature} &  Shows that Ar etching makes stronger bonds, shows no void from IR microscopy or KOH etching  \\
-              \hline
-     \cite{1996JaJAP..35.2102R} &  \emph{Characterization of Interfaces of Directly Bonded Silicon Wafers: A Comparative Study of Secondary Ion Mass Spectroscopy Multiple Internal Reflection Spectroscopy, and Transmission Electron Microscopy} &  They figure out what is the chemistry participating in Si bond formation\\
-              \hline
-     \cite{1998AnRMS..28..215G} &  \emph{Semiconductor wafer bonding} &  Annual Reviews article summarizing the state of the subfield\\ 
-              \hline
-     \cite{JJAP.37.4197} &  \emph{Effect of Surface Roughness on Room-Temperature Wafer Bonding by Ar Beam Surface Activation} &  Appears to only explore surface roughness of less than 1 nm RMS, which seems low compared to our wafers.\\
-              \hline
-     \cite{2000JAP....88.4404H} &  \emph{Modeling the dynamics of Si wafer bonding during annealing} &  Prescription for how long to run annealing experiments\\     
-              \hline
-     \cite{2000RScI...71.1869G} &  \emph{Photoacoustic imaging of voids in direct wafer bonding} &  Compares photoacoustic technique to the IR technique \\     
-              \hline
-     \cite{2001JOptA...3...85G} &  \emph{Optical contact and van der Waals interactions: the role of the surface topography in determining the bonding strength of thick glass plates} &  What is the effect of surface roughness on optical bonding; uses 1 nm roughness.\\
-              \hline
-     \cite{2001JAP....89.6013L} &  \emph{Modeling of hydrophilic wafer bonding by molecular dynamics simulations} &  Examines the effect of moisture and defects in forming bonds\\ 
-          \cite{2007ApOpt..46.6793H} &  \emph{Direct bonding and beyond} &  Excellent paper.  Demonstrates Si direct bonding on thick substrates.\\ 
-    \hline
-\end{longtable}
-\end{center}
-\end{spacing}
+The foundation paper on Si bonding came in 1986\cite{1986JAP....60.2987S}.
+After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.
+In some of cases, the bond propagation could be directly monitored\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G}.
+The bond process could be shown to be bubble-free\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L}.
+The bond surface energy was directly measured\cite{1989JaJAP..28L2141L}.
+The causes and remedies for bubble formation were discussed as early as 1990\cite{Mitani1990}.
+Specifically, the IR bubble density changes as a function of anneal temprature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.
+The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
+A robust picture of the chemistry as a function of anneal temperature was ultimate fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation \cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, moisture and defects \cite{2001JAP....89.6013L}.
+No where in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.
 

### 2014-03-04 AO_bondedSi.tex
Make power law dust distribution discussion more quantitative

@@
 A robust picture of the chemistry as a function of anneal temperature was ultimate fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
-Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation \cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, moisture and defects \cite{2001JAP....89.6013L}.
+Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
 No where in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.
+
+\section{Conceivable pitfalls in Si-Si direct bonding}
+
+\subsection{Particles}
+Consider the following two conceivable pitfalls in Si-Si direct bonding.
+The obvious pitfall is particle contamination.
+A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
+In the next section we show that smaller bubbles can unacceptably degrade the performance to as low as 10 nm.
+In a class 100 cleanroom, the volume density could be as high as $\rho=100$ particles/ft$^3$, with sizes larger than 0.5 $\mu$m.
+What we care about is not the volume number density of particles, but the delivered surface flux $F=\rho(D) v(D)$, where v(D) is the average freefall speed of particles, which is a function of particle size.
+Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$-2$ s$-1$) vs.
+particle size ($\mu$m).
+For a change from 1 $\mu$m to 10$-2$ $\mu$m, there is an increase in a factor of $10^4$ in the particle flux expected.
+In other words:
+$$\log(F) = -2\log(D) + const.$$
+
+So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles, which means for every detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
+
+\subsection{Surface roughness or aberrations}
+The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
+Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.
+The details of this aspect of the bonding process are outside the scope of this paper.
 
@@
 Popular optical materials with refractive indices in the range 1.5$-$1.7 with Fresnel reflection 4$-$7\% per surface will suffer less than Si from the effects described here.
-
-Consider the following two conceivable pitfalls in Si direct bonding.
-First, particles at the tens of nanometer size scale are trapped between the Si pucks during bonding.
-This potential problem is especially acute for silicon direct bonding compared to bonding of popular low-index optical materials.
-A typical particle surface flux is shown in Figure 5 of Cooper 1986\cite{Cooper_1986}.
-The flux of particles is a step function of particle size.
-For every micron sized particle causing Newton's rings, there will be as many as 100,000 more ten-nanometer sized particles.
-
-The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.
-The details of this aspect of the bonding process are outside the scope of this paper.
 

### 2014-03-05 AO_bondedSi.tex
add particle schematic and text about particles

@@
 \section{History of Si direct bonding}
-\ref{secHistory}
+\label{secHistory}
 
@@
 \subsection{Particles}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.5\columnwidth]{figs/SiBondParticle.pdf}}
+\caption{Schematic of a particle preventing bonding of two Si surfaces\label{figParticle}}
+\end{figure}
+
 Consider the following two conceivable pitfalls in Si-Si direct bonding.
 The obvious pitfall is particle contamination.
+Figure \ref{figParticle} shows a schematic of a particle wedged between two flat Si optical surfaces.
 A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
@@
 
-So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles, which means for every detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
+So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles, which means for every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
+One key idea that saves the day, is the intuition that the areal fill factor scales at least as fast as $D^2$.
+Imagine that the particles serve as miniature tentpoles.
+The volume beneath the tentpole looks like a cone.
+Since the number density goes down as roughly $D^-2$, (per the equation above) the total area covered by particles might be constant in $log(D)$.
+Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.
+It is further conceivable that the areal fill factor attributable to a tent pole particle scales even faster than $D^2$ since surface energy probably scales inversely with distance of the pieces (\`a la VanDerWaals).
+
+It has been shown\cite{1992JEMat..21..669M} that the gases seep through the Si$-$Si interface, so pressure builds around those particles that serve as catalysts.
+In this sense, the size of the particle causing an IR bubble cannot be uniquely identified from observations of the extent of IR bubble defects.
+In section \ref{secResults} we address the question of the origin of the IR bubbles.
+Specifically, are the particles more like \emph{tentpoles}, rigid structures that merely prevent the zippering up of the bond?
+Or are the particles more like \emph{catalysts} so that their size is immaterial- the particles merely accelerate the production of gases that bow out the bond?
+These are open questions.
 
@@
 \section{Results of infrared spectroscopy of bonded Si pucks}
+\label{secResults}
 In this section we describe the spectroscopy of bonded Si pucks.
@@
 100\% relative transmission).
-For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section ref{secTheory}, specifically equation \label{eqFP}.
+For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section ref{secTheory}, specifically equation \ref{eqFP}.
 For non-unit fill factor, we simply average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor.
@@
 No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we take the design fill factor as gospel.
-The depth of the meshes has an associated uncertainty listed in \label{tabbondexper}.
+The depth of the meshes has an associated uncertainty listed in \ref{tabbondexper}.
 We incorporated the uncertainty in the mesh depth in the following way.

### 2014-03-06 AO_bondedSi.tex
updates of figures, etc.

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.5\columnwidth]{figs/SiBondParticle.pdf}}
-\caption{Schematic of a particle preventing bonding of two Si surfaces\label{figParticle}}
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiBondTentCatFig.pdf}}
+\caption{Schematic of a particle preventing bonding of two Si surfaces\label{figParticle}.
+There are two conceivable scenarios in which particle contamination can cause gaps in the bond.
+\emph{Left-} The particle serves as a mechanical structure hindering the Si-Si bond from zippering up.
+The geometry is like a tent-pole.
+\emph{Right-} The particle serves as a catalyst for gas formation.
+There is evidence that hydrocarbons can serve as catalysts, the especially pernicious particles are ``thermally unstable hydrocarbon contaminants on the surfaces coming from plastic containers or simply from the ambient in a cleanroom.'' \cite{1998AnRMS..28..215G}.
+}
 \end{figure}
@@
 \section{Silicon-gap-silicon Fabry-P\`{e}rot theory}
+\label{secTheory}
 

### 2014-03-18 AO_bondedSi.tex
lots of changes- add 2 figures, copious prose, minor tweaks

@@
 
-\subsection{Surface roughness or aberrations}
+\subsection{Surface roughness or large-scale aberrations}
 The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
+This issue is much less of a problem for wafers compared to our unconventional thick ($\sim$ 30 mm) Si applications.
 Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.
 The details of this aspect of the bonding process are outside the scope of this paper.
-
-\section{Silicon-gap-silicon Fabry-P\`{e}rot theory}
-\label{secTheory}
-
-Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
-The Fresnel reflection is accordingly about 30\% per surface.
-Popular optical materials with refractive indices in the range 1.5$-$1.7 with Fresnel reflection 4$-$7\% per surface will suffer less than Si from the effects described here.
+In short, our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
+We expect such well-polished substrates to have no deleterious impact on the bonding process.
 
@@
 We protected the unexposed areas with thick photoresist.
-We achieved etch depths of 20$\pm$5 nm and 4000$\pm$100 nm, as measured with Wyko 2D noncontact white light microscopy for the small depth, and Dektak stylus profilometry for the large depth.
-For measurements taken with the Wyko, uncertainties were constructed by inspecting the histogram of topology.
+We achieved etch depths of 20$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
+For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology.
 The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
@@
 \section{Transmission measurement theory}
+\label{secTheory}
+
+% Figures (shared wavelength axis)
+%	1) Refractive index as a function of wavelength at room temp
+%	2) Fresnel transmission as a function of wavelength (normal incidence)
+%	3) Coefficient of finesse as a function of wavelength
+%	4) Example transmission through VG03 thick gap
+
+Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
+The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+The Fresnel reflection is accordingly about 30\% per surface.
+For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
+Figure X shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{fig} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
+The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
+The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
+\end{figure}
+
 The Si-air\footnote{We imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.} interfaces of the in the bond gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.
@@
 The substrates were silicon wafers.
-The surface roughness was typically about 2 nm, as measured with a Wyko 2D noncontact microscope.
+The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler .
 The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
@@
 
-Our sensitive measurement technique introduces a few challenges at the parts per thousand level baseline level.
-Specifically, we have been ambivalent about whether or not to use the double-beam capability of the Cary5000.
-We have shown that this mode reduces by a factor of 10, the random flicker attributable to lamp noise.
-In practice, the double beam mode can cause some systematic problems near the portions of wavelengths near lamp and filter changes.
-So of these issues might be related to polarization differences, and a way to mitigate these techniques is to follow the advice of the vendor and employ a depolarizer in the beam path.
-A depolarizer strategy is critical for applications in which a non-normal incidence angle is desired.
-A half-hearted attempt to measure non-normal incidence without the polarization accessory resulted in failure.
-In the next section we describe our interest in non-normal incidence.
-
 \subsection{Non-normal incidence angles}
-For many applications, normal incidence performance is the quantity of interest.
-The technique described here outlines an approach for measuring the gap size of bonded Si wafers at normal incidence.
+For the vast majority of applications, normal incidence performance is the quantity of interest.
+The technique laid out in this paper outlines an approach for measuring the gap size of bonded Si wafers at normal incidence.
 In principle, from there the performance at non-normal incidence can be worked out either by simulation or calculation.
 Here we briefly motivate the challenge of experiment and simulation at non-normal incidence.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/Sibondedprismcartoon.pdf}}
+\caption{Schematic of a bonded Si immersion grating\label{figSiPrism}.  }
+\end{figure}
 
@@
 This configuration is called an immersion grating.
+Figure \ref{figSiPrism} shows a schematic of the geometry of our bonded Si immersion grating.
 Our group has made monolithic immersion gratings for over a decade, but challenges in processing large (30 mm thick) Si wafers motivated an approach of patterning and processing thin (1-10 mm thick) Si wafers, and then bonding them to shaped and polished prisms.
-This technique has been pioneered at SRON.
+This technique has been pioneered at SRON \cite{2012SPIE.8450E..2TV}.
+In Figure \ref{figSiPrism} we show an example with a 10 mm thick Si puck bonded to a 25 mm thick, precision cut, shaped, and polished Si prism.
+The group at SRON has patterned Si immersion echelle gratings on commercial thin  ($< 1$ mm) Si wafers.
+Thin wafers have the drawback that they are conformal, so total-thickness-variations (TTV) directly translate into surface optical aberrations\cite{2012SPIE.8450E..2TV}.
 
-The consequence of our design is a non-normal incidence interface in double pass.
-In this case, the framework developed here fails to predict the spectrum of the interface loss attributable to a finite albeit small (few nm) gap.
-Specifically, at large incidence angles complete internal reflection and evanescent waves will matter.
+The consequence of our immersion grating design is a non-normal incidence interface in double pass.
+In this case, the Fabry-P\'{e}rot framework developed here fails to predict the spectrum of the interface loss attributable to a finite albeit small (few nm) gap.
+Specifically, complete internal reflection and evanescent waves will matter at large incidence angles, and the incoherent matrix technique is no longer accurate.
 Polarization effects are important.
+We do not have a mechanism for evaluating the bond performance until the final device is completed and its wavelength-dependent transmission is measured.
+We can detect the performance degradation on final devices in the following nuanced way.
+We will measure the efficiency of the bonded Si immersion grating as a function of both wavelength and angle.
+Our group built a custom tool to make these measurements- the Custom Robotic Order, Wavelength, and Blaze Angle Recorder (CROWBAR) \cite{2012SPIE.8450E..2SG}.
+Figure 3 in \cite{2012SPIE.8450E..2SG} shows the directly measured spectrum of a monolithic immersion grating.
+Notably, the spectrum in Figure 3 in \cite{2012SPIE.8450E..2SG}  shows dozens of blaze peaks where the efficiency should be maximized.
+A perfectly contacted Si$-$Si direct bonded optic would have identical performance to a monolithic counterpart.
+Bonded immersion gratings with finite gap sizes will superimpose a wavelength dependent transmission loss on top of the blaze peaks.
+In this way, the presence or absence of deleterious gaps can be identified in final immersion gratings.
+One drawback of this approach is that gaps can only be identified after the costly process of producing the final optical device.
+Large gaps identified this late in the game could render the costly optic useless, both in terms of efficiency and wavefront performance.
+Our group is exploring the effectiveness of high-temperature annealing to eliminate IR bubbles in completed devices.
 

### 2014-03-19 AO_bondedSi.tex
minor tweaks, I think
add new figures

@@
 The details of this aspect of the bonding process are outside the scope of this paper.
-In short, our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
+In short, most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
 We expect such well-polished substrates to have no deleterious impact on the bonding process.
+Some experimental substrates had larger surface error, of order a few waves of surface error.
+These substrates were reserved for process development, since we anticipated that bonding might progress either not at all, or with large interface gaps.
 
 \section{Imbedding gaps of known sizes}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/VG03patternSchematic.pdf}}
+\caption{Schematic of part VG03 \label{figVG03pattern}.
+The four inch diameter VG03 is comprised of two 3.3 mm thick silicon pucks.
+We patterned a one inch diameter hole on one of the pucks.
+The depth was 4.0 $\mu$m as indicated by Dektak profilometry.
+We also patterned a low-fill factor petal pattern surrounding the hole.
+The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.
+The technique can reveal the axial extent and fill factor of an embedded gap.}
+\end{figure}
+
 We evaluated our metrology method by imbedding gaps of known sizes in the interface.
+Here we describe the process of how we made the gaps and how we characterized their axial extent.
+The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.
 We bored divots in silicon with inductively coupled plasma etching.
@@
 We protected the unexposed areas with thick photoresist.
-We achieved etch depths of 20$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
-For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology.
+We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
+Figure \ref{figX} shows a schematic of VG03 and its pattern.
+In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates.
+For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{fig}.
 The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
 The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/VS20fineGapCrop.pdf}}
+\caption{Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section\label{figVS20pattern}.
+The four inch diameter VS20 is comprised of two thick silicon pucks.
+We patterned three meshes with coarse, medium, and fine boxes.
+The depth was 95 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
+The Veeco was operated in PSI mode.
+The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.
+The technique can reveal the axial extent and fill factor of an embedded gap.}
+\end{figure}
 
@@
     VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  23  & 2/13/2013 &  \\
-    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  23 & 2/14/2013 & 4 $\mu$m hole \\    
-    VG04 & Bonded & 100 & $1$ $\times$ 2 &  23 & 2/14/2013 & 15 nm gap \\        
+    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  23 & 2/14/2013 & 4 $\mu$m gap 100\% fill factor hole \& petal pattern\\    
+    VG04 & Bonded & 100 & $1$ $\times$ 2 &  23 & 2/14/2013 & 15 nm gap, $\sim50\%$ fill factor\\        
     VG05 & DSP & 100 & 1&  23 & 2/15/2013 & cleaved  \\
@@
     VG08 & - & - & - &  - & - & \\
-    VG09 & Bonded & 75 & 1.5? & 23 & 6/21/2013 & mesh C, 49 $\pm$6 nm \\
-    VG15 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 14 $\pm$2 nm\\
-    VG17 & Bonded & 75 & 6.0? &  23 & 6/21/2013 & mesh C, 64 $\pm$6 nm\\
-    VS20 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 95 $\pm$5 nm\\
+    VG09 & Bonded & 75 & 1.5? & 23 & 6/21/2013 & mesh C, 49 $\pm$6 nm, 50\% fill factor\\
+    VG15 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 14 $\pm$2 nm, 50\% fill factor\\
+    VG17 & Bonded & 75 & 6.0? &  23 & 6/21/2013 & mesh C, 64 $\pm$6 nm, 50\% fill factor\\
+    VS20 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 95 $\pm$5 nm, 50\% fill factor\\
     \hline
@@
 % Figures (shared wavelength axis)
-%	1) Refractive index as a function of wavelength at room temp
 %	2) Fresnel transmission as a function of wavelength (normal incidence)
-%	3) Coefficient of finesse as a function of wavelength
 %	4) Example transmission through VG03 thick gap

### 2014-03-20 AO_bondedSi.tex
some section rearrangement and edits for clarity

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/VS20fineGapCrop.pdf}}
+\centerline{\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}}
 \caption{Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section\label{figVS20pattern}.
-The four inch diameter VS20 is comprised of two thick silicon pucks.
-We patterned three meshes with coarse, medium, and fine boxes.
-The depth was 95 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
+The four inch diameter VS20 is comprised of two thick silicon wafers.
+We patterned three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+The depth was 95 $\pm$ 5 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
 The Veeco was operated in PSI mode.
 The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.
-The technique can reveal the axial extent and fill factor of an embedded gap.}
+The technique can reveal the axial extent and fill factor of an embedded gap.
+}
 \end{figure}
@@
 
-\section{Transmission measurement theory}
+\section{Substrate preparation}
+It is important to clean the surfaces before bonding to minimize the particle density.
+The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
+The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
+We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
+The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
+There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+We measured the large scale surface flatness of the pucks.
+The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.
+We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
+
+The detailed bonding procedure will be chronicled in a future paper.
+Briefly, we prepared the substrates with standard cleaning procedures of solvents in a megasonic.
+We then applied MHz frequency oxygen plasma ashing.
+We soaked the wafers in DI water then dried with an N$_2$ gun.
+We pressed the wafers together from the center to the outside with normal human force.
+
+\section{Si$-$air$-$Si Fabry P\`{e}rot measurement theory}
 \label{secTheory}
@@
 
+\subsection{The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon}
+
 Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
@@
 For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
-Figure X shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
+Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
 
@@
 \centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{fig} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
 The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
@@
 
-The Si-air\footnote{We imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.} interfaces of the in the bond gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.
-The transmission through a Fabry-Perot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We assume transmission is at normal incidence, the index of the gap is 1.0, and the material is at room temperature.
+The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.
+The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
+We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
 We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.
-We compute the reflectivity for an Si-air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}.
+We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
 The computation is simply Fresnel's law at normal incidence:
@@
 $$
-R = \frac{(n-1)^2}{(n+1)^2}
+R = \frac{(n-1)^2}{(n+1)^2} \\
+F \equiv \frac{4R}{(1-R)^2}
 $$
@@
 The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
-The modest wavelength dependence is shown in the bottom curve of the figure below.
-The coefficient of Finesse\cite{2007fuph.book.....S} encapsulates the Fabry-P\`erot etalon's dependence on the reflectivity.
-We compute a coefficient of Finesse in the range of about 2.7$-$2.9 depending on the wavelength.
+We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
+The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
 
@@
 Multiple incoherent reflections occur within the Si substrate.
-Treatment for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
+Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
 The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-I adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.
+We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.
 Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
@@
 If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
-We work out this behavior in the Appendix section X and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
+This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
+%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
 
-For the case of a pair of bonded Si pucks with a gap thickness $d$, we normalize the transmission with a gap to the transmission with no gap to isolate the effect of the gap.
-We call this normalized transmission the etalon transmission $T_{e}$.
-It is derived in Equation \ref{eqn:Tetalon} in the Appendix:
+Consider a pair of bonded Si pucks with a gap thickness $d$.
+We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
+We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.
+We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.
 
@@
 It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.  In the next section we compare measurements to by-eye fits of this equation.
-
-\section{Substrate preparation}
-It is important to clean the wafers to minimize the particle density, since particles cause gaps.
-The substrates were silicon wafers.
-The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler .
-The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
-We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
-The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
-We measured the large scale surface flatness of the pucks.
-The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.
-We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
-
-The detailed bonding procedure will be chronicled in a future paper.
-Briefly, we prepared the substrates with standard cleaning procedures.
-We used MHz frequency oxygen plasma ashing.
-We soaked the wafers in DI water then dried with an N$_2$ gun.
-We pressed the wafers together from the center.
 

### 2014-03-20 AO_bondedSi.tex
commit before I remove the double beam mode detailed discussion

@@
 The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
+Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.
 
@@
 \end{figure}
-
-Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.
 
@@
 
-\subsection{The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon}
+\subsection{The air cavity is a low Finesse Fabry-P\`{e}rot etalon}
 
@@
 
-The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+The air cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
 Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.
@@
 
-$$
+\begin{eqnarray}
 R = \frac{(n-1)^2}{(n+1)^2} \\
 F \equiv \frac{4R}{(1-R)^2}
-$$
+\end{eqnarray}
 
@@
 The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
+In the next subsection we work out the wavelength dependent transmission for the air etalon
 
-\subsection{Consideration of multiple incoherent reflections}
+\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
 \label{secTheory}
-Multiple incoherent reflections occur within the Si substrate.
+There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate.
 Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
@@
 
-Consider a pair of bonded Si pucks with a gap thickness $d$.
+Now consider a pair of bonded Si pucks with a gap thickness $d$.
 We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
@@
 \section{Measurement strategy, conditions, and details}
-The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \nocite{cary5000web}.
+The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \cite{cary5000web}.
 The typical measurement parameters and tool performance are listed below.
@@
 Briefly, the tool has a double monochromator.
-The double monochromator feature reduces scattered light, with the vendor reported linearity exceeding 40 dB.
+The double monochromator feature reduces scattered light.
+The vendor reported linearity exceeding 40 dB.
 

### 2014-03-20 AO_bondedSi.tex
some tweaks

@@
 The double monochromator feature reduces scattered light.
-The vendor reported linearity exceeding 40 dB.
+The vendor reported linearity exceeds 40 dB.
 
@@
 We found that the mean value in single mode drifted by about 0.2\%.
-The standard deviation of the featureless spectrum was typically 0.02\%, with some portions of the spectrum demonstrating increased uncertainty, perhaps attributable to atmospheric absorption.
+The standard deviation of the featureless spectrum was typically 0.02\%, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.
 In double beam mode the spectrograph performance was much better.
@@
 In this way the transmission spectrum of a target Si wafer is automatically normalized.
-The double beam mode effectively divides out variations in the mean transmission levels attributable to lamp drift at 0.2\% in single beam measurements.
-The drift in mean transmission level is merely 0.02\% in double beam mode.
+The double beam mode effectively divides out the 0.2\% drift in the mean transmission level attributable to lamp drift, as seen in single beam measurements.
+In double beam mode, the drift in mean transmission level is merely 0.02\%.
 These values are accurate for measurements over a half hour timescale or less.

### 2014-03-31 AO_bondedSi.tex
Add corrections from DTJ up to page 4

@@
 Air cavities in between the Si interface will have large Fresnel reflections which will interfere as a low-Finesse Fabry-P\`erot etalon.
-Even small 35 nm gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = $ 1.15 $\mu$m at normal incidence.
+Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = $ 1.15 $\mu$m at normal incidence.
 Stock near infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% photometric precision.
-We can detect measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
+We can use instruments like the Cary 5000 to detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
 Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
@@
 
-Direct optical bonding of silicon allows manufacturers to combine parts with no Fresnel loss at the interface.
+Direct optical bonding of silicon allows manufacturers to combine Si parts with no Fresnel loss at the interface.
 The high refractive index of Si makes this advantage particularly useful.
@@
 
-Issues with the quality of the bonds become particular acute when the optics are used in double-pass and/or when the incidence angle at the bonds is far from the normal.
+%Issues with the quality of the bonds become particularly acute when the optics are used in double-pass and/or when the incidence angle at the bonds is far from the normal.
 
@@
 Large, thick Si substrates are also prohibitively expensive to experiment on.
-Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by reducing the thickness and mass needed for the development of a thin (0.5$-$10 mm) Silicon diffractive optic that can later be bonded to a fine polished Si prism substrate.
-Other groups producing SI diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
+Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon puck that can later be bonded to a fine polished Si prism substrate.
+Other groups producing Si diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
 
-While it is possible that proper precautions to eliminate all micron-sized particles from both SI surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments (\cite{Cooper_1986}) means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
-Small particles could be a concern if they act as "tent poles", ti create gaps where the bonding is incomplete.
+While it is possible that proper precautions to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
+Small particles could be a concern if they act as "tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
+Authors have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
+They have also been referred to as voids, or interfacial voids, or gaps.
+In this paper we refer to them as gaps.
 
-There are two major concerns with Si direct bonding: the performance and robustness.
-Specifically, what happens if the bonding process is imperfect and there are small gaps between the Si parts?
-Does a gap degrade the performance of a final device, either through optical aberrations or transmission loss?
-How mechanically robust is the final unit?
-What is the strength of the bond?
-Can the interface be broken by shearing, torquing, pulling, or thermal stress?
-In practice, the figure of merit for robustness is whether the part can survive a shake test simulating the forcing experienced during a rocket launch or transport, or during cryogenic cycling, if the part will be used at low temperature.
-Optically, the relevant parameters are the loss caused by the gap and the effective distortion of the transmitted wavefront.
+Finite gaps will degrade the performance of a final device, either through optical aberrations or transmission loss.
+The effect of the gap size on the IR transmission is the topic of this paper.
+The effect of the gap on optical aberrations is beyond the scope of this paper.
+A gap might also degrade the mechanical integrity of a final compound optic.
+In practice, the part must survive thermal shocks and shake tests.
+The mechanical properties of the gap are also beyond the scope of this paper.
 
-There are two main solutions for dealing with bubbles.
+There are two main solutions for reducing risk attributable to gaps.
 We will refer to the two approaches as the \emph{strength approach} and the \emph{resilience approach}.
 The \emph{strength approach} is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
-Typically this approach demands disciplined handling protocols, and rigorous multi-stage cleaning steps, although unconventional micro-cleanroom approaches exist \cite{1989JaJAP..28L2141L}.
-The \emph{resilience approach} is simply to deal with the uncomfortable likelihood that bubbles will occur, and then do something to remove them.
-The most cited way to get rid of IR bubbles is to anneal them\cite{Mitani1990}.
-We describe the current understanding of both approaches in section \ref{secHistory}.
+Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.
+The reason the strength approach is so hard is that hydrocarbons serve as catalysts for gap formation\cite{1991JaJAP..30..615M}.
+Hydrocarbons are common in Si wafer handling equipment, so avoiding hydrocarbons is challenging.
+Unconventional micro-cleanroom approaches have been effective at meeting this goal \cite{1989JaJAP..28L2141L}.
+The \emph{resilience approach} is simply to accept the uncomfortable likelihood that bubbles will occur, and then do something to remove them.
+The most cited way to get rid of gaps is to anneal them\cite{Mitani1990}.
+We describe the current understanding of both approaches in Section \ref{secHistory}.
 
+Whatever technique one uses to remove or reuce the effects of bubbles, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.
 This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.
-Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}, including IR interference\cite{Mitani1990}, X-ray topography\cite{1994JaJAP..33....6H}, and ultrasound microscopy\cite{2000RScI...71.1869G}.
-The standard infrared interference technique is to examine the bond in reflection for the presence of Newton rings.
-The presence and number of such rings can yield information about both the lateral and axial extend of any gap, as long as the gap along the direction of propagation is of the order of the vacuum wavelength or larger.
+Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}.
+One key technique is IR Newton rings from IR imaging\cite{Mitani1990}.
+This paper gives deeper understanding of the infrared interference technique seen in IR images as Newton rings.
+The number and morphology of Newton rings yields information about the lateral and axial extent of gaps.
+Specifically, dark-light fringes manifest when the gap is comparable to the vacuum wavelength or larger.
+Our technique is IR interference in the limit of small gaps.
+
+There are other techniques for mapping gaps.
 X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.
@@
 
-Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared wavelength range to measure the axial extent of gaps at the few nanometer size scale.
+Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.
 In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.
@@
 It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore the optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.
+We do not pursue further the question of mechanical robustness.
 By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
-
-Our new technique for measurement of the axial extent of gaps exploits wavelength-dependence of reflection losses at the gap and the availability of commercial scanning monochromator systems with both high precision and good amplitude stability.
-We first describe the theory behind the technique and predicted outcomes for various gap dimensions.
-We then evaluate the technique experimentally on a variety of bonded samples.
 
@@
 After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.
-In some of cases, the bond propagation could be directly monitored\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G}.
-The bond process could be shown to be bubble-free\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L}.
-The bond surface energy was directly measured\cite{1989JaJAP..28L2141L}.
-The causes and remedies for bubble formation were discussed as early as 1990\cite{Mitani1990}.
-Specifically, the IR bubble density changes as a function of anneal temprature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.
+\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
+\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding.
+\cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
+\cite{Mitani1990} discussed the causes and remedies for bubble formation.
+Specifically, the IR bubble density changes as a function of anneal temperature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.
 The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
-A robust picture of the chemistry as a function of anneal temperature was ultimate fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
 Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
-No where in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.
+Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
+Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.
+Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.
 

### 2014-04-01 AO_bondedSi.tex
add DTJ edits from page 5

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiBondTentCatFig.pdf}}
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiTentCatFig2.pdf}}
 \caption{Schematic of a particle preventing bonding of two Si surfaces\label{figParticle}.
@@
 \emph{Right-} The particle serves as a catalyst for gas formation.
-There is evidence that hydrocarbons can serve as catalysts, the especially pernicious particles are ``thermally unstable hydrocarbon contaminants on the surfaces coming from plastic containers or simply from the ambient in a cleanroom.'' \cite{1998AnRMS..28..215G}.
+There is evidence that hydrocarbons can serve as catalysts.
+The especially pernicious particles are ``thermally unstable hydrocarbon contaminants on the surfaces coming from plastic containers or simply from the ambient in a cleanroom'' \cite{1998AnRMS..28..215G}.
 }
@@
 A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
-In the next section we show that smaller bubbles can unacceptably degrade the performance to as low as 10 nm.
+In the next section we show that smaller bubbles can unacceptably degrade the performance to as low as 35 nm.
 In a class 100 cleanroom, the volume density could be as high as $\rho=100$ particles/ft$^3$, with sizes larger than 0.5 $\mu$m.
-What we care about is not the volume number density of particles, but the delivered surface flux $F=\rho(D) v(D)$, where v(D) is the average freefall speed of particles, which is a function of particle size.
-Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$-2$ s$-1$) vs.
+What we care about is not the volume number density of particles, but the delivered surface flux $F=\rho(D) v(D)$, where $v(D)$ is the average freefall speed of particles, which is a function of particle size.
+Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$^{-2}$ s$^{-1}$) vs.
 particle size ($\mu$m).
-For a change from 1 $\mu$m to 10$-2$ $\mu$m, there is an increase in a factor of $10^4$ in the particle flux expected.
+For a change from 1 $\mu$m to 10$^{-2}$ $\mu$m, there is an increase in a factor of $10^4$ in the particle flux expected.
 In other words:
@@
 
-So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles, which means for every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
+So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
+For every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
 One key idea that saves the day, is the intuition that the areal fill factor scales at least as fast as $D^2$.
@@
 The volume beneath the tentpole looks like a cone.
-Since the number density goes down as roughly $D^-2$, (per the equation above) the total area covered by particles might be constant in $log(D)$.
+Since the number density goes down as roughly $D^{-2}$, (per the equation above) the total area covered by particles might be constant in $log(D)$.
 Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.
@@
 
-It has been shown\cite{1992JEMat..21..669M} that the gases seep through the Si$-$Si interface, so pressure builds around those particles that serve as catalysts.
-In this sense, the size of the particle causing an IR bubble cannot be uniquely identified from observations of the extent of IR bubble defects.
+It has been shown\cite{1992JEMat..21..669M} that the gases seep through the Si$-$Si interface.
+Pressure builds around catalysts.
+In this sense, the size of the particle causing a gap cannot be uniquely identified from observations of the extent of gaps.
 In section \ref{secResults} we address the question of the origin of the IR bubbles.

### 2014-04-15 AO_bondedSi.tex
incorporate DTJ edits up to page 9, overhaul table 1

@@
 Large, thick Si substrates are also prohibitively expensive to experiment on.
-Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon puck that can later be bonded to a fine polished Si prism substrate.
-Other groups producing Si diffractive optics are also experimenting with this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
+Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon wafer or puck that can later be bonded to a fine polished Si prism substrate.
+Other groups producing Si diffractive optics are also experimenting with or using this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
 
@@
 The effect of the gap on optical aberrations is beyond the scope of this paper.
+
 A gap might also degrade the mechanical integrity of a final compound optic.
 In practice, the part must survive thermal shocks and shake tests.
-The mechanical properties of the gap are also beyond the scope of this paper.
+The gap size scales we can probe with our optical technique are comparable to the range over which molecular (van der Waals) forces are acting.
+It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore our optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.
+By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
+We do not investigate the mechanical robustness further.
 
 There are two main solutions for reducing risk attributable to gaps.
-We will refer to the two approaches as the \emph{strength approach} and the \emph{resilience approach}.
-The \emph{strength approach} is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
+The first approach is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
 Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.
-The reason the strength approach is so hard is that hydrocarbons serve as catalysts for gap formation\cite{1991JaJAP..30..615M}.
-Hydrocarbons are common in Si wafer handling equipment, so avoiding hydrocarbons is challenging.
-Unconventional micro-cleanroom approaches have been effective at meeting this goal \cite{1989JaJAP..28L2141L}.
-The \emph{resilience approach} is simply to accept the uncomfortable likelihood that bubbles will occur, and then do something to remove them.
+The reason the approach is so hard is that hydrocarbons can serve as catalysts for gap formation\cite{1991JaJAP..30..615M}, and that these hydrocarbons are common in Si wafer handling equipment.
+Unconventional micro-cleanroom approaches have been effective at meeting the goal of bubble-free bonding\cite{1989JaJAP..28L2141L}.
+The second approach is to accept the likelihood that gaps will occur, and take steps to remove them.
 The most cited way to get rid of gaps is to anneal them\cite{Mitani1990}.
@@
 
-Whatever technique one uses to remove or reuce the effects of bubbles, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.
+Whatever technique one uses to remove or reduce the effects of gaps, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.
 This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.
-Other techniques for measuring and mapping gaps have already been described elsewhere \cite{1992JEMat..21..669M}.
+Mitani and G\"oselle review other techniques for measuring and mapping gaps\cite{1992JEMat..21..669M}.
 One key technique is IR Newton rings from IR imaging\cite{Mitani1990}.
 This paper gives deeper understanding of the infrared interference technique seen in IR images as Newton rings.
-The number and morphology of Newton rings yields information about the lateral and axial extent of gaps.
+The number and morphology of Newton rings yields information about the lateral and axial extent of large gaps.
 Specifically, dark-light fringes manifest when the gap is comparable to the vacuum wavelength or larger.
-Our technique is IR interference in the limit of small gaps.
+Our technique is IR interference in the limit of small gaps down to 20 nm or less.
 
@@
 X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.
-Ultrasound microscopy exploits the density change at the gap interface to produce reflections of sound waves and map the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
+In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
 
-Here we describe an exploitation of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.
+Here we describe an application of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.
 In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.
-Because the size scales probed are so small, they are comparable to the range over which molecular (van der Waals) forces are acting.
-It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore the optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.
-We do not pursue further the question of mechanical robustness.
-By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
 
@@
 \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
-\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding.
-\cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
+\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding in a micro-cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
+Our group has not attempted the micro-cleanroom technique owing to the challenge of spin-cleaning large (30 mm thick) Si substrates.
+It is not clear whether the custom micro-cleanroom technique has gained wide usage.
+Notably, the   \cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
 \cite{Mitani1990} discussed the causes and remedies for bubble formation.
@@
 It is further conceivable that the areal fill factor attributable to a tent pole particle scales even faster than $D^2$ since surface energy probably scales inversely with distance of the pieces (\`a la VanDerWaals).
+The small particles are even less of a problem than the large particles, since the small ($\sim10-60$ nm) particles will have lower average etalon reflectivity than larger (500-1000 nm) particles.
+This last point will become clearer in later sections.
 
@@
 
-\subsection{Surface roughness or large-scale aberrations}
+\subsection{Surface roughness or large-scale non-conformance}
+Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform.
 The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-This issue is much less of a problem for wafers compared to our unconventional thick ($\sim$ 30 mm) Si applications.
-Similarly, surface roughness will also limit the bonding strength  \cite{JJAP.37.4197,2001JOptA...3...85G}.
-The details of this aspect of the bonding process are outside the scope of this paper.
-In short, most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
-We expect such well-polished substrates to have no deleterious impact on the bonding process.
-Some experimental substrates had larger surface error, of order a few waves of surface error.
-These substrates were reserved for process development, since we anticipated that bonding might progress either not at all, or with large interface gaps.
+Unconventional thick ($\sim$ 30 mm) Si applications a not as conformal as wafers, so we expect thick substrates could have gaps even in the absence of particles.
+Most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
 
@@
 \caption{Schematic of part VG03 \label{figVG03pattern}.
-The four inch diameter VG03 is comprised of two 3.3 mm thick silicon pucks.
-We patterned a one inch diameter hole on one of the pucks.
+The 100 mm diameter VG03 is comprised of two 3.3 mm thick silicon pucks.
+We patterned a 25 mm diameter hole on one of the pucks.
 The depth was 4.0 $\mu$m as indicated by Dektak profilometry.
 We also patterned a low-fill factor petal pattern surrounding the hole.
-The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.
-The technique can reveal the axial extent and fill factor of an embedded gap.}
+The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.}
 \end{figure}
 
-We evaluated our metrology method by imbedding gaps of known sizes in the interface.
-Here we describe the process of how we made the gaps and how we characterized their axial extent.
+We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.
+
+The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
+We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
+The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
+There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+
 The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.
-We bored divots in silicon with inductively coupled plasma etching.
-The tool was a Unaxis.
+We bored holes in silicon with inductively coupled plasma etching.
 We used two different gases with low and high etch rates.
@@
 We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
-Figure \ref{figX} shows a schematic of VG03 and its pattern.
-In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates.
-For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{fig}.
+Table \ref{tabbondexper} has more information about the bonding samples.
+Figure \ref{figVG03pattern} shows a schematic of VG03 and its pattern, as described in Table \ref{tabbondexper}.
+In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
 The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
@@
     \hline
-    Name & Bonded/DSP & Diameter & Thickness & Anneal Temp & Date & Pattern notes \\ 
-    - & - & mm & mm & $^\circ$C & MM/DD/YYY & - \\ 
+    Name & Bonded/DSP & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
+    - & - & mm & mm & - & nm & - \\ 
         \hline
-    VG01 & Bonded  & 100  &$1$ $\times$ 2 & 23  & 2/13/2013 &  \\
-    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  23  & 2/13/2013 &  \\
-    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  23 & 2/14/2013 & 4 $\mu$m gap 100\% fill factor hole \& petal pattern\\    
-    VG04 & Bonded & 100 & $1$ $\times$ 2 &  23 & 2/14/2013 & 15 nm gap, $\sim50\%$ fill factor\\        
-    VG05 & DSP & 100 & 1&  23 & 2/15/2013 & cleaved  \\
-    VG06 & DSP & 100 & 3.5 &  23 & 2/15/2013 & \\
+    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & 2/13/2013 &  \\
+    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & 2/13/2013 &  \\
+    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
+    VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
+    VG05 & DSP & 100 & 1&  - & - & -  \\
+    VG06 & DSP & 100 & 3.5 &  - & - & - \\
     VG07 & - & - & - &  - & - & \\
     VG08 & - & - & - &  - & - & \\
-    VG09 & Bonded & 75 & 1.5? & 23 & 6/21/2013 & mesh C, 49 $\pm$6 nm, 50\% fill factor\\
-    VG15 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 14 $\pm$2 nm, 50\% fill factor\\
-    VG17 & Bonded & 75 & 6.0? &  23 & 6/21/2013 & mesh C, 64 $\pm$6 nm, 50\% fill factor\\
-    VS20 & Bonded & 100 & 1.5? &  23 & 6/21/2013 & mesh C, M, F, 95 $\pm$5 nm, 50\% fill factor\\
+    VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
+    VG15 & Bonded & 100 & 1.5? &  mesh C, M, F  &14 $\pm$2 & \\
+    VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
+    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & \\
     \hline
@@
 \section{Substrate preparation}
-It is important to clean the surfaces before bonding to minimize the particle density.
+We cleaned the surfaces before bonding to minimize the interfacial particle density.
 The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
-The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
-We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
-The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
 We measured the large scale surface flatness of the pucks.

### 2014-04-15 AO_bondedSi.tex
more tweaks from DTJ up to page 10

@@
 They have also been referred to as voids, or interfacial voids, or gaps.
-In this paper we refer to them as gaps.
+In this paper we will reserve the term ``IR bubbles'' for gaps that are large enough to be detected directly in conventional IR imaging.
+The term ``gap'' will apply more broadly to any finite gap, which exceeds the scale over which interfacial molecular forces are acting, roughly say 2 nm.
 
@@
 The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
 
@@
         \hline
-    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & 2/13/2013 &  \\
-    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & 2/13/2013 &  \\
+    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & - &  \\
+    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
     VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
     VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
-    VG05 & DSP & 100 & 1&  - & - & -  \\
-    VG06 & DSP & 100 & 3.5 &  - & - & - \\
+    VG05 & DSP & 100 & 1&  - & - & Reference Sample  \\
+    VG06 & DSP & 100 & 3.5 &  - & - & Reference Sample \\
     VG07 & - & - & - &  - & - & \\
@@
 We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
-
-The detailed bonding procedure will be chronicled in a future paper.
-Briefly, we prepared the substrates with standard cleaning procedures of solvents in a megasonic.
+We prepared the substrates with standard cleaning procedures of solvents in a megasonic.
 We then applied MHz frequency oxygen plasma ashing.
 We soaked the wafers in DI water then dried with an N$_2$ gun.
-We pressed the wafers together from the center to the outside with normal human force.
+We pressed the wafers together from the center to the outside.
 
@@
 We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
-We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.
+Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.
 We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
+In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.
 The computation is simply Fresnel's law at normal incidence:
@@
 The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
-We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
-The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
-In the next subsection we work out the wavelength dependent transmission for the air etalon
+We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
+The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
+In the next subsection we work out the wavelength dependent transmission for the air etalon.
 
@@
 \label{secTheory}
-There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate.
+There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.
 Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
@@
 
-which has an average value of about 53\%, as shown in the top curve of the Figure above.
+which has an average value of about 53\%.
 Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+If the Si wafer path is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
 This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
@@
 \end{center}
+
+\subsection{IR bubble detection and limitations}
 

### 2014-04-16 AO_bondedSi.tex
overhaul Table 2

@@
 \section{Measurement strategy, conditions, and details}
-The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \cite{cary5000web}.
+We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers.
+The spectrophotometer is a Agilent Cary 5000 UV-Vis-NIR.
 The typical measurement parameters and tool performance are listed below.
@@
 
-Table \ref{tab:tablenum3} shows some properties of the Cary5000.
-
-%\caption{Summary of Cary 5000 Measurement parameters \label{tab:tablenum3}}
+\begin{table}[h!]
+\caption{Summary of Cary 5000 Measurement parameters \label{tab:tablenum3}}
 \begin{center}
-    \begin{tabular}{l l}
+\begin{tabular}{ c c }
+\hline
         Parameter (Units) & Value \\ 
+\hline
         Spectral sampling interval (nm) & 2.0 \\
@@
         Directly measured linearity range (dB) & 24 \\
+    \hline
     \end{tabular}
 \end{center}
+\end{table}
 

### 2014-04-16 AO_bondedSi.tex
more DTJ edits from an old draft

@@
 The spectrophotometer is a Agilent Cary 5000 UV-Vis-NIR.
-The typical measurement parameters and tool performance are listed below.
+The typical measurement settings and tool performance are listed in Table \ref{tab:Cary5000pars}.
 More details about the tool are available at the vendor's website.
@@
 Double mode takes a reference spectrum simultaneously to the target spectrum.
-In both single and double beam mode a baseline scan is taken of air.
-The baseline spectrum is the raw spectral response of the system, which is automatically divided out of subsequent spectra.
-In single beam mode the baseline measurements must be repeated about every half hour.
+In both single and double beam mode, a baseline scan is taken with no sample in the measurement holder.
+The baseline spectrum is the raw spectral response of the system.
+We automatically divide sample spectra by the baseline spectrum.
+In single beam mode, the baseline measurements must be repeated about every half hour.
 We estimated the uncertainty and measurement repeatability of single and double modes in the following way.
-We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of the beam with no sample in a series of spectra taken immediately after a 100\% baseline scan.
+We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
 The predicted mean level should naturally be 100\%.
-We found that the mean value in single mode drifted by about 0.2\%.
-The standard deviation of the featureless spectrum was typically 0.02\%, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.
-In double beam mode the spectrograph performance was much better.
-We put a double side polished Si wafer in the reference beam, and no sample in the target beam during the baseline scan.
+We found that the mean value in single mode drifted by about 0.2\% over several measurements.
+The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption; we list the wavelengths of instrumental artifacts and above-average variance in Table \ref{tab:Cary5000pars}.
+In double beam mode, the spectrograph performance was much better.
+We put a double side polished (DSP) Si wafer in the reference beam, and no sample in the target beam during the baseline scan.
 In this way the transmission spectrum of a target Si wafer is automatically normalized.
@@
 These values are accurate for measurements over a half hour timescale or less.
-Over periods of hours the drift increases for both single and double beam mode, so a repeat baseline scan is sometimes necessary.
-The need for repeat baselines is more acute for single mode.
 
 The choice of spectral sampling was straightforward.
-We expect smoothly varying spectral features for small ($<$ 1 $\mu$m) size gaps in Si bonded wafers.
-Sampling theory dictates a few spectral resolution elements over the spectral feature of interest.
-The Fabry-P\`erot fringes will be well-sampled so long as the gap is less than $\sim$100 $\mu$m.
-Such large gaps would be easily detectable by other means.
-So we simply pick the largest slit width, and therefore the lowest spectral resolution that the Cary 5000 permits: 5 nm.
-We then sample the spectral resolution with $\sim$2.5 samples per spectral resolution element, 2.0 nm samples.
+The spectral sampling was 2.0 nm.
+The spectral resolution was 5.0 nm, the lowest spectral resolution the Cary5000 offers.
+We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
+Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
+Such large gaps would be easily detectable by other means, like IR imaging.
 
 \begin{table}[h!]
-\caption{Summary of Cary 5000 Measurement parameters \label{tab:tablenum3}}
+\caption{Summary of Cary 5000 Measurement parameters \label{tab:Cary5000pars}}
 \begin{center}
@@
         Typical measurement range (nm) & 1000-2500 \\
-	Single mode RMS  (percent) & 0.02 \\
-	Double mode RMS  (percent) & 0.002 \\
-	Single mode mean drift (percent) & 0.20 \\
-	Double mode mean drift (percent) & 0.02 \\
-	Mean 0\% baseline correction (percent) & 0.03 \\
- 	Beam size (mm x mm) & 2 x 8 \\
-	Wavelengths of instrumental artifacts (nm) & stuff \\
-        num1 & 1200 \\
-        num2 & 1350 \\
-        num3 & 1360-1450\\
-        num4 & 1850 \\
-        num5 & 1800-1950 \\
-        num6 & 2000 \\
-        Vendor-reported linearity range (dB) & 40 \\
-        Directly measured linearity range (dB) & 24 \\
+	Single mode RMS  (\%) & 0.02 \\
+	Double mode RMS  (\%) & 0.002 \\
+	Typical Single mode mean drift (\%) & 0.20 \\
+	Typical Double mode mean drift (\%) & 0.02 \\
+	No beam dark signal (\%) & 0.03 \\
+ 	Beam size (mm $\times$ mm) & $2 \times 8$ \\
+	Wavelengths of instrumental artifacts (nm) & - \\
+        		 & 1200 \\
+		 & 1350 \\
+		 & 1360-1450\\
+		 & 1850 \\
+		 & 1800-1950 \\
+		 & 2000 \\
+        Vendor-reported linearity range (dB) & $>40$ \\
+        Directly measured linearity range (dB) & $>24$ \\
     \hline
@@
 We detected IR bubbles in all 3 of the bonded Si pucks on which we recorded IR images.
-The bubble areal density varied from about 4 per wafer (100 mm diameter) to over 20 per wafer.
-Naturally, the sample of control solitary DSP (unbonded) wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrtes as discussed in the literature (see Section \label{secHistory}).
+The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
+Naturally, the unbonded DSP wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrates as discussed in the literature (see Section \ref{secHistory}).
 Figure \ref{IRbubble} shows an example of one of the IR bubbles in which several bright and dark fringes are clearly discernible.
-No attempt at quantitative measurement was made on the IR images due to anecdotal evidence that the detector is non-linear.
+No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.
 The morphology was clear regardless of the detector response.
@@
 The yellow circle in the photo highlights the outline of the location of an IR bubble we detected.
-The bubble shows a couple of fringes over the effective wavelength range 1.1-1.6 $\mu$m.
+The bubble shows $\sim2$ fringes over the effective wavelength range 1.1-1.6 $\mu$m.
 This fringe pattern is consistent with a gap comparable to this measurement wavelength.

### 2014-04-21 AO_bondedSi.tex
tweaks from DTJ edits, add table for meshes

@@
 Small particles could be a concern if they act as "tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
-Authors have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
-They have also been referred to as voids, or interfacial voids, or gaps.
+Some authors \cite{1991JaJAP..30..615M, 1992JEMat..21..669M, 1989JaJAP..28L2141L, Mitani1990} have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
+The gaps have also been referred to as voids\cite{2000RScI...71.1869G}, interfacial voids, or gaps.
 In this paper we will reserve the term ``IR bubbles'' for gaps that are large enough to be detected directly in conventional IR imaging.
-The term ``gap'' will apply more broadly to any finite gap, which exceeds the scale over which interfacial molecular forces are acting, roughly say 2 nm.
+The term ``gap'' will apply more broadly to any finite size gap, exceeding the scale over which interfacial molecular forces are acting, roughly $>$2 nm.
 
@@
 We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
-Table \ref{tabbondexper} has more information about the bonding samples.
+Table \ref{tabbondexper} has more information about the bonding samples, and Table \ref{tabPatternFills} lists the properties of the patterns.
+The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
 Figure \ref{figVG03pattern} shows a schematic of VG03 and its pattern, as described in Table \ref{tabbondexper}.
@@
     VG06 & DSP & 100 & 3.5 &  - & - & Reference Sample \\
-    VG07 & - & - & - &  - & - & \\
-    VG08 & - & - & - &  - & - & \\
     VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
@@
     VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
-    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & \\
+    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & See Fig. \ref{figVS20pattern}\\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}[h!]
+\caption{Patterned gap properties \label{tabPatternFills}}
+\begin{center}
+    \begin{tabular}{ c c c c }
+    \hline
+    Pattern name & fill factor & minimum feature size & bulk description \\ 
+    - & \% & $\mu$m & '- \\ 
+    Hole & 100 & - & 25 mm diameter hole \\     
+    Petal & $\sim1$ & $\sim$2000 & long petal lines, Fig. \ref{figVG03pattern} \\         
+    Mesh F & 50 & 40 &  100 $\mu$ square holes, Fig. \ref{figVS20pattern}\\ 
+    Mesh M & 50 & 200 &  500 $\mu$ square holes\\ 
+    Mesh C & 50 & 620 &  1500 $\mu$ square holes\\     
         \hline
@@
 We soaked the wafers in DI water then dried with an N$_2$ gun.
-We pressed the wafers together from the center to the outside.
+We pressed the Si substrates together from the center to the outside.
 
@@
 Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-If the Si wafer path is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
 This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
@@
 
-We also looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with the coarse IR imaging technique.
+We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with the coarse IR imaging technique.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 100\% relative transmission).
-For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section ref{secTheory}, specifically equation \ref{eqFP}.
-For non-unit fill factor, we simply average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor.
-Notably the meshes we designed were all 50\% fill factor.
-No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we take the design fill factor as gospel.
+For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section \ref{secTheory}.
+Specifically we apply Equation \ref{eqFP}.
+For non-unit fill factor, we simply average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor listed in Table \ref{tabPatternFills}.
+No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assume the delivered fill factor is equal to the designed fill factor with no uncertainty.
 The depth of the meshes has an associated uncertainty listed in \ref{tabbondexper}.
@@
 \centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1}}
-\caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
+\caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.
+The spectrum was taken at a position through a mesh gap of depth 95 $\pm$5 with 50\% fill factor.
+The predicted spectrum lines up with the measurement to within the measurement and model uncertainties.}
 \end{figure}
@@
 \begin{table}[h!]
-  \caption{Interpretation of the diversity of IR spectra}
+  \caption{Interpretation of the diversity of IR spectra \label{tabInterpretation}}
   \begin{center}
@@
 \subsection{Mitigation of measurement artifacts}
-The principle artifacts and their interpretation are listed in table \ref{1}.
+The principle artifacts and their interpretation are listed in Table \ref{tabInterpretation}.
 We mitigated the artifacts in the following ways.
@@
 Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.
-We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature (see the work of Mitani and collaborators from the mid 1990s).
-Namely, Mitani showed that the gas pressure of bonded wafers grew as the fill factor of voids decreased.
+We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
+Specifically, the gas pressure of bonded wafers grows as the fill factor of voids decreased.
 In other words, what matters is the fill factor of voids, not their relative sizes.
@@
 Large gaps identified this late in the game could render the costly optic useless, both in terms of efficiency and wavefront performance.
-Our group is exploring the effectiveness of high-temperature annealing to eliminate IR bubbles in completed devices.
+Our group is exploring the effectiveness of high-temperature annealing to eliminate interfacial gaps in completed devices.
 

### 2014-04-21 AO_bondedSi.tex
minor tweaks to bubble-gap-void word usage

@@
 The term ``gap'' will apply more broadly to any finite size gap, exceeding the scale over which interfacial molecular forces are acting, roughly $>$2 nm.
+We omit the adjective ``interfacial'' before gap for simplicity.
 
@@
 There are two main solutions for reducing risk attributable to gaps.
-The first approach is to prevent bubbles from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
+The first approach is to prevent gaps from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
 Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.
@@
 A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
-In the next section we show that smaller bubbles can unacceptably degrade the performance to as low as 35 nm.
+In the next section we show that small gaps can unacceptably degrade the performance to as low as 35 nm.
 In a class 100 cleanroom, the volume density could be as high as $\rho=100$ particles/ft$^3$, with sizes larger than 0.5 $\mu$m.
@@
 In this sense, the size of the particle causing a gap cannot be uniquely identified from observations of the extent of gaps.
-In section \ref{secResults} we address the question of the origin of the IR bubbles.
+In section \ref{secResults} we address the question of the origin of interfacial gaps.
 Specifically, are the particles more like \emph{tentpoles}, rigid structures that merely prevent the zippering up of the bond?
@@
 
-We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with the coarse IR imaging technique.
+We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.
-It was easy to detect IR bubbles despite the artifacts since the bubbles translate with the wafer during dithering.
+It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the wafer during dithering.
 We detected IR bubbles in all 3 of the bonded Si pucks on which we recorded IR images.
@@
 In this section we describe the spectroscopy of bonded Si pucks.
-In some instances IR imaging was available and was leveraged to target IR bubbles.
-In other cases the IR imaging was not available or simply ignored, and spectra were taken at random positions on the wafer surface.
+In some instances IR imaging was available and was leveraged to target or avoid locations demonstrating IR bubbles.
+In other cases the IR imaging was not available or ignored, and spectra were taken at random positions on the wafer surface.
 
@@
 We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
-Specifically, the gas pressure of bonded wafers grows as the fill factor of voids decreased.
-In other words, what matters is the fill factor of voids, not their relative sizes.
-Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the voids, and therefore no difference in the bonding properties.
+Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.
+In other words, what matters is the fill factor of gaps, not their relative sizes.
+Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.
 
@@
 \section{Conclusions}
-We have shown a method for indirectly measuring the size of voids in bonded Si wafers.
+We have shown a method for indirectly measuring the size of gaps in bonded Si wafers.
 The technique broadly applies to any compound Si optics measured at normal incidence.

### 2014-04-21 AO_bondedSi.tex
add in two theory figures, make some edits in the text for those figures

@@
 
+\section{Si$-$air$-$Si Fabry P\`{e}rot measurement theory}
+\label{secTheory}
+
+% Figures (shared wavelength axis)
+%	2) Fresnel transmission as a function of wavelength (normal incidence)
+%	4) Example transmission through VG03 thick gap
+
+\subsection{The Si-Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}
+
+Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
+The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+The Fresnel reflection is accordingly about 30\% per surface.
+For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
+Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
+The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
+The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
+\end{figure}
+
+The Si-Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
+We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
+Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.
+We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
+In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.
+The computation is simply Fresnel's law at normal incidence:
+
+\begin{eqnarray}
+R = \frac{(n-1)^2}{(n+1)^2} \\
+F \equiv \frac{4R}{(1-R)^2}
+\end{eqnarray}
+
+The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
+We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
+The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
+In the next subsection we work out the wavelength dependent transmission for the air etalon.
+
+\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
+\label{secTheory}
+There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.
+Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
+The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
+We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.
+Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
+The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
+We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.
+The DSP Si puck with no air gap has a transmission equal to:
+
+$$
+T_{DSP} = \frac{2n}{1+n^2} \label{eqnAbsDSPtrans}
+$$
+
+which has an average value of about 53\%.
+Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
+Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range 1200$-2500$ nm.
+If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.45\columnwidth]{figs/20140421_absolute.pdf}}
+\caption{Modeled absolute transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figAbsoluteTrans}.
+The absolute transmission through Si depends on the Si refractive index, which depends on wavelength.
+}
+\end{figure}
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.45\columnwidth]{figs/20140421_absoluteB.pdf}}
+\caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
+The curves are examples from a family defined in Equation \ref{eqFP}.
+The limit of no gap is 100\% transmissive, compared to a DSP reference.
+A 35 nm gap degrades the transmission over a DSP reference by 4\% at 1200 nm.}
+\end{figure}
+
+Now consider a pair of bonded Si pucks with a gap thickness $d$.
+We derive the absolute transmission of bonded Si substrates with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
+We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.
+We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.
+
+$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}$$
+
+It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.
+Figure \ref{figEtalonRelTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 20 nm and 35 nm.
+In the next section we compare measurements to by-eye fits of this equation.
+
 \section{Imbedding gaps of known sizes}
@@
 We pressed the Si substrates together from the center to the outside.
-
-\section{Si$-$air$-$Si Fabry P\`{e}rot measurement theory}
-\label{secTheory}
-
-% Figures (shared wavelength axis)
-%	2) Fresnel transmission as a function of wavelength (normal incidence)
-%	4) Example transmission through VG03 thick gap
-
-\subsection{The air cavity is a low Finesse Fabry-P\`{e}rot etalon}
-
-Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
-The Fresnel reflection is accordingly about 30\% per surface.
-For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
-Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
-The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
-The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
-\end{figure}
-
-The air cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
-Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.
-The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
-Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.
-We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
-In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.
-The computation is simply Fresnel's law at normal incidence:
-
-\begin{eqnarray}
-R = \frac{(n-1)^2}{(n+1)^2} \\
-F \equiv \frac{4R}{(1-R)^2}
-\end{eqnarray}
-
-The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
-We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
-The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
-In the next subsection we work out the wavelength dependent transmission for the air etalon.
-
-\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
-\label{secTheory}
-There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.
-Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
-The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.
-Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
-The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.
-The DSP Si puck with no air gap has a transmission equal to:
-
-$$
-T_{DSP} = \frac{2n}{1+n^2}
-$$
-
-which has an average value of about 53\%.
-Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
-This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
-%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
-
-Now consider a pair of bonded Si pucks with a gap thickness $d$.
-We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
-We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.
-We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.
-
-$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}$$
-
-It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.  In the next section we compare measurements to by-eye fits of this equation.
 
@@
 No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assume the delivered fill factor is equal to the designed fill factor with no uncertainty.
-The depth of the meshes has an associated uncertainty listed in \ref{tabbondexper}.
+The depth of the meshes has an associated uncertainty listed in Table \ref{tabbondexper}.
 We incorporated the uncertainty in the mesh depth in the following way.

### 2014-04-21 AO_bondedSi.tex
add another figure for large gap theory, some concommittant tweaks

@@
 
-\section{Si$-$air$-$Si Fabry P\`{e}rot measurement theory}
+\section{Si$-$Si Fabry P\`{e}rot measurement theory}
 \label{secTheory}
@@
 
-\subsection{The Si-Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}
+\subsection{The Si$-$Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}
 
-Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.
+Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
 The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
@@
 
-The Si-Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
 The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
@@
 Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.
-We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
+We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
 In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.45\columnwidth]{figs/20140421_absolute.pdf}}
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absolute.pdf}}
 \caption{Modeled absolute transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figAbsoluteTrans}.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.45\columnwidth]{figs/20140421_absoluteB.pdf}}
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absoluteB.pdf}}
 \caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
@@
 Figure \ref{figEtalonRelTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 20 nm and 35 nm.
-In the next section we compare measurements to by-eye fits of this equation.
+In Section \ref{secResults} we compare wavelength dependent transmission measurements to by-eye fits of Equation \ref{eqFP}.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absoluteC.pdf}}
+\caption{Modeled etalon transmission through a silicon substrate with gaps of axial extent 35 nm (red dotted curve), 350 nm (green dash-dot curve), and 3500 nm (blue long-dashed curve)\label{figEtalonRelTrans}.
+The curves are examples from a family defined in Equation \ref{eqFP}.
+The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.
+}
+\end{figure}
 

### 2014-04-21 AO_bondedSi.tex
tweak MGS edits in intro, etc

@@
 
-While it is possible that proper precautions to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
-Small particles could be a concern if they act as "tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
+While it is possible that proper precautions can be made to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
+Small particles could be a concern if they act as ``tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
 Some authors \cite{1991JaJAP..30..615M, 1992JEMat..21..669M, 1989JaJAP..28L2141L, Mitani1990} have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
@@
 
+%The following 3 figures were made with fabry_perot_Si_20140421.pro and other variants.
 \begin{figure}[htbp]
@@
 The curves are examples from a family defined in Equation \ref{eqFP}.
-The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.
-}
+The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.}
 \end{figure}

### 2014-04-21 AO_bondedSi.tex
minor tweaks from MGS edits

@@
 Small particles could be a concern if they act as ``tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
-Some authors \cite{1991JaJAP..30..615M, 1992JEMat..21..669M, 1989JaJAP..28L2141L, Mitani1990} have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
+Some authors \cite{1991JaJAP..30..615M, 1992JEMat..21..669M, 1989JaJAP..28L2141L, Mitani1990} have referred to large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
 The gaps have also been referred to as voids\cite{2000RScI...71.1869G}, interfacial voids, or gaps.
 In this paper we will reserve the term ``IR bubbles'' for gaps that are large enough to be detected directly in conventional IR imaging.
-The term ``gap'' will apply more broadly to any finite size gap, exceeding the scale over which interfacial molecular forces are acting, roughly $>$2 nm.
-We omit the adjective ``interfacial'' before gap for simplicity.
+The term ``gap'' will apply more broadly to any finite size gap.
 
@@
 Mitani and G\"oselle review other techniques for measuring and mapping gaps\cite{1992JEMat..21..669M}.
-One key technique is IR Newton rings from IR imaging\cite{Mitani1990}.
-This paper gives deeper understanding of the infrared interference technique seen in IR images as Newton rings.
+One key technique is identifying IR Newton rings from IR imaging\cite{Mitani1990}.
 The number and morphology of Newton rings yields information about the lateral and axial extent of large gaps.
@@
 There are other techniques for mapping gaps.
-X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight with gaps to allow precise and high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but provides only coarse estimates of the axial extent.
+X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers.
 In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
@@
 It is not clear whether the custom micro-cleanroom technique has gained wide usage.
-Notably, the   \cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
+\cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
 \cite{Mitani1990} discussed the causes and remedies for bubble formation.
@@
 Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.
-Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that wasn't widely available until this decade.
+Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that was not widely available until this decade.
 
@@
 
-Consider the following two conceivable pitfalls in Si-Si direct bonding.
+There are two major conceivable pitfalls in Si-Si direct bonding.
 The obvious pitfall is particle contamination.
@@
 A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
-In the next section we show that small gaps can unacceptably degrade the performance to as low as 35 nm.
-In a class 100 cleanroom, the volume density could be as high as $\rho=100$ particles/ft$^3$, with sizes larger than 0.5 $\mu$m.
+In a class 100 cleanroom, the particle volume density could be as high as $\rho=100$ particles/ft$^3$, for particle sizes larger than 0.5 $\mu$m.
 What we care about is not the volume number density of particles, but the delivered surface flux $F=\rho(D) v(D)$, where $v(D)$ is the average freefall speed of particles, which is a function of particle size.
@@
 particle size ($\mu$m).
-For a change from 1 $\mu$m to 10$^{-2}$ $\mu$m, there is an increase in a factor of $10^4$ in the particle flux expected.
+There will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
 In other words:
-$$\log(F) = -2\log(D) + const.$$
+$$\log(F) = -2\log(D) + const. \label{eqnParticleFlux}$$
 
-So there will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
 For every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
@@
 The volume beneath the tentpole looks like a cone.
-Since the number density goes down as roughly $D^{-2}$, (per the equation above) the total area covered by particles might be constant in $log(D)$.
+Since the number density scales as roughly $D^{-2}$, (per the Equation \ref{eqnParticleFlux} above) the total area covered by particles might be constant in $log(D)$.
 Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.
@@
 
-It has been shown\cite{1992JEMat..21..669M} that the gases seep through the Si$-$Si interface.
+It has been shown\cite{1992JEMat..21..669M} that gases produced at the Si$-$Si interface seep along the interface surfaces.
 Pressure builds around catalysts.
 In this sense, the size of the particle causing a gap cannot be uniquely identified from observations of the extent of gaps.
-In section \ref{secResults} we address the question of the origin of interfacial gaps.
+In Section \ref{secResults} we address the question of the origin of gaps.
 Specifically, are the particles more like \emph{tentpoles}, rigid structures that merely prevent the zippering up of the bond?
@@
 Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform.
-The second pitfall to consider is that the stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-Unconventional thick ($\sim$ 30 mm) Si applications a not as conformal as wafers, so we expect thick substrates could have gaps even in the absence of particles.
-Most of our thick substrates are optically polished to much less than $\lambda = 632$ nm peak to valley over the 3 to 4 inch diameter clear aperture.
+Stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
+Unconventional thick ($\sim$ 30 mm) Si pucks are less conformal than wafers are, so we expect thick substrates could have gaps even in the absence of particles.
+Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 3 to 4 inch diameter clear aperture.
+We assume that if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
+Subsequent annealing will strength the bond.
 
@@
 The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
-The Fresnel reflection is accordingly about 30\% per surface.
+The Fresnel reflection is about 30\% per surface.
 For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
@@
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
-In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.
+In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
 The computation is simply Fresnel's law at normal incidence:
@@
 The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.
+We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no gap, and second a pair of bonded Si pucks with a gap thickness $d$.
 The DSP Si puck with no air gap has a transmission equal to:
@@
 $$
-T_{DSP} = \frac{2n}{1+n^2} \label{eqnAbsDSPtrans}
+T_{DSP} = \frac{2n}{1+n^2} \label{eqnAbsDSPtrans} \\
 $$
@@
 
-$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}$$
+$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP} \\$$
 

### 2014-04-29 AO_bondedSi.tex
some tweaks

@@
 \caption{Modeled absolute transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figAbsoluteTrans}.
-The absolute transmission through Si depends on the Si refractive index, which depends on wavelength.
-}
+The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.
+This curve is only accurate for normal incidence transmission.
+Furthermore, the curve assumes the multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.}
 \end{figure}
@@
 \begin{center}
-    \begin{tabular}{ c c c c c c c}
+    \begin{tabular}{ c c c c c c}
     \hline
-    Name & Bonded/DSP & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
-    - & - & mm & mm & - & nm & - \\ 
+    Name & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
+    -  & mm & mm & - & nm & - \\ 
         \hline
-    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & - &  \\
-    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
-    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
-    VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
-    VG05 & DSP & 100 & 1&  - & - & Reference Sample  \\
-    VG06 & DSP & 100 & 3.5 &  - & - & Reference Sample \\
-    VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
-    VG15 & Bonded & 100 & 1.5? &  mesh C, M, F  &14 $\pm$2 & \\
-    VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
-    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & See Fig. \ref{figVS20pattern}\\
+    VG01  & 100  &$1$ $\times$ 2 & None  & - &  \\
+    VG02   & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
+    VG03   & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
+    VG04   & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
+    VG09   & 75 & $1$ $\times$ 2 & mesh C & 49 $\pm$6 & \\
+    VG15   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F  &14 $\pm$2 & \\
+    VS20   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F & 95 $\pm$5 & See Fig. \ref{figVS20pattern}\\
     \hline

### 2014-05-01 AO_bondedSi.tex
Add a huge table- Experiment Log for all experiments so far
corrections to Tables, inclusion of details on VG07 and VG08

@@
         \hline
+    HA1     & 75   & $10$ $\times$ 2& None & - & \\
     VG01   & 100 &$1$ $\times$ 2 & None  & - &  \\
@@
     VG03   & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
-    VG04   & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
+    VG04   & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15$\pm7$ & \\
+    VG07   & 100 & $1$ $\times$ 2 &  mesh C, M, F & 12$\pm7$ & \\ %see email from Victor on 5/16/13
+    VG08   & 100 & $1$ $\times$ 2 &  mesh C, M, F & 12$\pm7$ & \\ 
     VG09   & 75   & $1$ $\times$ 2 & mesh C & 49 $\pm$6 & \\
@@
     Pattern name & fill factor & minimum feature size & bulk description \\ 
-    - & \% & $\mu$m & '- \\ 
+    - & \% & $\mu$m & - \\ 
+    \hline
     Hole & 100 & - & 25 mm diameter hole \\     
@@
 \hline
-        Spectral sampling interval (nm) & 2.0 \\
+        Spectral sampling interval (nm) & 2.0$-$10.0 \\
         Spectral resolution (nm) & 5 \\
@@
 In other cases the IR imaging was not available or ignored, and spectra were taken at random positions on the wafer surface.
+
+% Table of experiments
+\begin{table}[h!]
+\caption{Cary5000 Experiment Log \label{tabExperiments}}
+\begin{center}
+    \begin{tabular}{c c c c c c c}
+    \hline
+	Date &  Mode  & $\Delta \lambda$ & SBW & Sampling & \# of spectra & Sample IDs \\
+	\emph{mm/dd/yyyy} &  -  & nm & nm & nm & - \\
+%     \hline		
+%	02/04/2013 & ? & ? 				& 5.0 & 10.0 &  16 &   -  \\
+%     \hline		
+%	02/06/2013 & ? & ? 				& 5.0 & 10.0 &  16 &   -  \\	     
+     \hline		
+     	02/18/2013 & Single & 1000$-$2500 & 5.0 & 10.0 &  16 &   -  \\
+     	                   &            &                       &       &         &    4 &  Baseline\\	
+     	                   &            &                       &       &         &    1 &  VG01\\
+	                   &            &                       &       &         &    4 &  VG02\\	                   
+	                   &            &                       &       &         &    2 &  VG03\\	                   
+	                   &            &                       &       &         &    2 &  VG04\\
+	                   &            &                       &       &         &    2 &  VG05\\	                   
+	                   &            &                       &       &         &    2 &  VG06\\
+    \hline
+%	05/02/2013 & Single & 1000$-$2500 & 5.0 & 5.0  &  30 &   -  \\
+%     	                   &            &                       &       &         &    12 &  Baseline\\	
+%     	                   &            &                       &       &         &    18 &  Other\\
+%    \hline
+	05/03/2013 & Single & 1000$-$2500 & 5.0 & 5.0  &  28 &   -  \\
+     	                   &            &                       &       &         &    8 &  Baseline\\	
+     	                   &            &                       &       &         &    1 &  VG05\\ %Consider replacing VG05 with "DSP reference sample"  or something
+     	                   &            &                       &       &         &    6 &  VG03\\
+     	                   &            &                       &       &         &    13 &  Other\\ %These had to do with beam size experiments
+    \hline
+	06/07/2013 & Mix & 1000$-$2500 & 5.0 & 5.0  &  43 &   -  \\
+	                   &            &                       &       &         &    11 &  Baseline\\	                   
+	                   &            &                       &       &         &    9 &  VG05\\	                   
+	                   &            &                       &       &         &    7 &  VG07\\
+	                   &            &                       &       &         &    6 &  VG08\\
+	                   &            &                       &       &         &    4 &  HA1\\
+	                   &            &                       &       &         &    6 &  other\\ % These are mostly just baselines without automatic division
+    \hline
+	08/02/2013 & Single & 1000$-$2500 & 5.0 & 2.0  &  17 &   -  \\	
+	                   &            &                       &       &         &    3 &  Baseline\\
+	                   &            &                       &       &         &    4 &  VG05\\	                   
+	                   &            &                       &       &         &    2 &  VG15\\
+	                   &            &                       &       &         &    2 &  VS20\\
+	                   &            &                       &       &         &    6 &  VG09\\
+    \hline
+    	09/11/2013 & Double & 1000$-$2500 & 5.0 & 2.0  &  20 &   -  \\	                   
+	                   &            &                       &       &         &    1 &  Baseline\\
+	                   &            &                       &       &         &    14 &  VS21\\	                   
+	                   &            &                       &       &         &    5 &  Blank\\ %This is really the same as a baseline
+    \hline
+    	09/16/2013 & Double & 1000$-$2500 & 5.0 & 2.0  &  21 &   -  \\
+	                   &            &                       &       &         &    2 &  Baseline\\
+	                   &            &                       &       &         &    7 &  VG07\\	                   
+	                   &            &                       &       &         &    4 &  VG05\\	
+	                   &            &                       &       &         &    4 &  VG12\\
+	                   &            &                       &       &         &    2 &  VG15\\	                   
+	                   &            &                       &       &         &    2 &  VG08\\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
 

### 2014-05-01 AO_bondedSi.tex
tweaks here and there, especially to tables

@@
 
-% Figures (shared wavelength axis)
-%	2) Fresnel transmission as a function of wavelength (normal incidence)
-%	4) Example transmission through VG03 thick gap
-
 \subsection{The Si$-$Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}
@@
     Petal & $\sim1$ & $\sim$2000 & long petal lines, Fig. \ref{figVG03pattern} \\         
-    Mesh F & 50 & 40 &  100 $\mu$ square holes, Fig. \ref{figVS20pattern}\\ 
-    Mesh M & 50 & 200 &  500 $\mu$ square holes\\ 
-    Mesh C & 50 & 620 &  1500 $\mu$ square holes\\     
+    Mesh F & 50 & 40 &  100 $\mu$m square holes, Fig. \ref{figVS20pattern}\\ 
+    Mesh M & 50 & 200 &  500 $\mu$m square holes\\ 
+    Mesh C & 50 & 620 &  1500 $\mu$m square holes\\     
         \hline
@@
 
-%\caption{Summary of X-ray CT measurements for VG03 \label{tab:tablenum4}}
+\begin{table}[h!]
+\caption{Summary of X-ray CT measurements for VG03 \label{tab:tablenum4}}
 \begin{center}
     \begin{tabular}{ll}
+    \hline
     \emph{Parameter (Units)} & \emph{Value} \\ 
+    \hline
     Instrument& Xradia MicroCT \\
@@
         Sample thickness (mm) & 6.6 \\
+    \hline
      \end{tabular}
 \end{center}
+\end{table}
 

### 2014-05-02 AO_bondedSi.tex
add Si absorption figure and discussion

@@
 We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
-Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
@@
 The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
+Si is negligibly absorbing longward of about $\lambda$ = 1200 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses, as shown in Figure \ref{figSiAbsorbfig}.
+Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si pucks of different heritage and thickness.
+The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.
+\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference only for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
+An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is roughly our measurement uncertainty.
 In the next subsection we work out the wavelength dependent transmission for the air etalon.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/fpAbsorbfig}}
+\caption{Ratio of transmissions of Si wafers of different thicknesses, specifically the transmission of a thin ($\sim0.5$ mm) Si wafer divided by that of a 3 mm thick Si puck.
+Both samples were double side polished.\label{figSiAbsorbfig}  The transmission scatters about 1.0 in wavelength ranges for which Si is negligibly absorbing for a path length ratio of $\sim$6, and exceeds 1.0 in wavelength ranges for which Si absorption is detected.
+The left panel shows that Si is firmly absorptive shortward of $\lambda \; =$ 1200 nm.
+The right panel shows the absolute deviation from unity on a log scale.
+The typical uncertainty and drift per measurement is about 0.2\%.
+Absorption is detected at 0.4\% at $\lambda \;=$ 1200 nm.}
+\end{figure}
 

### 2014-05-02 AO_bondedSi.tex
minor

@@
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.
-\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference only for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
-An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is roughly our measurement uncertainty.
+\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
+An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
 In the next subsection we work out the wavelength dependent transmission for the air etalon.

### 2014-10-22 AO_bondedSi.tex
closes #2 (puts equation numbers on all equations)

@@
 In other words:
-$$\log(F) = -2\log(D) + const. \label{eqnParticleFlux}$$
+\begin{eqnarray}
+\log(F) = -2\log(D) + const. \label{eqnParticleFlux}
+\end{eqnarray}
 
@@
 
-$$
-T_{DSP} = \frac{2n}{1+n^2} \label{eqnAbsDSPtrans} \\
-$$
+\begin{eqnarray}
+T_{DSP} = \frac{2n}{1+n^2} \label{eqnAbsDSPtrans}
+\end{eqnarray}
 
@@
 
-$$T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP} \\$$
+\begin{eqnarray}
+T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}
+\end{eqnarray}
 
@@
 
-Finally, we cascade the matrices together to compute the net transmission through the stack of abstractions.
-The result is a $2\times2$ transfer matrix: $$\boldsymbol{M_{net}}=\boldsymbol{M_n}\boldsymbol{M_e}\boldsymbol{M_n}$$  From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.
+Finally, we cascade the matrices together to compute the net transmission through the stack of abstractions.  The result is a $2\times2$ transfer matrix:
+
+\begin{eqnarray}
+\boldsymbol{M_{net}}=\boldsymbol{M_n}\boldsymbol{M_e}\boldsymbol{M_n}
+\end{eqnarray}
+
+From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.
 Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:
@@
 Taking the inverse of the $M_{22}$ element, we find:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 which is identical to the result obtained by directly summing the intensities from multiple reflections:
@@
 
-$$
+\begin{eqnarray}
  <T_e> = \lim_{\sin^2(\delta/2) \to 0.5} \frac{1}{1+F\sin^2(\delta/2)} \\
  <T_e> = \frac{1}{1+F/2}
-$$
+\end{eqnarray}
 

### 2014-10-22 AO_bondedSi.tex
closes #11- deletes figure 1

@@
 \subsection{Particles}
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiTentCatFig2.pdf}}
-\caption{Schematic of a particle preventing bonding of two Si surfaces\label{figParticle}.
-There are two conceivable scenarios in which particle contamination can cause gaps in the bond.
-\emph{Left-} The particle serves as a mechanical structure hindering the Si-Si bond from zippering up.
-The geometry is like a tent-pole.
-\emph{Right-} The particle serves as a catalyst for gas formation.
-There is evidence that hydrocarbons can serve as catalysts.
-The especially pernicious particles are ``thermally unstable hydrocarbon contaminants on the surfaces coming from plastic containers or simply from the ambient in a cleanroom'' \cite{1998AnRMS..28..215G}.
-}
-\end{figure}
 

### 2014-10-22 AO_bondedSi.tex
closes #18- delete figure 6 altogether- figure 6 was VG03 cartoon of petal pattern

@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/VG03patternSchematic.pdf}}
-\caption{Schematic of part VG03 \label{figVG03pattern}.
-The 100 mm diameter VG03 is comprised of two 3.3 mm thick silicon pucks.
-We patterned a 25 mm diameter hole on one of the pucks.
-The depth was 4.0 $\mu$m as indicated by Dektak profilometry.
-We also patterned a low-fill factor petal pattern surrounding the hole.
-The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.}
-\end{figure}
-
 We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.
@@
 The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
-Figure \ref{figVG03pattern} shows a schematic of VG03 and its pattern, as described in Table \ref{tabbondexper}.
 In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
@@
     VG02   & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
-    VG03   & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
+    VG03   & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & \\    
     VG04   & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15$\pm7$ & \\
@@
     Hole & 100 & - & 25 mm diameter hole \\     
-    Petal & $\sim1$ & $\sim$2000 & long petal lines, Fig. \ref{figVG03pattern} \\         
+    Petal & $\sim1$ & $\sim$2000 & long petal lines \\         
     Mesh F & 50 & 40 &  100 $\mu$m square holes, Fig. \ref{figVS20pattern}\\ 

### 2014-10-22 AO_bondedSi.tex
closes #21- cut Fig 8, which was photo of IR bubble and fringes

@@
 Naturally, the unbonded DSP wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrates as discussed in the literature (see Section \ref{secHistory}).
-Figure \ref{IRbubble} shows an example of one of the IR bubbles in which several bright and dark fringes are clearly discernible.
 No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.
 The morphology was clear regardless of the detector response.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/IR_bubble.pdf}}
-\caption{Photo of part VG09 with transmission image of a bubble highlighted\label{IRbubble}.
-The right side photo shows the 75 mm diameter wafer with blue Sharpie rectangles indicating the position of different spectral measurements (as detailed in the next section).
-The yellow circle in the photo highlights the outline of the location of an IR bubble we detected.
-The bubble shows $\sim2$ fringes over the effective wavelength range 1.1-1.6 $\mu$m.
-This fringe pattern is consistent with a gap comparable to this measurement wavelength.
-The pointlike measurement artifacts, Sharpie marker, and vignetting are perceptible in the raw IR image.
-Our interpretation of the IR bubbles is robust against these artifacts.
-The area surrounding the IR bubble appears to be mostly devoid of large (wavelength-scale) gaps, since the transmission is fairly uniform, and consistent with the brightest fringe of the IR bubble.
-}
-\end{figure}
 

### 2014-10-22 AO_bondedSi.tex
closes #23 - cut Table 4, which was the interpretation of diversity of IR spectra

@@
 
-\begin{table}[h!]
-  \caption{Interpretation of the diversity of IR spectra \label{tabInterpretation}}
-  \begin{center}
-    \begin{tabular}{ |p{8cm}| p{8cm} | p{2.2cm} |}
-    \hline
-    Appearance of spectra & Interpretation & Exemplars \\
-    \hline
-    Measurement within the prediction band & Life is good & VS21-M1, Figure \ref{VS21specm1}, VG15, VG07\\
-    \hline
-    Measured spectrum below the prediction band but shape is similar & Gap was slightly larger than predicted or bonding was incomplete& VS21-F1, Figure \ref{VS21specf1}, VS09, VG15, \\
-    \hline
-    Measured spectrum above the prediction band but shape is similar & Gap was slightly smaller than predicted & VS21-M2\\
-    \hline
-    Measured spectrum exceeds DSP reference puck & unanticipated anti-reflection coating on the surface of the Si sample (e.g. Sharpie marker residue) & VG08\\
-    \hline
-    High frequency, low amplitude wiggles superimposed upon smooth underlying trend & Large, unanticipated gap, with a small fill factor  & VS21-C3\\   
-     																			     & wavelength dependent diffraction from the mesh patterns & VS21-C3\\       
-     																			     & measurement artifact  & VS21-C3\\       												
-    \hline
-    Non-unit transmission constant with wavelength (gray spectrum) & Artifact of measurement- vignetting of beam, absorptive coating, zero-point drift & VS21-C4 \\
-    \hline
-    Completely different spectrum than anticipated with low absolute transmission & Large, inhomogeneous gap consistent with an IR bubble & VG09 \\
-    \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
 A key result of this work is to show that gaps of known dimensions are recovered in our spectroscopy.
@@
 \subsection{Mitigation of measurement artifacts}
-The principle artifacts and their interpretation are listed in Table \ref{tabInterpretation}.
-We mitigated the artifacts in the following ways.
+We mitigated measurement artifacts in the following ways.
 For the Sharpie marker behaving as an anti-reflection (AR) coating, we simply removed the sharpie with IPA.

### 2014-10-22 AO_bondedSi.tex
closes #24 - cut discussion of non-normal incidence, i.e. actual immersion gratings

@@
 
-\subsection{Non-normal incidence angles}
-For the vast majority of applications, normal incidence performance is the quantity of interest.
-The technique laid out in this paper outlines an approach for measuring the gap size of bonded Si wafers at normal incidence.
-In principle, from there the performance at non-normal incidence can be worked out either by simulation or calculation.
-Here we briefly motivate the challenge of experiment and simulation at non-normal incidence.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/Sibondedprismcartoon.pdf}}
-\caption{Schematic of a bonded Si immersion grating\label{figSiPrism}.  }
-\end{figure}
-
-Our final application that motivated this project is to make diffraction gratings on the hypotenuse of a silicon prism.
-The diffraction occurs inside the silicon material and is dispersed and reflected back along its path.
-This configuration is called an immersion grating.
-Figure \ref{figSiPrism} shows a schematic of the geometry of our bonded Si immersion grating.
-Our group has made monolithic immersion gratings for over a decade, but challenges in processing large (30 mm thick) Si wafers motivated an approach of patterning and processing thin (1-10 mm thick) Si wafers, and then bonding them to shaped and polished prisms.
-This technique has been pioneered at SRON \cite{2012SPIE.8450E..2TV}.
-In Figure \ref{figSiPrism} we show an example with a 10 mm thick Si puck bonded to a 25 mm thick, precision cut, shaped, and polished Si prism.
-The group at SRON has patterned Si immersion echelle gratings on commercial thin  ($< 1$ mm) Si wafers.
-Thin wafers have the drawback that they are conformal, so total-thickness-variations (TTV) directly translate into surface optical aberrations\cite{2012SPIE.8450E..2TV}.
-
-The consequence of our immersion grating design is a non-normal incidence interface in double pass.
-In this case, the Fabry-P\'{e}rot framework developed here fails to predict the spectrum of the interface loss attributable to a finite albeit small (few nm) gap.
-Specifically, complete internal reflection and evanescent waves will matter at large incidence angles, and the incoherent matrix technique is no longer accurate.
-Polarization effects are important.
-We do not have a mechanism for evaluating the bond performance until the final device is completed and its wavelength-dependent transmission is measured.
-We can detect the performance degradation on final devices in the following nuanced way.
-We will measure the efficiency of the bonded Si immersion grating as a function of both wavelength and angle.
-Our group built a custom tool to make these measurements- the Custom Robotic Order, Wavelength, and Blaze Angle Recorder (CROWBAR) \cite{2012SPIE.8450E..2SG}.
-Figure 3 in \cite{2012SPIE.8450E..2SG} shows the directly measured spectrum of a monolithic immersion grating.
-Notably, the spectrum in Figure 3 in \cite{2012SPIE.8450E..2SG}  shows dozens of blaze peaks where the efficiency should be maximized.
-A perfectly contacted Si$-$Si direct bonded optic would have identical performance to a monolithic counterpart.
-Bonded immersion gratings with finite gap sizes will superimpose a wavelength dependent transmission loss on top of the blaze peaks.
-In this way, the presence or absence of deleterious gaps can be identified in final immersion gratings.
-One drawback of this approach is that gaps can only be identified after the costly process of producing the final optical device.
-Large gaps identified this late in the game could render the costly optic useless, both in terms of efficiency and wavefront performance.
-Our group is exploring the effectiveness of high-temperature annealing to eliminate interfacial gaps in completed devices.
-
 \subsection{Spectra of pucks with no implanted gap}

### 2014-10-22 AO_bondedSi.tex
closes #20- Overhaul Table 3 to reflect marked-up paper draft, Table 3 is summary of Cary5000 measurement parameters

@@
         Time per sample (s) & 0.1 \\
-        Available measurement range (nm) & 175-3300 \\
         Typical measurement range (nm) & 1000-2500 \\
-	Single mode RMS  (\%) & 0.02 \\
-	Double mode RMS  (\%) & 0.002 \\
 	Typical Single mode mean drift (\%) & 0.20 \\
-	Typical Double mode mean drift (\%) & 0.02 \\
-	No beam dark signal (\%) & 0.03 \\
  	Beam size (mm $\times$ mm) & $2 \times 8$ \\
-	Wavelengths of instrumental artifacts (nm) & - \\
-        		 & 1200 \\
-		 & 1350 \\
-		 & 1360-1450\\
-		 & 1850 \\
-		 & 1800-1950 \\
-		 & 2000 \\
-        Vendor-reported linearity range (dB) & $>40$ \\
         Directly measured linearity range (dB) & $>24$ \\

### 2014-10-22 AO_bondedSi.tex
change the title as suggested by Dan Jaffe

@@
 
-\title{Characterization of gaps in directly bonded Si compound optics}
+\title{Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy}
 
@@
 \begin{abstract}
-Si direct bonding offers flexibility in the design and development of large Si optics.
+Si direct bonding offers flexibility in the design and development of large Si optics by allowing manufacturers to combine subcomponents with a lossless interface.
 The bonding process presents challenges in meeting the performance demands and physical robustness, in particular for cryogenic and space applications.

### 2014-10-22 AO_bondedSi.tex
Cut all mention of particle origin, replace with a reference (maybe not the best reference though)

@@
 
-\subsection{Particles}
+It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \cite{Mitani1990}.
 
-There are two major conceivable pitfalls in Si-Si direct bonding.
-The obvious pitfall is particle contamination.
-Figure \ref{figParticle} shows a schematic of a particle wedged between two flat Si optical surfaces.
-A single 0.6 $\mu$m particle can create a dark IR bubble from interference.
-In a class 100 cleanroom, the particle volume density could be as high as $\rho=100$ particles/ft$^3$, for particle sizes larger than 0.5 $\mu$m.
-What we care about is not the volume number density of particles, but the delivered surface flux $F=\rho(D) v(D)$, where $v(D)$ is the average freefall speed of particles, which is a function of particle size.
-Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$^{-2}$ s$^{-1}$) vs.
-particle size ($\mu$m).
-There will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
-In other words:
-\begin{eqnarray}
-\log(F) = -2\log(D) + const. \label{eqnParticleFlux}
-\end{eqnarray}
-
-For every obviously detectable IR bubble, there are thousands more particles that could unacceptably degrade the transmission performance.
-One key idea that saves the day, is the intuition that the areal fill factor scales at least as fast as $D^2$.
-Imagine that the particles serve as miniature tentpoles.
-The volume beneath the tentpole looks like a cone.
-Since the number density scales as roughly $D^{-2}$, (per the Equation \ref{eqnParticleFlux} above) the total area covered by particles might be constant in $log(D)$.
-Simply put, the big particles cover just as much area as the small particles, even though there are many more small particles.
-It is further conceivable that the areal fill factor attributable to a tent pole particle scales even faster than $D^2$ since surface energy probably scales inversely with distance of the pieces (\`a la VanDerWaals).
-The small particles are even less of a problem than the large particles, since the small ($\sim10-60$ nm) particles will have lower average etalon reflectivity than larger (500-1000 nm) particles.
-This last point will become clearer in later sections.
-
-It has been shown\cite{1992JEMat..21..669M} that gases produced at the Si$-$Si interface seep along the interface surfaces.
-Pressure builds around catalysts.
-In this sense, the size of the particle causing a gap cannot be uniquely identified from observations of the extent of gaps.
-In Section \ref{secResults} we address the question of the origin of gaps.
-Specifically, are the particles more like \emph{tentpoles}, rigid structures that merely prevent the zippering up of the bond?
-Or are the particles more like \emph{catalysts} so that their size is immaterial- the particles merely accelerate the production of gases that bow out the bond?
-These are open questions.
-
-\subsection{Surface roughness or large-scale non-conformance}
 Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform.

### 2014-10-28 AO_bondedSi.tex
overhaul the README.md file to include code and figure descriptions
complete overhaul of the introduction from DTJ email dated Sept 2, 2014

@@
 
-Direct optical bonding of silicon allows manufacturers to combine Si parts with no Fresnel loss at the interface.
-The high refractive index of Si makes this advantage particularly useful.
-Traditional figures of merit for the optical audits of bonds, however, are either insensitive to gaps that are small yet have a significant impact on throughput and phase coherence, or do not provide quantitative information about the extent of the gaps along the direction of propagation.
+\subsection{Why Silicon is good}
+Crystalline silicon is an excellent optical material for the infrared.
+Si has a high refractive index ($\sim$3.4, Frey et al.
+2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m (REF) and in the far-IR and submillimeter (REF).
+It also has adequate transmission  (cite loss per cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well (REF).
+There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
 
-%Issues with the quality of the bonds become particularly acute when the optics are used in double-pass and/or when the incidence angle at the bonds is far from the normal.
+\subsection{Why bonding is good and AR Coating is bad}
+As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
+One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \emph{cite the Dutch group, one of our papers with cross-dispersed but not contacted grisms, and Lumella here}.
+A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.
+The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
+Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choices even further.
+It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.
 
-The University of Texas diffractive optics group produces diffraction gratings on silicon using microlithographic techniques \cite{1998SPIE.3354..201J,2010SPIE.7739E.146W}.
-These gratings are used either in transmission (i.e.
-as grisms \cite{2008SPIE.7014E..77D, 2010SPIE.7739E.123G}), or in reflection (i.e.
-as immersion gratings, \cite{2007ApOpt..46.3400M, 2010SPIE.7739E.146W,2012SPIE.8450E..2SG}.
-These gratings operate throughout the infrared, from 1.2 to 38 $\mu$m, and find application in ground-based (IGRINS \cite{2010SPIE.7735E..54Y, 2012SPIE.8450E..2SG}), airborne (SOFIA FORCAST \cite{2008SPIE.7014E..77D}), and space (JWST NIRCAM \cite{2005SPIE.5904...21B,2010SPIE.7739E.123G}) applications.
-To date, our group has produced the gratings on large (up to 100 mm diameter and 30 mm thick) monolithic pieces of monocrystalline silicon and we shape the substrates by sawing, grinding, and polishing to form the final devices \cite{2010SPIE.7739E.146W}.
-The next generation of astronomy and earth science spectrographs, however, will need optical devices larger and more precise than what can be made with existing technology.
-The proposed Giant Magellan Telescope \cite{2012SPIE.8444E..1HJ} Near Infrared Spectrograph \cite{2006SPIE.6269E.143J,2010SPIE.7735E..87L} (GMTNIRS) will need immersion gratings with thicknesses up to 60 mm and lengths up to 200 mm.
-Substrates of the required size cannot fit into conventional processing equipment like reactive ion etchers and electron-beam lithography tools.
-Large, thick Si substrates are also prohibitively expensive to experiment on.
-Silicon direct bonding \cite{1986JAP....60.2987S,2012SPIE.8450E..2TV} offers advantages by allowing us to produce the diffraction grating on a thin (0.5$-$10 mm) low-mass Silicon wafer or puck that can later be bonded to a fine polished Si prism substrate.
-Other groups producing Si diffractive optics are also experimenting with or using this approach (see \cite{2012SPIE.8450E..2TV} and \url{www.lumella.com}).
+\subsection{Silicon Bonds- history and techniques}
 
-While it is possible that proper precautions can be made to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.
-Small particles could be a concern if they act as ``tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).
-Some authors \cite{1991JaJAP..30..615M, 1992JEMat..21..669M, 1989JaJAP..28L2141L, Mitani1990} have referred to large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.
-The gaps have also been referred to as voids\cite{2000RScI...71.1869G}, interfacial voids, or gaps.
-In this paper we will reserve the term ``IR bubbles'' for gaps that are large enough to be detected directly in conventional IR imaging.
-The term ``gap'' will apply more broadly to any finite size gap.
+There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems (Goesele and Tong, Mastreika et al.
+2014).
+The foundation paper on Si bonding came in 1986[14].
+After that, there was a sequence of papers that used IR imaging to look for evidence of defects in the bond.
+Stengl et al.
+(1988) and Goesele et al.
+(1995) [23, 24] directly monitored the bond propagation with IR imaging.
+Lehman et al.
+1989 and Goesele et al.
+[19, 23] demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
+Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 C (REFS including looking at some from the Mastreika paper).
 
-Finite gaps will degrade the performance of a final device, either through optical aberrations or transmission loss.
-The effect of the gap size on the IR transmission is the topic of this paper.
-The effect of the gap on optical aberrations is beyond the scope of this paper.
+The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy[25], which lent credibility to the idea of distinct temperature phases and evidence for Si?H bonds[26].
+A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out[27, 28].
+Other variables potentially affecting bond strength were studied, for example surface roughness[29], surface topography [30], surface preparation[31], annealing time [32], ambient pressure and substrate thickness [33, 34], and moisture and defects [35].
+Notably, [29] found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
+Results from studies of surface topography are inconclusive[30].
 
-A gap might also degrade the mechanical integrity of a final compound optic.
-In practice, the part must survive thermal shocks and shake tests.
-The gap size scales we can probe with our optical technique are comparable to the range over which molecular (van der Waals) forces are acting.
-It is conceivable that the strength of the bond energy can be constrained from the \emph{in situ} measurement of the separation and that therefore our optical technique can provide not only a measure of the transmission but also an indication of the mechanical robustness of the bond.
-By measuring both the spatial and axial extent of any gaps, the new metrology technique informs revisions of the bonding strategy that can improve both the optical and mechanical performance.
-We do not investigate the mechanical robustness further.
+\subsection{Requirement:  Do better than coatings averaged over the surface}
 
-There are two main solutions for reducing risk attributable to gaps.
-The first approach is to prevent gaps from forming in the first place, specifically by removing all possible contaminants from the environment before bonding.
-Typically this approach involves disciplined handling protocols, and rigorous multi-stage cleaning steps.
-The reason the approach is so hard is that hydrocarbons can serve as catalysts for gap formation\cite{1991JaJAP..30..615M}, and that these hydrocarbons are common in Si wafer handling equipment.
-Unconventional micro-cleanroom approaches have been effective at meeting the goal of bubble-free bonding\cite{1989JaJAP..28L2141L}.
-The second approach is to accept the likelihood that gaps will occur, and take steps to remove them.
-The most cited way to get rid of gaps is to anneal them\cite{Mitani1990}.
-We describe the current understanding of both approaches in Section \ref{secHistory}.
+If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than those of two antireflection coated surfaces over the operating wavelength (typically $< 2$\%).
+If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond.
+While completely bonded Si parts will be completely lossless at the interface, problems on different scales with different physical origins can lead to less-than-perfect bonds.
+On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation[17].
+Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance.
+Figure 5 of Cooper 1986[16] shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
+particle size ($\mu$m).
+There will be roughly 104 times more 10 nm sized particles than 1 ?m sized particles.
+On larger scale, gaseous hydrogen, which is a byproduct of the  hydrophilic bonding process, can migrate into the bond and form local bubbles (REFS).
+Most of the bonding literature concerns itself with wafer bonding.
+In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.
+When bonding two thicker and more rigid substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
 
-Whatever technique one uses to remove or reduce the effects of gaps, it is essential to have a quantitative method by which to assess accurately the optical impact of the gaps before and after treatment.
-This paper introduces a technique for directly measuring gaps at silicon-silicon interfaces.
-Mitani and G\"oselle review other techniques for measuring and mapping gaps\cite{1992JEMat..21..669M}.
-One key technique is identifying IR Newton rings from IR imaging\cite{Mitani1990}.
-The number and morphology of Newton rings yields information about the lateral and axial extent of large gaps.
-Specifically, dark-light fringes manifest when the gap is comparable to the vacuum wavelength or larger.
-Our technique is IR interference in the limit of small gaps down to 20 nm or less.
+\subsection{Existing detection schemes}
+Verification that an optical bond meets its requirements and possible diagnosis of problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
+Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.
+Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect and the transmission is 100\%.
+In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap [20].
+X-ray topography [18, 22] exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
+In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension [21].
 
-There are other techniques for mapping gaps.
-X-ray topography \cite{1992JEMat..21..669M,1994JaJAP..33....6H} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers.
-In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
+\subsection{Technique}
+We describe here a new technique for measuring the axial extent of gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
+Stock near-infrared spectrophotometer like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
+We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few nanometers in size.
+In order to verify the applicability and accuracy of the technique, we have used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
+Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that was not widely available until this decade.
 
-Here we describe an application of the IR interference technique with new precision instrumentation in the near infrared to measure the axial extent of gaps at the few nanometer size scale.
-In this paper we show that gaps as small as 35 nm cause an unacceptable degradation in transmission performance.
+\subsection{What level do you need to measure to}
+For all types of gaps and bubbles, our technique needs to provide an accurate assessment of the transmission loss at the operating wavelength.
+For finite-sized bubbles, measurement of the axial dimension helps us to assess the effects of high-temperature annealing on removal of hydrogen bubbles.
+A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to resolve spatially, as one might expect if there are many very small contaminants present along the interface.
+To meet these diagnostic goals, our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
+A 2\% loss at 1.15 $\mu$m corresponds to a gap of XX nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
+We therefore need to be able to measure transmission as a function of wavelength to a fraction of a percent, including all systematic errors, for devices operating near 1 $\mu$m and to a somewhat looser standard for devices operating at longer wavelengths.
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1}}
+\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1.pdf}}
 \caption{IR transmission spectrum of VS21 (M1) showing consistent gap\label{VS21specm1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}

### 2014-10-28 AO_bondedSi.tex
this and prev commit closes #9, DTJ edit to the Abstract

@@
 \begin{abstract}
-Si direct bonding offers flexibility in the design and development of large Si optics by allowing manufacturers to combine subcomponents with a lossless interface.
-The bonding process presents challenges in meeting the performance demands and physical robustness, in particular for cryogenic and space applications.
-Air cavities in between the Si interface will have large Fresnel reflections which will interfere as a low-Finesse Fabry-P\`erot etalon.
-Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = $ 1.15 $\mu$m at normal incidence.
-Stock near infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% photometric precision.
-We can use instruments like the Cary 5000 to detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
+Silicon direct bonding offers flexibility in the design and development of Si optics by allowing manufacturers to combine subcomponents with a potentially lossless and mechanically stable interface.
+The bonding process presents challenges in meeting the requirements for optical performance because air gaps at the Si interface cause large Fresnel reflections.
+Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = 1.15 \; \mu$m at normal incidence.
+Understanding the optical effects of such gaps and evaluation of various methods for preventing or eliminating them demands that we have a method for measuring not only the two-dimensional extent of the gaps but also their axial extent.
+Existing methods for studying gaps provide only poor measures of this axial extent.
+We can use such instruments to detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
+We describe a bond inspection method that makes use of precision transmission spectrophotometry.
+Our analysis of the spectra then produces accurate estimates of the axial depth of air gaps.
 Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
-The detection and morphology of minuscule gaps provides valuable feedback in the design of Si direct bonding strategies.
+We describe the measurement and analysis techniques and demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
 \end{abstract}

### 2014-11-24 AO_bondedSi.tex
add water mark to the draft for pre-publication circulation
Add 60 nm curve to the existing figure 4, closes #1

@@
 \usepackage{longtable}
+
+%\usepackage{draftwatermark}
+%\SetWatermarkText{DRAFT}
+%\SetWatermarkScale{6}
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absoluteB.pdf}}
-\caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_0-60nm.pdf}}
+\caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 60 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
 The curves are examples from a family defined in Equation \ref{eqFP}.
 The limit of no gap is 100\% transmissive, compared to a DSP reference.
-A 35 nm gap degrades the transmission over a DSP reference by 4\% at 1200 nm.}
+A 60 nm gap degrades the transmission over a DSP reference by 12\% at 1200 nm.}
 \end{figure}

### 2014-11-24 AO_bondedSi.tex
generalize the use of "puck" to include substrate and sample, closes #4

@@
 Si is negligibly absorbing longward of about $\lambda$ = 1200 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses, as shown in Figure \ref{figSiAbsorbfig}.
-Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si pucks of different heritage and thickness.
+Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si substrates of different heritage and thickness.
+Specifically DSP Si pucks and wafers have the indistinguishable transmission longward of 1250 nm.
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.
@@
 The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no gap, and second a pair of bonded Si pucks with a gap thickness $d$.
-The DSP Si puck with no air gap has a transmission equal to:
+We computed the transmission for two scenarios- first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.
+The DSP Si sample with no air gap has a transmission equal to:
 
@@
 
-Now consider a pair of bonded Si pucks with a gap thickness $d$.
+Now consider a pair of bonded Si samples with a gap thickness $d$.
 We derive the absolute transmission of bonded Si substrates with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
@@
 The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
-We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.
+We classify substrates with thicknesses larger than 1 mm as ``pucks''.
+Substrates with thickness less than 1.0 mm thick are called ``wafers''.
 The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
 There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+Our ultimate goal is to bond pucks with thicknesses as large as 30 mm thick.
+These pucks truly resemble over-sized hockey pucks.
 
@@
 It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the wafer during dithering.
-We detected IR bubbles in all 3 of the bonded Si pucks on which we recorded IR images.
+We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.
 The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
@@
 
-\section{Results of infrared spectroscopy of bonded Si pucks}
+\section{Results of infrared spectroscopy of directly bonded Si}
 \label{secResults}
-In this section we describe the spectroscopy of bonded Si pucks.
+In this section we describe the spectroscopy of directly bonded Si.
 In some instances IR imaging was available and was leveraged to target or avoid locations demonstrating IR bubbles.
@@
 
-\subsection{Spectra of pucks with known gap sizes}
+\subsection{Spectra of direct bonded Si with known gap sizes}
 Figures \ref{VS21specm1} to \ref{VS21specf1} show examples of the predicted and measured IR spectra.
@@
 
-\subsection{Spectra of pucks with no implanted gap}
+\subsection{Spectra of direct bonded Si with no intentionally implanted gap}
 We have shown that the IR spectroscopy technique combined with the Fabry-P\'{e}rot framework provide accurate knowledge of the gap size.
@@
 In other cases we have seen evidence for gaps, which is simply an indication that the bonding process introduces gaps.
-In a followup paper we will describe ways to mitigate bond formation in bonded Si pucks.
 
@@
 While we are at it, let's compute the matrix for the scenario with no intermediate etalon: there are simply two Fresnel interfaces with which light interacts incoherently.
-This scenario is the model for a single DSP reference puck.
+This scenario is the model for a single DSP reference sample.
 The revised matrix multiplication is simply $\boldsymbol{M_{DSP}}=\boldsymbol{M_n}\boldsymbol{M_n}$.

### 2014-11-24 AO_bondedSi.tex
Get rid of the 373 K temp on n and F plot, closes #7

@@
 \centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.
 The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.

### 2014-11-24 AO_bondedSi.tex
closes #15, stacking figures 4 and 5

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_0-60nm.pdf}}
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_stack.pdf}
 \caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 60 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
@@
 The limit of no gap is 100\% transmissive, compared to a DSP reference.
-A 60 nm gap degrades the transmission over a DSP reference by 12\% at 1200 nm.}
+A 60 nm gap degrades the transmission over a DSP reference by 12\% at 1200 nm.
+Modeled etalon transmission through a silicon substrate with gaps of axial extent 35 nm (red dotted curve), 350 nm (green dash-dot curve), and 3500 nm (blue long-dashed curve).
+The curves are examples from a family defined in Equation \ref{eqFP}.
+The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.}
 \end{figure}
@@
 It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.
-Figure \ref{figEtalonRelTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 20 nm and 35 nm.
+Figure \ref{figEtalonRelTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
 In Section \ref{secResults} we compare wavelength dependent transmission measurements to by-eye fits of Equation \ref{eqFP}.
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absoluteC.pdf}}
-\caption{Modeled etalon transmission through a silicon substrate with gaps of axial extent 35 nm (red dotted curve), 350 nm (green dash-dot curve), and 3500 nm (blue long-dashed curve)\label{figEtalonRelTrans}.
-The curves are examples from a family defined in Equation \ref{eqFP}.
-The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.}
-\end{figure}
 

### 2014-11-24 AO_bondedSi.tex
minor bugfix- missing end backet in includegraphics command
add references about Si absorption as a function of wavelength

@@
 Si has a high refractive index ($\sim$3.4, Frey et al.
-2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m (REF) and in the far-IR and submillimeter (REF).
-It also has adequate transmission  (cite loss per cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well (REF).
+2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
+It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
 There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_stack.pdf}
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_stack.pdf}}
 \caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 60 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.

### 2014-11-26 AO_bondedSi.tex
Add some references

@@
 As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
-One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \emph{cite the Dutch group, one of our papers with cross-dispersed but not contacted grisms, and Lumella here}.
+One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
 A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.

### 2014-11-26 AO_bondedSi.tex
fix tons of broken references, what a waste of time

@@
 
+Placeholder citation so the document will compile: see: \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
+
 \subsection{Why Silicon is good}
@@
 Si has a high refractive index ($\sim$3.4, Frey et al.
-2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
-It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
+2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \verb|\cite{PhysRev.108.268, PhysRev.78.178}| and in the far-IR \verb|\cite{doi:10.1117/12.323764}|.
+It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \verb|\cite{doi:10.1117/12.323764}|.
 There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
@@
 As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
-One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
+One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \verb|\cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}|.
 A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.
@@
 
-There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems (Goesele and Tong, Mastreika et al.
-2014).
-The foundation paper on Si bonding came in 1986[14].
+There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \verb|\cite{1998AnRMS..28..215G,Masteika2014}|.
+The foundation paper on Si bonding came in 1986 \verb|\cite|{1986JAP....60.2987S}|.
 After that, there was a sequence of papers that used IR imaging to look for evidence of defects in the bond.
@@
 (1988) and Goesele et al.
-(1995) [23, 24] directly monitored the bond propagation with IR imaging.
+(1995) \verb|\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G}| directly monitored the bond propagation with IR imaging.
 Lehman et al.
 1989 and Goesele et al.
-[19, 23] demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
-Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 C (REFS including looking at some from the Mastreika paper).
+\verb|\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G}| demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
+Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \verb|\cite{Horn2009, Masteika2014}|.
 
-The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy[25], which lent credibility to the idea of distinct temperature phases and evidence for Si?H bonds[26].
-A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out[27, 28].
-Other variables potentially affecting bond strength were studied, for example surface roughness[29], surface topography [30], surface preparation[31], annealing time [32], ambient pressure and substrate thickness [33, 34], and moisture and defects [35].
-Notably, [29] found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
-Results from studies of surface topography are inconclusive[30].
+The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
+A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
+Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
+Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.
 
@@
 While completely bonded Si parts will be completely lossless at the interface, problems on different scales with different physical origins can lead to less-than-perfect bonds.
-On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation[17].
+On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation.
 Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance.
-Figure 5 of Cooper 1986[16] shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
+Figure 5 of Cooper 1986\verb|\cite{doi:10.1080/02786828608959094}| shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
 particle size ($\mu$m).
-There will be roughly 104 times more 10 nm sized particles than 1 ?m sized particles.
-On larger scale, gaseous hydrogen, which is a byproduct of the  hydrophilic bonding process, can migrate into the bond and form local bubbles (REFS).
+There will be roughly 104 times more 10 nm sized particles than 1 $\mu$m sized particles.
+On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \verb|\cite{Masteika2014}|.
 Most of the bonding literature concerns itself with wafer bonding.
@@
 Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect and the transmission is 100\%.
-In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap [20].
-X-ray topography [18, 22] exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
-In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension [21].
+In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap.
+X-ray topography \verb|\cite{Mitani1990}| exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
+In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \verb|\cite{2000RScI...71.1869G}|.
 
@@
 
-\section{History of Si direct bonding}
-\label{secHistory}
-
-The foundation paper on Si bonding came in 1986\cite{1986JAP....60.2987S}.
-After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.
-\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
-\cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding in a micro-cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
-Our group has not attempted the micro-cleanroom technique owing to the challenge of spin-cleaning large (30 mm thick) Si substrates.
-It is not clear whether the custom micro-cleanroom technique has gained wide usage.
-\cite{1989JaJAP..28L2141L} directly measured the bond surface energy.
-\cite{Mitani1990} discussed the causes and remedies for bubble formation.
-Specifically, the IR bubble density changes as a function of anneal temperature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.
-The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
-A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
-Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
-Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
-Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.
-Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that was not widely available until this decade.
-
 \section{Conceivable pitfalls in Si-Si direct bonding}
 
-It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \cite{Mitani1990}.
+It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \verb|\cite{Mitani1990}|.
 
@@
 
-Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+Our measurement technique exploits the large Fresnel reflection \verb|\cite{2001opt4.book.....H}| at the Si$-$vacuum interface.
+The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \verb|\cite{2006SPIE.6273E..77F}|, one of the largest refractive indices of any conventional material.
 The Fresnel reflection is about 30\% per surface.
 For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
-Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
+Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \verb|\cite{2006SPIE.6273E..77F}|.
 
@@
 \centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of CITE\texttt{2006SPIE.6273E..77F}.
 The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
@@
 
-The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\verb|\cite{2007fuph.book.....S}|.
 The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
 We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
-We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
+We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \verb|\cite{2006SPIE.6273E..77F}|, as shown in Figure \ref{figSiIndexFinesse}.
 In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
@@
 The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
-We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
+We've defined the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
 The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
@@
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.
-\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
+\verb|\cite{2006SPIE.6273E..77F}| cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
 An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
@@
 There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.
-Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
+Treatments for multiple coherent reflections have been described in detail in the optics literature \verb|\cite{2007fuph.book.....S}|.
 The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.
+We adapted the wave transfer technique for incoherent interactions \verb|\cite{2002ApOpt..41.3978K}|.
 Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
@@
 The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
+There is also evidence that bond front propagation is slower in thick pucks \verb|\cite{2007ApOpt..46.6793H}| than in wafers.
 Our ultimate goal is to bond pucks with thicknesses as large as 30 mm thick.
@@
 
-We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
+We looked for the indication of large gaps detectable as ``IR bubbles'' \verb|\cite{1992JEMat..21..669M}| with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.
-We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
+We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \verb|\cite{1992JEMat..21..669M}|.
 Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.
@@
 
-The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.
+The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \verb|\cite{2007fuph.book.....S}|.
 Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
@@
 
-The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\cite{2007fuph.book.....S}.
+The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\verb|\cite{2007fuph.book.....S}|.
 Much of the literature on the transfer matrix method deals with applications related to thin films, in which the wavelength is comparable to the size of the dielectric layer.
 For thin films the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.
-The polarization state can be encapsulated in scattering matrix components\cite{2007fuph.book.....S}.
+The polarization state can be encapsulated in scattering matrix components\verb|\cite{2007fuph.book.....S}|.
 The intensities of the emergent spectrum can be computed from the absolute square of the complex amplitudes.
 For thick films, the vector components are the intensities of the emergent spectrum, since waves are incoherent.
-\cite{2002ApOpt..41.3978K} work out the general transfer-matrix method for optical multilayer systems with incoherent interference.
+\verb|\cite{2002ApOpt..41.3978K}| work out the general transfer-matrix method for optical multilayer systems with incoherent interference.
 The key idea for the incoherent transfer matrix method approach is to populate a scattering matrix with elements equal to the (wavelength dependent) transmitted $T_i$ and reflected $R_i$ intensities of an interface or set of interfaces that act together.
@@
 
-From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.
+From the matrix transformation equations \verb|\cite{2007fuph.book.....S}| we know $M_{22}=1/T$.
 Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:

### 2015-01-05 AO_bondedSi.tex
Incorporate DTJ edits on pages 1-8

@@
 Existing methods for studying gaps provide only poor measures of this axial extent.
-We can use such instruments to detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
-We describe a bond inspection method that makes use of precision transmission spectrophotometry.
-Our analysis of the spectra then produces accurate estimates of the axial depth of air gaps.
+We describe a bond inspection method that makes use of precision transmission spectroscopy.
+Using spectroscopy, we can detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
 Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
@@
 The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
-Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choices even further.
+Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choice of coatings even further.
 It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.
@@
 particle size ($\mu$m).
-There will be roughly 104 times more 10 nm sized particles than 1 $\mu$m sized particles.
+There will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
 On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \verb|\cite{Masteika2014}|.
@@
 \subsection{Existing detection schemes}
-Verification that an optical bond meets its requirements and possible diagnosis of problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
+Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
 Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.
@@
 \subsection{Technique}
-We describe here a new technique for measuring the axial extent of gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
-Stock near-infrared spectrophotometer like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
+We describe here a new technique for measuring the axial extent of small gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
+Stock near-infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
 We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few nanometers in size.
@@
 A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to resolve spatially, as one might expect if there are many very small contaminants present along the interface.
-To meet these diagnostic goals, our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
+An important figure of merit for a bond is whether any loss due to gaps is smaller than the loss one might expect from two passages through a reasonably effective broad-band anti-relfection coating in the infrared.
+To meet this diagnostic goal, our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
 A 2\% loss at 1.15 $\mu$m corresponds to a gap of XX nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
@@
 Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 3 to 4 inch diameter clear aperture.
-We assume that if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
+We assume that, if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
 Subsequent annealing will strength the bond.
@@
 The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
-The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
+We assume $T= 295\;$K for all calculations in this article.}
 \end{figure}
@@
 The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\verb|\cite{2007fuph.book.....S}|.
+The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
+For the narrow air gaps we consider here and for the modest reflectivity of a Si-vacuum interface, most gaps at Si-Si bonds will form effective etalons over most of their area.
 The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.
+We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at $T= 295\;$K.
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \verb|\cite{2006SPIE.6273E..77F}|, as shown in Figure \ref{figSiIndexFinesse}.

### 2015-01-05 AO_bondedSi.tex
DTJ edits on page 8

@@
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \verb|\cite{2006SPIE.6273E..77F}|, as shown in Figure \ref{figSiIndexFinesse}.
-In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
+We drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
 The computation is simply Fresnel's law at normal incidence:
@@
 \begin{eqnarray}
-R = \frac{(n-1)^2}{(n+1)^2} \\
-F \equiv \frac{4R}{(1-R)^2}
+R = \frac{(n-1)^2}{(n+1)^2} \label{Eq:FresnelR}\\
+F \equiv \frac{4R}{(1-R)^2} \label{Eq:coeffF}
 \end{eqnarray}
 
-The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.
-We've defined the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.
+The reflectivity $R$ (Equation \ref{Eq:FresnelR}) is about 30\% in the wavelength range 1100 to 5000 nm.
+We define the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
 The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.

### 2015-01-05 AO_bondedSi.tex
edits from DTJ to Figure 2 caption

@@
 We define the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
-The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.
-Si is negligibly absorbing longward of about $\lambda$ = 1200 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses, as shown in Figure \ref{figSiAbsorbfig}.
-Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si substrates of different heritage and thickness.
-Specifically DSP Si pucks and wafers have the indistinguishable transmission longward of 1250 nm.
-The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.
+The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength for $T = 295\;$K and $T = 77\;$K.
+Si absorbs negligibly longward of $\lambda$ = 1250 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses (Figure \ref{figSiAbsorbfig}).
+We computed the ratio of transmissions of double side polished Si wafers of different thicknesses.
+Specifically, we evaluated the transmission ratio at each wavelength bin from 1000 to 2500 nm, from the transmission of a thin ($\sim0.5$ mm) Si wafer divided by that of a 3 mm thick Si puck.
+Absorption is detected shortward of 1250 nm.
+Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness.
+Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level longward of 1250 nm.
+The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
 \verb|\cite{2006SPIE.6273E..77F}| cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
@@
 \centerline{\includegraphics[width=0.95\columnwidth]{figs/fpAbsorbfig}}
-\caption{Ratio of transmissions of Si wafers of different thicknesses, specifically the transmission of a thin ($\sim0.5$ mm) Si wafer divided by that of a 3 mm thick Si puck.
-Both samples were double side polished.\label{figSiAbsorbfig}  The transmission scatters about 1.0 in wavelength ranges for which Si is negligibly absorbing for a path length ratio of $\sim$6, and exceeds 1.0 in wavelength ranges for which Si absorption is detected.
-The left panel shows that Si is firmly absorptive shortward of $\lambda \; =$ 1200 nm.
-The right panel shows the absolute deviation from unity on a log scale.
-The typical uncertainty and drift per measurement is about 0.2\%.
-Absorption is detected at 0.4\% at $\lambda \;=$ 1200 nm.}
+\caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
+\label{figSiAbsorbfig} We detect Si absorption shortward of $\lambda \; =$ 1250 nm.
+We assume spectral regions longward of 1250 nm are non-absorptive.
+The correlated and heteroscedastic measurement errors are typically less than 0.2\%.}
 \end{figure}

### 2015-01-06 AO_bondedSi.tex
DTJ edits up to page 18

@@
 \label{secTheory}
-There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.
+A pair of bonded silicon wafers or pucks with a small gap actually forms there coupled cavities with the air gap as the the central cavity.
+The outer cavities, those within the Si, however, were not produced with coherence in mind.
+Their sides are not necessarily flat and parallel to the required level and the specifics of their deviation from perfect etalons are usually not known to the level necessary to treat them as oherent structures in the analysis.
+The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps where the gap size si small enough that the cavity it forms can always be treated as a coherent structure.
 Treatments for multiple coherent reflections have been described in detail in the optics literature \verb|\cite{2007fuph.book.....S}|.
 The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \verb|\cite{2002ApOpt..41.3978K}|.
+We adapted the wave transfer technique for incoherent interactions \verb|\cite{2002ApOpt..41.3978K}| to deal with the cavities within the Si.
 Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
@@
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range 1200$-2500$ nm.
-If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
-This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.
+If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
 
@@
 In Section \ref{secResults} we compare wavelength dependent transmission measurements to by-eye fits of Equation \ref{eqFP}.
+Depending upon the intended operating wavelength, gaps as small as 35-60 nm can cause losses comparable to or greater than the losses produced by antireflection coatings.
 
-\section{Imbedding gaps of known sizes}
+\section{Laboratory demonstration: Imbedding gaps of known sizes}
 
@@
 There is also evidence that bond front propagation is slower in thick pucks \verb|\cite{2007ApOpt..46.6793H}| than in wafers.
-Our ultimate goal is to bond pucks with thicknesses as large as 30 mm thick.
+Our ultimate goal is to bond optics with thicknesses as large as 30 mm.
 These pucks truly resemble over-sized hockey pucks.
@@
 CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
-The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.
+The area exposed to the process gases was either a 25 mm diameter circle, or a 15 mm $\times$ 20 mm rectangle.
 We protected the unexposed areas with thick photoresist.
@@
 The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
-In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+In a second round of experiments, we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
 For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
@@
 
-\section{Substrate preparation}
 We cleaned the surfaces before bonding to minimize the interfacial particle density.
@@
 We measured the large scale surface flatness of the pucks.
-The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.
-We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.
+The Fisba2 interferometer used for these measurements has a 50 mm diameter beam at $\lambda=$632.8 nm.
+We found a typical peak to valley surface flatness of 2 waves over the central 50 mm diameter.
 We prepared the substrates with standard cleaning procedures of solvents in a megasonic.
 We then applied MHz frequency oxygen plasma ashing.
-We soaked the wafers in DI water then dried with an N$_2$ gun.
+We soaked the wafers in DI water then dried with N$_2$.
 We pressed the Si substrates together from the center to the outside.
@@
 \section{Measurement strategy, conditions, and details}
-We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers.
-The spectrophotometer is a Agilent Cary 5000 UV-Vis-NIR.
-The typical measurement settings and tool performance are listed in Table \ref{tab:Cary5000pars}.
-More details about the tool are available at the vendor's website.
-Briefly, the tool has a double monochromator.
-The double monochromator feature reduces scattered light.
+We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
+Table \ref{tab:Cary500pars} lists the typical measurement settings and tool performance.
+The tool has a double monochromator, which reduces scattered light.
 The vendor reported linearity exceeds 40 dB.
@@
 We experimented with single front and double modes.
-Double mode takes a reference spectrum simultaneously to the target spectrum.
-In both single and double beam mode, a baseline scan is taken with no sample in the measurement holder.
+Double mode takes a reference spectrum simultaneously with the target spectrum.
+In both single and double mode, a baseline scan is taken with no sample in the measurement holder.
 The baseline spectrum is the raw spectral response of the system.
 We automatically divide sample spectra by the baseline spectrum.
-In single beam mode, the baseline measurements must be repeated about every half hour.
-We estimated the uncertainty and measurement repeatability of single and double modes in the following way.
-We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
-The predicted mean level should naturally be 100\%.
+In single beam mode, the baseline measurements must be repeated about every half hour for post-process division.
+In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.
+We estimated the uncertainty and measurement repeatability of single and double modes in the following way: we computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
 We found that the mean value in single mode drifted by about 0.2\% over several measurements.
-The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption; we list the wavelengths of instrumental artifacts and above-average variance in Table \ref{tab:Cary5000pars}.
-In double beam mode, the spectrograph performance was much better.
-We put a double side polished (DSP) Si wafer in the reference beam, and no sample in the target beam during the baseline scan.
-In this way the transmission spectrum of a target Si wafer is automatically normalized.
-The double beam mode effectively divides out the 0.2\% drift in the mean transmission level attributable to lamp drift, as seen in single beam measurements.
-In double beam mode, the drift in mean transmission level is merely 0.02\%.
+The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.
+In double beam mode, the spectrograph drift was much lower.
+The simultaneous reference in double beam mode divides out the 0.2\% drift in the mean transmission level attributable to lamp drift that we saw in single beam measurements.
+In double beam mode, the drift in mean transmission level is 0.02\%.
 These values are accurate for measurements over a half hour timescale or less.
+With accuracies and drift rates at this level, the curves in Figure \ref{figEtalonRelTrans} indicate that the axial extent of gaps as small as a few nm should be measurable.
 
-The choice of spectral sampling was straightforward.
 The spectral sampling was 2.0 nm.
-The spectral resolution was 5.0 nm, the lowest spectral resolution the Cary5000 offers.
+We used a spectral resolution of 5.0 nm, the lowest spectral resolution the Cary5000 offers.
 We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
 Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
-Such large gaps would be easily detectable by other means, like IR imaging.
+The axial extends of such large gaps would be easily measurable by other means, like IR imaging.
 
@@
 
-We looked for the indication of large gaps detectable as ``IR bubbles'' \verb|\cite{1992JEMat..21..669M}| with IR imaging.
+Before testing the bonded pieces from Table 1 using spectrophotometry, we looked for the indication of large gaps detectable as ``IR bubbles'' \verb|\cite{1992JEMat..21..669M}| with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
 We used a 50 mm F/2 lens, with a fiber optic white light illumination source.
-We estimate that the lamp-detector combination was sensitive to wavelengths in the range 1.1-1.6 $\mu$m.
-The field of view sub-sampled the 4 inch diameter wafers, so we dithered the sample.
+The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately 1.15$-$1.6 $\mu$m.
+The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.
 No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.
-It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the wafer during dithering.
+It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the sample when we dither its position.
 We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.
 The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
-Naturally, the unbonded DSP wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrates as discussed in the literature (see Section \ref{secHistory}).
+As expected, the unbonded DSP wafers show no such IR bubbles.
 No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.
@@
 \subsection{Constructing the predicted transmission}
-We constructed the predicted transmission in the following way.
-If IR imaging was unavailable, our optimistic prior belief was simply that the bond should be perfect and indistinguishable from a solitary DSP Si wafer, (i.e.
-100\% relative transmission).
-For the cases where we had bored gaps of known dimension and fill factors we could predict the transmission with the Fabry-Perot machinery described in section \ref{secTheory}.
-Specifically we apply Equation \ref{eqFP}.
-For non-unit fill factor, we simply average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor listed in Table \ref{tabPatternFills}.
-No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assume the delivered fill factor is equal to the designed fill factor with no uncertainty.
+We constructed the predicted transmission  for different artificially induced gaps of known dimension and fill factors using the Fabry-P\`{e}rot machinery described in Section \ref{secTheory}, in particular Equation \ref{eqFP}.
+For non-unit fill factor, we assume a perfect bond outside of the artifical gap region and average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor listed in Table \ref{tabPatternFills}.
+No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
 The depth of the meshes has an associated uncertainty listed in Table \ref{tabbondexper}.
@@
 We computed the transmission curve for the 1$\sigma$ lower and upper bounds, and plot a filled band marked as the prediction band.
-The uncertainties in the mesh depths were low enough that the band spread includes all intermediate lines of predicted transmission between plus and minus 1$\sigma$.
-In cases where the IR imaging was unavailable the mesh position was only coarsely known, so our prediction band had to incorporate our uncertainty about whether or not our measurement beam was on or off or overlapping the mesh border.
-If the beam was off the mesh we assumed 100\% relative transmission, so we simply set the lower bound of the predicted transmission to the equivalent of a zero gap.
+In some cases the IR imaging was unavailable and we had to place the spectrometer aperture on the mesh without being able to confirm the exact location of the meshed area of the substrate.
+Our prediction band had to incorporate our uncertainty about whether or not our measurement beam was overlapping the mesh border.
+This overlap behaves like a fill factor weighted towards 100\% bonded.
 
@@
 Other parts show less consistency with the prediction.
-A complete description of the diversity of the measured spectra requires an appeal to optical phenomena beyond the scope of this paper.
-Briefly we classify the diversity of measurements into the categories laid out in the table below.
 
@@
 We mitigated measurement artifacts in the following ways.
-For the Sharpie marker behaving as an anti-reflection (AR) coating, we simply removed the sharpie with IPA.
 We have mostly been able to rule out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.

### 2015-01-06 AO_bondedSi.tex
Remove Appendix A- x-ray CT scans

@@
 
-\section{Appendix: X-ray Computed Tomography}
-We acquired x-ray computed tomography of sample VG03.  The measurement details are listed in Table\ref{tab:tablenum4}.
-
-\begin{table}[h!]
-\caption{Summary of X-ray CT measurements for VG03 \label{tab:tablenum4}}
-\begin{center}
-    \begin{tabular}{ll}
-    \hline
-    \emph{Parameter (Units)} & \emph{Value} \\ 
-    \hline
-    Instrument& Xradia MicroCT \\
-     Objective & 0.7X \\
-    Accelerating voltage (kV) & 100 \\
-        Power (W) & 10 \\
-        Acquisition time (s) & 3.5 \\
-        Number of views & 2521 \\
-        Total slices & 469 \\
-        Field of view diameter (pixels) & 988 \\
-        Voxel lateral pixel resolution ($\mu$m) & 38.02 \\
-        Estimate field of view (mm) & 37.5 \\
-        Voxel axial pixel resolution ($\mu$m) & 0.5 \\
-        Scan date & 2013 August 8 \\
-        Sample thickness (mm) & 6.6 \\
-    \hline
-     \end{tabular}
-\end{center}
-\end{table}
-
-The axial resolution of the X-ray CT system is about 500 nm, which is much larger than the small ($\sim$ 15$-$90 nm) patterned gaps on many of our bonded samples.
-Part VG03 has a cylindrical patterned gap with a $\sim$20 mm diameter and 4000 nm height.
-We expected to see the cylindrical gap in about 8 resolution elements, given that the pattern was centered in the 25 mm diameter field of view.
-We saw a circular feature that extended over 297 pixels, which corresponds to 11.3 mm, assuming the voxel lateral dimension from the table above, which is reported in the X-ray CT lab measurement details.
-The feature...
-
 \section{Incoherent Multiple Reflections Transfer Matrix Method}

### 2015-01-06 AO_bondedSi.tex
all but two DTJ edits are incorporated.

@@
 The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\verb|\cite{2007fuph.book.....S}|.
-Much of the literature on the transfer matrix method deals with applications related to thin films, in which the wavelength is comparable to the size of the dielectric layer.
-For thin films the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.
+In thin films, the wavelength is comparable to the size of the dielectric layer and the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.
 The polarization state can be encapsulated in scattering matrix components\verb|\cite{2007fuph.book.....S}|.
@@
 
-The matrix for the Fabry-P\`erot etalon is calculated in the following way.
-The first key idea is that I had to treat the entire air gap as an abstraction, a Fabry-P\`erot etalon.
-We don't need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon.
+The matrix for the silicon air gap is calculated in the following way.
+The first key idea is to treat the entire air gap as an abstraction, a Fabry-P\`erot etalon.
+We do not need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon:
 
 \begin{eqnarray}
- \delta = \frac{2\pi}{\lambda}2d \\
+ \delta = \frac{2\pi}{\lambda}2d \label{eq:phase} \\
   F \equiv \frac{4R}{(1-R)^2} \\
@@
 The coefficient of finesse $F$ and the phase $\delta$ are the two parameters of the Fabry-P\`erot etalon.
-The coefficient of Finesse encapsulates the Fresnel reflection and depends only on the Si refractive index which has a minute wavelength (and temperature) dependence.
-The phase depends on the wavelength $\lambda$ and $d$ the air gap spacing: $\delta=\frac{2\pi}{\lambda}2d$ .
+The coefficient of Finesse encapsulates the Fresnel reflection and depends only on the Si refractive index which has only a small wavelength (and temperature) dependence.
+The phase depends (Equation \ref{eq:phase}) on the wavelength $\lambda$ and $d$ the air gap spacing.
 We assume the etalon is lossless, i.e.
-the etalon has no absorption $T_e+R_e=1$.
+$T_e+R_e=1$.
 So the incoherent scattering and transfer matrices for the etalon are:
@@
 
-I assembled the matrix for the Air-Si Fresnel interface in the following way.
+We assembled the matrix for the Air-Si Fresnel interface at the exterior of the bonded Si parts in the following way.
 First it is important to note that the matrix is the same whether the transmission is from Si to air or air to Si.
@@
 
-Throughout this article we have assumed that $L >> \lambda^2 / \delta \lambda$, so that the interference of the substrate is safely incoherent.
+Throughout this article we have assumed that the thickness $L$ of the wafers is much greater than the coherence length scale: $L >> \lambda^2 / \delta \lambda$, so that the interference of the substrate is safely incoherent.
 Another way of stating this criterion is that the spectral bandwidth is large enough that high spectral frequency fringes average out in a composite spectrum of low spectral resolution.

### 2015-01-06 AO_bondedSi.tex
delete the entire last appendix about high spectral resolution coherence

@@
 
-\section{Appendix: High spectral resolution coherent interference of thin substrates}
-
-Throughout this article we have assumed that the thickness $L$ of the wafers is much greater than the coherence length scale: $L >> \lambda^2 / \delta \lambda$, so that the interference of the substrate is safely incoherent.
-Another way of stating this criterion is that the spectral bandwidth is large enough that high spectral frequency fringes average out in a composite spectrum of low spectral resolution.
-If we relax this assumption so that the Si substrate thickness is comparable to or smaller than the coherence length, we would begin to observe high frequency fringes in the transmission spectra.
-In this appendix section we work out the transmission spectrum for two scenarios.
-First, we work out the case of a DSP Si wafer of thickness $L$, with no gap.
-Second we work out the case of two bonded Si wafers of thicknesses $L_1$ and $L_2$.
-For both cases we consider infinite spectral resolution, then convolve the spectra with a finite spectral resolution to show that the spectrum tends to that derived using an incoherent assumption.
-
-\subsection{DSP Si wafer at high spectral resolution}
-The Fabry-P\`erot etalon transmission is given in Equation \# X (currently, 5).  What we want to evaluate is:
-
-\begin{eqnarray}
- <T_e> = \lim_{\sin^2(\delta/2) \to 0.5} \frac{1}{1+F\sin^2(\delta/2)} \\
- <T_e> = \frac{1}{1+F/2}
-\end{eqnarray}
-
 \bibliographystyle{osajnl}

### 2015-01-19 AO_bondedSi.tex
add new figures from IPython notebooks- VG03, etc.

@@
 \usepackage{longtable}
+\usepackage{subcaption}
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/fpAbsorbfig}}
+\centerline{\includegraphics[width=0.95\columnwidth]{figs/fpAbsorbfig_alt}}
 \caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
-\label{figSiAbsorbfig} We detect Si absorption shortward of $\lambda \; =$ 1250 nm.
-We assume spectral regions longward of 1250 nm are non-absorptive.
-The correlated and heteroscedastic measurement errors are typically less than 0.2\%.}
+\label{figSiAbsorbfig} We detect Si absorption shortward of $\lambda$ = 1250 nm.
+The individual measurements have correlated uncertainties attributable to lamp drift at the level of 0.2\%.
+The random uncertainties are typically about 0.02\% per sample, though some wavelength regions (e.g.
+1700-1900 nm) demonstrate much greater per sample uncertainties.
+Absorption longward of 1250 nm (vertical dashed line) is negligible.}
 \end{figure}
@@
 
-%The following 3 figures were made with fabry_perot_Si_20140421.pro and other variants.
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/20140421_absolute.pdf}}
-\caption{Modeled absolute transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 35 nm (cyan triple-dot dashed curve)\label{figAbsoluteTrans}.
+\centerline{\includegraphics[width=0.65\columnwidth]{figs/etalon_trans.pdf}}
+\caption{\label{figAbsoluteTrans}The family of curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
+\emph{Top panel}: Absolute transmission through a Si substrate with no gap (top denim-blue solid curve), and small gaps of axial extent 20 nm (green dashed curve), 35 nm (red dash-dotted curve), and 60 nm (purple dotted curve).
 The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.
 This curve is only accurate for normal incidence transmission.
-Furthermore, the curve assumes the multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/Rel_etalon_trans_stack.pdf}}
-\caption{Modeled etalon transmission through a silicon substrate with no gap (top black solid curve), and small gaps of axial extent 20 nm (red dashed curve) and 60 nm (cyan triple-dot dashed curve)\label{figEtalonRelTrans}.
-The curves are examples from a family defined in Equation \ref{eqFP}.
-The limit of no gap is 100\% transmissive, compared to a DSP reference.
-A 60 nm gap degrades the transmission over a DSP reference by 12\% at 1200 nm.
-Modeled etalon transmission through a silicon substrate with gaps of axial extent 35 nm (red dotted curve), 350 nm (green dash-dot curve), and 3500 nm (blue long-dashed curve).
-The curves are examples from a family defined in Equation \ref{eqFP}.
-The 3500 nm gap curve has obvious Fabry-P\`erot fringes, whereas the 35 and 350 nm gap curves vary smoothly with wavelength.}
+Furthermore, the curve assumes the multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.
+\emph{Middle panel}: The ``Gap Transmission'', defined as the Absolute transmission through a sample normalized by the absolute transmission through a gapless sample.
+The gap sizes and colors are the same as in the top panel.
+\emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0-3500 nm.
+The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\`erot etalon fringes.}
 \end{figure}
@@
 \centerline{\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}}
-\caption{Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section\label{figVS20pattern}.
-The four inch diameter VS20 is comprised of two thick silicon wafers.
+\caption{
+\label{figVS20pattern}
+Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section.
+The 100 mm diameter VS20 is comprised of two thick silicon wafers.
 We patterned three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
 The depth was 95 $\pm$ 5 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
-The Veeco was operated in PSI mode.
-The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.
-The technique can reveal the axial extent and fill factor of an embedded gap.
-}
+The fill factor for this and all mesh patterns was 50\%.
+The purpose for these patterns was to have gaps of known axial extent against which we could test the optical metrology technique we describe in this article.}
 \end{figure}
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posM1.pdf}}
-\caption{IR transmission spectrum of VS21 (M1) showing consistent gap\label{VS21specm1}.  Example spectrum through part VS21 at a position through the mesh gap of known dimensions.}
+        \centering
+        \begin{subfigure}[b]{0.5\textwidth}
+              \includegraphics[width=\textwidth]{figs/VG03_corner.pdf}
+              \caption{Corner plot of our fitted gap size $d$ and fill factor $f$ for VG03}
+		\label{figVG03_f100}
+        \end{subfigure}
+
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\textwidth]{figs/VG03_f100.pdf}
+                \caption{Measured spectrum and model fit of VG03}
+                \label{figVG03_corner}
+        \end{subfigure}
+        \caption{Pictures of animals}\label{fig:animals}
+\caption{
+
+\emph{Top panel:} Corner plot of MCMC samples in our non-linear model fit of the transmission spectrum of Si optic part \# VG03.
+The physical parameters of the model are the axial extent of the gap $d$, and the areal fill factor $f$ of the gap over the measurement region.
+The MCMC model fit also includes nuisance parameters $a$ and $s$ (not shown) for the Gaussian Process covariance model.
+
+\emph{Bottom panel:}Transmission through bonded Si sample VG03, normalized by the transmission of a DSP Si wafer.
+VG03 has a gap of axial extent $\sim4\;\mu$m implanted in its interface.
+The measured spectrum (black stepped line) in this region shows characteristic Fabry-P\`erot fringes of constructive and destructive interference.
+The only tunable parameters of the model are the gap axial extent, $d$ and fill factor $f$.
+No scaling has been applied.
+The red dashed line shows the best model fit with $d$ = 3960 $\pm$ 2 nm, consistent with 100\% fill factor.
+The blue band shows the collective locations of 60 random draws from the Gaussian process model, described in the text.
+Wavelengths shortward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.
+}
 \end{figure}
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.8\columnwidth]{figs/20130911_VS21posF1}}
-\caption{IR transmission spectrum of part VS21 (F1) showing evidence for an excess gap\label{VS21specf1}.
-The spectrum was taken at a position through a mesh gap of depth 95 $\pm$5 with 50\% fill factor.
-The predicted spectrum lines up with the measurement to within the measurement and model uncertainties.}
+\begin{figure}
+        \centering
+        \begin{subfigure}[b]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{gull}
+                \caption{A gull}
+                \label{fig:gull}
+        \end{subfigure}%
+        ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
+          %(or a blank line to force the subfigure onto a new line)
+        \begin{subfigure}[b]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{tiger}
+                \caption{A tiger}
+                \label{fig:tiger}
+        \end{subfigure}
+        ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
+          %(or a blank line to force the subfigure onto a new line)
+        \begin{subfigure}[b]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{mouse}
+                \caption{A mouse}
+                \label{fig:mouse}
+        \end{subfigure}
+        \caption{Pictures of animals}\label{fig:animals}
 \end{figure}

### 2015-01-19 AO_bondedSi.tex
revise and add figure about VG12

@@
 \usepackage{setspace}
-\usepackage{longtable}
+\usepackage{caption}
 \usepackage{subcaption}
+%\usepackage{longtable}
 
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/fpAbsorbfig_alt}}
+\centerline{\includegraphics[width=0.40\columnwidth]{figs/fpAbsorbfig_alt}}
 \caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.65\columnwidth]{figs/etalon_trans.pdf}}
+\centerline{\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}}
 \caption{\label{figAbsoluteTrans}The family of curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
@@
               \includegraphics[width=\textwidth]{figs/VG03_corner.pdf}
-              \caption{Corner plot of our fitted gap size $d$ and fill factor $f$ for VG03}
-		\label{figVG03_f100}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 position 1}
+		\label{figVG03_corner}
         \end{subfigure}
@@
                 \includegraphics[width=\textwidth]{figs/VG03_f100.pdf}
-                \caption{Measured spectrum and model fit of VG03}
-                \label{figVG03_corner}
+                \caption{Measured spectrum and fit of VG03 position 1}
+                \label{figVG03_f100}
         \end{subfigure}
-        \caption{Pictures of animals}\label{fig:animals}
-\caption{
-
-\emph{Top panel:} Corner plot of MCMC samples in our non-linear model fit of the transmission spectrum of Si optic part \# VG03.
+\caption{ \emph{Top panel:} Corner plot of MCMC samples in our non-linear model fit of the transmission spectrum of Si optic part \# VG03.
 The physical parameters of the model are the axial extent of the gap $d$, and the areal fill factor $f$ of the gap over the measurement region.
 The MCMC model fit also includes nuisance parameters $a$ and $s$ (not shown) for the Gaussian Process covariance model.
-
 \emph{Bottom panel:}Transmission through bonded Si sample VG03, normalized by the transmission of a DSP Si wafer.
@@
 The blue band shows the collective locations of 60 random draws from the Gaussian process model, described in the text.
-Wavelengths shortward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.
-}
+Wavelengths shortward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.\label{figVG03full} }
 \end{figure}
 
-\begin{figure}
+\begin{figure}[htbp]
         \centering
-        \begin{subfigure}[b]{0.3\textwidth}
-                \includegraphics[width=\textwidth]{gull}
-                \caption{A gull}
-                \label{fig:gull}
-        \end{subfigure}%
-        ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.3\textwidth}
-                \includegraphics[width=\textwidth]{tiger}
-                \caption{A tiger}
-                \label{fig:tiger}
+        \begin{subfigure}[b]{0.5\textwidth}
+              \includegraphics[width=\textwidth]{figs/VG03p2_corner.pdf}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 position 2}
+		\label{figVG03p2_corner}
         \end{subfigure}
-        ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
-          %(or a blank line to force the subfigure onto a new line)
-        \begin{subfigure}[b]{0.3\textwidth}
-                \includegraphics[width=\textwidth]{mouse}
-                \caption{A mouse}
-                \label{fig:mouse}
+
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\textwidth]{figs/VG03_f045.pdf}
+                \caption{Measured spectrum and fit of VG03 position 2}
+                \label{figVG03_f045}
         \end{subfigure}
-        \caption{Pictures of animals}\label{fig:animals}
+\caption{ Same as Figure \ref{figVG03full}, but the measurement was in a region where less than $<5\%$ of the area exhibited the $\sim4\; \mu$m gap.
+The measured spectrum in this region shows subdued Fabry-P\`erot fringes of constructive and destructive interference.
+The model derived axial extent is 4094 $\pm$ 5 nm covering a fill factor of 4.58 $\pm$ 0.06$\%$.
+\label{figVG03part}}
+\end{figure}
+
+\begin{figure}[htbp]
+        \centering
+        \begin{subfigure}[b]{0.5\textwidth}
+              \includegraphics[width=\textwidth]{figs/VG12_corner.pdf}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG12}
+		\label{figVG12_corner}
+        \end{subfigure}
+
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\textwidth]{figs/VG12_f50.pdf}
+                \caption{Measured spectrum and fit of VG12 position 2}
+                \label{figVG12_f50}
+        \end{subfigure}
+\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
+The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.
+The measured spectrum also shows a deficit of transmission compared to the predicted transmission of a 49 $\pm6\;$nm gap with 50\% fill factor (green apple shaded color band).
+The discord between prediction and measurement could be due to several factors including an underestimate of the intrinsic gap size, an underestimate of the intrinsic fill factor, or an additional gap in excess of the prediction.
+We favor the latter interpretation, since undesired gaps are common in these direct-bonded Si samples that have not undergone high-temperature thermal annealing.
+\label{figVG12}}
 \end{figure}

### 2015-01-20 AO_bondedSi.tex
Properly typeset the references by removing \verb prefixes

@@
 
-Placeholder citation so the document will compile: see: \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
-
 \subsection{Why Silicon is good}
 Crystalline silicon is an excellent optical material for the infrared.
-Si has a high refractive index ($\sim$3.4, Frey et al.
-2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \verb|\cite{PhysRev.108.268, PhysRev.78.178}| and in the far-IR \verb|\cite{doi:10.1117/12.323764}|.
-It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \verb|\cite{doi:10.1117/12.323764}|.
+Si has a high refractive index ($\sim$3.4, \cite{2006SPIE.6273E..77F}) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
+It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
 There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
@@
 As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
-One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \verb|\cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}|.
+One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
 A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.
@@
 
-There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \verb|\cite{1998AnRMS..28..215G,Masteika2014}|.
-The foundation paper on Si bonding came in 1986 \verb|\cite|{1986JAP....60.2987S}|.
+There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \cite{1998AnRMS..28..215G,Masteika2014}.
+The foundation paper on Si bonding came in 1986 \cite{1986JAP....60.2987S}.
 After that, there was a sequence of papers that used IR imaging to look for evidence of defects in the bond.
@@
 (1988) and Goesele et al.
-(1995) \verb|\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G}| directly monitored the bond propagation with IR imaging.
+(1995) \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
 Lehman et al.
 1989 and Goesele et al.
-\verb|\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G}| demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
-Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \verb|\cite{Horn2009, Masteika2014}|.
+\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
+Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \cite{Horn2009, Masteika2014}.
 
@@
 Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance.
-Figure 5 of Cooper 1986\verb|\cite{doi:10.1080/02786828608959094}| shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
+Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
 particle size ($\mu$m).
 There will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
-On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \verb|\cite{Masteika2014}|.
+On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \cite{Masteika2014}.
 Most of the bonding literature concerns itself with wafer bonding.
@@
 In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap.
-X-ray topography \verb|\cite{Mitani1990}| exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
-In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \verb|\cite{2000RScI...71.1869G}|.
+X-ray topography \cite{Mitani1990} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
+In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
 
@@
 
-It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \verb|\cite{Mitani1990}|.
+It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \cite{Mitani1990}.
 
@@
 
-Our measurement technique exploits the large Fresnel reflection \verb|\cite{2001opt4.book.....H}| at the Si$-$vacuum interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \verb|\cite{2006SPIE.6273E..77F}|, one of the largest refractive indices of any conventional material.
+Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
+The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
 The Fresnel reflection is about 30\% per surface.
 For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
-Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \verb|\cite{2006SPIE.6273E..77F}|.
+Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
 
@@
 
-The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\verb|\cite{2007fuph.book.....S}|.
+The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
 The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
@@
 We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at $T= 295\;$K.
-We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \verb|\cite{2006SPIE.6273E..77F}|, as shown in Figure \ref{figSiIndexFinesse}.
+We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
 We drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
@@
 The reflectivity $R$ (Equation \ref{Eq:FresnelR}) is about 30\% in the wavelength range 1100 to 5000 nm.
-We define the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
+We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
 The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength for $T = 295\;$K and $T = 77\;$K.
@@
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
-\verb|\cite{2006SPIE.6273E..77F}| cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
+\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
 An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
@@
 The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps where the gap size si small enough that the cavity it forms can always be treated as a coherent structure.
-Treatments for multiple coherent reflections have been described in detail in the optics literature \verb|\cite{2007fuph.book.....S}|.
+Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
 The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \verb|\cite{2002ApOpt..41.3978K}| to deal with the cavities within the Si.
+We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K} to deal with the cavities within the Si.
 Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
@@
 The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is slower in thick pucks \verb|\cite{2007ApOpt..46.6793H}| than in wafers.
+There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
 Our ultimate goal is to bond optics with thicknesses as large as 30 mm.
@@
     VG15   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F  &14 $\pm$2 & \\
-    VS20   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F & 95 $\pm$5 & See Fig. \ref{figVS20pattern}\\
+    VS20   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F & 95 $\pm$5 & \\
     \hline
@@
     Petal & $\sim1$ & $\sim$2000 & long petal lines \\         
-    Mesh F & 50 & 40 &  100 $\mu$m square holes, Fig. \ref{figVS20pattern}\\ 
+    Mesh F & 50 & 40 &  100 $\mu$m square holes, Fig. 
+    VS20pattern}\\ 
     Mesh M & 50 & 200 &  500 $\mu$m square holes\\ 
@@
 
-Before testing the bonded pieces from Table 1 using spectrophotometry, we looked for the indication of large gaps detectable as ``IR bubbles'' \verb|\cite{1992JEMat..21..669M}| with IR imaging.
+Before testing the bonded pieces from Table 1 using spectrophotometry, we looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.
-We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \verb|\cite{1992JEMat..21..669M}|.
+We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
 Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.
@@
 
-The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \verb|\cite{2007fuph.book.....S}|.
+The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.
 Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
@@
 
-The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\verb|\cite{2007fuph.book.....S}|.
+The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\cite{2007fuph.book.....S}.
 In thin films, the wavelength is comparable to the size of the dielectric layer and the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.
-The polarization state can be encapsulated in scattering matrix components\verb|\cite{2007fuph.book.....S}|.
+The polarization state can be encapsulated in scattering matrix components\cite{2007fuph.book.....S}.
 The intensities of the emergent spectrum can be computed from the absolute square of the complex amplitudes.
 For thick films, the vector components are the intensities of the emergent spectrum, since waves are incoherent.
-\verb|\cite{2002ApOpt..41.3978K}| work out the general transfer-matrix method for optical multilayer systems with incoherent interference.
+\cite{2002ApOpt..41.3978K} work out the general transfer-matrix method for optical multilayer systems with incoherent interference.
 The key idea for the incoherent transfer matrix method approach is to populate a scattering matrix with elements equal to the (wavelength dependent) transmitted $T_i$ and reflected $R_i$ intensities of an interface or set of interfaces that act together.
@@
 
-From the matrix transformation equations \verb|\cite{2007fuph.book.....S}| we know $M_{22}=1/T$.
+From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.
 Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:

### 2015-01-20 AO_bondedSi.tex
break out the tables into \input{} in their own files

@@
 \label{secTheory}
-A pair of bonded silicon wafers or pucks with a small gap actually forms there coupled cavities with the air gap as the the central cavity.
+A pair of bonded silicon wafers or pucks with a small gap actually forms their coupled cavities with the air gap as the the central cavity.
 The outer cavities, those within the Si, however, were not produced with coherence in mind.
@@
 It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.
-Figure \ref{figEtalonRelTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
+Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
 In Section \ref{secResults} we compare wavelength dependent transmission measurements to by-eye fits of Equation \ref{eqFP}.
@@
 
-\begin{table}[h!]
-\caption{UTexas Si bonding experiments \label{tabbondexper}}
-\begin{center}
-    \begin{tabular}{ c c c c c c}
-    \hline
-    Name & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
-    -  & mm & mm & - & nm & - \\ 
-        \hline
-    HA1     & 75   & $10$ $\times$ 2& None & - & \\
-    VG01   & 100 &$1$ $\times$ 2 & None  & - &  \\
-    VG02   & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
-    VG03   & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & \\    
-    VG04   & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15$\pm7$ & \\
-    VG07   & 100 & $1$ $\times$ 2 &  mesh C, M, F & 12$\pm7$ & \\ %see email from Victor on 5/16/13
-    VG08   & 100 & $1$ $\times$ 2 &  mesh C, M, F & 12$\pm7$ & \\ 
-    VG09   & 75   & $1$ $\times$ 2 & mesh C & 49 $\pm$6 & \\
-    VG15   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F  &14 $\pm$2 & \\
-    VS20   & 100 & $0.8$ $\times$ 2 &  mesh C, M, F & 95 $\pm$5 & \\
-    \hline
-    \end{tabular}
-\end{center}
-\end{table}
+\input{tbls/tbl_experiments.tex}
 
-\begin{table}[h!]
-\caption{Patterned gap properties \label{tabPatternFills}}
-\begin{center}
-    \begin{tabular}{ c c c c }
-    \hline
-    Pattern name & fill factor & minimum feature size & bulk description \\ 
-    - & \% & $\mu$m & - \\ 
-    \hline
-    Hole & 100 & - & 25 mm diameter hole \\     
-    Petal & $\sim1$ & $\sim$2000 & long petal lines \\         
-    Mesh F & 50 & 40 &  100 $\mu$m square holes, Fig. 
-    VS20pattern}\\ 
-    Mesh M & 50 & 200 &  500 $\mu$m square holes\\ 
-    Mesh C & 50 & 620 &  1500 $\mu$m square holes\\     
-        \hline
-    \end{tabular}
-\end{center}
-\end{table}
+\input{tbls/tbl_meshPatterns.tex}
 
@@
 These values are accurate for measurements over a half hour timescale or less.
-With accuracies and drift rates at this level, the curves in Figure \ref{figEtalonRelTrans} indicate that the axial extent of gaps as small as a few nm should be measurable.
+With accuracies and drift rates at this level, the curves in Figure \ref{figAbsoluteTrans} indicate that the axial extent of gaps as small as a few nm should be measurable.
 
@@
 
-\begin{table}[h!]
-\caption{Summary of Cary 5000 Measurement parameters \label{tab:Cary5000pars}}
-\begin{center}
-\begin{tabular}{ c c }
-\hline
-        Parameter (Units) & Value \\ 
-\hline
-        Spectral sampling interval (nm) & 2.0$-$10.0 \\
-        Spectral resolution (nm) & 5 \\
-        Time per sample (s) & 0.1 \\
-        Typical measurement range (nm) & 1000-2500 \\
-	Typical Single mode mean drift (\%) & 0.20 \\
- 	Beam size (mm $\times$ mm) & $2 \times 8$ \\
-        Directly measured linearity range (dB) & $>24$ \\
-    \hline
-    \end{tabular}
-\end{center}
-\end{table}
+\input{tbls/tbl_cary5000Pars.tex}
 
@@
 
-% Table of experiments
-\begin{table}[h!]
-\caption{Cary5000 Experiment Log \label{tabExperiments}}
-\begin{center}
-    \begin{tabular}{c c c c c c c}
-    \hline
-	Date &  Mode  & $\Delta \lambda$ & SBW & Sampling & \# of spectra & Sample IDs \\
-	\emph{mm/dd/yyyy} &  -  & nm & nm & nm & - \\
-%     \hline		
-%	02/04/2013 & ? & ? 				& 5.0 & 10.0 &  16 &   -  \\
-%     \hline		
-%	02/06/2013 & ? & ? 				& 5.0 & 10.0 &  16 &   -  \\	     
-     \hline		
-     	02/18/2013 & Single & 1000$-$2500 & 5.0 & 10.0 &  16 &   -  \\
-     	                   &            &                       &       &         &    4 &  Baseline\\	
-     	                   &            &                       &       &         &    1 &  VG01\\
-	                   &            &                       &       &         &    4 &  VG02\\	                   
-	                   &            &                       &       &         &    2 &  VG03\\	                   
-	                   &            &                       &       &         &    2 &  VG04\\
-	                   &            &                       &       &         &    2 &  VG05\\	                   
-	                   &            &                       &       &         &    2 &  VG06\\
-    \hline
-%	05/02/2013 & Single & 1000$-$2500 & 5.0 & 5.0  &  30 &   -  \\
-%     	                   &            &                       &       &         &    12 &  Baseline\\	
-%     	                   &            &                       &       &         &    18 &  Other\\
-%    \hline
-	05/03/2013 & Single & 1000$-$2500 & 5.0 & 5.0  &  28 &   -  \\
-     	                   &            &                       &       &         &    8 &  Baseline\\	
-     	                   &            &                       &       &         &    1 &  VG05\\ %Consider replacing VG05 with "DSP reference sample"  or something
-     	                   &            &                       &       &         &    6 &  VG03\\
-     	                   &            &                       &       &         &    13 &  Other\\ %These had to do with beam size experiments
-    \hline
-	06/07/2013 & Mix & 1000$-$2500 & 5.0 & 5.0  &  43 &   -  \\
-	                   &            &                       &       &         &    11 &  Baseline\\	                   
-	                   &            &                       &       &         &    9 &  VG05\\	                   
-	                   &            &                       &       &         &    7 &  VG07\\
-	                   &            &                       &       &         &    6 &  VG08\\
-	                   &            &                       &       &         &    4 &  HA1\\
-	                   &            &                       &       &         &    6 &  other\\ % These are mostly just baselines without automatic division
-    \hline
-	08/02/2013 & Single & 1000$-$2500 & 5.0 & 2.0  &  17 &   -  \\	
-	                   &            &                       &       &         &    3 &  Baseline\\
-	                   &            &                       &       &         &    4 &  VG05\\	                   
-	                   &            &                       &       &         &    2 &  VG15\\
-	                   &            &                       &       &         &    2 &  VS20\\
-	                   &            &                       &       &         &    6 &  VG09\\
-    \hline
-    	09/11/2013 & Double & 1000$-$2500 & 5.0 & 2.0  &  20 &   -  \\	                   
-	                   &            &                       &       &         &    1 &  Baseline\\
-	                   &            &                       &       &         &    14 &  VS21\\	                   
-	                   &            &                       &       &         &    5 &  Blank\\ %This is really the same as a baseline
-    \hline
-    	09/16/2013 & Double & 1000$-$2500 & 5.0 & 2.0  &  21 &   -  \\
-	                   &            &                       &       &         &    2 &  Baseline\\
-	                   &            &                       &       &         &    7 &  VG07\\	                   
-	                   &            &                       &       &         &    4 &  VG05\\	
-	                   &            &                       &       &         &    4 &  VG12\\
-	                   &            &                       &       &         &    2 &  VG15\\	                   
-	                   &            &                       &       &         &    2 &  VG08\\
-    \hline
-    \end{tabular}
-\end{center}
-\end{table}
+\input{tbls/tbl_experimentLog.tex}
 

### 2015-01-21 AO_bondedSi.tex
add VS20-21 IR image, add text about correlated errors,

@@
 
-\section{Measurement strategy, conditions, and details}
+\section{Measurement technique and correlated measurement errors}
 We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
-Table \ref{tab:Cary500pars} lists the typical measurement settings and tool performance.
+Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
 The tool has a double monochromator, which reduces scattered light.
@@
 In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.
+
 We estimated the uncertainty and measurement repeatability of single and double modes in the following way: we computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
@@
 These values are accurate for measurements over a half hour timescale or less.
-With accuracies and drift rates at this level, the curves in Figure \ref{figAbsoluteTrans} indicate that the axial extent of gaps as small as a few nm should be measurable.
+Table \ref{tabCary5000pars} summarizes the amplitudes of the drifts and individual measurement errors in both single and double beam modes.\emph{Correlated errors dominate the measurement uncertainty of our transmission spectra.}  The measurement uncertainty correlations are apparent in in Figure \ref{figSiAbsorbfig}.
+In some parts of the spectrum, the mean drift is about ten times larger than the standard deviation of the spectrum.
+Double beam mode has about ten times lower lamp drift but suffers from other instrumental artifacts, like disjointed spectra at the wavelengths of known grating changeover points.
 
@@
 
-Before testing the bonded pieces from Table 1 using spectrophotometry, we looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
+We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 The morphology was clear regardless of the detector response.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=0.3\columnwidth]{figs/VS2021_IR_image.bmp}}
+\caption{
+\label{figVS2021_IR_image}
+IR image of part VS20-21.
+Three rectangular mesh areas are apparent (Table \ref{tabPatternFills}).
+The coarsest mesh, C, shows substructure, while the medium and fine meshes are indistinguishable at the low resolution of the image.
+The three dark spots are gaps at the inferface of the direct bonded Si samples.}
+\end{figure}
 
@@
 
+\subsectio{Inferring axial extent of gaps from noisy transmission spectra}
+We infer the axial extent of the gaps with correlated measurement uncertainties using Gaussian Processes.
+
 \subsection{Spectra of direct bonded Si with known gap sizes}
-Figures \ref{VS21specm1} to \ref{VS21specf1} show examples of the predicted and measured IR spectra.
-Figure \ref{VS21specm1} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
-Meanwhile, Figure \ref{VS21specf1} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 2 \sigma$).
-Other parts show less consistency with the prediction.
+Figures \ref{figVG03full} to \ref{figVG12_f50} show examples of the predicted and measured IR spectra.
+Figure \ref{figVG03_f100} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
+Meanwhile, Figure \ref{figVG12_f50} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 3 \sigma$).
+Other parts show less consistency with the prediction due to the difficulty of direct-bonding Si in the absence of thermal annealing.
 

### 2015-01-21 AO_bondedSi.tex
convert VS20-21 image from bitmap to pdf, fix bug in subsection heading

@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.3\columnwidth]{figs/VS2021_IR_image.bmp}}
+\centerline{\includegraphics[width=0.4\columnwidth]{figs/VS2021_IR_image.pdf}}
 \caption{
@@
 
-\subsectio{Inferring axial extent of gaps from noisy transmission spectra}
+\subsection{Inferring axial extent of gaps from noisy transmission spectra}
 We infer the axial extent of the gaps with correlated measurement uncertainties using Gaussian Processes.

### 2015-01-21 AO_bondedSi.tex
write up the Gaussian Process regression stuff

@@
 Table \ref{tabbondexper} has more information about the bonding samples, and Table \ref{tabPatternFills} lists the properties of the patterns.
+
 The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
 In a second round of experiments, we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
 For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
@@
 
-\subsection{Constructing the predicted transmission}
-We constructed the predicted transmission for different artificially induced gaps of known dimension and fill factors using the Fabry-P\`{e}rot machinery described in Section \ref{secTheory}, in particular Equation \ref{eqFP}.
-For non-unit fill factor, we assume a perfect bond outside of the artifical gap region and average 100\% transmission with the predicted gap transmission, weighted by the estimated fill factor listed in Table \ref{tabPatternFills}.
-No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
-The depth of the meshes has an associated uncertainty listed in Table \ref{tabbondexper}.
-We incorporated the uncertainty in the mesh depth in the following way.
-We computed the transmission curve for the 1$\sigma$ lower and upper bounds, and plot a filled band marked as the prediction band.
-In some cases the IR imaging was unavailable and we had to place the spectrometer aperture on the mesh without being able to confirm the exact location of the meshed area of the substrate.
-Our prediction band had to incorporate our uncertainty about whether or not our measurement beam was overlapping the mesh border.
-This overlap behaves like a fill factor weighted towards 100\% bonded.
+\subsection{Inferring axial extent of gaps from noisy transmission spectra}
+We infer the axial extent of gaps from their noisy transmission spectra.
+We use a Bayesian inference technique.
+First we define a mixture model in which the observed normalized gap spectrum $T_{obs}$ is composed of a sum of a spectrum through a gap of axial extent $d$, and a spectrum through a perfect bond:
 
-\subsection{Inferring axial extent of gaps from noisy transmission spectra}
-We infer the axial extent of the gaps with correlated measurement uncertainties using Gaussian Processes.
+\begin{equation}
+	T_{mix} = f T_{e} + (1-f) \label{eqnMix}
+\end{equation}
+
+where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.
+$T_e$ is defined in equation \ref{eqn:Tetalon}.
+The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
+Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tabPatternFills}.
+These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
+
+Equipped with an observed spectrum and our mixture model, we constructed the Likelihood function \verb|\cite{SomeStatisticsBook}|, $\mathcal{L} = p(T_{obs} | d, f)$.
+The measurement errors are not assumed to be independent and identically distributed, and in fact are strongly correlated as discussed in Section \verb|\ref{sectionAboutErrors}|.
+To handle the correlated measurement errors we use Gaussian Processes \verb|\cite{RasmussenAndWilliams, DFM}|.
+The Gaussian Process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.
+We parameterize the covariance matrix elements in the following way:
+
+\begin{equation}
+	C_{ij} = \sigma^2_{i}\delta_{ij}+a^2\exp{(-\frac{(\lambda_i-\lambda_j)^2}{2l^2})} \label{eqnGPkernel}
+\end{equation}
+
+where $\delta_{ij}$ is the Kronecker delta, and $\sigma_i$ are the independent measurement uncertainties on the $i^{\mathrm{th}}$ data point.
+For simplicity, we adopt values from Table \ref{tabCary5000pars} for single and double beam modes respectively.
+This assumption of homoscedasticity is fairly safe, since the correlated errors dominate in comparison to individual measurement errors.
+The parameter $l$ controls the correlation length, and $a$ controls the amplitude of the noise.
+We do not know $a$ nor $l$ and they are, in general, different for each measurement.
+The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.
+The value of $l$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (thousands of nm) to capture global features attributable to lamp drift.
+We allow both $a$ and $s$ to be free parameters in our fitting procedure.
+
+We set our priors \verb|\cite{someStatisticsBook}| on $\ln{a}$, $\ln{s}$, $d$, and $f$ as uniform over a wide range of values consistent with visual inspection of each spectrum.
+The final un-normalized posterior probability function, expressed as a natural logarithm is:
+
+\begin{equation}
+	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -0.5\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r}} -0.5 \ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
+\end{equation}
+
+where the residual vector $\boldsymbol{r}$ is the data $T_{obs}$ minus the model $T_{mix}$.
+We produce posterior samples with Markov Chain Monte Carlo (MCMC).
+Specifically, we implemented \texttt{emcee}\footnote{https://github.com/dfm/emcee}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.
+We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the observed normalized Transmission spectra.
+Figures \ref{figVG03full, figVG03part, figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
 
 \subsection{Spectra of direct bonded Si with known gap sizes}
+\label{secKnownGaps}
 Figures \ref{figVG03full} to \ref{figVG12_f50} show examples of the predicted and measured IR spectra.

### 2015-01-22 AO_bondedSi.tex
document compiles with GP section and long equations, added the IR image pdf
Add the derived Gap Sizes table

@@
 The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.
-No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.
-It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the sample when we dither its position.
+Figure \ref{figVS2021_IR_image} shows an IR image of sample VS20-21.
+For this image we coarsely flat fielded the detector by observing a homogeneously illuminated white screen.
+Despite the flat fielding, detector non-uniformities and vignetting are perceptible in our image as vertical stripes.
+Mosaic-stitching errors are also perceptible.
+It was easy to detect IR bubbles despite detector artifacts since the bubbles translate with the sample when we dither its position.
 We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.
 The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
+The individual IR bubble diameters were as small as a few square millimeters up to 400 mm$^2$.
+The morphologies of the IR bubbles ranged from perfectly circular to irregular.
+The structure within the IR bubbles was sometimes indicative of Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
 As expected, the unbonded DSP wafers show no such IR bubbles.
-No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.
-The morphology was clear regardless of the detector response.
+No attempt at quantitative measurement was made on the IR images due to evidence of a non-linear detector response.
+The bulk morphologies were clear regardless of the detector response.
+When IR imaging was available, it was leveraged to target IR specectrscopy in locations demonstrating presence or absense of IR bubbles, as desired.
+In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.
 
@@
 In this section we describe the spectroscopy of directly bonded Si.
-In some instances IR imaging was available and was leveraged to target or avoid locations demonstrating IR bubbles.
-In other cases the IR imaging was not available or ignored, and spectra were taken at random positions on the wafer surface.
 
@@
 
-We set our priors \verb|\cite{someStatisticsBook}| on $\ln{a}$, $\ln{s}$, $d$, and $f$ as uniform over a wide range of values consistent with visual inspection of each spectrum.
+We set our prior probability distribution functions \verb|\cite{someStatisticsBook}|, $\ln{p(d,f,a,s)}$ as uniform over a wide range of values for $\ln{a}$, $\ln{s}$, $d$, and $f$ consistent with visual inspection of each spectrum.
 The final un-normalized posterior probability function, expressed as a natural logarithm is:
@@
 \begin{equation}
-	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -0.5\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r}} -0.5 \ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
+	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -0.5\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -0.5\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
 \end{equation}
@@
 We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the observed normalized Transmission spectra.
-Figures \ref{figVG03full, figVG03part, figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
+Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
 
@@
 \label{secKnownGaps}
-Figures \ref{figVG03full} to \ref{figVG12_f50} show examples of the predicted and measured IR spectra.
+Figures \ref{figVG03full} to \ref{figVG12} show examples of the predicted and measured IR spectra.
 Figure \ref{figVG03_f100} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
@@
 A key result of this work is to show that gaps of known dimensions are recovered in our spectroscopy.
-The cases in which predicted spectroscopy differs from measurement are due to either a bona-fide detection of a gap above and beyond the implanted gap, or some measurement artifacts.
-In the next section we address mitigation of artifacts.
+Table \verb|\ref{tbl_DerivedGapSizes}| summarizes the outcome of some measurements in the areas with known gaps and fill factors.
+The table lists the median sample and 68\% confidence intervals.
 
-\subsection{Mitigation of measurement artifacts}
+\subsection{TODO: Comparison of Transmission-spectroscopy derived gap sizes with lab-measured gap size}
+\subsection{TODO: Spectra of direct bonded Si with no intentionally implanted gaps}
+\section{TODO: Discussion}
+
+\subsection{TODO: Mitigation of measurement artifacts}
 We mitigated measurement artifacts in the following ways.
@@
 Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.
-
-\subsection{Spectra of direct bonded Si with no intentionally implanted gap}
-We have shown that the IR spectroscopy technique combined with the Fabry-P\'{e}rot framework provide accurate knowledge of the gap size.
-In the cases where we expected no gap (i.e.
-perfect Si-Si bond), we have seen spectra consistent with 100\% transmission relative to a DSP Si wafer.
-In other cases we have seen evidence for gaps, which is simply an indication that the bonding process introduces gaps.
 

### 2015-01-23 AO_bondedSi.tex
First few pages of DTJ edits from Jan 23, 2015

@@
 \cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
-Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \cite{Horn2009, Masteika2014}.
+Other authors investigated the growth of bubbles at bond interfaces with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \cite{Horn2009, Masteika2014}.
 
 The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
-A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+These results ultimately helped to flesh out a robust picture of the chemistry as a function of anneal temperature\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
 Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
@@
 
-If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than those of two antireflection coated surfaces over the operating wavelength (typically $< 2$\%).
+If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than the concatenated losses of two antireflection coated surfaces over the operating wavelength (typically $< 4$\%).
+For Si-air gaps, where the full Fresnel loss is $\sim30\%$ per interface this requirement translates to physical gaps or bubbles having an axial extent of $ < 35\;$nm.
 If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond.
@@
 In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.
-When bonding two thicker and more rigid substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
+In optics, however, both of the bonding partners will likely be thicker and more mechanically rigid.
+When bonding two thick substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
 
@@
 Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.
-Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect and the transmission is 100\%.
+Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect.
 In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap.
-X-ray topography \cite{Mitani1990} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself.
-In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \cite{2000RScI...71.1869G}.
+X-ray topography \cite{Mitani1990} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not of the axial extent itself.
+In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but also produces only limited information about the axial dimension \cite{2000RScI...71.1869G}.
 
@@
 We describe here a new technique for measuring the axial extent of small gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
-Stock near-infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
-We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few nanometers in size.
+A recent generation of stock near-infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
+We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few tens of nanometers in size.
 In order to verify the applicability and accuracy of the technique, we have used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
-Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that was not widely available until this decade.
 

### 2015-01-23 AO_bondedSi.tex
more DTJ edits including a typo in Title of Figure 1

@@
 For all types of gaps and bubbles, our technique needs to provide an accurate assessment of the transmission loss at the operating wavelength.
-For finite-sized bubbles, measurement of the axial dimension helps us to assess the effects of high-temperature annealing on removal of hydrogen bubbles.
+For finite-sized bubbles, measurement of the axial dimension helps us to assess the effects of high-temperature annealing \cite{Horn2009, Masteika2014} on removal of hydrogen bubbles.
 A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to resolve spatially, as one might expect if there are many very small contaminants present along the interface.
-An important figure of merit for a bond is whether any loss due to gaps is smaller than the loss one might expect from two passages through a reasonably effective broad-band anti-relfection coating in the infrared.
-To meet this diagnostic goal, our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
+Our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
 A 2\% loss at 1.15 $\mu$m corresponds to a gap of XX nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
@@
 The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
-The Fresnel reflection is about 30\% per surface.
-For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.
-Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.
+The Fresnel reflection is about 30\% per surface, much larger than the $4-7\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
+Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature,  computed using the coefficients from \cite{2006SPIE.6273E..77F}.
 

### 2015-01-23 AO_bondedSi.tex
more DTJ edits, Section 3, before addendum A-in-a-circle

@@
 
-The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+We can treat a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
 The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
@@
 
-The reflectivity $R$ (Equation \ref{Eq:FresnelR}) is about 30\% in the wavelength range 1100 to 5000 nm.
 We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
@@
 Si absorbs negligibly longward of $\lambda$ = 1250 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses (Figure \ref{figSiAbsorbfig}).
-We computed the ratio of transmissions of double side polished Si wafers of different thicknesses.
-Specifically, we evaluated the transmission ratio at each wavelength bin from 1000 to 2500 nm, from the transmission of a thin ($\sim0.5$ mm) Si wafer divided by that of a 3 mm thick Si puck.
-Absorption is detected shortward of 1250 nm.
+We measured the ratio of transmissions of a 0.5 mm thick double side polished Si wafer and a 3.0 mm Si puck.
 Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness.
@@
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
-\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
+\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as a reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
 An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
@@
 \label{secTheory}
-A pair of bonded silicon wafers or pucks with a small gap actually forms their coupled cavities with the air gap as the the central cavity.
+A pair of bonded silicon wafers or pucks with a small gap actually forms three coupled cavities with the air gap as the the central cavity.
 The outer cavities, those within the Si, however, were not produced with coherence in mind.
-Their sides are not necessarily flat and parallel to the required level and the specifics of their deviation from perfect etalons are usually not known to the level necessary to treat them as oherent structures in the analysis.
-The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps where the gap size si small enough that the cavity it forms can always be treated as a coherent structure.
+Their sides are not necessarily flat and parallel to the required level and the specifics of their deviation from perfect etalons are usually not known to the level necessary to treat them as coherent structures in the analysis.
+The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps whereas the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
 Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
 The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K} to deal with the cavities within the Si.
-Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
-The details are worked out in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios- first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.
+We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K} to deal with the cavities within the thick Si substrates.
+Specifically we constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
+The details appear in Appendix \ref{sec:Append-IMRTMM}.
+We computed the transmission for two scenarios: first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.
 The DSP Si sample with no air gap has a transmission equal to:
@@
 Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range 1200$-2500$ nm.
+Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
 If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
-This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
+We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.
 
@@
 
-Now consider a pair of bonded Si samples with a gap thickness $d$.
+Consider a pair of bonded Si samples with a gap thickness $d$.
 We derive the absolute transmission of bonded Si substrates with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
 We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.
-We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.
+We call this normalized transmission the etalon transmission $T_{e}$.
 
@@
 
-It is clear that in the limit $\lim_{d \rightarrow 0}$, the etalon approaches 100\% transmission.
+In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP}} $ and the etalon approaches 100\% transmission.
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
-In Section \ref{secResults} we compare wavelength dependent transmission measurements to by-eye fits of Equation \ref{eqFP}.
-Depending upon the intended operating wavelength, gaps as small as 35-60 nm can cause losses comparable to or greater than the losses produced by antireflection coatings.
 

### 2015-01-23 AO_bondedSi.tex
add circle A

@@
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
+The bottom two panels show the ``gap'' transmission, the value of the throughput with the losses from the outer Si cavities divided out.
+For the very largest gap, large fringes appear as the cavity resonance of the air gap modulates the transmission.
+Below a $\lambda/2$ gap size the effect of the gap manifests itself as a wavelength-dependent diminution of the transmission.
+For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from 2400 nm down to 1250 nm.
+It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.
 

### 2015-01-23 AO_bondedSi.tex

@@
 Below a $\lambda/2$ gap size the effect of the gap manifests itself as a wavelength-dependent diminution of the transmission.
-For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from 2400 nm down to 1250 nm.
+For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.
 It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.
 
-\section{Laboratory demonstration: Imbedding gaps of known sizes}
+\section{Laboratory demonstration: Embedding gaps of known sizes}
 
-We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.
-
-The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.
-We classify substrates with thicknesses larger than 1 mm as ``pucks''.
-Substrates with thickness less than 1.0 mm thick are called ``wafers''.
-The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
+We evaluated our metrology method by embedding gaps of known sizes in the interface and measuring the effects of these gaps on IR transmission.
+The substrate thicknesses ranged from 0.5 to 3.3 mm.
+We denote substrates with thicknesses larger than 1 mm as ``pucks'' and substrates with thickness less than 1.0 mm thick as ``wafers''.
+We distinguish the two classes of substrate thickness based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
 There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
 Our ultimate goal is to bond optics with thicknesses as large as 30 mm.
-These pucks truly resemble over-sized hockey pucks.
 
-The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.
-We bored holes in silicon with inductively coupled plasma etching.
-We used two different gases with low and high etch rates.
-CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
-The area exposed to the process gases was either a 25 mm diameter circle, or a 15 mm $\times$ 20 mm rectangle.
-We protected the unexposed areas with thick photoresist.
-We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
-Table \ref{tabbondexper} has more information about the bonding samples, and Table \ref{tabPatternFills} lists the properties of the patterns.
+The three important characteristics of gaps are their axial extent, their lateral areal extent, and their areal fill factor.
+To fabricate our test samples, we used photoresist lithography to pattern Si substrates, and we bored holes with inductively coupled plasma etching.
+We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
+Tables \ref{tabbondexper} and \ref{tblPatternFills} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
 
@@
 The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
-The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
+The stylus profilometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
 Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.
@@
 \label{figVS20pattern}
-Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section.
-The 100 mm diameter VS20 is comprised of two thick silicon wafers.
+Veeco optical profiler surface plot and histogram for part VS20-21 in the fine pattern section.
+The 100 mm diameter VS20-21 is comprised of two thick silicon wafers.
 We patterned three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
@@
 In both single and double mode, a baseline scan is taken with no sample in the measurement holder.
-The baseline spectrum is the raw spectral response of the system.
 We automatically divide sample spectra by the baseline spectrum.
@@
 
-The spectral sampling was 2.0 nm.
-We used a spectral resolution of 5.0 nm, the lowest spectral resolution the Cary5000 offers.
+The spectral sampling was ranged from 2.0 to 10.0 nm.
+We used a spectral resolution of 5.0 nm.
 We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
 Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
-The axial extends of such large gaps would be easily measurable by other means, like IR imaging.
+The axial extents of such large gaps would be easily measurable by other means, like IR imaging.
 
@@
 
-We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
+We looked for large gaps, detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.
-Figure \ref{figVS2021_IR_image} shows an IR image of sample VS20-21.
-For this image we coarsely flat fielded the detector by observing a homogeneously illuminated white screen.
-Despite the flat fielding, detector non-uniformities and vignetting are perceptible in our image as vertical stripes.
+Figure \ref{figVS2021_IR_image} shows an IR image of bonded sample VS20-21.
+For this image, we coarsely flat-fielded the detector by observing a homogeneously illuminated white screen.
+Despite the flat-fielding, detector non-uniformities and vignetting are perceptible in our image as vertical stripes.
 Mosaic-stitching errors are also perceptible.
 It was easy to detect IR bubbles despite detector artifacts since the bubbles translate with the sample when we dither its position.
-We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.
-The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
+
+We also used IR imaging to inspect VG09-12 and VG15-13.
+We detected IR bubbles in all 3 of these bonded Si samples.
+The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
 The individual IR bubble diameters were as small as a few square millimeters up to 400 mm$^2$.
-The morphologies of the IR bubbles ranged from perfectly circular to irregular.
+The IR bubbles had varied morphologies.
 The structure within the IR bubbles was sometimes indicative of Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
@@
 The bulk morphologies were clear regardless of the detector response.
-When IR imaging was available, it was leveraged to target IR specectrscopy in locations demonstrating presence or absense of IR bubbles, as desired.
+When IR imaging was available, it was leveraged to target IR specectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.
 In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.
@@
 The coarsest mesh, C, shows substructure, while the medium and fine meshes are indistinguishable at the low resolution of the image.
-The three dark spots are gaps at the inferface of the direct bonded Si samples.}
+The three dark spots are gaps at the interface of the direct bonded Si samples.}
 \end{figure}
@@
 \label{secResults}
-In this section we describe the spectroscopy of directly bonded Si.
 
@@
 
-\subsection{Inferring axial extent of gaps from noisy transmission spectra}
-We infer the axial extent of gaps from their noisy transmission spectra.
+The gaps in our test pairs should be a combination of the deliberately induced gaps and any inadvertent additional spaces caused by imperfections in the bonds.
+We infer the axial extent of gaps in our samples with etched holes from their measured transmission spectra.
 We use a Bayesian inference technique.
@@
               \includegraphics[width=\textwidth]{figs/VG12_corner.pdf}
-              \caption{Fitted gap size $d$ and fill factor $f$ for VG12}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG09-12}
 		\label{figVG12_corner}
@@
                 \includegraphics[width=\textwidth]{figs/VG12_f50.pdf}
-                \caption{Measured spectrum and fit of VG12 position 2}
+                \caption{Measured spectrum and fit of VG09-12 position 2}
                 \label{figVG12_f50}
         \end{subfigure}
-\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
+\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
 The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.

### 2015-01-26 AO_bondedSi.tex
add circle-C from DTJ edits, required making new figures

@@
 \begin{equation}
-	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -0.5\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -0.5\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
+	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
 \end{equation}
@@
 \label{secKnownGaps}
+
+Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap.
+The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.
+Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$ nm deep, 2 mm wide trench crosses the field field scanned by the Cary 5000.
+Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6\% \pm 0.1\;\%$, both consistent with the part geometry.
+These two results show that the formalism performs as expected on resonant gaps several wavelengths deep.
+
+In our test samples, the nominal gap we have induced in the sample represents a lower limit to the actual gap since poor adhesion of the non-continuous patterns may result in gaps larger than intended.
+In this case, we modify our mixture model \ref{eqnMix} to fix $f=0.5$ but allow two different depths:
+
+\begin{equation}
+	T_{mix} =  0.5 \times T_{e}(d_1) + 0.5 \times T_{e}(d_2) \label{eqnMix2Gaps}
+\end{equation}
+
+where the subscripts on $d$ distinguish the two depths.
+
+Figure \ref{figVG12} shows the results for a bonded pair with a 50\% fill factor mesh where the grooves in the mesh had a measured depth of $49\pm6\;$nm.
+The green band shows the predicted range for the transmission curve.
+The measurements lie below this band.
+Figure \ref{figVG12_corner} demonstrates that $d_1$ and $d_2$ are degenerate, but consistent with a $49 \pm 6\;$nm step function in the gap size distribution.
+
+The presence of a number of bubbles both on and off the meshed areas of this and other samples (Figure \ref{figVS2021_IR_image}) lend credence to the idea that an additional gap of this magnitude was present at the position where we made the measurements but was not detected in the preparatory imaging.
+
+\input{tbl_DerivedGapSizes}
+
 Figures \ref{figVG03full} to \ref{figVG12} show examples of the predicted and measured IR spectra.

### 2015-01-26 AO_bondedSi.tex
change figure VG12 to show two gaps instead of one.

@@
               \includegraphics[width=\textwidth]{figs/VG03_corner.pdf}
-              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 position 1}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 hole region}
 		\label{figVG03_corner}
@@
                 \includegraphics[width=\textwidth]{figs/VG03_f100.pdf}
-                \caption{Measured spectrum and fit of VG03 position 1}
+                \caption{Measured spectrum and fit of VG03 hole region}
                 \label{figVG03_f100}
@@
               \includegraphics[width=\textwidth]{figs/VG03p2_corner.pdf}
-              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 position 2}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG03 petal region}
 		\label{figVG03p2_corner}
@@
                 \includegraphics[width=\textwidth]{figs/VG03_f045.pdf}
-                \caption{Measured spectrum and fit of VG03 position 2}
+                \caption{Measured spectrum and fit of VG03 petal region}
                 \label{figVG03_f045}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-              \includegraphics[width=\textwidth]{figs/VG12_corner.pdf}
-              \caption{Fitted gap size $d$ and fill factor $f$ for VG09-12}
+              \includegraphics[width=\textwidth]{figs/VG12_twoGaps_corner.pdf}
+              \caption{Fitted gap sizes $d_1$ and $d_2$ for VG09-12}
 		\label{figVG12_corner}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG12_f50.pdf}
-                \caption{Measured spectrum and fit of VG09-12 position 2}
+                \includegraphics[width=\textwidth]{figs/VG12_twoGaps_f50.pdf}
+                \caption{Measured spectrum and fit of VG09-12}
                 \label{figVG12_f50}
@@
 A key result of this work is to show that gaps of known dimensions are recovered in our spectroscopy.
-Table \verb|\ref{tbl_DerivedGapSizes}| summarizes the outcome of some measurements in the areas with known gaps and fill factors.
+Table \ref{tbl_DerivedGapSizes} summarizes the outcome of some measurements in the areas with known gaps and fill factors.
 The table lists the median sample and 68\% confidence intervals.

### 2015-01-26 AO_bondedSi.tex
fix some bugs, and overhaul the conclusion

@@
 
-In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP}} $ and the etalon approaches 100\% transmission.
+In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP} $ and the etalon approaches 100\% transmission.
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
@@
 In our test samples, the nominal gap we have induced in the sample represents a lower limit to the actual gap since poor adhesion of the non-continuous patterns may result in gaps larger than intended.
-In this case, we modify our mixture model \ref{eqnMix} to fix $f=0.5$ but allow two different depths:
+In this case, we modify our mixture model (Equation \ref{eqnMix}) to fix $f=0.5$ but allow two different gap depths:
 
@@
 
-\input{tbl_DerivedGapSizes}
+\input{tbls/tbl_DerivedGapSizes}
 
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG12_twoGaps_f50.pdf}
+                \includegraphics[width=\textwidth]{figs/VG12_twoGaps.pdf}
                 \caption{Measured spectrum and fit of VG09-12}
@@
 \section{Conclusions}
-We have shown a method for indirectly measuring the size of gaps in bonded Si wafers.
-The technique broadly applies to any compound Si optics measured at normal incidence.
-We showed lab IR spectra with consistent predictions, and addressed ways to mitigate measurement artifacts.
+We have described a measurement technique and laid out an analysis formalism for the detection and measurement of sub-wavelength gaps at Si-Si interfaces.
+Experimental measurements show that we can precisely recover the extent of wavelength-scale gaps of known size, account for the presence of small gaps, and capture results consistent with no gap when measuring a monolithic part.
+
+The technique and formalism we describe here can detect and measure the axial extent of gaps in Si-Si optical bonds down to a level of $20-30\;$nm.
+It makes use of widely available off-the-shelf metrology equipment and simple to implement analysis.
+Armed with this technique, optical fabricators have a new diagnostic tool to assess the quality of Si-Si bonds.
 

### 2015-01-26 AO_bondedSi.tex
remove the experiment log, minor final tweaks from DTJ

@@
 \centerline{\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}}
-\caption{\label{figAbsoluteTrans}The family of curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
+\caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
 \emph{Top panel}: Absolute transmission through a Si substrate with no gap (top denim-blue solid curve), and small gaps of axial extent 20 nm (green dashed curve), 35 nm (red dash-dotted curve), and 60 nm (purple dotted curve).
 The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.
-This curve is only accurate for normal incidence transmission.
-Furthermore, the curve assumes the multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.
+The curves assume that multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.
 \emph{Middle panel}: The ``Gap Transmission'', defined as the Absolute transmission through a sample normalized by the absolute transmission through a gapless sample.
@@
 \label{secResults}
-
-\input{tbls/tbl_experimentLog.tex}
 

### 2015-01-26 AO_bondedSi.tex
edit the draft, update notebooks, cleanup tables, references

@@
 \section{Measurement technique and correlated measurement errors}
+\label{sec_aboutErrors}
 We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
@@
 
-Equipped with an observed spectrum and our mixture model, we constructed the Likelihood function \verb|\cite{SomeStatisticsBook}|, $\mathcal{L} = p(T_{obs} | d, f)$.
-The measurement errors are not assumed to be independent and identically distributed, and in fact are strongly correlated as discussed in Section \verb|\ref{sectionAboutErrors}|.
-To handle the correlated measurement errors we use Gaussian Processes \verb|\cite{RasmussenAndWilliams, DFM}|.
+Equipped with an observed spectrum and our mixture model, we constructed the Likelihood function \cite{2013sdmm.book.....I}.
+The measurement errors are not assumed to be independent and identically distributed, and in fact are strongly correlated as discussed in Section \ref{sec_aboutErrors}.
+To handle the correlated measurement errors we use Gaussian Processes \cite{rasmussen2006gaussian, DFMgp}.
 The Gaussian Process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.
@@
 where $\delta_{ij}$ is the Kronecker delta, and $\sigma_i$ are the independent measurement uncertainties on the $i^{\mathrm{th}}$ data point.
-For simplicity, we adopt values from Table \ref{tabCary5000pars} for single and double beam modes respectively.
-This assumption of homoscedasticity is fairly safe, since the correlated errors dominate in comparison to individual measurement errors.
+We experimented with different values for $\sigma_i$.
+When repeated DSP reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.
+In other cases we simply adopted mean values from Table \ref{tabCary5000pars} for single and double beam modes respectively.
+The latter assumption of homoscedasticity is fairly safe, since the correlated errors dominate in comparison to individual measurement errors.
 The parameter $l$ controls the correlation length, and $a$ controls the amplitude of the noise.
@@
 The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.
-The value of $l$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (thousands of nm) to capture global features attributable to lamp drift.
+The value of $l$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.
 We allow both $a$ and $s$ to be free parameters in our fitting procedure.
 
-We set our prior probability distribution functions \verb|\cite{someStatisticsBook}|, $\ln{p(d,f,a,s)}$ as uniform over a wide range of values for $\ln{a}$, $\ln{s}$, $d$, and $f$ consistent with visual inspection of each spectrum.
+We set our prior probability distribution functions \cite{2013sdmm.book.....I}, $\ln{p(d,f,a,s)}$ as uniform over a wide range of values for $\ln{a}$, $\ln{s}$, $d$, and $f$ consistent with visual inspection of each spectrum.
 The final un-normalized posterior probability function, expressed as a natural logarithm is:
@@
 \begin{equation}
-	T_{mix} =  0.5 \times T_{e}(d_1) + 0.5 \times T_{e}(d_2) \label{eqnMix2Gaps}
+	T_{mix} =  0.5 \times T_{e}(d_M + \epsilon) + 0.5 \times T_{e}(\epsilon) \label{eqnMix2Gaps}
 \end{equation}
 
-where the subscripts on $d$ distinguish the two depths.
+where the $d_M$ is the lab-measured mesh depth and $\epsilon$ is an undesired gap in excess of the mesh gap.
+For $\epsilon =0$, Equation \ref{eqnMix2Gaps} reduces to Equation \ref{eqnMix} with $f=0.5$.
+We capture the uncertainty in the lab-measured mesh depth by setting the prior on $d_M$ to a Normal distribution with standard deviation set by the values listed in Table \ref{tbl_experiments}.
 
 Figure \ref{figVG12} shows the results for a bonded pair with a 50\% fill factor mesh where the grooves in the mesh had a measured depth of $49\pm6\;$nm.
-The green band shows the predicted range for the transmission curve.
+The apple green band shows the predicted range for the transmission curve.
 The measurements lie below this band.
-Figure \ref{figVG12_corner} demonstrates that $d_1$ and $d_2$ are degenerate, but consistent with a $49 \pm 6\;$nm step function in the gap size distribution.
-
+Figure \ref{figVG12_corner} demonstrates that $d_M$ and $\epsilon$ are degenerate, but consistent with a small $\sim 12\;$nm gap in excess of the mesh.
 The presence of a number of bubbles both on and off the meshed areas of this and other samples (Figure \ref{figVS2021_IR_image}) lend credence to the idea that an additional gap of this magnitude was present at the position where we made the measurements but was not detected in the preparatory imaging.
@@
 \subsection{TODO: Comparison of Transmission-spectroscopy derived gap sizes with lab-measured gap size}
+
 \subsection{TODO: Spectra of direct bonded Si with no intentionally implanted gaps}

### 2015-01-26 AO_bondedSi.tex
tweak table formating, change VG12 figures,

@@
 We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
-Tables \ref{tabbondexper} and \ref{tblPatternFills} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
+Tables \ref{tbl_experiments} and \ref{tblPatternFills} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
 
@@
 The stylus profilometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
-Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.
+Table \ref{tbl_experiments} lists part numbers and descriptions for our bonded substrates.
 
@@
 We produce posterior samples with Markov Chain Monte Carlo (MCMC).
-Specifically, we implemented \texttt{emcee}\footnote{https://github.com/dfm/emcee}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.
+Specifically, we implemented \texttt{emcee}\footnote{\url{https://github.com/dfm/emcee}}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.
 We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the observed normalized Transmission spectra.
@@
               \includegraphics[width=\textwidth]{figs/VG12_twoGaps_corner.pdf}
-              \caption{Fitted gap sizes $d_1$ and $d_2$ for VG09-12}
+              \caption{Fitted gap sizes $d_M$ and $\epsilon$ for VG09-12}
 		\label{figVG12_corner}
@@
         \end{subfigure}
-\caption{ Same as Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
+\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
 The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.
@@
 
-The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.
+The wave transfer matrix method is described in detail in Chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.
 Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:

### 2015-01-27 AO_bondedSi.tex
change format of numbers, units, and spacings

@@
 \usepackage{subcaption}
+\usepackage{hyperref}
 %\usepackage{longtable}
@@
 The bonding process presents challenges in meeting the requirements for optical performance because air gaps at the Si interface cause large Fresnel reflections.
-Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 5\% at $\lambda = 1.15 \; \mu$m at normal incidence.
+Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 4\% at $\lambda = 1.25 \; \mu$m at normal incidence.
 Understanding the optical effects of such gaps and evaluation of various methods for preventing or eliminating them demands that we have a method for measuring not only the two-dimensional extent of the gaps but also their axial extent.
@@
 As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
-One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
+One advantage of this direct Si$-$Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
 A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.
@@
 Stengl et al.
-(1988) and Goesele et al.
+(1988) and G{\"o}esele et al.
 (1995) \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
@@
 \cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
-Other authors investigated the growth of bubbles at bond interfaces with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \cite{Horn2009, Masteika2014}.
+Other authors investigated the growth of bubbles at bond interfaces with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of $800-1100\;^\circ$C \cite{Horn2009, Masteika2014}.
 
@@
 
-If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than the concatenated losses of two antireflection coated surfaces over the operating wavelength (typically $< 4$\%).
+If Si$-$Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than the concatenated losses of two antireflection coated surfaces over the operating wavelength (typically $< 4$\%).
 For Si-air gaps, where the full Fresnel loss is $\sim30\%$ per interface this requirement translates to physical gaps or bubbles having an axial extent of $ < 35\;$nm.
@@
 On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation.
-Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance.
-Figure 5 of Cooper 1986\cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
+Even in the best of cleanrooms, particles at the optically significant $10-20\;$nm scale can be present in relatively high abundance.
+Figure 5 of Cooper 1986 \cite{doi:10.1080/02786828608959094} shows the surface flux (particles cm$^2$ s$^{-1}$) vs.
 particle size ($\mu$m).
@@
 We describe here a new technique for measuring the axial extent of small gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
-A recent generation of stock near-infrared spectrophotometers like the Cary5000 from Agilent offer 0.2\% precision and good repeatability.
+A recent generation of stock near-infrared spectrophotometers like the Cary5000 from Agilent offer $0.2\;\%$ precision and good repeatability.
 We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few tens of nanometers in size.
@@
 Our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
-A 2\% loss at 1.15 $\mu$m corresponds to a gap of XX nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
-We therefore need to be able to measure transmission as a function of wavelength to a fraction of a percent, including all systematic errors, for devices operating near 1 $\mu$m and to a somewhat looser standard for devices operating at longer wavelengths.
+A $2\;\%$ loss at $1.25\;\mu$m corresponds to a gap of $>25\;$nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
+We therefore need to be able to measure transmission as a function of wavelength to a fraction of a percent, including all systematic errors, for devices operating near $1\;\mu$m and to a somewhat looser standard for devices operating at longer wavelengths.
 
-\section{Conceivable pitfalls in Si-Si direct bonding}
+\section{Conceivable pitfalls in Si$-$Si direct bonding}
 
@@
 Stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-Unconventional thick ($\sim$ 30 mm) Si pucks are less conformal than wafers are, so we expect thick substrates could have gaps even in the absence of particles.
-Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 3 to 4 inch diameter clear aperture.
+Unconventional thick ($\sim30\;$mm) Si pucks are less conformal than wafers are, so we expect thick substrates could have gaps even in the absence of particles.
+Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 75 to 100 mm diameter clear aperture.
 We assume that, if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
-Subsequent annealing will strength the bond.
+Subsequent annealing will strengthen the bond.
 
@@
 Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
-The Fresnel reflection is about 30\% per surface, much larger than the $4-7\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
+The refractive index of silicon at room temperature ranges from $3.55-3.45$ from $\lambda = 1150-2500\;$nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+The Fresnel reflection is about $30\;\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
 Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature,  computed using the coefficients from \cite{2006SPIE.6273E..77F}.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of CITE\texttt{2006SPIE.6273E..77F}.
+\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}
+\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.
 The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
@@
 The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
-For the narrow air gaps we consider here and for the modest reflectivity of a Si-vacuum interface, most gaps at Si-Si bonds will form effective etalons over most of their area.
+For the narrow air gaps we consider here and for the modest reflectivity of a Si-vacuum interface, most gaps at Si$-$Si bonds will form effective etalons over most of their area.
 The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
@@
 Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness.
-Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level longward of 1250 nm.
+Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level longward of $1250\;$nm.
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.40\columnwidth]{figs/fpAbsorbfig_alt}}
+\includegraphics[width=0.40\columnwidth]{figs/fpAbsorbfig_alt}
 \caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}}
+\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
 \caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
@@
 We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
-Tables \ref{tbl_experiments} and \ref{tblPatternFills} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
+Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
 
 The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
-In a second round of experiments, we achieved gap sizes of 14$-$95 nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+In a second round of experiments, we achieved gap sizes of $14-95\;$nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
 No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}}
+\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}
 \caption{
@@
 We patterned three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
-The depth was 95 $\pm$ 5 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
+The depth was $95 \pm 5\;$nm as indicated by the well-separated peaks in the Veeco profilometry histogram.
 The fill factor for this and all mesh patterns was 50\%.
@@
 We measured the large scale surface flatness of the pucks.
-The Fisba2 interferometer used for these measurements has a 50 mm diameter beam at $\lambda=$632.8 nm.
+The Fisba2 interferometer used for these measurements has a 50 mm diameter beam at $\lambda=632.8\;$nm.
 We found a typical peak to valley surface flatness of 2 waves over the central 50 mm diameter.
@@
 We used a 50 mm F/2 lens, with a fiber optic white light illumination source.
-The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately 1.15$-$1.6 $\mu$m.
+The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately $1.15-1.6\;\mu$m.
 The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.
@@
 The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
-The individual IR bubble diameters were as small as a few square millimeters up to 400 mm$^2$.
+The individual IR bubble diameters were as small as a few square millimeters up to $400\;mm^2$.
 The IR bubbles had varied morphologies.
@@
 \begin{figure}[htbp]
-\centerline{\includegraphics[width=0.4\columnwidth]{figs/VS2021_IR_image.pdf}}
+\includegraphics[width=0.4\columnwidth]{figs/VS2021_IR_image.pdf}
 \caption{
@@
 \begin{equation}
-	T_{mix} = f T_{e} + (1-f) \label{eqnMix}
+	T_{mix} = f\;T_{e} + (1-f) \label{eqnMix}
 \end{equation}
@@
 The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.
-Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$ nm deep, 2 mm wide trench crosses the field field scanned by the Cary 5000.
-Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6\% \pm 0.1\;\%$, both consistent with the part geometry.
+Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field field scanned by the Cary 5000.
+Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6 \pm 0.1\;\%$, both consistent with the part geometry.
 These two results show that the formalism performs as expected on resonant gaps several wavelengths deep.
@@
 \begin{equation}
-	T_{mix} =  0.5 \times T_{e}(d_M + \epsilon) + 0.5 \times T_{e}(\epsilon) \label{eqnMix2Gaps}
+	T_{mix} =  0.5 \; T_{e}(d_M + \epsilon) + 0.5 \; T_{e}(\epsilon) \label{eqnMix2Gaps}
 \end{equation}
@@
 \section{Conclusions}
-We have described a measurement technique and laid out an analysis formalism for the detection and measurement of sub-wavelength gaps at Si-Si interfaces.
+We have described a measurement technique and laid out an analysis formalism for the detection and measurement of sub-wavelength gaps at Si$-$Si interfaces.
 Experimental measurements show that we can precisely recover the extent of wavelength-scale gaps of known size, account for the presence of small gaps, and capture results consistent with no gap when measuring a monolithic part.
 
-The technique and formalism we describe here can detect and measure the axial extent of gaps in Si-Si optical bonds down to a level of $20-30\;$nm.
+The technique and formalism we describe here can detect and measure the axial extent of gaps in Si$-$Si optical bonds down to a level of $20-30\;$nm.
 It makes use of widely available off-the-shelf metrology equipment and simple to implement analysis.
-Armed with this technique, optical fabricators have a new diagnostic tool to assess the quality of Si-Si bonds.
+Armed with this technique, optical fabricators have a new diagnostic tool to assess the quality of Si$-$Si bonds.
 

### 2015-01-28 AO_bondedSi.tex
loads of new text for the draft discussion section.  some review of older notebooks

@@
 
-\subsection{TODO: Comparison of Transmission-spectroscopy derived gap sizes with lab-measured gap size}
+\subsection{Measurement of a sample with no implanted gaps}
+We measured direct bonded Si$-$Si sample (part VG02, Table \ref{tbl_experiments}) with no intentionally embedded gaps.
+We measured transmission spectra in four randomly chosen regions of the $100\;$mm bonded wafers, with no preparatory imaging available.
+Two of the regions showed evidence of large ($>\lambda$) gaps.
+The other two regions showed almost 100\% transmission, when normalized by measurements of $T_{DSP}$.
+We apply our formalism to one of these regions to get an estimate for the smallest gap that we can constrain.
+Figure \verb|\ref{fig_xxx}| shows a zoom in of the measured normalized transmission spectrum, which shows strongly correlated noise, especially in the wavelength range $2000-2500\;$nm.
 
-\subsection{TODO: Spectra of direct bonded Si with no intentionally implanted gaps}
-\section{TODO: Discussion}
+The variance of individual measurement uncertainties was large enough to consider for this measurement.
+So rather than assume homoscedastic uncertainty as we did for the other measurements, we estimated the heteroscedastic uncertainties in the following way.
+We computed the mean-subtracted transmission for four DSP measurements bookending the those of VG02.
+The DSP measurements were detrended, and the measurement-to-measurement variance computed.
+The result was a variance spectrum which peaks at $1850\;$nm at 0.25\%, but with a median value equal to 0.03\%.
+Figure \verb|\ref{fig_xxx}| includes the $1\;\sigma$ errorbars computed in this way.
 
-\subsection{TODO: Mitigation of measurement artifacts}
-We mitigated measurement artifacts in the following ways.
-We have mostly been able to rule out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.
-Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.
+\begin{figure}[htbp]
+        \centering
+        \begin{subfigure}[b]{0.5\textwidth}
+              \includegraphics[width=\textwidth]{figs/VG02p3_corner.pdf}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG02}
+		\label{fig_VG02p3_corner}
+        \end{subfigure}
+
+        \begin{subfigure}[b]{0.5\textwidth}
+                \includegraphics[width=\textwidth]{figs/VG02_noGap.pdf}
+                \caption{Measured spectrum and fit of VG02}
+                \label{figVG02_noGap}
+        \end{subfigure}
+\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG02, which has no intentionally imbedded gap.
+The measured spectrum in this region shows no appreciable deficit of transmission from the gapless counterparts.
+\label{figVG12}}
+\end{figure}
+
+Figure \ref{fig_VG02p3_corner} shows the degeneracy between the axial extent of a gap and its fractional fill factor over the measurement beam.
+Simply put, a larger gap can be hidden if it doesn't cover much area.
+The red dashed line shows the normalized transmission spectrum of an 18 nm gap over 30\% of the measurement area.
+Gaps larger than 60 nm would have to cover less than 10\% of the measurement area in order to be consistent with the measurement.
+On this zoomed-in vertical scale, the individual draws from the Gaussian process model show morphologies consistent with the correlated measurement error (Figure \ref{figVG02_noGap}).
+
+\section{Discussion}
+
+\subsection{Questioning assumptions of the model}
+
+\emph{More complex underlying gap size distributions-} One key assumption of our model is that the distribution of gap sizes over our measurement area is bimodal.
+Specifically, Equation \ref{eqnMix} assumes that there is either a gap of size $d$, or else it is 100\% bonded.
+If there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
+The full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
+Solving for the relative distributions of gaps is highly degenerate, in the same way as we saw in the 2-parameter corner plots of gap depth and fill factor.
+We experimented with by-eye fitting parametric distributions of gap sizes, characterized by a maximum and minimum gap size, and a monotonic slope of the distribution.
+We found that measurements of very large ($>10\lambda$) gaps with a large gradient offered surprisingly strong constraints on the maximum gap size, because of the conspicuous high frequency Fabry-P\`erot fringes.
+For sub-wavelength gaps, the gap size distibution is too degenerate to derive much useful insight on the slope of the distribution.
+Specifically, for $d_{max}<\lambda/4$, our simple mixture model will converge on the spectrum of the average gap size, even if the \emph{True} distribution is more complex.
+Reducing the measurement area could break some of these degeneracies, if more precision is desired.
+
+\emph{Origin and properties of correlated noise- } Another key assumption of the model is that correlated errors dominate the measurement uncertainty.
+This is the same thing as saying that there is an instrumental bias, or that our measurements are \emph{precise} but \emph{inaccurate}.
+The extent to which the measurement is inaccurate is tabulated in \ref{tbl_cary5000Pars}, as already mentioned.
+We have assumed that the Gaussian Process model can capture the effect of the observed correlated errors.
+But to what extent does the Gaussian Process emulate the \emph{True} covariance matrix, which is unknown?
+One way to gauge the goodness of fit of the covariance matrix is to generate synthetic data from a best-fit model.
+Many of these synthetic draws are shown in the spectra in Figures \rev|\ref{a thru f}| as transparent blue lines.
+When overplotted these lines look like a halo of uncertainty enveloping the model and measurement.
+The bulk agreement is encouraging.
+We also scrutinized individual synthethic spectra with the measured data, and found that the noise model tended to under-predict the length scale of the correlation.
+For example, we typically derive $0.15<a<0.4\%$ and $20<s<200$.
+The measured data appear to prefer larger values of $s$.
+The discrepancy probably comes from the fact that, in reality, the \emph{True} covariance matrix probably has local structures that cannot be replicated by a simple 2-parameter (\emph{i.e.} $a$ and $s$) estimate of the covariance matrix.
+A model employing local covariance structures \verb|\cite{Czekala et al.
+in press}| could better capture some of the residual structure, but would introduce additional computational complexity with little gain.
+Overall the Gaussian Process technique provides much better agreement than the na\"{i}ve assumption of zero correlated errors.
+
+\emph{Validating the MCMC with synthetic data, and addressing heteroscedasticity-}  To evaluate our MCMC fitting method and assumption of homoscedasticity, we generated synthetic data with a known parameters, $d$, $f$, $a$, and $s$ and $\sigma_i$.
+We then fit the parameters $d$, $f$, $a$, and $s$ assuming homoscedastic errors, ($\sigma_i = \sigma = 0.02\%$).
+The derived physical parameters were consistent with the \emph{\Truth} to within the $1\;\sigma$ confidence interval, but we significantly underestimated $s$ (8 nm instead of 200 nm) and marginally underestimated $a$ (0.0013 instead of 0.0016).
+This result is reassuring for our estimation of physical parameters, and shows that the Gaussian Process technique captures realistic uncertainty in the physical parameters, despite an imperfect characterization of the noise correlation length.
+
+\emph{Polarization and lamp jumps- } As mentioned, the Cary 5000 double beam mode offers considerably lower amplitude of correlated errors than single beam mode \ref{tbl_cary5000Pars}.
+However, the double beam mode was found to have enhanced artifacts at wavelengths at which the Cary 5000 monochromator changes its diffraction grating and light sources.
+The sharp jumps in transmission were up to a few percent.
+The origin of these jumps could be polarization-dependent transmission losses.
+Whatever the origin, they cause large systematic errors.
+These are still captured by the Gaussian Process structure.
+
+\emph{Normal incidence- } We have assumed that the transmission measurement is performed at normal incidence.
+If the measurement beam forms a non-zero angle with the entrance face, bond interface, or exit face, several things happen to make this technique invalid.
+Notably, the polarization effects, Fresnel reflection losses, and the incoherent multiple reflections change the emergent transmission spectrum in non-trivial ways.
+
+\emph{DSP vs AR coated front faces- } Another key assumption of the model is that all the Si$-$air interfaces are well-approximated by Fresnel reflection losses.
+If the entrance and exit faces of the bonded samples are AR coated, the incoherent multiple reflection model will have to be modified.
+AR coated front and exit faces will result in a larger interface loss for a given gap size than the interface loss computed here.
+
+\emph{Refactive index- } We have assumed that the refractive index is identical to that given by \cite{2006SPIE.6273E..77F}.
+Double beam mode measurements can reach a precision of comparable to the part-to-part variation of refractive index.
+We have normalized spectra by their singleton counterparts from the same heritage (i.e.
+same wafer box).
+
+\emph{Diffraction from mesh patterns}
+It's conceivable that the mesh patterns Figure \ref{figVS20pattern}.
+We ruled out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.
+To eliminate this concern, we measured in the coarse mesh rather than the medium or fine meshes, when possible.
 We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
@@
 Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.
+
+\subsection{What is the smallest gap size this technique can measure?}
+In this section we quantify the limits of how far the technique can be pushed.
+There are three key ideas when trying to answer the question ``What is the smallest gap size this technique can measure?''.
+The answer for single beam mode is evidently $\sim20 nm$ (Figure \ref{fig_VG02p3_corner}).
+In principle this number could be pushed down by averaging many spectra, all normalized by different reference spectra.
+Alternatively, double beam mode could be used.
+The grating jumps noted above would also limit double beam mode analysis.
+If these could be rectified, a very low uncertainty would be obtained, and it is conceivable that gaps as small as a few nanometers could be inferred.
 

### 2015-01-28 AO_bondedSi.tex
bug fixes to table formatting and reference and labels

@@
 IR image of part VS20-21.
-Three rectangular mesh areas are apparent (Table \ref{tabPatternFills}).
+Three rectangular mesh areas are apparent (Table \ref{tbl_meshPatterns}).
 The coarsest mesh, C, shows substructure, while the medium and fine meshes are indistinguishable at the low resolution of the image.
@@
 The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
-Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tabPatternFills}.
+Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tab_meshPatterns}.
 These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
@@
 One way to gauge the goodness of fit of the covariance matrix is to generate synthetic data from a best-fit model.
-Many of these synthetic draws are shown in the spectra in Figures \rev|\ref{a thru f}| as transparent blue lines.
+Many of these synthetic draws are shown in the spectra in Figures \verb|\ref{a thru f}| as transparent blue lines.
 When overplotted these lines look like a halo of uncertainty enveloping the model and measurement.
@@
 We then fit the parameters $d$, $f$, $a$, and $s$ assuming homoscedastic errors, ($\sigma_i = \sigma = 0.02\%$).
-The derived physical parameters were consistent with the \emph{\Truth} to within the $1\;\sigma$ confidence interval, but we significantly underestimated $s$ (8 nm instead of 200 nm) and marginally underestimated $a$ (0.0013 instead of 0.0016).
+The derived physical parameters were consistent with the \emph{Truth} to within the $1\;\sigma$ confidence interval, but we significantly underestimated $s$ (8 nm instead of 200 nm) and marginally underestimated $a$ (0.0013 instead of 0.0016).
 This result is reassuring for our estimation of physical parameters, and shows that the Gaussian Process technique captures realistic uncertainty in the physical parameters, despite an imperfect characterization of the noise correlation length.

### 2015-02-10 AO_bondedSi.tex
add some text about the sample transport accessory, realization about coarse mesh comparable to beam size

@@
 
+\subsection{Spatially resolved measurements with Sample Transport Accessory}
+We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory.
+This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
+The 2 mm step size is comparable to the monochromator's slit size.
+We sampled 50 mm across the mesh pattern and into the off-mesh pattern.
+Figure \verb|\ref{photo-of-VG09-12}| shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with black dots.
+The colored dots match the color of the spectra in Figure \verb|\ref{fig-STA-obs}| The approximate size of the measurement beam is shown for scale.
+The conspicuous mesh pattern has a lab measured depth of $49\pm6\;$nm, and fill factor of 50\%.
+The prediction band for a transmision spectrum with those properties, and zero extra gap is shown as the green apple colored swath.
+If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \verb|\ref{fig-STA-obs}|.
+Instead we see that some spectra are inside the swath, and others show a deficit of flux, indicating an extra gap as in the mixture model in Equation \verb|\ref{mixture-model-equation}|.
+
 \begin{figure}[htbp]

### 2015-02-10 AO_bondedSi.tex
add text about sample transport

@@
 \subsection{Spatially resolved measurements with Sample Transport Accessory}
-We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory.
+We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory\verb|\ref{link-to-STA-website}|.
 This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
@@
 We sampled 50 mm across the mesh pattern and into the off-mesh pattern.
+The large number of consecutive measurements prevented the requisite near-contemporaneous baseline measurements for normalization.
+We fit a line to before-and-after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
+We reduced the measured wavelength range to $1250-1775\;$nm.
+
 Figure \verb|\ref{photo-of-VG09-12}| shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with black dots.

### 2015-02-16 AO_bondedSi.tex
add illustration and spectra from sample transport accessory measurements.

--- 2015-01-28 AO_bondedSi.tex
@@
 
+\subsection{Spatially resolved measurements with Sample Transport Accessory}
+We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory\verb|\ref{link-to-STA-website}|.
+This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
+The 2 mm step size is comparable to the monochromator's slit size.
+We sampled 50 mm across the mesh pattern and into the off-mesh pattern.
+The large number of consecutive measurements prevented the requisite near-contemporaneous baseline measurements for normalization.
+We fit a line to before-and-after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
+We reduced the measured wavelength range to $1250-1775\;$nm.
+
+%Figure \verb|\ref{photo-of-VG09-12}| shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
+The colored dots match the color of the spectra in Figure \verb|\ref{fig-STA-obs}| The approximate size of the measurement beam is shown for scale.
+The conspicuous mesh pattern has a lab measured depth of $49\pm6\;$nm, and fill factor of 50\%.
+The prediction band for a transmision spectrum with those properties, and zero extra gap is shown as the green apple colored swath.
+If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \verb|\ref{fig-STA-obs}|.
+Instead we see that some spectra are inside the swath, and others show a deficit of flux, indicating an extra gap as in the mixture model in Equation \verb|\ref{mixture-model-equation}|.
+
+Sample VG09-12 has a single coarse mesh.
+The mesh is composed of boxes plasma-etched to a depth of $49\pm6\;$nm.
+The boxes have an edge length of 1.5 mm.
+The edges of adjacent boxes are spaced 0.62 mm apart, to yield an overall fill factor of 50\%.
+The Cary5000 measurement beam size is $\sim$1 mm $\times$ 10 mm.
+Since the beam size is comparable to the mesh size, the fill factor at the position of the measurement will will differ from 50\%.
+The slit measurement was aligned with the mesh grid.
+The highest fill factor acheivable is about 75\%, assuming the slit fit perfectly within a column of boxes.
+The slit is larger than 0.62 mm, so the measurement cannot exhibit a 0\% fill factor of voids.
+The precise value of the minimum fill factor depends on the details of the slit width, which is not known precisely.
+We assume that a minimum fill factor of 30\%, but caution that this boundary is fuzzy.
+Figure \verb|\ref{VG09-12 coarse mesh STA}| shows
+
 \begin{figure}[htbp]
@@
         \begin{subfigure}[b]{0.5\textwidth}
-              \includegraphics[width=\textwidth]{figs/VG12_twoGaps_corner.pdf}
-              \caption{Fitted gap sizes $d_M$ and $\epsilon$ for VG09-12}
-		\label{figVG12_corner}
+              \includegraphics[width=\textwidth]{figs/VG09_12_STA_illustration.pdf}
+              \caption{Illustration of VG09-12 sample transport accessory measurement positions}
+		\label{figVG12_STA_corner}
         \end{subfigure}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG12_twoGaps.pdf}
-                \caption{Measured spectrum and fit of VG09-12}
+                \includegraphics[width=\textwidth]{figs/VG0912_STA_scan_coarse.pdf}
+                \caption{Measured gap transmission spectra of VG09-12 on- and off- mesh}
                 \label{figVG12_f50}
         \end{subfigure}
-\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12, which has an imbedded gap of axial extent $d=49\pm6\;$nm with fill factor $f=0.5$.
-The measured spectrum in this region shows a clear deficit of transmission from the gapless counterparts.
-The measured spectrum also shows a deficit of transmission compared to the predicted transmission of a 49 $\pm6\;$nm gap with 50\% fill factor (green apple shaded color band).
-The discord between prediction and measurement could be due to several factors including an underestimate of the intrinsic gap size, an underestimate of the intrinsic fill factor, or an additional gap in excess of the prediction.
-We favor the latter interpretation, since undesired gaps are common in these direct-bonded Si samples that have not undergone high-temperature thermal annealing.
+\caption{Predictions and measurements of the coarse mesh region in sample VG09-12.
 \label{figVG12}}

### 2015-02-16 AO_bondedSi.tex
SiGap_17 finalized, sample transport accessory text added, new pdfs made.
Analysis for zero gap, SiGaps_18 and 19.

@@
         \centering
-        \begin{subfigure}[b]{0.5\textwidth}
+        \begin{subfigure}[b]{0.35\textwidth}
               \includegraphics[width=\textwidth]{figs/VG09_12_STA_illustration.pdf}
@@
         \end{subfigure}
-
-        \begin{subfigure}[b]{0.5\textwidth}
+        \begin{subfigure}[b]{0.65\textwidth}
                 \includegraphics[width=\textwidth]{figs/VG0912_STA_scan_coarse.pdf}
@@
 
-\subsection{Measurement of a sample with no implanted gaps}
-We measured direct bonded Si$-$Si sample (part VG02, Table \ref{tbl_experiments}) with no intentionally embedded gaps.
-We measured transmission spectra in four randomly chosen regions of the $100\;$mm bonded wafers, with no preparatory imaging available.
-Two of the regions showed evidence of large ($>\lambda$) gaps.
-The other two regions showed almost 100\% transmission, when normalized by measurements of $T_{DSP}$.
-We apply our formalism to one of these regions to get an estimate for the smallest gap that we can constrain.
-Figure \verb|\ref{fig_xxx}| shows a zoom in of the measured normalized transmission spectrum, which shows strongly correlated noise, especially in the wavelength range $2000-2500\;$nm.
+\subsection{What is the smallest gap this technique can detect?}
+We devised two experiments to approach the question, ``What is the smallest gap this technique can detect?''.
+We applied our MCMC model fitting formalism to an unbonded DSP wafer.
+The measured transmission spectrum was normalized by a transmission spectrum of the same sample taken only a few minutes earlier.
+Any difference of this normalized transmission from unity is attributable entirely to measurement uncertainty, and represents the smallest possible signal one could extract.
+We also measured bonded sample VG09-12 in a region away from the intentionally implanted gap.
+This spectrum was processed in the same way as the spectra of the VG09-12 mesh area, exhibiting a mean value of 0.9998 and a standard deviation $\sigma=0.0008$.
+This spectrum is consistent with no gap.
+We applied our MCMC formalism to further constrain the range of gap sizes consistent with the spectrum.
+Table \verb|\ref{tbl summary of gap fit extents}| summarizes the sizes of gaps.
+The well-calibrated DSP noise spectrum rules out ($>3\sigma$) gap sizes larger than about 8.0 nm over $>80$\% of the measurement area.
+Similarly, the VG09-12 off-mesh spectrum rules out gaps greater than 13 nm over $>80$\% of the measurement area.
 

### 2015-02-16 AO_bondedSi.tex
minor edits to first 17 pages of the text

@@
 We describe a bond inspection method that makes use of precision transmission spectroscopy.
-Using spectroscopy, we can detect and measure gaps as small as 8 nm by measuring transmission as a function of wavelength.
+Using spectroscopy, we can detect and measure gaps as small as 12 nm by measuring transmission as a function of wavelength.
 Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
+We infer the gap axial extent and fill factor with Monte Carlo Markov Chain (MCMC) and a flexible Gaussian process noise model.
 We describe the measurement and analysis techniques and demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
@@
 
-\subsection{The Si$-$Si interface gap is a low Finesse Fabry-P\`{e}rot etalon}
+\subsection{The Si$-$Si interface gap is a low finesse Fabry-P\`{e}rot etalon}
 
@@
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
-If the Si wafer path thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+If the Si wafer thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
 We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.
 
-\begin{figure}[htbp]
+\begin{figure}[!htbp]
 \includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
@@
 We automatically divide sample spectra by the baseline spectrum.
-In single beam mode, the baseline measurements must be repeated about every half hour for post-process division.
+In single beam mode, the baseline measurements must be repeated about every 10 minutes for post-process division.
 In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.
@@
 
-The spectral sampling was ranged from 2.0 to 10.0 nm.
+The spectral sampling ranged from 2.0 to 10.0 nm.
 We used a spectral resolution of 5.0 nm.
@@
 The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
-The individual IR bubble diameters were as small as a few square millimeters up to $400\;mm^2$.
+The individual IR bubble areas were as small as a few square millimeters up to $400\;\mathrm{mm}^2$.
 The IR bubbles had varied morphologies.
@@
 The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
-Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tab_meshPatterns}.
+Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
 These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
 
-Equipped with an observed spectrum and our mixture model, we constructed the Likelihood function \cite{2013sdmm.book.....I}.
-The measurement errors are not assumed to be independent and identically distributed, and in fact are strongly correlated as discussed in Section \ref{sec_aboutErrors}.
-To handle the correlated measurement errors we use Gaussian Processes \cite{rasmussen2006gaussian, DFMgp}.
-The Gaussian Process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.
+Equipped with an observed spectrum and our mixture model, we constructed the likelihood function \cite{2013sdmm.book.....I}.
+The measurement errors are strongly correlated as discussed in Section \ref{sec_aboutErrors}.
+To handle the correlated measurement errors we use Gaussian process regression \cite{rasmussen2006gaussian, DFMgp}.
+The Gaussian process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.
 We parameterize the covariance matrix elements in the following way:
@@
 \begin{equation}
-	C_{ij} = \sigma^2_{i}\delta_{ij}+a^2\exp{(-\frac{(\lambda_i-\lambda_j)^2}{2l^2})} \label{eqnGPkernel}
+	C_{ij} = \sigma^2_{i}\delta_{ij}+a^2\exp{(-\frac{(\lambda_i-\lambda_j)^2}{2s^2})} \label{eqnGPkernel}
 \end{equation}
@@
 When repeated DSP reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.
-In other cases we simply adopted mean values from Table \ref{tabCary5000pars} for single and double beam modes respectively.
-The latter assumption of homoscedasticity is fairly safe, since the correlated errors dominate in comparison to individual measurement errors.
-The parameter $l$ controls the correlation length, and $a$ controls the amplitude of the noise.
-We do not know $a$ nor $l$ and they are, in general, different for each measurement.
+The parameter $s$ controls the correlation length, and $a$ controls the amplitude of the noise.
+We do not know $a$ nor $s$ and they are, in general, different for each measurement.
 The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.
-The value of $l$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.
+The value of $s$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.
 We allow both $a$ and $s$ to be free parameters in our fitting procedure.
@@
 \begin{equation}
-	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
+	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma_i)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
 \end{equation}
@@
 Specifically, we implemented \texttt{emcee}\footnote{\url{https://github.com/dfm/emcee}}\cite{emcee}, with 32 walkers, hundreds of burn-in samples, and 600-1000 iterations.
-We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the observed normalized Transmission spectra.
+We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the spectra.
 Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
@@
 The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.
-Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field field scanned by the Cary 5000.
+Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field scanned by the Cary 5000.
 Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6 \pm 0.1\;\%$, both consistent with the part geometry.

### 2015-02-16 AO_bondedSi.tex
delete major swath of extra gap mixture model and discussion of extra gap in VG09

@@
 
-In our test samples, the nominal gap we have induced in the sample represents a lower limit to the actual gap since poor adhesion of the non-continuous patterns may result in gaps larger than intended.
-In this case, we modify our mixture model (Equation \ref{eqnMix}) to fix $f=0.5$ but allow two different gap depths:
-
-\begin{equation}
-	T_{mix} =  0.5 \; T_{e}(d_M + \epsilon) + 0.5 \; T_{e}(\epsilon) \label{eqnMix2Gaps}
-\end{equation}
-
-where the $d_M$ is the lab-measured mesh depth and $\epsilon$ is an undesired gap in excess of the mesh gap.
-For $\epsilon =0$, Equation \ref{eqnMix2Gaps} reduces to Equation \ref{eqnMix} with $f=0.5$.
-We capture the uncertainty in the lab-measured mesh depth by setting the prior on $d_M$ to a Normal distribution with standard deviation set by the values listed in Table \ref{tbl_experiments}.
-
-Figure \ref{figVG12} shows the results for a bonded pair with a 50\% fill factor mesh where the grooves in the mesh had a measured depth of $49\pm6\;$nm.
-The apple green band shows the predicted range for the transmission curve.
-The measurements lie below this band.
-Figure \ref{figVG12_corner} demonstrates that $d_M$ and $\epsilon$ are degenerate, but consistent with a small $\sim 12\;$nm gap in excess of the mesh.
-The presence of a number of bubbles both on and off the meshed areas of this and other samples (Figure \ref{figVS2021_IR_image}) lend credence to the idea that an additional gap of this magnitude was present at the position where we made the measurements but was not detected in the preparatory imaging.
-
 \input{tbls/tbl_DerivedGapSizes}

### 2015-02-16 AO_bondedSi.tex
overhaul section on smallest gap this technique can detect

@@
 
-Figures \ref{figVG03full} to \ref{figVG12} show examples of the predicted and measured IR spectra.
-Figure \ref{figVG03_f100} shows an example spectrum in which the measurement is consistent with the prediction for a gap with the known conditions.
-Meanwhile, Figure \ref{figVG12_f50} shows a spectrum in which the measurement marginally exceeds the prediction (by about $\sim 3 \sigma$).
-Other parts show less consistency with the prediction due to the difficulty of direct-bonding Si in the absence of thermal annealing.
-
 \begin{figure}[htbp]
@@
 \subsection{Spatially resolved measurements with Sample Transport Accessory}
-We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory\verb|\ref{link-to-STA-website}|.
+We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory \footnote{\url{http://www.chem.agilent.com/en-US/products-services/instruments-systems/molecular-spectroscopy/sample-transport-accessory/}}.
 This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
-The 2 mm step size is comparable to the monochromator's slit size.
-We sampled 50 mm across the mesh pattern and into the off-mesh pattern.
+We chose a 2 mm step size, which is slightly larger than the $\sim1$ mm beam width.
+We sampled 50 mm across the mesh pattern and into the off-mesh pattern (Figure \ref{figVG0912_STA_illus}).
 The large number of consecutive measurements prevented the requisite near-contemporaneous baseline measurements for normalization.
@@
 
-%Figure \verb|\ref{photo-of-VG09-12}| shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
-The colored dots match the color of the spectra in Figure \verb|\ref{fig-STA-obs}| The approximate size of the measurement beam is shown for scale.
-The conspicuous mesh pattern has a lab measured depth of $49\pm6\;$nm, and fill factor of 50\%.
-The prediction band for a transmision spectrum with those properties, and zero extra gap is shown as the green apple colored swath.
-If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \verb|\ref{fig-STA-obs}|.
-Instead we see that some spectra are inside the swath, and others show a deficit of flux, indicating an extra gap as in the mixture model in Equation \verb|\ref{mixture-model-equation}|.
-
-Sample VG09-12 has a single coarse mesh.
-The mesh is composed of boxes plasma-etched to a depth of $49\pm6\;$nm.
-The boxes have an edge length of 1.5 mm.
-The edges of adjacent boxes are spaced 0.62 mm apart, to yield an overall fill factor of 50\%.
-The Cary5000 measurement beam size is $\sim$1 mm $\times$ 10 mm.
-Since the beam size is comparable to the mesh size, the fill factor at the position of the measurement will will differ from 50\%.
-The slit measurement was aligned with the mesh grid.
-The highest fill factor acheivable is about 75\%, assuming the slit fit perfectly within a column of boxes.
-The slit is larger than 0.62 mm, so the measurement cannot exhibit a 0\% fill factor of voids.
-The precise value of the minimum fill factor depends on the details of the slit width, which is not known precisely.
-We assume that a minimum fill factor of 30\%, but caution that this boundary is fuzzy.
-Figure \verb|\ref{VG09-12 coarse mesh STA}| shows
+Figure \ref{figV09G12_STA_illus}) shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
+The colored dots match the color of the spectra in Figure \ref{figVG0912} The approximate size of the measurement beam is shown for scale.
+The conspicuous mesh pattern has a lab measured depth of $49\pm6\;$nm, and global fill factor of 50\%.
+However, since the beam size is comparable to mesh grid size, the achieved fill factor can vary from about 30\% to 75\%.
+The gray band in Figure \ref{figVG0912} represents the allowable range of predictions, taking into account both the uncertainty in the mesh depth and fill factor.
+If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \ref{figVG0912}.
+We see excellent consistency between prediction and measurement, though we cannot rule out the presence of extra gaps in the mesh area.
+Outside the mesh area, the bonding is excellent, as demonstrated by the off-mesh transmission close to unity transmission.
 
@@
               \caption{Illustration of VG09-12 sample transport accessory measurement positions}
-		\label{figVG12_STA_corner}
+		\label{figVG0912_STA_illus}
         \end{subfigure}
@@
                 \caption{Measured gap transmission spectra of VG09-12 on- and off- mesh}
-                \label{figVG12_f50}
+                \label{figVG0912}
         \end{subfigure}
 \caption{Predictions and measurements of the coarse mesh region in sample VG09-12.
+We took 50 measurements across the face of VG09-12, including both on- and off- mesh regions.
+The measurements are consistent with the predicted transmission, shown as the gray band.
+The off-mesh regions demonstrate near-perfect bonding.
 \label{figVG12}}
@@
 We applied our MCMC formalism to further constrain the range of gap sizes consistent with the spectrum.
-Table \verb|\ref{tbl summary of gap fit extents}| summarizes the sizes of gaps.
-The well-calibrated DSP noise spectrum rules out ($>3\sigma$) gap sizes larger than about 8.0 nm over $>80$\% of the measurement area.
-Similarly, the VG09-12 off-mesh spectrum rules out gaps greater than 13 nm over $>80$\% of the measurement area.
-
-The variance of individual measurement uncertainties was large enough to consider for this measurement.
-So rather than assume homoscedastic uncertainty as we did for the other measurements, we estimated the heteroscedastic uncertainties in the following way.
-We computed the mean-subtracted transmission for four DSP measurements bookending the those of VG02.
-The DSP measurements were detrended, and the measurement-to-measurement variance computed.
-The result was a variance spectrum which peaks at $1850\;$nm at 0.25\%, but with a median value equal to 0.03\%.
-Figure \verb|\ref{fig_xxx}| includes the $1\;\sigma$ errorbars computed in this way.
+Table \ref{tbl_DerivedGapSizes} summarizes the sizes of gaps.
+The well-calibrated DSP noise spectrum rules out (at $>3\sigma$) gap sizes larger than about 9.0 nm over $>80$\% of the measurement area.
+Similarly, the VG09-12 off-mesh spectrum rules out gaps greater than 14 nm over $>80$\% of the measurement area.
+Figure \ref{} shows the corner plots and spectra for the
 

### 2015-02-16 AO_bondedSi.tex
MGS edits down to the end of the paper

@@
 We describe a bond inspection method that makes use of precision transmission spectroscopy.
-Using spectroscopy, we can detect and measure gaps as small as 12 nm by measuring transmission as a function of wavelength.
+Using spectroscopy, we can detect and measure gaps as small as 14 nm by measuring transmission as a function of wavelength.
 Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
@@
               \includegraphics[width=\textwidth]{figs/VG02p3_corner.pdf}
-              \caption{Fitted gap size $d$ and fill factor $f$ for VG02}
+              \caption{Fitted gap size $d$ and fill factor $f$ for VG09-12 off-mesh}
 		\label{fig_VG02p3_corner}
@@
 \emph{Origin and properties of correlated noise- } Another key assumption of the model is that correlated errors dominate the measurement uncertainty.
-This is the same thing as saying that there is an instrumental bias, or that our measurements are \emph{precise} but \emph{inaccurate}.
-The extent to which the measurement is inaccurate is tabulated in \ref{tbl_cary5000Pars}, as already mentioned.
-We have assumed that the Gaussian Process model can capture the effect of the observed correlated errors.
-But to what extent does the Gaussian Process emulate the \emph{True} covariance matrix, which is unknown?
-One way to gauge the goodness of fit of the covariance matrix is to generate synthetic data from a best-fit model.
-Many of these synthetic draws are shown in the spectra in Figures \verb|\ref{a thru f}| as transparent blue lines.
-When overplotted these lines look like a halo of uncertainty enveloping the model and measurement.
-The bulk agreement is encouraging.
-We also scrutinized individual synthethic spectra with the measured data, and found that the noise model tended to under-predict the length scale of the correlation.
-For example, we typically derive $0.15<a<0.4\%$ and $20<s<200$.
-The measured data appear to prefer larger values of $s$.
-The discrepancy probably comes from the fact that, in reality, the \emph{True} covariance matrix probably has local structures that cannot be replicated by a simple 2-parameter (\emph{i.e.} $a$ and $s$) estimate of the covariance matrix.
-A model employing local covariance structures \verb|\cite{Czekala et al.
-in press}| could better capture some of the residual structure, but would introduce additional computational complexity with little gain.
-Overall the Gaussian Process technique provides much better agreement than the na\"{i}ve assumption of zero correlated errors.
-
-\emph{Validating the MCMC with synthetic data, and addressing heteroscedasticity-}  To evaluate our MCMC fitting method and assumption of homoscedasticity, we generated synthetic data with a known parameters, $d$, $f$, $a$, and $s$ and $\sigma_i$.
-We then fit the parameters $d$, $f$, $a$, and $s$ assuming homoscedastic errors, ($\sigma_i = \sigma = 0.02\%$).
-The derived physical parameters were consistent with the \emph{Truth} to within the $1\;\sigma$ confidence interval, but we significantly underestimated $s$ (8 nm instead of 200 nm) and marginally underestimated $a$ (0.0013 instead of 0.0016).
-This result is reassuring for our estimation of physical parameters, and shows that the Gaussian Process technique captures realistic uncertainty in the physical parameters, despite an imperfect characterization of the noise correlation length.
-
-\emph{Polarization and lamp jumps- } As mentioned, the Cary 5000 double beam mode offers considerably lower amplitude of correlated errors than single beam mode \ref{tbl_cary5000Pars}.
-However, the double beam mode was found to have enhanced artifacts at wavelengths at which the Cary 5000 monochromator changes its diffraction grating and light sources.
-The sharp jumps in transmission were up to a few percent.
-The origin of these jumps could be polarization-dependent transmission losses.
-Whatever the origin, they cause large systematic errors.
-These are still captured by the Gaussian Process structure.
+We have assumed that the Gaussian Process can capture the effect of the observed correlated errors.
+We tested the need for Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
+We applied ordinary least squares regression to the synthetic spectra and derived values for $d$ and $f$ that were biased by up to $20\sigma$ or more.
+Applying Gaussian process regression recovered the input values to within $1\sigma$.
 
@@
 Double beam mode measurements can reach a precision of comparable to the part-to-part variation of refractive index.
-We have normalized spectra by their singleton counterparts from the same heritage (i.e.
-same wafer box).
+We have attempted to normalize spectra by their singleton counterparts when possible.
+We do not present measurements made with double beam mode in this article.
 
 \emph{Diffraction from mesh patterns}
-It's conceivable that the mesh patterns Figure \ref{figVS20pattern}.
+It's conceivable that the edges of the mesh patterns diffract or scatter the beam (Figure \ref{figVS20pattern}).
 We ruled out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.
@@
 
-\subsection{What is the smallest gap size this technique can measure?}
-In this section we quantify the limits of how far the technique can be pushed.
-There are three key ideas when trying to answer the question ``What is the smallest gap size this technique can measure?''.
-The answer for single beam mode is evidently $\sim20 nm$ (Figure \ref{fig_VG02p3_corner}).
-In principle this number could be pushed down by averaging many spectra, all normalized by different reference spectra.
-Alternatively, double beam mode could be used.
-The grating jumps noted above would also limit double beam mode analysis.
-If these could be rectified, a very low uncertainty would be obtained, and it is conceivable that gaps as small as a few nanometers could be inferred.
-
 \section{Conclusions}
@@
 
-The technique and formalism we describe here can detect and measure the axial extent of gaps in Si$-$Si optical bonds down to a level of $20-30\;$nm.
+The technique and formalism we describe here can detect and measure the axial extent of gaps in Si$-$Si optical bonds down to a level of $\sim14\;$nm.
 It makes use of widely available off-the-shelf metrology equipment and simple to implement analysis.

### 2015-03-23 AO_bondedSi.tex
wtf I had uncommitted changes from weeks ago.  looks like changes to figures mostly

@@
 We initialized the walkers in the vicinity of our best guess for parameters based on visual inspection of the spectra.
-Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12_corner} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
+Figures \ref{figVG03full}, \ref{figVG03part}, and \ref{figVG12} show examples of posterior samples including corner plots of the interesting physical parameters $d$ and $f$.
 
@@
 Similarly, the VG09-12 off-mesh spectrum rules out gaps greater than 14 nm over $>80$\% of the measurement area.
-Figure \ref{} shows the corner plots and spectra for the
+Figure \ref{figVG12} shows the corner plots and spectra for the off-mesh region in VG09-12.
 
@@
         \begin{subfigure}[b]{0.5\textwidth}
-              \includegraphics[width=\textwidth]{figs/VG02p3_corner.pdf}
+              \includegraphics[width=\textwidth]{figs/VG0912_0gap_corner.pdf}
               \caption{Fitted gap size $d$ and fill factor $f$ for VG09-12 off-mesh}
-		\label{fig_VG02p3_corner}
+		\label{VG0912_0gap_corner}
         \end{subfigure}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG02_noGap.pdf}
-                \caption{Measured spectrum and fit of VG02}
-                \label{figVG02_noGap}
+                \includegraphics[width=\textwidth]{figs/VG0912_0gap.pdf}
+                \caption{Measured spectrum of VG09-12}
+                \label{VG0912_0gap}
         \end{subfigure}
-\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG02, which has no intentionally imbedded gap.
-The measured spectrum in this region shows no appreciable deficit of transmission from the gapless counterparts.
+\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12 in the off-mesh region.
+The measured spectrum in this region shows no appreciable deficit of transmission from a DSP reference sample.
 \label{figVG12}}
@@
 Simply put, a larger gap can be hidden if it doesn't cover much area.
-The red dashed line shows the normalized transmission spectrum of an 18 nm gap over 30\% of the measurement area.
+The red dashed line shows the normalized transmission spectrum of a 14 nm gap over 100\% of the measurement area.
 Gaps larger than 60 nm would have to cover less than 10\% of the measurement area in order to be consistent with the measurement.
-On this zoomed-in vertical scale, the individual draws from the Gaussian process model show morphologies consistent with the correlated measurement error (Figure \ref{figVG02_noGap}).
 

### 2015-03-23 AO_bondedSi.tex
include DTJ edits up to page 10

@@
 Existing methods for studying gaps provide only poor measures of this axial extent.
-We describe a bond inspection method that makes use of precision transmission spectroscopy.
-Using spectroscopy, we can detect and measure gaps as small as 14 nm by measuring transmission as a function of wavelength.
-Our approach involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
+We describe a bond inspection method that makes use of precision slit spectroscopy.
+Using spectroscopy, we detect and measure gaps as small as 14 nm by measuring transmission as a function of wavelength.
+Our analysis involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
 We infer the gap axial extent and fill factor with Monte Carlo Markov Chain (MCMC) and a flexible Gaussian process noise model.
@@
 
-\subsection{Why Silicon is good}
 Crystalline silicon is an excellent optical material for the infrared.
-Si has a high refractive index ($\sim$3.4, \cite{2006SPIE.6273E..77F}) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
+Si has a high refractive index ($3.55-3.45$ from $\lambda = 1150-2500\;$nm at room temperature, \cite{2006SPIE.6273E..77F}) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
 It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
@@
 
-\subsection{Why bonding is good and AR Coating is bad}
 As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
@@
 
-\subsection{Silicon Bonds- history and techniques}
-
 There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \cite{1998AnRMS..28..215G,Masteika2014}.
@@
 (1988) and G{\"o}esele et al.
-(1995) \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.
+(1995) \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with time-lapse IR imaging.
 Lehman et al.
@@
 Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
-Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.
-
-\subsection{Requirement:  Do better than coatings averaged over the surface}
+Greco \emph{et al.} examined the role of surface topography in determining the the bonding strength of thick optical surfaces.
+These authors showed that van der Waals interactions by themselves were consistent with their available experimental evidence, but they could not rule out other forces and phenomena.
+The dearth of measurements of the mean separation between contacted surfaces hindered the unambiguous identification of the forces at play in bonded, rough surfaces \cite{2001JOptA...3...85G}.
 
@@
 While completely bonded Si parts will be completely lossless at the interface, problems on different scales with different physical origins can lead to less-than-perfect bonds.
-On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation.
+On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart \cite{Mitani1990} or where hydrocarbons can serve as catalysts for gap formation.
 Even in the best of cleanrooms, particles at the optically significant $10-20\;$nm scale can be present in relatively high abundance.
@@
 There will be roughly $10^4$ times more 10 nm sized particles than 1 $\mu$m sized particles.
-On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \cite{Masteika2014}.
-Most of the bonding literature concerns itself with wafer bonding.
+Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform \cite{2001JOptA...3...85G}.
+On larger scales, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \cite{Masteika2014}.
+
+Most of the bonding literature concerns itself with wafer bonding, as opposed to thicker optical substrates.
 In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.
@@
 When bonding two thick substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
+Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 75 to 100 mm diameter clear aperture.
+We assume that, if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
+Subsequent annealing will strengthen the bond.
 
-\subsection{Existing detection schemes}
 Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
@@
 Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect.
-In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap.
+In cases where the axial extent of the gap is a half-wavelength or more, Newton rings appear in the image of the gap.
 X-ray topography \cite{Mitani1990} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not of the axial extent itself.
@@
 
-\subsection{Technique}
 We describe here a new technique for measuring the axial extent of small gaps, founded upon the ability to measure transmission as a function of wavelength with high accuracy.
@@
 
-\subsection{What level do you need to measure to}
 For all types of gaps and bubbles, our technique needs to provide an accurate assessment of the transmission loss at the operating wavelength.
@@
 
-\section{Conceivable pitfalls in Si$-$Si direct bonding}
-
-It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \cite{Mitani1990}.
-
-Another possible source of gaps at the bond is inherent roughness in the initial surfaces that leads to a failure to conform.
-Stiff non-flat Si surfaces will not conform upon contact leaving the complement of the aberrations as an air gap.
-Unconventional thick ($\sim30\;$mm) Si pucks are less conformal than wafers are, so we expect thick substrates could have gaps even in the absence of particles.
-Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 75 to 100 mm diameter clear aperture.
-We assume that, if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
-Subsequent annealing will strengthen the bond.
-
 \section{Si$-$Si Fabry P\`{e}rot measurement theory}
@@
 Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon at room temperature ranges from $3.55-3.45$ from $\lambda = 1150-2500\;$nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.
+The refractive index of silicon is one of the highest of any commonly used optical material.
+Figure \ref{figSiIndexFinesse} shows the refractive index at room temperature \cite{2006SPIE.6273E..77F}, which we assume through this article.
 The Fresnel reflection is about $30\;\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
-Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature,  computed using the coefficients from \cite{2006SPIE.6273E..77F}.
 
@@
 \caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.
-The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.
-We assume $T= 295\;$K for all calculations in this article.}
+The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.}
 \end{figure}
@@
 The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at $T= 295\;$K.
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
@@
 We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
-The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength for $T = 295\;$K and $T = 77\;$K.
+The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength.
+In this work, we assume transmission is at normal incidence and the refractive index of the gap is 1.0.
 Si absorbs negligibly longward of $\lambda$ = 1250 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses (Figure \ref{figSiAbsorbfig}).
@@
 An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
-In the next subsection we work out the wavelength dependent transmission for the air etalon.
+In the next subsection we work out the wavelength dependent transmission for the air etalon\footnote{A note on nomenclature- we use the term \emph{gap} for the physical structure, and \emph{etalon} for the model of that structure.}
 
@@
 \label{secTheory}
-A pair of bonded silicon wafers or pucks with a small gap actually forms three coupled cavities with the air gap as the the central cavity.
+A pair of bonded silicon wafers or pucks with a small gap forms three coupled cavities with the air gap as the the central cavity.
 The outer cavities, those within the Si, however, were not produced with coherence in mind.
 Their sides are not necessarily flat and parallel to the required level and the specifics of their deviation from perfect etalons are usually not known to the level necessary to treat them as coherent structures in the analysis.
-The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps whereas the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
+The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps.
+In the air gap, however, the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
 Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
@@
 Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
-If the Si wafer thickness is less than the coherence length for the given spectral bandwidth, then our assumptions break down and the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.
+The top line in the upper panel of Figure \ref{figAbsoluteTrans} shows the value derived from Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
+If the Si substrate thickness is less than the coherence length for the given spectral bandwidth, and if its surface roughness is much smaller than the wavelength, our assumptions break down and the multiple reflections would interfere coherently, and the outer Si cavities would behave as a Fabry-P\`erot etalon.
 We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.
@@
 We derive the absolute transmission of bonded Si substrates with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
-We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.
-We call this normalized transmission the etalon transmission $T_{e}$.
+We normalize the absolute transmission by the transmission of a double-sided Si structure with no gap to isolate the effect of the gap.
+We call this normalized transmission the gap transmission $T_{g}$.
 
 \begin{eqnarray}
-T_{e} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}
+T_{g} = \frac{n^2+1}{2 n F \sin ^2(2\pi \frac{d}{\lambda})+n^2+1} \label{eqFP}
 \end{eqnarray}
 
-In the limit $\lim_{d \to 0} T_e \rightarrow T_{DSP} $ and the etalon approaches 100\% transmission.
+In the limit $\lim_{d \to 0} T_g \rightarrow T_{DSP} $ and the gap approaches 100\% transmission.
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
@@
 \begin{equation}
-	T_{mix} = f\;T_{e} + (1-f) \label{eqnMix}
+	T_{mix} = f\;T_{g} + (1-f) \label{eqnMix}
 \end{equation}
@@
 where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.
-$T_e$ is defined in equation \ref{eqn:Tetalon}.
+$T_g$ is defined in equation \ref{eqn:Tetalon}.
 The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
@@
 The first key idea is to treat the entire air gap as an abstraction, a Fabry-P\`erot etalon.
-We do not need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon:
+We do not need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon model for the gap:
 
@@
   F \equiv \frac{4R}{(1-R)^2} \\
- T_e = \frac{1}{1+F\sin^2(\delta/2)}  \label{eq:FabPerot}
+ T_g = \frac{1}{1+F\sin^2(\delta/2)}  \label{eq:FabPerot}
 \end{eqnarray}
@@
 The phase depends (Equation \ref{eq:phase}) on the wavelength $\lambda$ and $d$ the air gap spacing.
-We assume the etalon is lossless, i.e.
-$T_e+R_e=1$.
-So the incoherent scattering and transfer matrices for the etalon are:
+We assume the gap is lossless, i.e.
+$T_g+R_g=1$.
+So the incoherent scattering and transfer matrices for the gap are:
 
 \begin{eqnarray}
-\boldsymbol{S_e}&=&\frac{1}{1+F\sin^2{\delta/2}} \left(
+\boldsymbol{S_g}&=&\frac{1}{1+F\sin^2{\delta/2}} \left(
 \begin{array}{cc}
@@
 \nonumber \\
-\boldsymbol{M_e}&=&\left(
+\boldsymbol{M_g}&=&\left(
 \begin{array}{cc}
@@
 \begin{eqnarray}
-\boldsymbol{M_{net}}=\boldsymbol{M_n}\boldsymbol{M_e}\boldsymbol{M_n}
+\boldsymbol{M_{net}}=\boldsymbol{M_n}\boldsymbol{M_g}\boldsymbol{M_n}
 \end{eqnarray}
@@
 The revised transmission has picked up a few factors of 2 and $n$.
-While we are at it, let's compute the matrix for the scenario with no intermediate etalon: there are simply two Fresnel interfaces with which light interacts incoherently.
+While we are at it, let's compute the matrix for the scenario with no intermediate gap: there are simply two Fresnel interfaces with which light interacts incoherently.
 This scenario is the model for a single DSP reference sample.
@@
 It is informative to isolate the effect of the gap by dividing the measured bonded wafer transmission by the transmission of a reference DSP Si part.
-We call this normalized transmission the etalon transmission $T_{e}$:
+We call this normalized transmission the gap transmission $T_{g}$:
 \begin{eqnarray}
-T_{e} = T_{net}/T_{DSP} \\
-T_{e} = \frac{n^2+1}{2 n F \sin ^2(\delta/2)+n^2+1} \label{eqn:Tetalon}
+T_{g} = T_{net}/T_{DSP} \\
+T_{g} = \frac{n^2+1}{2 n F \sin ^2(\delta/2)+n^2+1} \label{eqn:Tetalon}
 \end{eqnarray}

### 2015-03-23 AO_bondedSi.tex
incorporate DTJ edits up to page 17

@@
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
-The bottom two panels show the ``gap'' transmission, the value of the throughput with the losses from the outer Si cavities divided out.
+The bottom two panels show the ``gap'' transmission based on Equation \ref{eqFP}, the value of the throughput with the losses from the outer Si cavities divided out.
 For the very largest gap, large fringes appear as the cavity resonance of the air gap modulates the transmission.
-Below a $\lambda/2$ gap size the effect of the gap manifests itself as a wavelength-dependent diminution of the transmission.
-For the smallest $35-60$ nm gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.
+Below a $\lambda/2$ gap size, the gap manifests itself as a wavelength-dependent diminution of the transmission.
+For the smallest ($35-60$ nm) gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.
 It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.
@@
 
-We evaluated our metrology method by embedding gaps of known sizes in the interface and measuring the effects of these gaps on IR transmission.
+We evaluated our metrology method by embedding gaps of known sizes in the interface between two bonded Si substrates and measuring the effects of these gaps on IR transmission.
 The substrate thicknesses ranged from 0.5 to 3.3 mm.
@@
 
-The three important characteristics of gaps are their axial extent, their lateral areal extent, and their areal fill factor.
+The three important characteristics of gaps are their axial extent, their lateral areal extent, and the areal fill factor of ensembles of such gaps.
 To fabricate our test samples, we used photoresist lithography to pattern Si substrates, and we bored holes with inductively coupled plasma etching.
-We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.
+We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 $\mu$m/minute to produce gaps with small and large depths.
 Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
 
-The fill factor is defined as the pattern area covered by gaps relative to the total pattern area.
-In a second round of experiments, we achieved gap sizes of $14-95\;$nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
-No attempt was made to verify the delivered fill factor, but since high precision lithographic processes were employed, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
+The fill factor is the pattern area covered by gaps relative to the total pattern area.
+We achieved gap sizes of $14-95\;$nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+We did not verify the delivered fill factor, but since high precision lithography is very reliable, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
 For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
 The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
-The stylus profilometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.
+The stylus profilometry measured values were assigned an uncertainty of 5\%, our experience with other etched parts.
 Table \ref{tbl_experiments} lists part numbers and descriptions for our bonded substrates.
@@
 The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
-We measured the large scale surface flatness of the pucks.
+We also measured the large scale surface flatness of the pucks.
 The Fisba2 interferometer used for these measurements has a 50 mm diameter beam at $\lambda=632.8\;$nm.
@@
 
-\section{Measurement technique and correlated measurement errors}
-\label{sec_aboutErrors}
-We took transmission spectra of bonded Si wafers and pucks, and double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
-Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
-The tool has a double monochromator, which reduces scattered light.
-The vendor reported linearity exceeds 40 dB.
+\subsection{IR bubble imaging and limitations}
 
-The Cary 5000 has four operation modes: single front, single back, double, and double reverse.
-We experimented with single front and double modes.
-Double mode takes a reference spectrum simultaneously with the target spectrum.
-In both single and double mode, a baseline scan is taken with no sample in the measurement holder.
-We automatically divide sample spectra by the baseline spectrum.
-In single beam mode, the baseline measurements must be repeated about every 10 minutes for post-process division.
-In double beam mode the front beam is automatically divided by the rear beam, which corrects out lamp luminosity drift, for example.
-
-We estimated the uncertainty and measurement repeatability of single and double modes in the following way: we computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
-We found that the mean value in single mode drifted by about 0.2\% over several measurements.
-The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance, perhaps attributable to atmospheric absorption.
-In double beam mode, the spectrograph drift was much lower.
-The simultaneous reference in double beam mode divides out the 0.2\% drift in the mean transmission level attributable to lamp drift that we saw in single beam measurements.
-In double beam mode, the drift in mean transmission level is 0.02\%.
-These values are accurate for measurements over a half hour timescale or less.
-Table \ref{tabCary5000pars} summarizes the amplitudes of the drifts and individual measurement errors in both single and double beam modes.\emph{Correlated errors dominate the measurement uncertainty of our transmission spectra.}  The measurement uncertainty correlations are apparent in in Figure \ref{figSiAbsorbfig}.
-In some parts of the spectrum, the mean drift is about ten times larger than the standard deviation of the spectrum.
-Double beam mode has about ten times lower lamp drift but suffers from other instrumental artifacts, like disjointed spectra at the wavelengths of known grating changeover points.
-
-The spectral sampling ranged from 2.0 to 10.0 nm.
-We used a spectral resolution of 5.0 nm.
-We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
-Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
-The axial extents of such large gaps would be easily measurable by other means, like IR imaging.
-
-\input{tbls/tbl_cary5000Pars.tex}
-
-\subsection{IR bubble detection and limitations}
-
-We looked for large gaps, detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.
+We first looked for large gaps in our bonded substrate pairs detectable as ``IR bubbles'' \cite{1992JEMat..21..669M}, with IR imaging.
 The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
@@
 The individual IR bubble areas were as small as a few square millimeters up to $400\;\mathrm{mm}^2$.
-The IR bubbles had varied morphologies.
-The structure within the IR bubbles was sometimes indicative of Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
+The IR bubbles had various morphologies.
+The structure within some of the images of IR bubbles showed Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
 As expected, the unbonded DSP wafers show no such IR bubbles.
@@
 
+\section{Measurement technique and correlated measurement errors}
+\label{sec_aboutErrors}
+
+We used the test pieces with known gaps to validate our measurement and analysis techniques.
+We took transmission spectra of the bonded Si wafers and pucks, and the double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
+Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
+The tool has a double monochromator, which reduces scattered light.
+The vendor reported linearity exceeds 40 dB.
+
+The Cary 5000 has four operation modes: single front, single back, double, and double reverse.
+Double mode takes a reference spectrum simultaneously with the target spectrum.
+We experimented with single front and double modes.
+In principle double beam mode is even more precise than single beam mode.
+In practice we found that souble beam mode showed evidence for lamp-jump artifacts attributable to polarization differences between the two arms.
+So in this article we focus our results to those obtained in single beam mode.
+In both single and double mode, the spectrometer takes a baseline scan with no sample in the measurement holder.
+We automatically divide sample spectra by this baseline spectrum.
+In single beam mode, we must repeat the baseline measurements about every 10-30 minutes for post-process division.
+
+We estimated the uncertainty and measurement repeatability experimentally.
+We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
+We found that the mean value in single mode drifted by about 0.2\% over several measurements.
+The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance.
+Table \ref{tabCary5000pars} summarizes the amplitudes of the drifts and individual measurement errors.
+\emph{Correlated errors dominate the measurement uncertainty of our transmission spectra.}  The measurement uncertainty correlations are apparent in Figure \ref{figSiAbsorbfig}.
+In some parts of the spectrum, the mean drift is about ten times larger than the local standard deviation of the spectrum.
+
+The spectral sampling ranged from 2.0 to 10.0 nm.
+We set the spectrometer slit width to produce a spectral resolution of 5.0 nm.
+We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
+Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
+This maximum gap limitation brings our technique to the point where it overlaps in capability with other techniques since the axial extents of such large gaps would be easily measurable by other means, like IR imaging and fringe-counting.
+
+We also verified the linearity of the detector response my measuring the transmission through stacks of filters.
+We constained the linearity to at least $>24$ dB.
+The vendor reports linearity exceeding 40 dB, which is consistent with our measurement.
+
+\input{tbls/tbl_cary5000Pars.tex}
+
 \section{Results of infrared spectroscopy of directly bonded Si}
@@
 We infer the axial extent of gaps in our samples with etched holes from their measured transmission spectra.
-We use a Bayesian inference technique.
+We use a Bayesian inference technique in our analysis.
 First we define a mixture model in which the observed normalized gap spectrum $T_{obs}$ is composed of a sum of a spectrum through a gap of axial extent $d$, and a spectrum through a perfect bond:
@@
 where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.
-$T_g$ is defined in equation \ref{eqn:Tetalon}.
+$T_g$ is defined in Equation \ref{eqFP}.
 The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.

### 2015-03-23 AO_bondedSi.tex
incorporate DTJ edits up to page 21

@@
 where $\delta_{ij}$ is the Kronecker delta, and $\sigma_i$ are the independent measurement uncertainties on the $i^{\mathrm{th}}$ data point.
+The parameter $s$ controls the correlation length, and $a$ controls the amplitude of the correlated noise.
 We experimented with different values for $\sigma_i$.
-When repeated DSP reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.
-The parameter $s$ controls the correlation length, and $a$ controls the amplitude of the noise.
-We do not know $a$ nor $s$ and they are, in general, different for each measurement.
+When repeated double side polished reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.
+We do not know $a$ nor $s$ and they are, in general, different for each measurement-- they are \emph{nuisance parameters}, and are ultimately marginalized out.
+They arise from the uninteresting properties of the spectrometer, and their interpretation is immaterial.
 The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.
@@
 
-We set our prior probability distribution functions \cite{2013sdmm.book.....I}, $\ln{p(d,f,a,s)}$ as uniform over a wide range of values for $\ln{a}$, $\ln{s}$, $d$, and $f$ consistent with visual inspection of each spectrum.
+We set our prior probability distribution functions \cite{2013sdmm.book.....I}, $\ln{p(d,f,a,s)}$ as uniform over a wide range of values for $\ln{a}$, $\ln{s}$, $d$, and $f$, consistent with visual inspection of each spectrum.
 The final un-normalized posterior probability function, expressed as a natural logarithm is:
@@
 
-Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap.
-The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physicsal measurement.
+Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap in part VG03.
+The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physical measurement (Table \ref{tbl_experiments}).
+The spectrum in this region shows characteristic Fabry-P\`erot fringes of constructive and destructive interference.
+The only tunable parameters of the model are the gap axial extent, $d$ and fill factor $f$.
+No scaling has been applied to the curves in Figure \ref{figVG03full}.
 Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field scanned by the Cary 5000.
@@
 The physical parameters of the model are the axial extent of the gap $d$, and the areal fill factor $f$ of the gap over the measurement region.
-The MCMC model fit also includes nuisance parameters $a$ and $s$ (not shown) for the Gaussian Process covariance model.
+The MCMC model fit also includes nuisance parameters $a$ and $s$ (not shown) for the Gaussian process covariance model.
 \emph{Bottom panel:}Transmission through bonded Si sample VG03, normalized by the transmission of a DSP Si wafer.
-VG03 has a gap of axial extent $\sim4\;\mu$m implanted in its interface.
-The measured spectrum (black stepped line) in this region shows characteristic Fabry-P\`erot fringes of constructive and destructive interference.
-The only tunable parameters of the model are the gap axial extent, $d$ and fill factor $f$.
-No scaling has been applied.
+See the text for details on the sample, setup, and analysis.
+The measured spectrum is the black stepped line.
 The red dashed line shows the best model fit with $d$ = 3960 $\pm$ 2 nm, consistent with 100\% fill factor.
-The blue band shows the collective locations of 60 random draws from the Gaussian process model, described in the text.
+The blue band shows the collective locations of 60 random draws from the Gaussian process model.
 Wavelengths shortward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.\label{figVG03full} }
@@
         \end{subfigure}
-\caption{ Same as Figure \ref{figVG03full}, but the measurement was in a region where less than $<5\%$ of the area exhibited the $\sim4\; \mu$m gap.
-The measured spectrum in this region shows subdued Fabry-P\`erot fringes of constructive and destructive interference.
+\caption{ Same as Figure \ref{figVG03full}, but the measurement was in a region where $<5\%$ of the area exhibited the $\sim4\; \mu$m gap.
+The measured spectrum in this region shows subdued fringes.
 The model derived axial extent is 4094 $\pm$ 5 nm covering a fill factor of 4.58 $\pm$ 0.06$\%$.
@@
 \subsection{Spatially resolved measurements with Sample Transport Accessory}
-We acquired spatially resolved measurements of sample VG09-12 with the Cary 5000 sample transport accessory \footnote{\url{http://www.chem.agilent.com/en-US/products-services/instruments-systems/molecular-spectroscopy/sample-transport-accessory/}}.
+We acquired spatially resolved measurements of the pair of bonded parts VG09-12, (Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns}) with the Cary 5000 sample transport accessory \footnote{\url{http://www.chem.agilent.com/en-US/products-services/instruments-systems/molecular-spectroscopy/sample-transport-accessory/}}.
 This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
@@
 We fit a line to before-and-after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
-We reduced the measured wavelength range to $1250-1775\;$nm.
+We measured over a wavelength range from 1250 to 1775 nm.
 
-Figure \ref{figV09G12_STA_illus}) shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
+Figure \ref{figV09G12_STA_illus} shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
 The colored dots match the color of the spectra in Figure \ref{figVG0912} The approximate size of the measurement beam is shown for scale.
@@
               \includegraphics[width=\textwidth]{figs/VG09_12_STA_illustration.pdf}
-              \caption{Illustration of VG09-12 sample transport accessory measurement positions}
-		\label{figVG0912_STA_illus}
+              \caption{Illustration of VG09-12 sample transport accessory measurement positions \label{figVG0912_STA_illus}}
         \end{subfigure}
@@
                 \includegraphics[width=\textwidth]{figs/VG0912_STA_scan_coarse.pdf}
-                \caption{Measured gap transmission spectra of VG09-12 on- and off- mesh}
-                \label{figVG0912}
+                \caption{Measured gap transmission spectra of VG09-12 on- and off- mesh \label{figVG0912}}
         \end{subfigure}

### 2015-03-24 AO_bondedSi.tex
incorporate all DTJ edits except for (old) section 7

@@
 
-Figure \ref{figV09G12_STA_illus} shows an illustration of sample VG09-12, with the approximate center locations of measurements marked with colored dots.
-The colored dots match the color of the spectra in Figure \ref{figVG0912} The approximate size of the measurement beam is shown for scale.
-The conspicuous mesh pattern has a lab measured depth of $49\pm6\;$nm, and global fill factor of 50\%.
-However, since the beam size is comparable to mesh grid size, the achieved fill factor can vary from about 30\% to 75\%.
+The colored dots in the position marked in Figure \ref{figVG0912_STA_illus} match the color of the spectra in Figure \ref{figVG0912}.
+The approximate size of the measurement beam is shown for scale.
+The conspicuous mesh pattern has a depth of $49\pm6\;$nm, as measured with an optical profiler, and global fill factor of 50\%.
+However, since the beam size is comparable to mesh grid size, the fill factor for any given measurement can vary from about 30\% to 75\%.
 The gray band in Figure \ref{figVG0912} represents the allowable range of predictions, taking into account both the uncertainty in the mesh depth and fill factor.
 If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \ref{figVG0912}.
-We see excellent consistency between prediction and measurement, though we cannot rule out the presence of extra gaps in the mesh area.
-Outside the mesh area, the bonding is excellent, as demonstrated by the off-mesh transmission close to unity transmission.
+We see excellent consistency between prediction and measurement, with one measured spectrum hinting at a small gap ($\lesssim15$) in one measurement position.
+Outside the mesh area, the bonding is excellent, as demonstrated by the off-mesh transmission with transmission consistent with 100\% to within the measurement uncertainty of 0.2\%.
 
@@
         \includegraphics[width=\textwidth]{figs/VG09_12_STA_illustration.pdf}
-              \caption{Illustration of VG09-12 sample transport accessory measurement positions \label{figVG0912_STA_illus}}
+        \caption{Illustration of VG09-12 sample transport accessory measurement positions. \label{figVG0912_STA_illus} }
     \end{subfigure}
@@
 The measurements are consistent with the predicted transmission, shown as the gray band.
-The off-mesh regions demonstrate near-perfect bonding.
-\label{figVG12}}
+The off-mesh regions demonstrate near-perfect bonding.\label{figVG12}}
 \end{figure}
 
-A key result of this work is to show that gaps of known dimensions are recovered in our spectroscopy.
+A key result of these test measurements is to demonstrate that gaps of known dimensions are recovered in our spectroscopy, both for large and small gaps.
 Table \ref{tbl_DerivedGapSizes} summarizes the outcome of some measurements in the areas with known gaps and fill factors.
 The table lists the median sample and 68\% confidence intervals.
+The results from the scans shown in Figure \ref{figVG12} further support our conclusion.
 
@@
     \end{subfigure}
-
     \begin{subfigure}[b]{0.5\textwidth}
@@
     \end{subfigure}
-\caption{ Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12 in the off-mesh region.
+\caption{Similar to Figure \ref{figVG03full} and \ref{figVG03part}, but for sample VG09-12 in the off-mesh region.
 The measured spectrum in this region shows no appreciable deficit of transmission from a DSP reference sample.
@@
 
-Figure \ref{fig_VG02p3_corner} shows the degeneracy between the axial extent of a gap and its fractional fill factor over the measurement beam.
-Simply put, a larger gap can be hidden if it doesn't cover much area.
+Figure \ref{VG0912_0gap_corner} shows that there is a degeneracy between the axial extent of a gap and its fractional fill factor over the measurement beam; a larger gap can be hidden if it covers only a small fraction of the measured area.
 The red dashed line shows the normalized transmission spectrum of a 14 nm gap over 100\% of the measurement area.
@@
 
-\emph{Origin and properties of correlated noise- } Another key assumption of the model is that correlated errors dominate the measurement uncertainty.
-We have assumed that the Gaussian Process can capture the effect of the observed correlated errors.
+\emph{Origin and properties of correlated noise- } Another key assumption of our analysis is that correlated errors dominate the measurement uncertainty.
+We have assumed that the Gaussian process can capture the effect of the observed correlated errors.
 We tested the need for Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
@@
 
-The wave transfer matrix method is described in detail in Chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.
-Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
+Saleh and Teich (see Chapter 7 ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}) describe the wave transfer matrix method.
+Their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
 \begin{eqnarray}
@@
 Where $t$ and $r$ stand for transmission and reflection respectively.
-The order of subscripts is the order of origin and destination of the wave in regards to the interface, so we know something about the direction of the wave and how it got there.
-The $\boldsymbol{S}$ matrix encapsulates all the information we need to know about how light waves interact with the interface.
-The real power of the matrix technique comes from a cousin of the scattering matrix, the so-called wave transfer matrix, $\boldsymbol{M}$.
+The order of subscripts is the order of origin and destination of the wave with respect to the interface, so we know the origin and direction of the wave.
+The $\boldsymbol{S}$ matrix encapsulates all information about how light waves interact with the interface.
+The power of the technique comes from the wave transfer matrix, $\boldsymbol{M}$.
 The matrix $\boldsymbol{M}$ has the convenient property that its output can be used as the input for another matrix.
@@
 The matrix for the silicon air gap is calculated in the following way.
-The first key idea is to treat the entire air gap as an abstraction, a Fabry-P\`erot etalon.
-We do not need to bother to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon model for the gap:
+We treat the air gap as a Fabry-P\`erot etalon.
+We do not need to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon model for the gap:
 

### 2015-03-24 AO_bondedSi.tex
spruce up the discussion section to be more general from a reader point of view

@@
 
-\subsection{Questioning assumptions of the model}
+The experiments and measurements in Section \ref{secResults} demonstrate that we can detect and measure the axial extent of gaps down to 14 nm.
+The technique is \emph{non-destructive} and can be implemented with an off-the-shelf comercial product, namely the Cary 5000 or equivalent, or any existing high precision spectrograph.
+The spectrograph need not be high resolution-- our 5 nm spectral resolution was not a limiting factor.
+The limiting factor is spectrophotometric \emph{accuracy} and \emph{precision}.
+In this Section we question the assumptions of our technique, identify its limitations, and mention some opportunities.
 
-\emph{More complex underlying gap size distributions-} One key assumption of our model is that the distribution of gap sizes over our measurement area is bimodal.
-Specifically, Equation \ref{eqnMix} assumes that there is either a gap of size $d$, or else it is 100\% bonded.
-If there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
+\emph{More complex underlying gap size distributions-} One key assumption of our gap-as-etalon modeling strategy is that the distribution of gap sizes over the measurement area is bimodal- a gap of size $d$ or zero gap.
+This choice was simply a matter of computational simplicity, and any more complicated gap size distribution could easily be dropped in as additional terms in Equation \ref{eqnMix}.
+For example, if there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
 The full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
@@
 
-\emph{Origin and properties of correlated noise- } Another key assumption of our analysis is that correlated errors dominate the measurement uncertainty.
-We have assumed that the Gaussian process can capture the effect of the observed correlated errors.
-We tested the need for Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
+\emph{The benefit of the Gaussian process-} We chose to estimate the gap size with a Gaussian process regression, but this is not necessary.
+Any model fitting strategy, like least squares or overplotting a model to data by-hand, will yield an estimate for a gap size.
+We tested the benefit of Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
 We applied ordinary least squares regression to the synthetic spectra and derived values for $d$ and $f$ that were biased by up to $20\sigma$ or more.
@@
 
-\emph{DSP vs AR coated front faces- } Another key assumption of the model is that all the Si$-$air interfaces are well-approximated by Fresnel reflection losses.
+\emph{DSP vs AR coated front faces- } We have assumed that the Si$-$air interfaces are approximated by Fresnel reflection losses.
 If the entrance and exit faces of the bonded samples are AR coated, the incoherent multiple reflection model will have to be modified.
@@
 
-\emph{Refactive index- } We have assumed that the refractive index is identical to that given by \cite{2006SPIE.6273E..77F}.
-Double beam mode measurements can reach a precision of comparable to the part-to-part variation of refractive index.
-We have attempted to normalize spectra by their singleton counterparts when possible.
-We do not present measurements made with double beam mode in this article.
+\emph{How far can this technique go?}  The spectrophotometric precision of our instrument flows down directly to limit the smallest measureable gap size.
+This limitation will vary from instrument to instrument.
+Our instrument can achieve 10 times better accuracy and precision in double beam mode.
+However, the slightly different polarization sensitivity in the different beams caused systematic jumps at the lamp/grating change-over wavelengths.
+Agilent offers an adaptor to reduce the polarization sensitivity, which mitigates these lamp jumps.
+If precisions of 0.02\% could be achieved, gaps $<4$ nm could be reliably measured.
+This level begins to be comparable to the effect of interesting physics, like surface roughness and refractive index.
+In fact, the van der Waals bond energy could be constrained from these types of measurements \cite{2001JOptA...3...85G}.
 
-\emph{Diffraction from mesh patterns}
-It's conceivable that the edges of the mesh patterns diffract or scatter the beam (Figure \ref{figVS20pattern}).
-We ruled out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.
-To eliminate this concern, we measured in the coarse mesh rather than the medium or fine meshes, when possible.
-We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.
-Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.
-In other words, what matters is the fill factor of gaps, not their relative sizes.
-Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.
+\emph{Can this technique be applied to imaging?}  The analysis performed in this article informs smart strategies for using IR imaging for gap detection in bonded Si optics.
+First, Equation \ref{eqFP} shows that short wavelengths have the most discerning power for detecting small gaps.
+So an IR imaging approach should use the shortest possible wavelength without getting close to the Si absorption cutoff.
+This work recommends a 1250$-$1300 nm filter.
+Images can then be normalized to the transmission through a DSP sample under identical, nearly collimated lighting conditions.
+Lastly, any deficits in flux of a bonded sample can be converted to a gap size $d$, using Equation \ref{eqFP}.
+IR imaging will lack the spectral information available in the spectra shown in this article.
+The imaging will require exceptional quality of flat-fielding or dithering of the images to reach the precision achieved here with relative ease.
 

### 2015-03-24 AO_bondedSi.tex
run spell check

@@
 Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness.
-Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level longward of $1250\;$nm.
+Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level long-ward of $1250\;$nm.
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
@@
 The bulk morphologies were clear regardless of the detector response.
-When IR imaging was available, it was leveraged to target IR specectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.
+When IR imaging was available, it was leveraged to target IR spectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.
 In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.
@@
 We also verified the linearity of the detector response my measuring the transmission through stacks of filters.
-We constained the linearity to at least $>24$ dB.
+We constrained the linearity to at least $>24$ dB.
 The vendor reports linearity exceeding 40 dB, which is consistent with our measurement.
@@
 The blue band shows the collective locations of 60 random draws from the Gaussian process model.
-Wavelengths shortward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.\label{figVG03full} }
+Wavelengths short-ward of 1250 nm (gray band) were not included in the fit, since Si is absorptive there.\label{figVG03full} }
 \end{figure}
@@
 The experiments and measurements in Section \ref{secResults} demonstrate that we can detect and measure the axial extent of gaps down to 14 nm.
-The technique is \emph{non-destructive} and can be implemented with an off-the-shelf comercial product, namely the Cary 5000 or equivalent, or any existing high precision spectrograph.
+The technique is \emph{non-destructive} and can be implemented with an off-the-shelf commercial product, namely the Cary 5000 or equivalent, or any existing high precision spectrograph.
 The spectrograph need not be high resolution-- our 5 nm spectral resolution was not a limiting factor.
@@
 We found that measurements of very large ($>10\lambda$) gaps with a large gradient offered surprisingly strong constraints on the maximum gap size, because of the conspicuous high frequency Fabry-P\`erot fringes.
-For sub-wavelength gaps, the gap size distibution is too degenerate to derive much useful insight on the slope of the distribution.
+For sub-wavelength gaps, the gap size distribution is too degenerate to derive much useful insight on the slope of the distribution.
 Specifically, for $d_{max}<\lambda/4$, our simple mixture model will converge on the spectrum of the average gap size, even if the \emph{True} distribution is more complex.
@@
 \emph{The benefit of the Gaussian process-} We chose to estimate the gap size with a Gaussian process regression, but this is not necessary.
-Any model fitting strategy, like least squares or overplotting a model to data by-hand, will yield an estimate for a gap size.
+Any model fitting strategy, like least squares or over-plotting a model to data by-hand, will yield an estimate for a gap size.
 We tested the benefit of Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
@@
 However, the slightly different polarization sensitivity in the different beams caused systematic jumps at the lamp/grating change-over wavelengths.
-Agilent offers an adaptor to reduce the polarization sensitivity, which mitigates these lamp jumps.
+Agilent offers an adapter to reduce the polarization sensitivity, which mitigates these lamp jumps.
 If precisions of 0.02\% could be achieved, gaps $<4$ nm could be reliably measured.

### 2015-04-17 AO_bondedSi.tex
incorporate final DTJ edits, add Acknowledgment section

@@
 \emph{More complex underlying gap size distributions-} One key assumption of our gap-as-etalon modeling strategy is that the distribution of gap sizes over the measurement area is bimodal- a gap of size $d$ or zero gap.
-This choice was simply a matter of computational simplicity, and any more complicated gap size distribution could easily be dropped in as additional terms in Equation \ref{eqnMix}.
+This choice was simply a matter of computational simplicity, and we could easily account for any more complicated gap size distribution by inserting additional terms in Equation \ref{eqnMix}.
 For example, if there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
-The full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
+Since the gap sizes are small and the interface takes place within a subregion of constant gap size, the full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
 Solving for the relative distributions of gaps is highly degenerate, in the same way as we saw in the 2-parameter corner plots of gap depth and fill factor.
 We experimented with by-eye fitting parametric distributions of gap sizes, characterized by a maximum and minimum gap size, and a monotonic slope of the distribution.
-We found that measurements of very large ($>10\lambda$) gaps with a large gradient offered surprisingly strong constraints on the maximum gap size, because of the conspicuous high frequency Fabry-P\`erot fringes.
 For sub-wavelength gaps, the gap size distribution is too degenerate to derive much useful insight on the slope of the distribution.
@@
 
-\emph{The benefit of the Gaussian process-} We chose to estimate the gap size with a Gaussian process regression, but this is not necessary.
-Any model fitting strategy, like least squares or over-plotting a model to data by-hand, will yield an estimate for a gap size.
-We tested the benefit of Gaussian process regression by generating synthetic data spectra with known gap size and fill factor, and known covariance structure.
+\emph{The benefit of the Gaussian process-} We chose to estimate the gap size and fill factor with Gaussian process regression.
+Gaussian process regression is not necessary if random errors are much larger than the systematic errors, in which case any model fitting strategy---like least squares or over-plotting a model to data by-hand---will yield an unbiased estimate for the gap size.
+We tested the benefit of Gaussian process regression by generating synthetic transmission spectra with known gap size and fill factor, and known covariance structure.
 We applied ordinary least squares regression to the synthetic spectra and derived values for $d$ and $f$ that were biased by up to $20\sigma$ or more.
@@
 If the entrance and exit faces of the bonded samples are AR coated, the incoherent multiple reflection model will have to be modified.
-AR coated front and exit faces will result in a larger interface loss for a given gap size than the interface loss computed here.
+\emph{Equation \ref{eqn:Tetalon} is not valid if the entrance and exit Si interfaces are AR coated.}  Equation \ref{eqn:SiAirMatrix} would have to be altered with the the AR coating's wavelength-dependent transmission, which might not have a simple analytical form.
 
@@
 This limitation will vary from instrument to instrument.
-Our instrument can achieve 10 times better accuracy and precision in double beam mode.
+Our instrument can achieve 10 times better accuracy and precision in double beam mode than in single beam mode.
 However, the slightly different polarization sensitivity in the different beams caused systematic jumps at the lamp/grating change-over wavelengths.
@@
 \emph{Can this technique be applied to imaging?}  The analysis performed in this article informs smart strategies for using IR imaging for gap detection in bonded Si optics.
-First, Equation \ref{eqFP} shows that short wavelengths have the most discerning power for detecting small gaps.
-So an IR imaging approach should use the shortest possible wavelength without getting close to the Si absorption cutoff.
+In IR images (Figure \ref{figVS2021_IR_image}), interfacial gaps appear as dark patches.
+These dark patches are darker at shorter wavelengths as evinced in Equation \ref{eqFP}, which shows that observations at short wavelengths have the most sensitivity to small gaps.
+Therefore an IR imaging approach should use the shortest possible wavelength without getting too close to the Si absorption cutoff.
 This work recommends a 1250$-$1300 nm filter.
-Images can then be normalized to the transmission through a DSP sample under identical, nearly collimated lighting conditions.
-Lastly, any deficits in flux of a bonded sample can be converted to a gap size $d$, using Equation \ref{eqFP}.
+Table \ref{tbl_FB1250} lists the predicted integrated transmission through a commercially available narrow bandpass filter, FB1250-10, which has center wavelength 1250 nm and full width at half maximum 10 nm.
+We calculated the integrated flux by integrating the product of the etalon transmission (Equation \ref{eqn:Tetalon}) with the FB1250-10 filter curve downloaded from the Thorlabs website\footnote{\url{www.thorlabs.com}}.
+The filter curve and example gap spectra are shown in Figure \ref{figFB1250-10}.
+
+\begin{figure}[htbp]
+\includegraphics[width=0.8\columnwidth]{figs/F1250_10_filter.pdf}
+\caption{\label{figFB1250-10} Filter curve and gap spectra.
+The Thorlabs FB1250-10 filter transmission curve is shown as the dray shaded region.
+We integrate the gap spectra with this filter curve to relate the deficit of flux in IR imaging with the axial extent of interfacial gaps.}
+\end{figure}
+
+\input{tbls/tbl_FB1250.tex}
+
+Figure \ref{figFB1250-10_integ} shows the integrated transmission through the FB1250-10 filter for small gaps ($0 < d(\mathrm{nm}) < 50$).
+Table \ref{tbl_FB1250} and Figure \ref{figFB1250-10_integ} can be used as a look-up table for converting between the dark patches observed in IR imaging and the axial extent of the interface gap, if the FB1250-10 filter is used during IR imaging.
+IR images must be normalized to the transmission through a DSP sample under identical, nearly collimated lighting conditions.
+Any deficits in flux of a bonded sample can be converted to a gap size $d$, using Table \ref{tbl_FB1250}, if the FB1250-10 filter is used.
 IR imaging will lack the spectral information available in the spectra shown in this article.
-The imaging will require exceptional quality of flat-fielding or dithering of the images to reach the precision achieved here with relative ease.
+The curvature of the spectra provides assurance of the physical origin of the dark patches seen in IR imaging.
+IR imaging will require exceptional quality of flat-fielding or dithering of the images to reach the precision achieved here with relative ease.
+
+\begin{figure}[htbp]
+\includegraphics[width=0.8\columnwidth]{figs/FB1250-10_integ_trans.pdf}
+\caption{\label{figFB1250-10_integ} Integrated transmission through the FB1250-10 filter as a function of interface gap size.
+The 99.8\% accuracy achieved in this article is shown as the horizontal dashed line.
+Imagers that can achieve 5\% photometric accuracy can detect 40 nm gaps.
+This figure assumes uncoated Si substrates.}
+\end{figure}
 
@@
 Armed with this technique, optical fabricators have a new diagnostic tool to assess the quality of Si$-$Si bonds.
+
+Funding information should be listed in a separate block preceding any acknowledgments. List just the funding agencies and any associated grants or project numbers, as shown in the example below:
+
+MGS acknowledges support from Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.  Portions of this research were supported by NASA ACT grant NNX12AC31G,
+
+The authors thank Amanda Turbyfill, Mike Pavel, and Carleen Boyer for assistance with IR imaging and specroscopy.
+MGS thanks the organizers and participants of the \emph{Astro Data Hack Week} (2014), especially Dan Foreman-Mackey for sharing tutorials and code for implementing the Gaussian process technique.
+This research has made use of NASA's Astrophysics Data System.
+Many of the figures in the paper were made with Matplotlib \citep{Hunter:2007} in IPython Notebooks \citep{PER-GRA:2007}.
+The Notebooks with complete version history are available on GitHub\footnote{\url{https://github.com/Echelle/AO_bonding_paper}}.
 

### 2015-04-21 AO_bondedSi.tex
add the new figures, and tables about the integrating a filter curve with the Si gap transmission.
incorporate most of DTJ comments from his April 19, 2015 email

@@
 Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 75 to 100 mm diameter clear aperture.
-We assume that, if enough pressure is placed on the substrates during bonding, then the surfaces will come in contact and zipper up.
+If enough pressure is placed on the substrates and if the bond strength is high enough relative to the stiffness of one or the other component, however, the surfaces will conform and bond.
 Subsequent annealing will strengthen the bond.
@@
 Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
-Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.
+Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image them in the infrared.
 Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect.
@@
 A recent generation of stock near-infrared spectrophotometers like the Cary5000 from Agilent offer $0.2\;\%$ precision and good repeatability.
-We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot to measure axial gaps down to a few tens of nanometers in size.
+We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot etalon to measure axial gaps down to a few tens of nanometers in size.
 In order to verify the applicability and accuracy of the technique, we have used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
@@
 The refractive index of silicon is one of the highest of any commonly used optical material.
-Figure \ref{figSiIndexFinesse} shows the refractive index at room temperature \cite{2006SPIE.6273E..77F}, which we assume through this article.
+Figure \ref{figSiIndexFinesse} shows the refractive index at room temperature \cite{2006SPIE.6273E..77F} that we assume through this article.
 The Fresnel reflection is about $30\;\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
@@
 We measured the ratio of transmissions of a 0.5 mm thick double side polished Si wafer and a 3.0 mm Si puck.
-Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness.
-Specifically, double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level long-ward of $1250\;$nm.
+Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness: double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level long-ward of $1250\;$nm.
 The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
@@
 \includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
-\caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of wavelength for various gap sizes in bonded Si optics.
+\caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of vacuum wavelength for various gap sizes in bonded Si optics.
 \emph{Top panel}: Absolute transmission through a Si substrate with no gap (top denim-blue solid curve), and small gaps of axial extent 20 nm (green dashed curve), 35 nm (red dash-dotted curve), and 60 nm (purple dotted curve).
@@
 
-We evaluated our metrology method by embedding gaps of known sizes in the interface between two bonded Si substrates and measuring the effects of these gaps on IR transmission.
+We evaluated our metrology method experimentally by embedding gaps of known sizes in the interface between two bonded Si substrates and measuring the effects of these gaps on IR transmission.
 The substrate thicknesses ranged from 0.5 to 3.3 mm.
@@
 To fabricate our test samples, we used photoresist lithography to pattern Si substrates, and we bored holes with inductively coupled plasma etching.
-We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 $\mu$m/minute to produce gaps with small and large depths.
-Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.
+We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 17 nm/second to produce gaps with small and large depths.
+Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depths, and Dektak stylus profilometry for the large depths.
 
 The fill factor is the pattern area covered by gaps relative to the total pattern area.
-We achieved gap sizes of $14-95\;$nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
+We achieved gap sizes of $49-4000\;$nm on four substrates with three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.
 We did not verify the delivered fill factor, but since high precision lithography is very reliable, we assign no uncertainty to the difference between the delivered fill factor and the designed fill factor.
@@
 We soaked the wafers in DI water then dried with N$_2$.
+The patterned Si substrate is bonded to a blank (unpatterned) Si substrate.
 We pressed the Si substrates together from the center to the outside.
@@
 Mosaic-stitching errors are also perceptible.
-It was easy to detect IR bubbles despite detector artifacts since the bubbles translate with the sample when we dither its position.
+It was easy to detect IR bubbles despite detector artifacts since the bubbles translate with the sample when we dither its position, while camera defects remain stationary.
 
 We also used IR imaging to inspect VG09-12 and VG15-13.
-We detected IR bubbles in all 3 of these bonded Si samples.
+We detected IR bubbles in all bonded Si samples.
 The observable bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.
@@
 The IR bubbles had various morphologies.
-The structure within some of the images of IR bubbles showed Newton's rings, exposing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
+The structure within some of the images of IR bubbles showed Newton's rings, revealing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
 As expected, the unbonded DSP wafers show no such IR bubbles.
@@
 The bulk morphologies were clear regardless of the detector response.
-When IR imaging was available, it was leveraged to target IR spectroscopy in locations demonstrating presence or absense of IR bubbles, as desired.
+When IR imaging was available, it was used to target IR spectroscopy toward locations displaying the presence or absense of IR bubbles, as desired.
 In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.
@@
 The coarsest mesh, C, shows substructure, while the medium and fine meshes are indistinguishable at the low resolution of the image.
-The three dark spots are gaps at the interface of the direct bonded Si samples.}
+The three dark spots are bubbles at the interface of the direct bonded Si samples.}
 \end{figure}
@@
 Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
-The tool has a double monochromator, which reduces scattered light.
-The vendor reported linearity exceeds 40 dB.
+The measurement beam is about 1 mm $\times$ 10 mm The tool has a double monochromator, which reduces scattered light.
 
@@
 We experimented with single front and double modes.
-In principle double beam mode is even more precise than single beam mode.
-In practice we found that souble beam mode showed evidence for lamp-jump artifacts attributable to polarization differences between the two arms.
-So in this article we focus our results to those obtained in single beam mode.
+In principle double beam mode is 10 times more precise than single beam mode.
+In practice, we found that double beam mode showed evidence for artifacts attributable to polarization differences between the two arms.
+In this article, we therefore restrict the discussion to results obtained in single beam mode.
 In both single and double mode, the spectrometer takes a baseline scan with no sample in the measurement holder.
@@
 This maximum gap limitation brings our technique to the point where it overlaps in capability with other techniques since the axial extents of such large gaps would be easily measurable by other means, like IR imaging and fringe-counting.
-
-We also verified the linearity of the detector response my measuring the transmission through stacks of filters.
-We constrained the linearity to at least $>24$ dB.
-The vendor reports linearity exceeding 40 dB, which is consistent with our measurement.
 
@@
 No scaling has been applied to the curves in Figure \ref{figVG03full}.
-Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field scanned by the Cary 5000.
+Figure \ref{figVG03part} shows a fit for a region of the same bonded pair where a nominally $4000\;$nm deep, 2 mm wide trench crosses the field (about 1 mm $\times$ 10 mm) scanned by the Cary 5000.
 Analysis of this transmission curve gave a best fit depth of $4094 \pm 4\;$nm and a pattern fill factor of $4.6 \pm 0.1\;\%$, both consistent with the part geometry.
@@
 This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
-We chose a 2 mm step size, which is slightly larger than the $\sim1$ mm beam width.
+We chose a 2 mm step size, which is slightly larger than the $\sim1$ mm beam width, but less than the 10 mm beam height.
 We sampled 50 mm across the mesh pattern and into the off-mesh pattern (Figure \ref{figVG0912_STA_illus}).
 The large number of consecutive measurements prevented the requisite near-contemporaneous baseline measurements for normalization.
-We fit a line to before-and-after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
+We linearly interpolated between before and after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
 We measured over a wavelength range from 1250 to 1775 nm.
@@
 The measurements are consistent with the predicted transmission, shown as the gray band.
-The off-mesh regions demonstrate near-perfect bonding.\label{figVG12}}
+The off-mesh regions demonstrate near-perfect bonding.\label{figVG12sta}}
 \end{figure}
@@
 The table lists the median sample and 68\% confidence intervals.
-The results from the scans shown in Figure \ref{figVG12} further support our conclusion.
+The results from the scans shown in Figure \ref{figVG12sta} further support our conclusion.
 
@@
 This spectrum is consistent with no gap.
-We applied our MCMC formalism to further constrain the range of gap sizes consistent with the spectrum.
-Table \ref{tbl_DerivedGapSizes} summarizes the sizes of gaps.
+We applied our MCMC formalism to both the unbonded DSP wafer and the VG09-12 off-mesh region.
+Table \ref{tbl_DerivedGapSizes} summarizes the formal derivation of gap sizes and fill factors for both the double-side polished puck and the off-mesh region of the VG09-12 bonded pair.
 The well-calibrated DSP noise spectrum rules out (at $>3\sigma$) gap sizes larger than about 9.0 nm over $>80$\% of the measurement area.
@@
 The measured spectrum in this region shows no appreciable deficit of transmission from a DSP reference sample.
+The blue lines represent individual draws from the Gaussian process model.
 \label{figVG12}}
@@
 
-Figure \ref{VG0912_0gap_corner} shows that there is a degeneracy between the axial extent of a gap and its fractional fill factor over the measurement beam; a larger gap can be hidden if it covers only a small fraction of the measured area.
+The actual measurement we are making is of the change in transmission through the sample as a function of wavelength.
+This change is larger for larger gaps but becomes diluted when the fill factor is less than unity.
+This dilution leads to a degeneracy between the axial extent of the gap and its fractional fill factor over the measurement beam,  one that is broken by the harder-to-measure curvature of the transmission spectrum.
+Figure \ref{VG0912_0gap_corner} shows illustrates the degeneracy.
 The red dashed line shows the normalized transmission spectrum of a 14 nm gap over 100\% of the measurement area.
@@
 The experiments and measurements in Section \ref{secResults} demonstrate that we can detect and measure the axial extent of gaps down to 14 nm.
-The technique is \emph{non-destructive} and can be implemented with an off-the-shelf commercial product, namely the Cary 5000 or equivalent, or any existing high precision spectrograph.
+The technique is \emph{non-destructive} and can be implemented with any high precision near-IR spectrograph.
 The spectrograph need not be high resolution-- our 5 nm spectral resolution was not a limiting factor.
@@
 For example, if there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
-Since the gap sizes are small and the interface takes place within a subregion of constant gap size, the full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
+Since the gap sizes for which we intend to use the model are small, and the interface takes place within a subregion of constant gap size, the full solution becomes an areal integral of Equation \ref{eqFP} normalized by the total measurement area.
 Solving for the relative distributions of gaps is highly degenerate, in the same way as we saw in the 2-parameter corner plots of gap depth and fill factor.
@@
 This work recommends a 1250$-$1300 nm filter.
-Table \ref{tbl_FB1250} lists the predicted integrated transmission through a commercially available narrow bandpass filter, FB1250-10, which has center wavelength 1250 nm and full width at half maximum 10 nm.
+Table \ref{tbl_FB1250} lists the predicted integrated transmission through a commercially available narrow bandpass filter, for example the FB1250-10 from Thorlabs, which has center wavelength 1250 nm and full width at half maximum 10 nm.
 We calculated the integrated flux by integrating the product of the etalon transmission (Equation \ref{eqn:Tetalon}) with the FB1250-10 filter curve downloaded from the Thorlabs website\footnote{\url{www.thorlabs.com}}.
@@
 \caption{\label{figFB1250-10} Filter curve and gap spectra.
-The Thorlabs FB1250-10 filter transmission curve is shown as the dray shaded region.
+The Thorlabs FB1250-10 filter transmission curve is shown as the gray shaded region.
 We integrate the gap spectra with this filter curve to relate the deficit of flux in IR imaging with the axial extent of interfacial gaps.}
@@
 
-Funding information should be listed in a separate block preceding any acknowledgments. List just the funding agencies and any associated grants or project numbers, as shown in the example below:
-
-MGS acknowledges support from Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.  Portions of this research were supported by NASA ACT grant NNX12AC31G,
-
+MGS acknowledges support from Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.
+Portions of this research were supported by NASA ACT grant NNX12AC31G.
 The authors thank Amanda Turbyfill, Mike Pavel, and Carleen Boyer for assistance with IR imaging and specroscopy.
@@
 $T_g+R_g=1$.
-So the incoherent scattering and transfer matrices for the gap are:
+The incoherent scattering and transfer matrices for the gap are then:
 
@@
 \end{eqnarray}
-For clarity I will drop the subscripts from $n_{Si}$, since we have already set $n_{air}=1$ and there are no other dielectric interfaces to think about.
-So the scattering and transfer matrices for the air-Si Fresnel boundary are:
+For clarity we will drop the subscripts from $n_{Si}$, since we have already set $n_{air}=1$ and there are no other dielectric interfaces to think about.
+The scattering and transfer matrices for the air-Si Fresnel boundary are then:
 
@@
 
-Compare equations \ref{eqn:FPmatTrans} and \ref{eq:FabPerot}.
-The revised transmission has picked up a few factors of 2 and $n$.
-While we are at it, let's compute the matrix for the scenario with no intermediate gap: there are simply two Fresnel interfaces with which light interacts incoherently.
-This scenario is the model for a single DSP reference sample.
-The revised matrix multiplication is simply $\boldsymbol{M_{DSP}}=\boldsymbol{M_n}\boldsymbol{M_n}$.
+The matrix technique recovers the usual relation for transmission through a double-side polished dielectric (Equation \ref{eqnAbsDSPtrans}).
+For this case, the matrix multiplication is simply $\boldsymbol{M_{DSP}}=\boldsymbol{M_n}\boldsymbol{M_n}$.
 Taking the inverse of the $M_{22}$ element, we find:

### 2015-04-22 AO_bondedSi.tex
Elucidate reason we can treat the Si blocks incoherently, frame the statement in the positive.  Closes #31

@@
 A pair of bonded silicon wafers or pucks with a small gap forms three coupled cavities with the air gap as the the central cavity.
-The outer cavities, those within the Si, however, were not produced with coherence in mind.
-Their sides are not necessarily flat and parallel to the required level and the specifics of their deviation from perfect etalons are usually not known to the level necessary to treat them as coherent structures in the analysis.
+The axial extent of the silicon cavities are large enough---and typical measurement spectral resolution low enough---that the the Si cavities reflect and transmit light incoherently at their entrance and exit interfaces.
 The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps.

### 2015-09-21 AO_bondedSi.tex
use the two-column format.  near final version
futz with the column formatting, and prepare content for tarball archive
respond to referee- typos and extraneous line

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %\documentclass[osajnl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1} %% use 11pt for Applied Optics
@@
 Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
-Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image them in the infrared.
+Since silicon is opaque in the optical range, the most straightforward technique for detecting defects in the bond is to image them in the infrared.
 Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect.
@@
 A recent generation of stock near-infrared spectrophotometers like the Cary5000 from Agilent offer $0.2\;\%$ precision and good repeatability.
-We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-Perot etalon to measure axial gaps down to a few tens of nanometers in size.
+We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-P\'{e}rot etalon to measure axial gaps down to a few tens of nanometers in size.
 In order to verify the applicability and accuracy of the technique, we have used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
@@
 For finite-sized bubbles, measurement of the axial dimension helps us to assess the effects of high-temperature annealing \cite{Horn2009, Masteika2014} on removal of hydrogen bubbles.
-A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to resolve spatially, as one might expect if there are many very small contaminants present along the interface.
+A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to be resolved spatially, as one might expect if there are many very small contaminants present along the interface.
 Our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
@@
 
-\section{Si$-$Si Fabry P\`{e}rot measurement theory}
+\section{Si$-$Si Fabry P\'{e}rot measurement theory}
 \label{secTheory}
 
-\subsection{The Si$-$Si interface gap is a low finesse Fabry-P\`{e}rot etalon}
+\subsection{The Si$-$Si interface gap is a low finesse Fabry-P\'{e}rot etalon}
 
@@
 
-We can treat a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.
+We can treat a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\'{e}rot etalon\cite{2007fuph.book.....S}.
 The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
 For the narrow air gaps we consider here and for the modest reflectivity of a Si-vacuum interface, most gaps at Si$-$Si bonds will form effective etalons over most of their area.
-The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
+The transmission through a Fabry-P\'{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
 We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
@@
 
-We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on reflectivity.
+We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\'{e}rot etalon's dependence on reflectivity.
 The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength.
@@
 The top line in the upper panel of Figure \ref{figAbsoluteTrans} shows the value derived from Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
-If the Si substrate thickness is less than the coherence length for the given spectral bandwidth, and if its surface roughness is much smaller than the wavelength, our assumptions break down and the multiple reflections would interfere coherently, and the outer Si cavities would behave as a Fabry-P\`erot etalon.
+If the Si substrate thickness is less than the coherence length for the given spectral bandwidth, and if its surface roughness is much smaller than the wavelength, our assumptions break down and the multiple reflections would interfere coherently, and the outer Si cavities would behave as a Fabry-P\'{e}rot etalon.
 We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.
@@
 \emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0-3500 nm.
-The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\`erot etalon fringes.}
+The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\'{e}rot etalon fringes.}
 \end{figure}
@@
 There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
-Our ultimate goal is to bond optics with thicknesses as large as 30 mm.
 
@@
 We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
-Specifically, Fabry-P\`erot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
+Specifically, Fabry-P\'{e}rot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
 This maximum gap limitation brings our technique to the point where it overlaps in capability with other techniques since the axial extents of such large gaps would be easily measurable by other means, like IR imaging and fringe-counting.
@@
 The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physical measurement (Table \ref{tbl_experiments}).
-The spectrum in this region shows characteristic Fabry-P\`erot fringes of constructive and destructive interference.
+The spectrum in this region shows characteristic Fabry-P\'{e}rot fringes of constructive and destructive interference.
 The only tunable parameters of the model are the gap axial extent, $d$ and fill factor $f$.
@@
 The matrix for the silicon air gap is calculated in the following way.
-We treat the air gap as a Fabry-P\`erot etalon.
-We do not need to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\`erot etalon model for the gap:
+We treat the air gap as a Fabry-P\'{e}rot etalon.
+We do not need to consider the microscopic coherent interactions with the etalon transmission, all of that information is encapsulated in these equations for a Fabry-P\'{e}rot etalon model for the gap:
 
@@
 with $\lambda$ the vacuum wavelength, $R$ the Fresnel reflection of silicon, and $F$ the coefficient of finesse.
-The coefficient of finesse $F$ and the phase $\delta$ are the two parameters of the Fabry-P\`erot etalon.
+The coefficient of finesse $F$ and the phase $\delta$ are the two parameters of the Fabry-P\'{e}rot etalon.
 The coefficient of Finesse encapsulates the Fresnel reflection and depends only on the Si refractive index which has only a small wavelength (and temperature) dependence.

### 2015-09-21 AO_bondedSi.tex
reduce abstract length by about 40 percent

@@
 Even small (35 nm) gaps reduce transmission through a direct bonded Si compound optic by 4\% at $\lambda = 1.25 \; \mu$m at normal incidence.
-Understanding the optical effects of such gaps and evaluation of various methods for preventing or eliminating them demands that we have a method for measuring not only the two-dimensional extent of the gaps but also their axial extent.
-Existing methods for studying gaps provide only poor measures of this axial extent.
-We describe a bond inspection method that makes use of precision slit spectroscopy.
-Using spectroscopy, we detect and measure gaps as small as 14 nm by measuring transmission as a function of wavelength.
-Our analysis involves modeling multiple incoherent reflections within the substrate interfaces with a wave transfer matrix, modified for intensities rather than complex amplitudes.
-We infer the gap axial extent and fill factor with Monte Carlo Markov Chain (MCMC) and a flexible Gaussian process noise model.
-We describe the measurement and analysis techniques and demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
+We describe a bond inspection method that makes use of precision slit spectroscopy to detect and measure gaps as small as 14 nm.
+Our method compares low finesse Fabry-P\'{e}rot models to high precision measurements of transmission as a function of wavelength.
+We demonstrate the validity of the approach by measuring bond gaps of known depths produced by microlithography.
 \end{abstract}

### 2015-09-21 AO_bondedSi.tex
Remove some paragraphs from Section 4, make some prose more concise.

@@
 
-\section{Laboratory demonstration: Embedding gaps of known sizes}
+\section{Embedding gaps of known axial and lateral extent}
 
@@
 
-We used the test pieces with known gaps to validate our measurement and analysis techniques.
+We used the test pieces with known gap sizes to validate our measurement and analysis techniques.
 We took transmission spectra of the bonded Si wafers and pucks, and the double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
 Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
-The measurement beam is about 1 mm $\times$ 10 mm The tool has a double monochromator, which reduces scattered light.
 
-The Cary 5000 has four operation modes: single front, single back, double, and double reverse.
-Double mode takes a reference spectrum simultaneously with the target spectrum.
-We experimented with single front and double modes.
-In principle double beam mode is 10 times more precise than single beam mode.
-In practice, we found that double beam mode showed evidence for artifacts attributable to polarization differences between the two arms.
-In this article, we therefore restrict the discussion to results obtained in single beam mode.
-In both single and double mode, the spectrometer takes a baseline scan with no sample in the measurement holder.
-We automatically divide sample spectra by this baseline spectrum.
-In single beam mode, we must repeat the baseline measurements about every 10-30 minutes for post-process division.
+The measurement uncertainty in our experiments can be characterized by both its \emph{bias} and \emph{variance}.
+The bias is equivalent to a mean-level drift, which may be attributable to monochromator lamp luminosity drift.
+The variance is the random sample-to-sample measurement residual remaining once the bias term is removed.
+In our experiments, the bias term has an amplitude of 0.2\%, whereas the variance has an amplitude of 0.02\%.
+Our measurement uncertainty is dominated by bias, not variance.
 
-We estimated the uncertainty and measurement repeatability experimentally.
-We computed the RMS error and measurement repeatability in a single measurement by computing the standard deviation of a spectrum taken immediately after a baseline scan with no sample present.
-We found that the mean value in single mode drifted by about 0.2\% over several measurements.
-The standard deviation of the featureless spectrum was typically 0.02\% from point to point, with some portions of the spectrum demonstrating increased variance.
-Table \ref{tabCary5000pars} summarizes the amplitudes of the drifts and individual measurement errors.
-\emph{Correlated errors dominate the measurement uncertainty of our transmission spectra.}  The measurement uncertainty correlations are apparent in Figure \ref{figSiAbsorbfig}.
-In some parts of the spectrum, the mean drift is about ten times larger than the local standard deviation of the spectrum.
-
-The spectral sampling ranged from 2.0 to 10.0 nm.
-We set the spectrometer slit width to produce a spectral resolution of 5.0 nm.
-We expect smoothly varying spectral features for all conceivable interface gaps in Si bonded wafers.
-Specifically, Fabry-P\'{e}rot fringes attributable to interfacial gaps in bonded Si will be well-sampled with 5 nm spectral resolution, so long as the interface gap is less than $\sim$100 $\mu$m.
-This maximum gap limitation brings our technique to the point where it overlaps in capability with other techniques since the axial extents of such large gaps would be easily measurable by other means, like IR imaging and fringe-counting.
+%New prose here:
+At this stage, \emph{any} regression technique can be employed to compare the model to data.
+Least squares regression, and other popular methods can be applied, for example.
+However, since least squares regression assumes zero bias term, we elected to apply a Gaussian process regression technique that can handle biased measurements.
+Given the relatively small bias amplitude we measured, least squares regression and Gaussian process regression would likely yield similar solutions for the gap size, but least squares regression would drastically underestimate the uncertainty associated with the gap size\citep{DFMgp}.
 

### 2015-09-23 AO_bondedSi.tex
be concise.  also respond to comment about missing Tdsp

@@
 We normalize the absolute transmission by the transmission of a double-sided Si structure with no gap to isolate the effect of the gap.
-We call this normalized transmission the gap transmission $T_{g}$.
+We call this \emph{normalized} transmission the gap transmission $T_{g}$.
 
@@
 
-In the limit $\lim_{d \to 0} T_g \rightarrow T_{DSP} $ and the gap approaches 100\% transmission.
+In the limit $\lim_{d \to 0} T_g \rightarrow \frac{n^2 + 1}{n^2 + 1} = 1 $ and the gap approaches 100\% transmission.
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
@@
 
-\subsection{Spatially resolved measurements with Sample Transport Accessory}
-We acquired spatially resolved measurements of the pair of bonded parts VG09-12, (Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns}) with the Cary 5000 sample transport accessory \footnote{\url{http://www.chem.agilent.com/en-US/products-services/instruments-systems/molecular-spectroscopy/sample-transport-accessory/}}.
-This accessory translates the sample through the measurement beam and takes a transmission spectrum at each position.
-We chose a 2 mm step size, which is slightly larger than the $\sim1$ mm beam width, but less than the 10 mm beam height.
-We sampled 50 mm across the mesh pattern and into the off-mesh pattern (Figure \ref{figVG0912_STA_illus}).
-The large number of consecutive measurements prevented the requisite near-contemporaneous baseline measurements for normalization.
-We linearly interpolated between before and after baseline measurements to infer the mean-level drift, which we divided out of each spectrum.
-We measured over a wavelength range from 1250 to 1775 nm.
+\subsection{Spatially resolved measurements across the face of a sample}
+We acquired spatially resolved measurements of the pair of bonded parts VG09-12, (Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns}).
+We translated the sample through the measurement beam and measured a transmission spectrum at 25 unique adjacent positions.
+We chose a 2 mm step size, which is comparable to the beam width.
+We sampled across the mesh pattern and into the off-mesh pattern (Figure \ref{figVG0912_STA_illus}).
 

### 2015-09-23 AO_bondedSi.tex
remove Figure 1 and discussion of temperature dependence

@@
 Lehman et al.
-1989 and Goesele et al.
+1989 and G{\"o}esele et al.
 \cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
@@
 Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon is one of the highest of any commonly used optical material.
-Figure \ref{figSiIndexFinesse} shows the refractive index at room temperature \cite{2006SPIE.6273E..77F} that we assume through this article.
+The refractive index of silicon is one of the highest of any commonly used optical material, at about 3.5 at a wavelength $\lambda = 1.2 \;\mu$m at room temperature \cite{2006SPIE.6273E..77F}.
 The Fresnel reflection is about $30\;\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
-
-\begin{figure}[htbp]
-\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}
-\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.
-The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.}
-\end{figure}
 
@@
 The transmission through a Fabry-P\'{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.
+We compute the Si$-$vacuum interface reflectivity, which is a modest function of wavelength, through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, where $T$ is temperature.
+We adopt room temperature values through this article.
 We drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
@@
 We chose a 2 mm step size, which is comparable to the beam width.
-We sampled across the mesh pattern and into the off-mesh pattern (Figure \ref{figVG0912_STA_illus}).
+We sampled across the mesh pattern and into the off-mesh pattern of sample VG09-12 (Figure \ref{figVG0912_STA_illus}).
 

### 2015-09-23 AO_bondedSi.tex
having problems with latex-diff, something about bib file

@@
 The transmission through a Fabry-P\'{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We compute the Si$-$vacuum interface reflectivity, which is a modest function of wavelength, through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, where $T$ is temperature.
+We compute the Si$-$vacuum interface reflectivity, which is a modest function of wavelength, through the refractive index of Silicon, $n(\lambda, T)$, where $T$ is temperature \cite{2006SPIE.6273E..77F}.
 We adopt room temperature values through this article.
@@
 We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\'{e}rot etalon's dependence on reflectivity.
-The right panel of Figure \ref{figSiIndexFinesse} shows the coefficient of finesse as a function of wavelength.
+We calculate a coefficient of finesse in the range 2.5$-$2.6, depending on wavelength.
 In this work, we assume transmission is at normal incidence and the refractive index of the gap is 1.0.
@@
 However, since least squares regression assumes zero bias term, we elected to apply a Gaussian process regression technique that can handle biased measurements.
-Given the relatively small bias amplitude we measured, least squares regression and Gaussian process regression would likely yield similar solutions for the gap size, but least squares regression would drastically underestimate the uncertainty associated with the gap size\citep{DFMgp}.
+Given the relatively small bias amplitude we measured, least squares regression and Gaussian process regression would likely yield similar solutions for the gap size, but least squares regression would drastically underestimate the uncertainty associated with the gap size\cite{DFMgp}.
 
@@
 This research has made use of NASA's Astrophysics Data System.
-Many of the figures in the paper were made with Matplotlib \citep{Hunter:2007} in IPython Notebooks \citep{PER-GRA:2007}.
+Many of the figures in the paper were made with Matplotlib \cite{Hunter:2007} in IPython Notebooks \cite{PER-GRA:2007}.
 The Notebooks with complete version history are available on GitHub\footnote{\url{https://github.com/Echelle/AO_bonding_paper}}.

### 2015-09-25 AO_bondedSi.tex
closes #35, fix address of JPL to be consistent with utexas
seriously greening (i.e. cutting) the fat from this manuscript.

@@
 \author{Victor White}
-\affiliation{NASA Jet Propulsion Laboratory, Pasadena, CA}
+\affiliation{NASA Jet Propulsion Laboratory, Pasadena, CA, 91109, USA}
 
@@
 Si has a high refractive index ($3.55-3.45$ from $\lambda = 1150-2500\;$nm at room temperature, \cite{2006SPIE.6273E..77F}) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
-It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
+It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18$-$30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
 There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
@@
 One advantage of this direct Si$-$Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
-A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.
+A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because close lossless bonds can eliminate concerns about optical ghosts.
 The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
@@
 If Si$-$Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than the concatenated losses of two antireflection coated surfaces over the operating wavelength (typically $< 4$\%).
-For Si-air gaps, where the full Fresnel loss is $\sim30\%$ per interface this requirement translates to physical gaps or bubbles having an axial extent of $ < 35\;$nm.
+For Si$-$air gaps, the full Fresnel loss is $\sim30\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
+So a 4\% transmission loss specification translates to physical gaps or bubbles having an axial extent of merely $ < 35\;$nm.
 If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond.
@@
 In optics, however, both of the bonding partners will likely be thicker and more mechanically rigid.
-When bonding two thick substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
-Most of our thick substrates are optically polished to less than 60 nm peak to valley over the 75 to 100 mm diameter clear aperture.
+When bonding two thick substrates, small differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
 If enough pressure is placed on the substrates and if the bond strength is high enough relative to the stiffness of one or the other component, however, the surfaces will conform and bond.
@@
 Our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
-A $2\;\%$ loss at $1.25\;\mu$m corresponds to a gap of $>25\;$nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
+A $2\%$ loss at $1.25\;\mu$m corresponds to a gap of $>25\;$nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
 We therefore need to be able to measure transmission as a function of wavelength to a fraction of a percent, including all systematic errors, for devices operating near $1\;\mu$m and to a somewhat looser standard for devices operating at longer wavelengths.
 
-\section{Si$-$Si Fabry P\'{e}rot measurement theory}
+\section{Theory: Modeling the Si$-$Si interface gap as a low finesse Fabry-P\'{e}rot etalon}
 \label{secTheory}
 
-\subsection{The Si$-$Si interface gap is a low finesse Fabry-P\'{e}rot etalon}
-
-Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon is one of the highest of any commonly used optical material, at about 3.5 at a wavelength $\lambda = 1.2 \;\mu$m at room temperature \cite{2006SPIE.6273E..77F}.
-The Fresnel reflection is about $30\;\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
-
-We can treat a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\'{e}rot etalon\cite{2007fuph.book.....S}.
+Our modeling technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
+The refractive index of silicon is one of the highest of any commonly used optical material\cite{2006SPIE.6273E..77F}.
+We model a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\'{e}rot etalon \cite{2007fuph.book.....S}.
 The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
-For the narrow air gaps we consider here and for the modest reflectivity of a Si-vacuum interface, most gaps at Si$-$Si bonds will form effective etalons over most of their area.
+For the narrow air gaps we consider here and for the relatively modest reflectivity of a Si$-$vacuum interface, most gaps at Si$-$Si bonds will form effective etalons over most of their area.
 The transmission through a Fabry-P\'{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.
-We compute the Si$-$vacuum interface reflectivity, which is a modest function of wavelength, through the refractive index of Silicon, $n(\lambda, T)$, where $T$ is temperature \cite{2006SPIE.6273E..77F}.
-We adopt room temperature values through this article.
-We drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.
-The computation is simply Fresnel's law at normal incidence:
+The reflectivity is given by Fresnel's law at normal incidence:
 
@@
 We calculate a coefficient of finesse in the range 2.5$-$2.6, depending on wavelength.
+We assume room temperature for all calculations of refractive index.
 In this work, we assume transmission is at normal incidence and the refractive index of the gap is 1.0.
-Si absorbs negligibly longward of $\lambda$ = 1250 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses (Figure \ref{figSiAbsorbfig}).
-We measured the ratio of transmissions of a 0.5 mm thick double side polished Si wafer and a 3.0 mm Si puck.
-Figure \ref{figSiAbsorbfig} verifies a key assumption that we make when comparing the transmission of scrap Si of different provenance and thickness: double side polished 3 mm thick Si pucks and 0.5 mm thick Si wafers have indistinguishable transmission at the $<0.2$\% level long-ward of $1250\;$nm.
-The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1250 nm.
-\cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as a reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.
-An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.
-In the next subsection we work out the wavelength dependent transmission for the air etalon\footnote{A note on nomenclature- we use the term \emph{gap} for the physical structure, and \emph{etalon} for the model of that structure.}
+Si absorbs negligibly longward of $\lambda$ = 1250 nm, which we verified experimentally by comparing the transmission of Si reference samples of different thicknesses (Figure \ref{figSiAbsorbfig}).
 
@@
 
-\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
-\label{secTheory}
 A pair of bonded silicon wafers or pucks with a small gap forms three coupled cavities with the air gap as the the central cavity.
@@
 In the air gap, however, the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
-Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.
-The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.
-We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K} to deal with the cavities within the thick Si substrates.
-Specifically we constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.
+We adapted coherent wave transfer techniques \cite{2007fuph.book.....S} for use with incoherent interactions \cite{2002ApOpt..41.3978K} to deal with the cavities within thick Si substrates.
 The details appear in Appendix \ref{sec:Append-IMRTMM}.
-We computed the transmission for two scenarios: first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.
+We computed analytic equations for total transmission (including multiple internal reflections) for two scenarios: first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.
 The DSP Si sample with no air gap has a transmission equal to:
@@
 
-which has an average value of about 53\%.
-Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.
-The top line in the upper panel of Figure \ref{figAbsoluteTrans} shows the value derived from Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
-If the Si substrate thickness is less than the coherence length for the given spectral bandwidth, and if its surface roughness is much smaller than the wavelength, our assumptions break down and the multiple reflections would interfere coherently, and the outer Si cavities would behave as a Fabry-P\'{e}rot etalon.
-We do not treat this case here since our focus is on thick substrates with small gaps between the interfaces.
+which has an average value of about 53\%.  The top line in the upper panel of Figure \ref{figAbsoluteTrans} shows the plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
 
@@
 The curves assume that multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.
-\emph{Middle panel}: The ``Gap Transmission'', defined as the Absolute transmission through a sample normalized by the absolute transmission through a gapless sample.
+\emph{Middle panel}: The ``Gap Transmission'', $T_g$.
 The gap sizes and colors are the same as in the top panel.
-\emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0-3500 nm.
+\emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0$-$3500 nm.
 The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\'{e}rot etalon fringes.}
@@
 
-Consider a pair of bonded Si samples with a gap thickness $d$.
 We derive the absolute transmission of bonded Si substrates with a gap in Equation \ref{eqn:Tetalon} in the Appendix.
-We normalize the absolute transmission by the transmission of a double-sided Si structure with no gap to isolate the effect of the gap.
-We call this \emph{normalized} transmission the gap transmission $T_{g}$.
+We divide the absolute transmission by $T_{DSP}$ to isolate the effect of the gap.
+We call this \emph{DSP-normalized} transmission the gap transmission $T_{g}$.
 
@@
 Figure \ref{figAbsoluteTrans} shows a plot of Equation \ref{eqFP} for gap sizes, $d$, of 0, 20, 35, 60, 350, and 3500 nm.
-The bottom two panels show the ``gap'' transmission based on Equation \ref{eqFP}, the value of the throughput with the losses from the outer Si cavities divided out.
+The bottom two panels show the gap transmission based on Equation \ref{eqFP}, the value of the throughput with the losses from the outer Si cavities divided out.
 For the very largest gap, large fringes appear as the cavity resonance of the air gap modulates the transmission.
@@
 For the smallest ($35-60$ nm) gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.
-It is the magnitude and wavelength dependence of this transmission spectrum that provides information about the axial extent of the gap.
+The magnitude and wavelength dependence of this transmission spectrum provide information about the axial extent of the gap.
 
-\section{Embedding gaps of known axial and lateral extent}
+\section{Fabricating and embedded gaps of known axial and lateral extent}
 
@@
 The substrate thicknesses ranged from 0.5 to 3.3 mm.
-We denote substrates with thicknesses larger than 1 mm as ``pucks'' and substrates with thickness less than 1.0 mm thick as ``wafers''.
-We distinguish the two classes of substrate thickness based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.
-There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.
 
@@
 To fabricate our test samples, we used photoresist lithography to pattern Si substrates, and we bored holes with inductively coupled plasma etching.
-We used two different gases with low and high etch rates, CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 17 nm/second to produce gaps with small and large depths.
 Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns} give information about the hole patterns and depths, as measured with Veeco NT9100 Optical Profiler for the small depths, and Dektak stylus profilometry for the large depths.

### 2015-09-25 AO_bondedSi.tex
Some major restructuing: Model, data, comparison of model to data.

@@
 
+We also define a mixture model in which the observed normalized gap spectrum $T_{obs}$ is composed of a sum of a spectrum through a gap of axial extent $d$, and a spectrum through a perfect bond:
+
+\begin{equation}
+	T_{mix} = f\;T_{g} + (1-f) \label{eqnMix}
+\end{equation}
+
+where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.
+$T_g$ is defined in Equation \ref{eqFP}.
+The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
+Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
+These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
+
 \section{Fabricating and embedded gaps of known axial and lateral extent}
@@
 For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
-The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.
-The stylus profilometry measured values were assigned an uncertainty of 5\%, our experience with other etched parts.
-Table \ref{tbl_experiments} lists part numbers and descriptions for our bonded substrates.
+The stylus profilometry measured values were assigned an uncertainty of 5\%, which accorded with our experience with other etched parts.
+Table \ref{tbl_experiments} lists descriptions for our bonded substrates.
 
@@
 The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
-We also measured the large scale surface flatness of the pucks.
-The Fisba2 interferometer used for these measurements has a 50 mm diameter beam at $\lambda=632.8\;$nm.
-We found a typical peak to valley surface flatness of 2 waves over the central 50 mm diameter.
+We also measured the large scale surface flatness of the pucks with an interferometer.
+We found a typical peak to valley surface flatness of 2 waves ($\lambda=632.8\;$nm) over the central 50 mm diameter.
 We prepared the substrates with standard cleaning procedures of solvents in a megasonic.
@@
 
-\subsection{IR bubble imaging and limitations}
-
 We first looked for large gaps in our bonded substrate pairs detectable as ``IR bubbles'' \cite{1992JEMat..21..669M}, with IR imaging.
-The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.
-We used a 50 mm F/2 lens, with a fiber optic white light illumination source.
 The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately $1.15-1.6\;\mu$m.
@@
 Figure \ref{figVS2021_IR_image} shows an IR image of bonded sample VS20-21.
-For this image, we coarsely flat-fielded the detector by observing a homogeneously illuminated white screen.
+For this image, we flat-fielded the detector by imaging and dividing out a homogeneously illuminated white screen.
 Despite the flat-fielding, detector non-uniformities and vignetting are perceptible in our image as vertical stripes.
 Mosaic-stitching errors are also perceptible.
-It was easy to detect IR bubbles despite detector artifacts since the bubbles translate with the sample when we dither its position, while camera defects remain stationary.
+The IR bubbles were conspicuous in the images despite detector artifacts.
 
@@
 The individual IR bubble areas were as small as a few square millimeters up to $400\;\mathrm{mm}^2$.
-The IR bubbles had various morphologies.
 The structure within some of the images of IR bubbles showed Newton's rings, revealing the presence of gaps larger than about $\lambda/2$ up to $\sim 15 \lambda$.
 As expected, the unbonded DSP wafers show no such IR bubbles.
-No attempt at quantitative measurement was made on the IR images due to evidence of a non-linear detector response.
-The bulk morphologies were clear regardless of the detector response.
+No attempt at quantitative measurement was made on the IR images due to evidence for a non-linear detector response.
 When IR imaging was available, it was used to target IR spectroscopy toward locations displaying the presence or absense of IR bubbles, as desired.
-In other cases when IR imaging was not available or ignored, spectra were taken at random positions on the wafer.
+In other cases when IR imaging was not available, spectra were taken at random positions on the wafer.
 
@@
 
-We used the test pieces with known gap sizes to validate our measurement and analysis techniques.
-We took transmission spectra of the bonded Si wafers and pucks, and the double side polished Si reference wafers using an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
+We measured transmission spectra through the bonded Si samples with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer.
 Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
 
-The measurement uncertainty in our experiments can be characterized by both its \emph{bias} and \emph{variance}.
+Measurement uncertainty can be characterized by both its \emph{bias} and \emph{variance} \cite{2013sdmm.book.....I}.
 The bias is equivalent to a mean-level drift, which may be attributable to monochromator lamp luminosity drift.
@@
 
-%New prose here:
-At this stage, \emph{any} regression technique can be employed to compare the model to data.
-Least squares regression, and other popular methods can be applied, for example.
-However, since least squares regression assumes zero bias term, we elected to apply a Gaussian process regression technique that can handle biased measurements.
-Given the relatively small bias amplitude we measured, least squares regression and Gaussian process regression would likely yield similar solutions for the gap size, but least squares regression would drastically underestimate the uncertainty associated with the gap size\cite{DFMgp}.
+\section{Comparing the Fabry-P\'{e}rot model to data}
+
+At this stage, \emph{any} regression technique can be employed to compare the Fabry-P model to data.
+Least squares regression and other popular methods can be applied easily.
+However, since least squares regression assumes zero bias, we elected to apply a Gaussian process regression technique that can handle biased measurements \cite{2013sdmm.book.....I,DFMgp,rasmussen2006gaussian}.
+Given the relatively small bias amplitude (0.2\%), least squares regression and Gaussian process regression might yield similar solutions for the gap size, but least squares regression would drastically underestimate the \emph{uncertainty} value associated with the derived value of gap size\cite{DFMgp}.
 
@@
 \label{secResults}
-
-The gaps in our test pairs should be a combination of the deliberately induced gaps and any inadvertent additional spaces caused by imperfections in the bonds.
-We infer the axial extent of gaps in our samples with etched holes from their measured transmission spectra.
-We use a Bayesian inference technique in our analysis.
-First we define a mixture model in which the observed normalized gap spectrum $T_{obs}$ is composed of a sum of a spectrum through a gap of axial extent $d$, and a spectrum through a perfect bond:
-
-\begin{equation}
-	T_{mix} = f\;T_{g} + (1-f) \label{eqnMix}
-\end{equation}
-
-where $f$ is the areal fill factor of the region exhibiting a finite gap size, and $1-f$ is therefore the areal fill factor of the perfectly bonded area.
-$T_g$ is defined in Equation \ref{eqFP}.
-The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
-Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
-These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
 

### 2015-09-25 AO_bondedSi.tex
end of first round of reviewer comments, a little more to do though

@@
 Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
-These synthetic imbedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secKnownGaps}).
+Synthetic embedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secResults}).
 
@@
 
-\section{Results of infrared spectroscopy of directly bonded Si}
-\label{secResults}
+\subsection{MCMC Gaussian process regression}
 
-Equipped with an observed spectrum and our mixture model, we constructed the likelihood function \cite{2013sdmm.book.....I}.
-The measurement errors are strongly correlated as discussed in Section \ref{sec_aboutErrors}.
-To handle the correlated measurement errors we use Gaussian process regression \cite{rasmussen2006gaussian, DFMgp}.
+We constructed the likelihood function \cite{2013sdmm.book.....I}.
 The Gaussian process takes into account the unknown but finite covariance structure by introducing a covariance matrix, $\boldsymbol{C}$ whose elements represent the correlations of each sample $i$ with every other sample $j$.
@@
 We experimented with different values for $\sigma_i$.
-When repeated double side polished reference sample measurements were available, we estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra.
+We estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra of DSP samples.
 We do not know $a$ nor $s$ and they are, in general, different for each measurement-- they are \emph{nuisance parameters}, and are ultimately marginalized out.
@@
 
-\subsection{Spectra of direct bonded Si with known gap sizes}
-\label{secKnownGaps}
+\section{Results of infrared spectroscopy of directly bonded Si}
+\label{secResults}
 
@@
 The conspicuous mesh pattern has a depth of $49\pm6\;$nm, as measured with an optical profiler, and global fill factor of 50\%.
-However, since the beam size is comparable to mesh grid size, the fill factor for any given measurement can vary from about 30\% to 75\%.
+However, since the beam size is comparable to the mesh grid size, the fill factor for any given measurement can vary from about 30\% to 75\%.
 The gray band in Figure \ref{figVG0912} represents the allowable range of predictions, taking into account both the uncertainty in the mesh depth and fill factor.
 If the mesh pattern were perfectly bonded, we would expect all the spectra to fall inside the color swath in Figure \ref{figVG0912}.
-We see excellent consistency between prediction and measurement, with one measured spectrum hinting at a small gap ($\lesssim15$) in one measurement position.
+We see excellent consistency between prediction and measurement, with one measured spectrum hinting at a small gap ($\lesssim15$ nm) in one measurement position.
 Outside the mesh area, the bonding is excellent, as demonstrated by the off-mesh transmission with transmission consistent with 100\% to within the measurement uncertainty of 0.2\%.
@@
 A key result of these test measurements is to demonstrate that gaps of known dimensions are recovered in our spectroscopy, both for large and small gaps.
-Table \ref{tbl_DerivedGapSizes} summarizes the outcome of some measurements in the areas with known gaps and fill factors.
+Table \ref{tbl_DerivedGapSizes} summarizes the outcome of measurements in the areas with known gaps and fill factors.
 The table lists the median sample and 68\% confidence intervals.
@@
 This dilution leads to a degeneracy between the axial extent of the gap and its fractional fill factor over the measurement beam,  one that is broken by the harder-to-measure curvature of the transmission spectrum.
-Figure \ref{VG0912_0gap_corner} shows illustrates the degeneracy.
+Figure \ref{VG0912_0gap_corner} illustrates the degeneracy.
 The red dashed line shows the normalized transmission spectrum of a 14 nm gap over 100\% of the measurement area.
@@
 
-\emph{The benefit of the Gaussian process-} We chose to estimate the gap size and fill factor with Gaussian process regression.
-Gaussian process regression is not necessary if random errors are much larger than the systematic errors, in which case any model fitting strategy---like least squares or over-plotting a model to data by-hand---will yield an unbiased estimate for the gap size.
-We tested the benefit of Gaussian process regression by generating synthetic transmission spectra with known gap size and fill factor, and known covariance structure.
+\emph{The benefit of the Gaussian process-} We chose to compare models to data with MCMC Gaussian process regression.
+Gaussian process regression is not necessary if random errors are much larger than the systematic errors, in which case any model fitting strategy---like least squares or over-plotting a model to data by-hand---will suffice.
+We quantified the benefit of Gaussian process regression by generating synthetic transmission spectra with known gap size and fill factor, and known covariance structure.
 We applied ordinary least squares regression to the synthetic spectra and derived values for $d$ and $f$ that were biased by up to $20\sigma$ or more.
@@
 This limitation will vary from instrument to instrument.
-Our instrument can achieve 10 times better accuracy and precision in double beam mode than in single beam mode.
+Our instrument can achieve 10 times better accuracy and precision in ``double beam'' mode than in ``single beam'' mode.
 However, the slightly different polarization sensitivity in the different beams caused systematic jumps at the lamp/grating change-over wavelengths.
@@
 
-MGS acknowledges support from Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.
+MGS acknowledges support from the NASA Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.
 Portions of this research were supported by NASA ACT grant NNX12AC31G.

### 2015-09-25 AO_bondedSi.tex
some typos and re-arrange sections 4, 5.  not sure if I love it tho.

@@
 
-\section{Fabricating and embedded gaps of known axial and lateral extent}
+\section{Embedded gaps of known axial and lateral extent}
 
@@
 
-\section{Measurement technique and correlated measurement errors}
-\label{sec_aboutErrors}
+\section{Data}
 
@@
 Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
+\input{tbls/tbl_cary5000Pars.tex}
 
-Measurement uncertainty can be characterized by both its \emph{bias} and \emph{variance} \cite{2013sdmm.book.....I}.
-The bias is equivalent to a mean-level drift, which may be attributable to monochromator lamp luminosity drift.
+\subsection{Comparing the Fabry-P\'{e}rot model to data}
+
+At this stage, \emph{any} regression technique can be employed to compare the Fabry-P\'{e}rot to data.  Least squares regression and other popular methods can be applied easily.
+
+Our measurement uncertainty is dominated by \emph{bias}, not \emph{variance} \cite{2013sdmm.book.....I}.
+The bias is similar to a mean-level drift, which may be attributable to monochromator lamp luminosity drift.
 The variance is the random sample-to-sample measurement residual remaining once the bias term is removed.
 In our experiments, the bias term has an amplitude of 0.2\%, whereas the variance has an amplitude of 0.02\%.
-Our measurement uncertainty is dominated by bias, not variance.
 
-\section{Comparing the Fabry-P\'{e}rot model to data}
-
-At this stage, \emph{any} regression technique can be employed to compare the Fabry-P model to data.
-Least squares regression and other popular methods can be applied easily.
-However, since least squares regression assumes zero bias, we elected to apply a Gaussian process regression technique that can handle biased measurements \cite{2013sdmm.book.....I,DFMgp,rasmussen2006gaussian}.
+Since least squares regression assumes zero bias, we instead elected to apply a Gaussian process regression technique that can handle biased data \cite{2013sdmm.book.....I,DFMgp,rasmussen2006gaussian}.
 Given the relatively small bias amplitude (0.2\%), least squares regression and Gaussian process regression might yield similar solutions for the gap size, but least squares regression would drastically underestimate the \emph{uncertainty} value associated with the derived value of gap size\cite{DFMgp}.
-
-\input{tbls/tbl_cary5000Pars.tex}
 

### 2015-09-30 AO_bondedSi.tex
Incorporate DTJ edits to my revisions from the referee comments, up to Section 3.

@@
 
-Crystalline silicon is an excellent optical material for the infrared.
+Crystalline silicon is an excellent material for infrared optics.
 Si has a high refractive index ($3.55-3.45$ from $\lambda = 1150-2500\;$nm at room temperature, \cite{2006SPIE.6273E..77F}) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.
-It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18$-$30 $\mu$m region as well \cite{doi:10.1117/12.323764}.
-There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.
+It also has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18$-$30 $\mu$m region \cite{doi:10.1117/12.323764}.
+There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching.
 
-As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.
-One advantage of this direct Si$-$Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
-A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because close lossless bonds can eliminate concerns about optical ghosts.
-The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
-Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choice of coatings even further.
+As with a number of glasses and crystalline optical materials, two planar or conformal silicon surfaces can form strong physical bonds.
+One advantage of this direct Si$-$Si bonding of optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.
+A sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because close lossless bonds can eliminate concerns about optical ghosts.
+Other advantages of bonding stem from the difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
+Systems often are used at cryogenic temperatures and problems with differential thermal expansion, brittleness and hygroscopic properties can restrict the choice of coatings even further.
 It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.
 
-There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \cite{1998AnRMS..28..215G,Masteika2014}.
+There are several good reviews of Si bonding that recount the history of the field and describe process alternatives, metrology, and problems \cite{1998AnRMS..28..215G,Masteika2014}.
 The foundation paper on Si bonding came in 1986 \cite{1986JAP....60.2987S}.
@@
 1989 and G{\"o}esele et al.
-\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.
+\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G} demonstrated gap-free bonding under a vacuum in a micro- cleanroom, in which bonding occurs immediately after a rinse-spin-vacuum cycle.
 Other authors investigated the growth of bubbles at bond interfaces with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of $800-1100\;^\circ$C \cite{Horn2009, Masteika2014}.
@@
 The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
-These results ultimately helped to flesh out a robust picture of the chemistry as a function of anneal temperature\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+A robust picture of the chemistry as a function of anneal temperature emerged \cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
 Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
@@
 If Si$-$Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than the concatenated losses of two antireflection coated surfaces over the operating wavelength (typically $< 4$\%).
-For Si$-$air gaps, the full Fresnel loss is $\sim30\%$ per surface, much larger than the $4-7\;\%$ per surface of the many optical materials with refractive indices in the range $1.5-1.7$.
-So a 4\% transmission loss specification translates to physical gaps or bubbles having an axial extent of merely $ < 35\;$nm.
+For Si$-$air gaps, the full Fresnel loss is $\sim30\%$ per surface, much larger than the $4-7\;\%$ per surface of optical materials with refractive indices $1.5-1.7$.
+A 4\% transmission loss specification for Si translates to physical gaps or bubbles with an axial extent of $ < 35\;$nm.
 If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond.
@@
 
-Most of the bonding literature concerns itself with wafer bonding, as opposed to thicker optical substrates.
+Most of the bonding literature concerns itself with wafer bonding, and not thicker optical substrates.
 In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.
-In optics, however, both of the bonding partners will likely be thicker and more mechanically rigid.
-When bonding two thick substrates, small differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.
+In optics, however, both of the bonding partners will be thicker and more rigid.
+When bonding two thick substrates, small differences in the shapes of the polished surfaces could lead to topological inconsistencies that prohibit some areas from bonding.
 If enough pressure is placed on the substrates and if the bond strength is high enough relative to the stiffness of one or the other component, however, the surfaces will conform and bond.
@@
 
-Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.
-Since silicon is opaque in the optical range, the most straightforward technique for detecting defects in the bond is to image them in the infrared.
+Verification that an optical bond meets its requirements and diagnosis of possible problems in the bonding technique require accurate measurement of the lateral and axial dimensions of any defects.
+Since silicon is opaque in the optical range, the most straightforward way to detect defects in the bond is to take infrared images.
 Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect.
-In cases where the axial extent of the gap is a half-wavelength or more, Newton rings appear in the image of the gap.
+Newton rings appear in the image where the axial extent of the gap is a half-wavelength or more.
 X-ray topography \cite{Mitani1990} exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not of the axial extent itself.
@@
 We use the wavelength dependence of the transmission and a model of the gap as a closely-spaced Fabry-P\'{e}rot etalon to measure axial gaps down to a few tens of nanometers in size.
-In order to verify the applicability and accuracy of the technique, we have used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
+In order to verify the applicability and accuracy of the technique, we used microlithography to create small artificial gaps of known size in wafers that we then bond and measure.
 
 For all types of gaps and bubbles, our technique needs to provide an accurate assessment of the transmission loss at the operating wavelength.
-For finite-sized bubbles, measurement of the axial dimension helps us to assess the effects of high-temperature annealing \cite{Horn2009, Masteika2014} on removal of hydrogen bubbles.
+For finite-sized bubbles, measurement of the axial dimension helps us to assess the efficacy of high-temperature annealing \cite{Horn2009, Masteika2014} in removal of hydrogen bubbles.
 A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to be resolved spatially, as one might expect if there are many very small contaminants present along the interface.
-Our technique must be able to characterize gaps with losses of a few percent at the operating wavelength.
+Our technique must be able to characterize gaps with losses of a few percent.
 A $2\%$ loss at $1.25\;\mu$m corresponds to a gap of $>25\;$nm at normal incidence and the gap size for a loss of this magnitude scales linearly with wavelength.
@@
 Our modeling technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.
-The refractive index of silicon is one of the highest of any commonly used optical material\cite{2006SPIE.6273E..77F}.
-We model a small Si$-$Si interface gap between two almost-bonded silicon parts as a low-Finesse Fabry-P\'{e}rot etalon \cite{2007fuph.book.....S}.
-The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse of the cavity.
+We model a small Si$-$Si interface gap between two almost-bonded parts as a low-Finesse Fabry-P\'{e}rot etalon \cite{2007fuph.book.....S}.
+The key property of an etalon is that it has a cavity enclosed between two smooth and parallel reflective surfaces where the requirements for smoothness and parallelism depend on the reflective finesse.
 For the narrow air gaps we consider here and for the relatively modest reflectivity of a Si$-$vacuum interface, most gaps at Si$-$Si bonds will form effective etalons over most of their area.
@@
 We define the coefficient of finesse\cite{2007fuph.book.....S}, $F$ (Equation \ref{Eq:coeffF}), in the customary way to encapsulate the Fabry-P\'{e}rot etalon's dependence on reflectivity.
-We calculate a coefficient of finesse in the range 2.5$-$2.6, depending on wavelength.
+$F$ has a value of 2.5$-$2.6, depending on wavelength.
 We assume room temperature for all calculations of refractive index.
@@
 The axial extent of the silicon cavities are large enough---and typical measurement spectral resolution low enough---that the the Si cavities reflect and transmit light incoherently at their entrance and exit interfaces.
-The multiple internal reflections within the Si therefore present a hurdle to an analysis of the air gaps.
 In the air gap, however, the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
@@
 
-which has an average value of about 53\%.  The top line in the upper panel of Figure \ref{figAbsoluteTrans} shows the plot of Equation \ref{eqnAbsDSPtrans} in the wavelength range $1200-2500\;$nm.
+which has an average value of about 53\%.  The top line in the upper panel of Figure \ref{figAbsoluteTrans} plots the value of $T_{DSP}$ from 1200 to 2500 nm.
 
@@
 $T_g$ is defined in Equation \ref{eqFP}.
-The only free parameters in Equation \ref{eqnMix} are $d$ and $f$.
-Both of these parameters are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
-Synthetic embedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secResults}).
+The only free parameters in Equation \ref{eqnMix} are $d$ (included in $T_g$) and $f$.
 
@@
 The substrate thicknesses ranged from 0.5 to 3.3 mm.
+Both $d$ and $f$ are fixed in our direct bonded Si wafers with synthetic gaps, and are listed in Table \ref{tbl_meshPatterns}.
+Synthetic embedded gaps therefore offer an excellent test of our ability to recover the gap axial extent $d$ and areal fill factor $f$ (Section \ref{secResults}).
 
@@
 For measurements taken with the Veeco NT9100 Optical Profiler, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{figVS20pattern}.
-The stylus profilometry measured values were assigned an uncertainty of 5\%, which accorded with our experience with other etched parts.
+The stylus profilometry measured values were assigned an uncertainty of 5\%, based on our experience with other etched parts.
 Table \ref{tbl_experiments} lists descriptions for our bonded substrates.
@@
 
-We cleaned the surfaces before bonding to minimize the interfacial particle density.
+To minimize the interfacial particle density, we cleaned the surfaces before bonding with standard cleaning procedures of solvents in a megasonic.
 The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
@@
 We found a typical peak to valley surface flatness of 2 waves ($\lambda=632.8\;$nm) over the central 50 mm diameter.
-We prepared the substrates with standard cleaning procedures of solvents in a megasonic.
-We then applied MHz frequency oxygen plasma ashing.
+After cleaning, we applied MHz frequency oxygen plasma ashing.
 We soaked the wafers in DI water then dried with N$_2$.

### 2015-09-30 AO_bondedSi.tex
Complete all edits from DTJ

@@
 No attempt at quantitative measurement was made on the IR images due to evidence for a non-linear detector response.
-When IR imaging was available, it was used to target IR spectroscopy toward locations displaying the presence or absense of IR bubbles, as desired.
+When IR imaging was available, it was used to target IR spectroscopy toward locations displaying the presence or absence of IR bubbles, as desired.
 In other cases when IR imaging was not available, spectra were taken at random positions on the wafer.
@@
 
-At this stage, \emph{any} regression technique can be employed to compare the Fabry-P\'{e}rot to data.  Least squares regression and other popular methods can be applied easily.
+At this stage, \emph{any} regression technique can be employed to compare the Fabry-P\'{e}rot model to data.  Least squares regression and other popular methods can be applied easily.
 
@@
 Since least squares regression assumes zero bias, we instead elected to apply a Gaussian process regression technique that can handle biased data \cite{2013sdmm.book.....I,DFMgp,rasmussen2006gaussian}.
-Given the relatively small bias amplitude (0.2\%), least squares regression and Gaussian process regression might yield similar solutions for the gap size, but least squares regression would drastically underestimate the \emph{uncertainty} value associated with the derived value of gap size\cite{DFMgp}.
+Given the relatively small bias amplitude (0.2\%), least squares regression and Gaussian process regression might yield similar solutions for the gap size, but least squares regression would drastically underestimate the \emph{uncertainty} associated with the derived value of gap size\cite{DFMgp}.
 
@@
 We estimated the $\sigma_i$ from the dispersion around mean-subtracted, detrended transmission spectra of DSP samples.
-We do not know $a$ nor $s$ and they are, in general, different for each measurement-- they are \emph{nuisance parameters}, and are ultimately marginalized out.
-They arise from the uninteresting properties of the spectrometer, and their interpretation is immaterial.
+We do not know $a$ or $s$ and they are, in general, different for each measurement-- they are \emph{nuisance parameters}, and are ultimately marginalized out.
+They arise from the behavior of the spectrometer, and their interpretation is immaterial.
 The value of $a$ should be close to the mean drift reported in Table \ref{tabCary5000pars}.
 The value of $s$ could be small (tens of nm) to capture local variation attributable to atmospheric air absorption, or large (hundreds or thousands of nm) to capture global features attributable to lamp drift.
-We allow both $a$ and $s$ to be free parameters in our fitting procedure.
+We allow both $a$ and $s$ to be free parameters.
 
@@
 The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physical measurement (Table \ref{tbl_experiments}).
-The spectrum in this region shows characteristic Fabry-P\'{e}rot fringes of constructive and destructive interference.
+The spectrum in this region shows characteristic Fabry-P\'{e}rot interference fringes.
 The only tunable parameters of the model are the gap axial extent, $d$ and fill factor $f$.
@@
 We acquired spatially resolved measurements of the pair of bonded parts VG09-12, (Tables \ref{tbl_experiments} and \ref{tbl_meshPatterns}).
-We translated the sample through the measurement beam and measured a transmission spectrum at 25 unique adjacent positions.
+We translated the sample through the measurement beam and measured a transmission spectrum at 25 adjacent positions.
 We chose a 2 mm step size, which is comparable to the beam width.
@@
 
-The colored dots in the position marked in Figure \ref{figVG0912_STA_illus} match the color of the spectra in Figure \ref{figVG0912}.
+The colored dots in the position marked in Figure \ref{figVG0912_STA_illus} match the colors of the spectra in Figure \ref{figVG0912}.
 The approximate size of the measurement beam is shown for scale.
@@
 \emph{More complex underlying gap size distributions-} One key assumption of our gap-as-etalon modeling strategy is that the distribution of gap sizes over the measurement area is bimodal- a gap of size $d$ or zero gap.
-This choice was simply a matter of computational simplicity, and we could easily account for any more complicated gap size distribution by inserting additional terms in Equation \ref{eqnMix}.
+This choice was a matter of computational simplicity, and we could easily account for any more complicated gap size distribution by inserting additional terms in Equation \ref{eqnMix}.
 For example, if there is a gradient across the bond, the emergent transmission spectrum will be an admixture of many gaps with various areal coverage fractions.
@@
 In IR images (Figure \ref{figVS2021_IR_image}), interfacial gaps appear as dark patches.
-These dark patches are darker at shorter wavelengths as evinced in Equation \ref{eqFP}, which shows that observations at short wavelengths have the most sensitivity to small gaps.
+These dark patches are darker at shorter wavelengths as we see from Equation \ref{eqFP}, which shows that observations at short wavelengths have the most sensitivity to small gaps.
 Therefore an IR imaging approach should use the shortest possible wavelength without getting too close to the Si absorption cutoff.

### 2015-09-30 AO_bondedSi.tex
Apply a few grammar corrections from grammarly

@@
 Other advantages of bonding stem from the difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.
-Systems often are used at cryogenic temperatures and problems with differential thermal expansion, brittleness and hygroscopic properties can restrict the choice of coatings even further.
+Systems often are used at cryogenic temperatures and problems with differential thermal expansion, brittleness, and hygroscopic properties can restrict the choice of coatings even further.
 It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.
@@
 The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.
-A robust picture of the chemistry as a function of anneal temperature emerged \cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
+A robust picture of the chemistry as a function of annealing temperature emerged \cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.
 Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.
 Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.
-Greco \emph{et al.} examined the role of surface topography in determining the the bonding strength of thick optical surfaces.
+Greco \emph{et al.} examined the role of surface topography in determining the bonding strength of thick optical surfaces.
 These authors showed that van der Waals interactions by themselves were consistent with their available experimental evidence, but they could not rule out other forces and phenomena.
@@
 For all types of gaps and bubbles, our technique needs to provide an accurate assessment of the transmission loss at the operating wavelength.
-For finite-sized bubbles, measurement of the axial dimension helps us to assess the efficacy of high-temperature annealing \cite{Horn2009, Masteika2014} in removal of hydrogen bubbles.
+For finite-sized bubbles, measurement of the axial dimension helps us to assess the efficacy of high-temperature annealing \cite{Horn2009, Masteika2014} in the removal of hydrogen bubbles.
 A wavelength-dependent transmission variation can also indicate the presence of large numbers of gaps that are too small to be resolved spatially, as one might expect if there are many very small contaminants present along the interface.
@@
 
-A pair of bonded silicon wafers or pucks with a small gap forms three coupled cavities with the air gap as the the central cavity.
-The axial extent of the silicon cavities are large enough---and typical measurement spectral resolution low enough---that the the Si cavities reflect and transmit light incoherently at their entrance and exit interfaces.
+A pair of bonded silicon wafers or pucks with a small gap forms three coupled cavities with the air gap as the central cavity.
+The axial extent of the silicon cavities are large enough---and typical measurement spectral resolution low enough---that the Si cavities reflect and transmit light incoherently at their entrance and exit interfaces.
 In the air gap, however, the gap size is small enough that the cavity it forms can always be treated as a coherent structure.
@@
 For the smallest ($35-60$ nm) gaps, there is a $2-6\%$ decrease in gap transmission from $2400\;$nm down to $1250\;$nm.
-The magnitude and wavelength dependence of this transmission spectrum provide information about the axial extent of the gap.
+The magnitude and wavelength-dependence of this transmission spectrum provide information about the axial extent of the gap.
 
@@
 The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.
-We also measured the large scale surface flatness of the pucks with an interferometer.
+We also measured the large-scale surface flatness of the pucks with an interferometer.
 We found a typical peak to valley surface flatness of 2 waves ($\lambda=632.8\;$nm) over the central 50 mm diameter.
@@
 We first looked for large gaps in our bonded substrate pairs detectable as ``IR bubbles'' \cite{1992JEMat..21..669M}, with IR imaging.
-The combination of the absorption of Si, the lamp spectrum and the detector responsivity led to an effective bandpass from approximately $1.15-1.6\;\mu$m.
+The combination of the absorption of Si, the lamp spectrum, and the detector responsivity led to an effective bandpass from approximately $1.15-1.6\;\mu$m.
 The field of view was smaller than the 100 mm diameter wafers, so we dithered the sample.
@@
 Figure \ref{figVG03full} shows a parameter determination for the measurement of a bond over a deep ($4000 \pm 200\;$nm) 100\% fill factor gap in part VG03.
-The analysis yields a best fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physical measurement (Table \ref{tbl_experiments}).
+The analysis yields a best-fit consistent with 100\% fill factor, and a depth of $3960 \pm 2\;$nm, well within the uncertainty of the physical measurement (Table \ref{tbl_experiments}).
 The spectrum in this region shows characteristic Fabry-P\'{e}rot interference fringes.
@@
 The limiting factor is spectrophotometric \emph{accuracy} and \emph{precision}.
-In this Section we question the assumptions of our technique, identify its limitations, and mention some opportunities.
+In this Section, we question the assumptions of our technique, identify its limitations, and mention some opportunities.
 
@@
 Specifically, for $d_{max}<\lambda/4$, our simple mixture model will converge on the spectrum of the average gap size, even if the \emph{True} distribution is more complex.
-Reducing the measurement area could break some of these degeneracies, if more precision is desired.
+Reducing the measurement area could break some of these degeneracies if more precision is desired.
 
 \emph{The benefit of the Gaussian process-} We chose to compare models to data with MCMC Gaussian process regression.
-Gaussian process regression is not necessary if random errors are much larger than the systematic errors, in which case any model fitting strategy---like least squares or over-plotting a model to data by-hand---will suffice.
+Gaussian process regression is not necessary if random errors are much larger than the systematic errors, in which case any model fitting strategy---like least squares or over-plotting a model to data by hand---will suffice.
 We quantified the benefit of Gaussian process regression by generating synthetic transmission spectra with known gap size and fill factor, and known covariance structure.
@@
 If the entrance and exit faces of the bonded samples are AR coated, the incoherent multiple reflection model will have to be modified.
-\emph{Equation \ref{eqn:Tetalon} is not valid if the entrance and exit Si interfaces are AR coated.}  Equation \ref{eqn:SiAirMatrix} would have to be altered with the the AR coating's wavelength-dependent transmission, which might not have a simple analytical form.
+\emph{Equation \ref{eqn:Tetalon} is not valid if the entrance and exit Si interfaces are AR coated.}  Equation \ref{eqn:SiAirMatrix} would have to be altered with the AR coating's wavelength-dependent transmission, which might not have a simple analytical form.
 
-\emph{How far can this technique go?}  The spectrophotometric precision of our instrument flows down directly to limit the smallest measureable gap size.
+\emph{How far can this technique go?}  The spectrophotometric precision of our instrument flows down directly to limit the smallest measurable gap size.
 This limitation will vary from instrument to instrument.
@@
 These dark patches are darker at shorter wavelengths as we see from Equation \ref{eqFP}, which shows that observations at short wavelengths have the most sensitivity to small gaps.
-Therefore an IR imaging approach should use the shortest possible wavelength without getting too close to the Si absorption cutoff.
+Therefore, an IR imaging approach should use the shortest possible wavelength without getting too close to the Si absorption cutoff.
 This work recommends a 1250$-$1300 nm filter.
@@
 IR images must be normalized to the transmission through a DSP sample under identical, nearly collimated lighting conditions.
-Any deficits in flux of a bonded sample can be converted to a gap size $d$, using Table \ref{tbl_FB1250}, if the FB1250-10 filter is used.
+Any deficits in transmission through a bonded sample can be converted to a gap size $d$, using Table \ref{tbl_FB1250}, if the FB1250-10 filter is used.
 IR imaging will lack the spectral information available in the spectra shown in this article.
@@
 Portions of this research were supported by NASA ACT grant NNX12AC31G.
-The authors thank Amanda Turbyfill, Mike Pavel, and Carleen Boyer for assistance with IR imaging and specroscopy.
+The authors thank Amanda Turbyfill, Mike Pavel, and Carleen Boyer for assistance with IR imaging and spectroscopy.
 MGS thanks the organizers and participants of the \emph{Astro Data Hack Week} (2014), especially Dan Foreman-Mackey for sharing tutorials and code for implementing the Gaussian process technique.

### 2015-09-30 referee_report/response_to_referee.txt
Clean up the repo structure by making referee report directory, which contains the .doc file.

From: Dr. Michael Gully-Santiago, on behalf of all coauthors
To: Anonymous referee and topical editor, Applied Optics
Date: Monday, September 21 - Friday September 25, 2015
Re: Response to referee report

Thank you for your referee report and detailed reading of our manuscript, "Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy".  Below I respond to the referee's feedback.

I corrected the accent mark in Fabry-Pérot for all instances of the name.

I corrected these grammatical errors:
- p5. Since silicon is opaque in the optical (range?), the...
- p6. are too small to (be) resolve(d) spatially

I removed the sentence "Our ultimate goal is to bond optics with thicknesses as large as 30 mm.", which we agree is too specific to our application.

I reduced the abstract length by about 40% to make it more concise.

I corrected the title of Section 3.

I entirely removed the paragraph in Section 4, "The Cary 5000 has four operation modes..."

I entirely removed the paragraph in Section 4, "The spectral sampling ranged from..."  (This uncontroversial information is in the nearby table.)

I added a discussion of least squares regression versus Gaussian process regression.

In regards to the comment, "For me Tdsp is missing in eq 5. ( Tmix= f Tg&#43;(1-f)Tdsp )", the referee is correct that the last term in eq. 5 represents the contribution of DSP wafer, however the Tdsp factor has been divided out as described in eq 4.  So the last term could be thought of as (1-f)*Tdsp/Tdsp = (1-f)*1.  The reviewers comment did, however, cause me to revise my description of equation 4 to clarify this point.

In regards to the comment, '5.B: the term "Sample Transport Accessory" is really not adapted...', I have removed the mention of the sample transport accessory, and instead simply described the process of stepping across the sample.

I have now defined temperature, T, before its first use, and have eliminated the Figure 1, which was a plot of the temperature dependence of the refractive index.

In response to the referee's comments on concision, I have cut many lines of text, and in some cases entire paragraphs that I agreed were immaterial to the main idea of the paper.  The manuscript lost at least 2 pages out of the raw 32.  I have motivated more clearly the interest in the MCMC formalism over a conventional least squares technique, now located in section 4- "comparing models to data."  Indeed, I agree that a least squares technique can be applied, but I also note the advantages of the Gaussian process regression technique.  I have posted several references to help the reader reproduce the Gaussian Process regression technique if desired.

Upon inspecting this final submission, it is clear that the manuscript has been much improved from the reviewer's comments, particularly with a clearer 'flow' to the paper.

Best regards,
Dr. Gully-Santiago

### 2015-10-11 MGS_AO_archive/AO_bondedSi.tex
first commit of the two-column version in the latex archive.  begin formatting for ArXiV

--- 2015-09-30 AO_bondedSi.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\documentclass[osajnl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1} %% use 11pt for Applied Optics
-\documentclass[osajnl,preprint,showpacs,superscriptaddress,12pt]{revtex4-1} %% use 12pt for preprint option
+\documentclass[osajnl,twocolumn,showpacs,superscriptaddress,11pt]{revtex4-1} %% use 11pt for Applied Optics
+%\documentclass[osajnl,preprint,showpacs,superscriptaddress,12pt]{revtex4-1} %% use 12pt for preprint option
 \usepackage{amsmath,amssymb,graphicx}
@@
 \begin{figure}[htbp]
-\includegraphics[width=0.40\columnwidth]{figs/fpAbsorbfig_alt}
+\includegraphics[width=0.40\columnwidth]{fpAbsorbfig_alt}
 \caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
@@
 \begin{figure}[!htbp]
-\includegraphics[width=0.40\columnwidth]{figs/etalon_trans.pdf}
+\includegraphics[width=0.40\columnwidth]{etalon_trans.pdf}
 \caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of vacuum wavelength for various gap sizes in bonded Si optics.
@@
 \begin{figure}[htbp]
-\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}
+\includegraphics[width=1.0\columnwidth]{VS20fineGapCrop.pdf}
 \caption{
@@
 
-\input{tbls/tbl_experiments.tex}
+\begin{table}[h!]
+\caption{UTexas Si bonding experiments \label{tbl_experiments}}
+\begin{center}
+    \begin{tabular}{ c c c c c}
+    \hline
+    Name & Diameter & Thickness & Pattern & Pattern Depth\\ 
+    -  & mm & mm & - & nm \\
+        \hline
+    VG02   & 100 & 3.3 $\times$ 2 &  None  & - \\
+    VG03   & 100 & 3.3 $\times$ 2 &  Hole \& Petal & 4000$\pm$200 \\
+    VG09-12   & 75   & 1 $\times$ 2 & Mesh C & 49 $\pm$6 \\
+    VS20-21   & 100 & 0.8 $\times$ 2 &  Mesh C, M, F & 95 $\pm$5 \\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
 
-\input{tbls/tbl_meshPatterns.tex}
+\begin{table}[h!]
+\caption{Patterned gap properties \label{tbl_meshPatterns}}
+\begin{center}
+    \begin{tabular}{ c c c c }
+    \hline
+    Pattern & Fill Factor & feature size & bulk description \\ 
+    - & \% & $\mu$m & - \\ 
+    \hline
+    Hole   & 100     &  -         & 25 mm diameter hole \\     
+    Petal  & $\sim5$ & $\sim$2000 & 2 mm wide lines \\         
+    Mesh F & 50      &         40 & 100 $\mu$m square holes, Fig. \ref{figVS20pattern}\\ 
+    Mesh M & 50      & 200        & 500 $\mu$m square holes\\ 
+    Mesh C & 50      & 620        & 1500 $\mu$m square holes\\     
+        \hline
+    \end{tabular}
+\end{center}
+\end{table}
 
@@
 \begin{figure}[htbp]
-\includegraphics[width=0.4\columnwidth]{figs/VS2021_IR_image.pdf}
+\includegraphics[width=0.4\columnwidth]{VS2021_IR_image.pdf}
 \caption{
@@
 Table \ref{tabCary5000pars} lists the typical measurement settings and tool performance.
-\input{tbls/tbl_cary5000Pars.tex}
+
+\begin{table}[h!]
+\caption{Summary of Cary 5000 measurement parameters \label{tabCary5000pars}}
+\begin{center}
+\begin{tabular}{ c c }
+\hline
+        Parameter (Units) & Value \\ 
+\hline
+        Spectral sampling interval (nm) & 2.0,10.0 \\
+        Spectral resolution (nm) & 5 \\
+        Typical measurement range (nm) & 1200-2500 \\
+		Mean level drift uncertainty (\%) & 0.20 \\
+		Random uncertainty (\%) & 0.02 \\
+ 		Beam size (mm $\times$ mm) & $2 \times 8$ \\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\input{tbls/tbl_DerivedGapSizes}
+\begin{table}[h!]
+\caption{Inferred gap sizes and fill factors \label{tbl_DerivedGapSizes}}
+\begin{center}
+    \begin{tabular}{ c c c c c }
+    \hline
+    Name & Region & Predicted $d$ & Measured $d$ & $f$ \\
+    -  & - & nm & nm & - \\
+    \hline
+    DSP & -    &   0.0  & $3^{+8}_{-2}$ & $0.260^{+0.456}_{-0.230}$\\
+    VG09-12 & Off-Mesh    &   0.0  & $11^{+21}_{-7}$ & $0.25^{+0.46}_{-0.22}$\\
+    VG03 & Hole    &   4000 $\pm$ 200  & $3960^{+2}_{-2}$ &  $0.999^{+0.001}_{-0.002}$\\
+    VG03 & Petal   &   4000 $\pm$ 200  & $4094^{+4}_{-4}$ &  $0.046^{+0.001}_{-0.001}$\\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
 
@@
         \begin{subfigure}[b]{0.5\textwidth}
-              \includegraphics[width=\textwidth]{figs/VG03_corner.pdf}
+              \includegraphics[width=\textwidth]{VG03_corner.pdf}
               \caption{Fitted gap size $d$ and fill factor $f$ for VG03 hole region}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG03_f100.pdf}
+                \includegraphics[width=\textwidth]{VG03_f100.pdf}
                 \caption{Measured spectrum and fit of VG03 hole region}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-              \includegraphics[width=\textwidth]{figs/VG03p2_corner.pdf}
+              \includegraphics[width=\textwidth]{VG03p2_corner.pdf}
               \caption{Fitted gap size $d$ and fill factor $f$ for VG03 petal region}
@@
         \begin{subfigure}[b]{0.5\textwidth}
-                \includegraphics[width=\textwidth]{figs/VG03_f045.pdf}
+                \includegraphics[width=\textwidth]{VG03_f045.pdf}
                 \caption{Measured spectrum and fit of VG03 petal region}
@@
     \begin{subfigure}[b]{0.35\textwidth}
-        \includegraphics[width=\textwidth]{figs/VG09_12_STA_illustration.pdf}
+        \includegraphics[width=\textwidth]{VG09_12_STA_illustration.pdf}
         \caption{Illustration of VG09-12 sample transport accessory measurement positions. \label{figVG0912_STA_illus} }
@@
     \begin{subfigure}[b]{0.65\textwidth}
-        \includegraphics[width=\textwidth]{figs/VG0912_STA_scan_coarse.pdf}
+        \includegraphics[width=\textwidth]{VG0912_STA_scan_coarse.pdf}
         \caption{Measured gap transmission spectra of VG09-12 on- and off- mesh \label{figVG0912}}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{figs/VG0912_0gap_corner.pdf}
+        \includegraphics[width=\textwidth]{VG0912_0gap_corner.pdf}
         \caption{Fitted gap size $d$ and fill factor $f$ for VG09-12 off-mesh}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{figs/VG0912_0gap.pdf}
+        \includegraphics[width=\textwidth]{VG0912_0gap.pdf}
         \caption{Measured spectrum of VG09-12}
@@
 \begin{figure}[htbp]
-\includegraphics[width=0.8\columnwidth]{figs/F1250_10_filter.pdf}
+\includegraphics[width=0.8\columnwidth]{F1250_10_filter.pdf}
 \caption{\label{figFB1250-10} Filter curve and gap spectra.
@@
 
-\input{tbls/tbl_FB1250.tex}
+\begin{table}[h!]
+\caption{FB1250-10 integrated gap transmission \label{tbl_FB1250}}
+\begin{center}
+    \begin{tabular}{ r r}
+    \hline
+    $d$ & $T_{1250}$\\
+    nm  & - \\
+    \hline
+      0 &                   1.000 \\
+      8 &                   0.998 \\
+     16 &                   0.991 \\
+     24 &                   0.981 \\
+     32 &                   0.966 \\
+     40 &                   0.948 \\
+     48 &                   0.928 \\
+    \hline
+    \end{tabular}
+\end{center}
+\end{table}
 
@@
 \begin{figure}[htbp]
-\includegraphics[width=0.8\columnwidth]{figs/FB1250-10_integ_trans.pdf}
+\includegraphics[width=0.8\columnwidth]{FB1250-10_integ_trans.pdf}
 \caption{\label{figFB1250-10_integ} Integrated transmission through the FB1250-10 filter as a function of interface gap size.
@@
 
+The authors thank the anonymous referee, whose comments improved the manuscript.
 MGS acknowledges support from the NASA Graduate Student Research Program (GSRP) Fellowship, grant NNX10AK82H.
@@
 \bibliographystyle{osajnl}
-\bibliography{AO_bondedSi}
+%\bibliography{AO_bondedSi}
 
-%\begin{thebibliography}{99}
+\begin{thebibliography}{10}
+\newcommand{\enquote}[1]{``#1''}
 
-%% Do not include separate BibTeX files; if BibTeX is used,
-%% paste the output (contents of .bbl file) here.
+\bibitem{2006SPIE.6273E..77F}
+B.~J. {Frey}, D.~B. {Leviton}, and T.~J. {Madison},
+  \enquote{{Temperature-dependent refractive index of silicon and germanium},}
+  (2006), vol. 6273 of \emph{Society of Photo-Optical Instrumentation Engineers
+  (SPIE) Conference Series}.
 
-%\end{thebibliography}
+\bibitem{PhysRev.108.268}
+W.~Spitzer and H.~Y. Fan, \enquote{Infrared absorption in $n$-type silicon,}
+  Phys. Rev. \textbf{108}, 268--271 (1957).
+
+\bibitem{PhysRev.78.178}
+H.~Y. Fan and M.~Becker, \enquote{Infra-red absorption of silicon,} Phys. Rev.
+  \textbf{78}, 178--179 (1950).
+
+\bibitem{doi:10.1117/12.323764}
+J.~E. Peters, P.~D. Ownby, C.~R. Poznich, and J.~C. Richter,
+  \enquote{Far-infrared absorption of czochralski germanium and silicon,} Proc. SPIE
+  \textbf{3424}, 98--105 (1998).
+
+\bibitem{2012SPIE.8450E..2TV}
+A.~H. {van Amerongen}, T.~{Agocs}, H.~{van Brug}, G.~{Nieuwland}, L.~{Venema},
+  and R.~W.~M. {Hoogeveen}, \enquote{{Development of silicon immersed grating
+  for METIS on E-ELT},}  (2012), vol. 8450 of \emph{Society of Photo-Optical
+  Instrumentation Engineers (SPIE) Conference Series}.
+
+\bibitem{2010SPIE.7739E.123G}
+M.~{Gully-Santiago}, W.~{Wang}, C.~{Deen}, D.~{Kelly}, T.~P. {Greene},
+  J.~{Bacon}, and D.~T. {Jaffe}, \enquote{{High-performance silicon grisms for
+  1.2-8.0 {$\mu$}m: detailed results from the JWST-NIRCam devices},}  (2010),
+  vol. 7739 of \emph{Society of Photo-Optical Instrumentation Engineers (SPIE)
+  Conference Series}.
+
+\bibitem{1998AnRMS..28..215G}
+U.~{G{\"o}sele} and Q.-Y. {Tong}, \enquote{{Semiconductor Wafer Bonding},}
+  Annual Review of Materials Science \textbf{28}, 215--241 (1998).
+
+\bibitem{Masteika2014}
+V.~Masteika, J.~Kowal, N.~S.~J. Braithwaite, and T.~Rogers, \enquote{A review
+  of hydrophilic silicon wafer bonding,} {ECS} Journal of Solid State Science
+  and Technology \textbf{3}, Q42--Q54 (2014).
+
+\bibitem{1986JAP....60.2987S}
+M.~{Shimbo}, K.~{Furukawa}, K.~{Fukuda}, and K.~{Tanzawa},
+  \enquote{{Silicon-to-silicon direct bonding method},} Journal of Applied
+  Physics \textbf{60}, 2987--2989 (1986).
+
+\bibitem{1988JaJAP..27L2364S}
+R.~{Stengl}, K.-Y. {Ahn}, and U.~{G{\"o}sele}, \enquote{{Bubble-Free Silicon
+  Wafer Bonding in a Non-Cleanroom Environment},} Japanese Journal of Applied
+  Physics \textbf{27}, L2364--L2366 (1988).
+
+\bibitem{1995ApPhL..67.3614G}
+U.~{G{\"o}sele}, H.~{Stenzel}, T.~{Martini}, J.~{Steinkirchner}, D.~{Conrad},
+  and K.~{Scheerschmidt}, \enquote{{Self-propagating room-temperature silicon
+  wafer bonding in ultrahigh vacuum},} Applied Physics Letters \textbf{67},
+  3614--3616 (1995).
+
+\bibitem{1989JaJAP..28L2141L}
+V.~{Lehmann}, K.~{Mitani}, R.~{Stengl}, T.~{Mii}, and U.~{G{\"o}sele},
+  \enquote{{Bubble-Free Wafer Bonding of GaAs and InP on Silicon in a
+  Microcleanroom},} Japanese Journal of Applied Physics \textbf{28},
+  L2141--L2143 (1989).
+
+\bibitem{Horn2009}
+G.~Horn, M.~Gabriel, J.~Lesniak, and T.~J. Mackin, \enquote{Detection of
+  interfacial gas bubbles in wafer bonded silicon with different surface
+  treatments,} J. Electrochem. Soc. \textbf{156}, H27 (2009).
+
+\bibitem{feijoo1994}
+D.~Feijo\`o, Y.~J. Chabal, and S.~B. Christman, \enquote{Silicon wafer bonding
+  studied by infrared absorption spectroscopy,} Applied Physics Letters
+  \textbf{65}, 2548--2550 (1994).
+
+\bibitem{1995ApPhA..61..101R}
+M.~{Reiche}, S.~{Hopfe}, U.~{G{\"o}sele}, and Q.~Y. {Tong}, \enquote{{Multiple
+  internal reflection spectroscopy of bonded silicon wafers},} Applied Physics
+  A: Materials Science \& Processing \textbf{61}, 101--105 (1995).
+
+\bibitem{1996JaJAP..35.2102R}
+M.~{Reiche}, S.~{Hopfe}, U.~{G{\"o}sele}, H.~{Strutzberg}, and Q.-Y. {Tong},
+  \enquote{{Characterization of Interfaces of Directly Bonded Silicon Wafers: A
+  Comparative Study of Secondary Ion Mass Spectroscopy Multiple Internal
+  Reflection Spectroscopy, and Transmission Electron Microscopy},} Japanese
+  Journal of Applied Physics \textbf{35}, 2102--2107 (1996).
+
+\bibitem{JJAP.37.4197}
+H.~Takagi, R.~Maeda, T.~R. Chung, N.~Hosoda, and T.~Suga, \enquote{Effect of
+  surface roughness on room-temperature wafer bonding by ar beam surface
+  activation,} Japanese Journal of Applied Physics \textbf{37}, 4197--4203
+  (1998).
+
+\bibitem{2001JOptA...3...85G}
+V.~{Greco}, F.~{Marchesini}, and G.~{Molesini}, \enquote{{Optical contact and
+  van der Waals interactions: the role of the surface topography in determining
+  the bonding strength of thick glass plates},} Journal of Optics A: Pure and
+  Applied Optics \textbf{3}, 85--88 (2001).
+
+\bibitem{1996ApPhL..68.2222T}
+H.~{Takagi}, K.~{Kikuchi}, R.~{Maeda}, T.~R. {Chung}, and T.~{Suga},
+  \enquote{{Surface activated bonding of silicon wafers at room temperature},}
+  Applied Physics Letters \textbf{68}, 2222--2224 (1996).
+
+\bibitem{2000JAP....88.4404H}
+W.~{Han}, J.~{Yu}, and Q.~{Wang}, \enquote{{Modeling the dynamics of Si wafer
+  bonding during annealing},} Journal of Applied Physics \textbf{88},
+  4404--4406 (2000).
+
+\bibitem{1995ApPhL..67..863G}
+U.~{G{\"o}sele}, S.~{Hopfe}, S.~{Li}, S.~{Mack}, T.~{Martini}, M.~{Reiche},
+  E.~{Schmidt}, H.~{Stenzel}, and Q.-Y. {Tong}, \enquote{{What determines the
+  lateral bonding speed in silicon wafer bonding?}} Applied Physics Letters
+  \textbf{67}, 863--865 (1995).
+
+\bibitem{2007ApOpt..46.6793H}
+J.~Haisma, N.~Hattu, J.~T. C. M.~D. Pulles, E.~Steding, and J.~C.~G. Vervest,
+  \enquote{Direct bonding and beyond,} Appl. Opt. \textbf{46}, 6793--6803
+  (2007).
+
+\bibitem{2001JAP....89.6013L}
+D.~A. {Litton} and S.~H. {Garofalini}, \enquote{{Modeling of hydrophilic wafer
+  bonding by molecular dynamics simulations},} Journal of Applied Physics
+  \textbf{89}, 6013--6023 (2001).
+
+\bibitem{Mitani1990}
+K.~Mitani, V.~Lehmann, and U.~Gosele, \enquote{Bubble formation during silicon
+  wafer bonding: causes and remedies,} in \enquote{Solid-State Sensor and
+  Actuator Workshop, 1990. 4th Technical Digest., IEEE,}  (1990), pp. 74--77.
+
+\bibitem{doi:10.1080/02786828608959094}
+D.~W. Cooper, \enquote{Particulate contamination and microelectronics
+  manufacturing: An introduction,} Aerosol Science and Technology \textbf{5},
+  287--299 (1986).
+
+\bibitem{2000RScI...71.1869G}
+A.~C. {Gracias}, C.~{Kuranaga}, J.~R. {Senna}, and M.~D. {Silva},
+  \enquote{{Photoacoustic imaging of voids in direct wafer bonding},} Review of
+  Scientific Instruments \textbf{71}, 1869--1872 (2000).
+
+\bibitem{2001opt4.book.....H}
+E.~{Hecht}, \emph{{Optics 4th edition}} (Addison-Wesley, 2001).
+
+\bibitem{2007fuph.book.....S}
+B.~E.~A. {Saleh} and M.~C. {Teich}, \emph{{Fundamentals of Photonics: Second
+  Edition}} (John Wiley, 2007).
+
+\bibitem{2002ApOpt..41.3978K}
+C.~C. Katsidis and D.~I. Siapkas, \enquote{General transfer-matrix method for
+  optical multilayer systems with coherent, partially coherent, and incoherent
+  interference,} Appl. Opt. \textbf{41}, 3978--3987 (2002).
+
+\bibitem{1992JEMat..21..669M}
+K.~Mitani and U.~G\"osele, \enquote{Wafer bonding technology for
+  silicon-on-lnsulator applications: A review,} Journal of Electronic Materials
+  \textbf{21}, 669--676 (1992).
+
+\bibitem{2013sdmm.book.....I}
+{\.Z}.~{Ivezi{\'c}}, A.~{Connolly}, J.~{VanderPlas}, and A.~{Gray},
+  \emph{{Statistics, Data Mining, and Machine Learning in Astronomy}} (2013).
+
+\bibitem{DFMgp}
+D.~{F}oreman Mackey, \enquote{Lab section for a course on gaussian process
+  regression,} \url{https://github.com/dfm/gp} (2014).
+
+\bibitem{rasmussen2006gaussian}
+C.~Rasmussen, \emph{Gaussian processes for machine learning} (MIT Press,
+  Cambridge, Mass, 2006).
+
+\bibitem{Note1}
+\protect \url {https://github.com/dfm/emcee}.
+
+\bibitem{emcee}
+D.~{Foreman-Mackey}, D.~W. {Hogg}, D.~{Lang}, and J.~{Goodman}, \enquote{emcee:
+  The mcmc hammer,} PASP \textbf{125}, 306--312 (2013).
+
+\bibitem{Note2}
+\protect \url {www.thorlabs.com}.
+
+\bibitem{Hunter:2007}
+J.~D. Hunter, \enquote{Matplotlib: A 2d graphics environment,} Computing In
+  Science \& Engineering \textbf{9}, 90--95 (2007).
+
+\bibitem{PER-GRA:2007}
+F.~P\'erez and B.~E. Granger, \enquote{{IP}ython: a system for interactive
+  scientific computing,} Computing in Science and Engineering \textbf{9},
+  21--29 (2007).
+
+\bibitem{Note3}
+\protect \url {https://github.com/Echelle/AO_bonding_paper}.
+
+\end{thebibliography}
 

### 2015-11-04 MGS_AO_archive/AO_bondedSi.tex
change table and figure formatting for two columns.
Fix equation length, reposition figures in place, submit to AstroPH

@@
+\pdfoutput=1
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[htbp]
-\includegraphics[width=0.40\columnwidth]{fpAbsorbfig_alt}
+\begin{figure}[!htbp]
+\includegraphics[width=0.95\columnwidth]{fpAbsorbfig_alt}
 \caption{Ratio of transmission of a 0.5 mm thick Si wafer to that of a 3.0 mm thick Si puck.
@@
 \begin{figure}[!htbp]
-\includegraphics[width=0.40\columnwidth]{etalon_trans.pdf}
+\includegraphics[width=0.85\columnwidth]{etalon_trans.pdf}
 \caption{\label{figAbsoluteTrans}The family of normal-incidence curves for transmission as a function of vacuum wavelength for various gap sizes in bonded Si optics.
@@
 The absolute transmission through Si depends on wavelength since the refractive index depends minutely on wavelength.
-The curves assume that multiple reflections within the Si substrates add incoherently as discussed in detail in Appendix \ref{sec:Append-IMRTMM}.
 \emph{Middle panel}: The ``Gap Transmission'', $T_g$.
@@
 \emph{Bottom panel}: ``Gap Transmission'' for larger gaps, spanning 0$-$3500 nm.
-The 3500 nm sized gap (cyan solid line), demonstrates characteristic Fabry-P\'{e}rot etalon fringes.}
+The 3500 nm sized gap (cyan solid line), demonstrates Fabry-P\'{e}rot etalon fringes.}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-\includegraphics[width=1.0\columnwidth]{VS20fineGapCrop.pdf}
+\begin{figure*}[!htbp]
+\includegraphics[width=0.75\textwidth]{VS20fineGapCrop.pdf}
 \caption{
@@
 The purpose for these patterns was to have gaps of known axial extent against which we could test the optical metrology technique we describe in this article.}
-\end{figure}
+\end{figure*}
 
-\begin{table}[h!]
+\begin{table*}[!htbp]
 \caption{UTexas Si bonding experiments \label{tbl_experiments}}
@@
 \end{center}
-\end{table}
+\end{table*}
 
-\begin{table}[h!]
+\begin{table*}[!htbp]
 \caption{Patterned gap properties \label{tbl_meshPatterns}}
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[htbp]
-\includegraphics[width=0.4\columnwidth]{VS2021_IR_image.pdf}
+\begin{figure*}[!htbp]
+\includegraphics[width=0.5\textwidth]{VS2021_IR_image.pdf}
 \caption{
@@
 The three dark spots are bubbles at the interface of the direct bonded Si samples.}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}[h!]
+\begin{table}[!htbp]
 \caption{Summary of Cary 5000 measurement parameters \label{tabCary5000pars}}
@@
 
-\begin{equation}
-	\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma_i)} \propto \ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
-\end{equation}
+\begin{eqnarray}
+	& &\ln{p(d,f,a,s|\lambda, T_{obs}, \sigma_i)} \propto  \nonumber\\
+	& &\,\,\,\ln{p(d,f,a,s)} -\frac{1}{2}\;\boldsymbol{r^\intercal}\boldsymbol{C^{-1}}\boldsymbol{r} -\frac{1}{2}\;\ln{\det{\boldsymbol{C}}} \label{eqnPosterior}
+\end{eqnarray}
 
@@
 
-\begin{table}[h!]
+\begin{table*}[!htbp]
 \caption{Inferred gap sizes and fill factors \label{tbl_DerivedGapSizes}}
@@
 \end{center}
-\end{table}
+\end{table*}
 
-\begin{figure}[htbp]
+\begin{figure}[!htbp]
         \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!htbp]
         \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!htbp]
     \centering
@@
     \end{subfigure}
-    \begin{subfigure}[b]{0.65\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \includegraphics[width=\textwidth]{VG0912_STA_scan_coarse.pdf}
@@
 
-\begin{figure}[htbp]
-\includegraphics[width=0.8\columnwidth]{F1250_10_filter.pdf}
+\begin{figure}[!htbp]
+\includegraphics[width=0.95\columnwidth]{F1250_10_filter.pdf}
 \caption{\label{figFB1250-10} Filter curve and gap spectra.
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
 \caption{FB1250-10 integrated gap transmission \label{tbl_FB1250}}
@@
 
-\begin{figure}[htbp]
-\includegraphics[width=0.8\columnwidth]{FB1250-10_integ_trans.pdf}
+\begin{figure}[!htbp]
+\includegraphics[width=0.95\columnwidth]{FB1250-10_integ_trans.pdf}
 \caption{\label{figFB1250-10_integ} Integrated transmission through the FB1250-10 filter as a function of interface gap size.
